---
title: "Simulating from a GLM"
author: "Carter Sevick"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(MASS)
```

In a generalized linear model (GLM) we have three main components, which are?




In the case of a logistic regression we have: 

1. a random variable (our outcome)
$$y_i\sim Bernouli(p_i), i.i.d.$$ 




2. a linear predictor
$$\nu_i= \beta_0 + \beta_1x_{1i}+\dots$$

3. the link function
$$log(\frac{p_i}{1-p_1}) = \nu_i$$

and its inverse:

$$p_i = \Phi(\nu_i) = \frac{1}{1+e^{-\nu_i}}$$




To get inspiration for a simulation let's build a model for the effect of smoking in pregnancy on low birth weight delivery using the birth weight data in the MASS package:

```{r}
lb_fit <- glm(low~smoke, data = MASS::birthwt, family = binomial())
s_lb_fit <- summary(lb_fit)
s_lb_fit 

# inverse logit function:
invlogit <- function(v) 1/(1+exp(-v))

#risk in non-smokers:
invlogit(-1.0871)

#risk in smokers:
invlogit(-1.0871 + 0.7041)

# check
by(MASS::birthwt$low, MASS::birthwt$smoke, function(x){mean(x)})

# proportion of smokers
mean(MASS::birthwt$smoke)
```


Construct the data generating mechanism

```{r}
sim_lowbwt <- function(n = 189 # sample size to generate
                       , B = c(-1.0871, 0.7041) # parameter effects for the distribution
                       , p_e = 0.4 # proportion of subjects to generate as exposed
                       ) {
  # generate exposure
  x <- ifelse(
    qnorm(1-p_e) < rnorm(n=n) # strange... but you will see why...
    , 1
    , 0)
  
  # assemble the predictor model matrix
  x_matrix <- cbind(intercept = rep(1, n), x1 = x)
  
  # the linear predictor
  v = as.vector(x_matrix %*% B)
  
  # generate the outcome
  low = sapply(v, function(x){rbinom(1,1,invlogit(x))})
  
  # return a simulated data set
  simdat <- data.frame(
    low = low,
    smoke = x
  )

  return(simdat)
}

```

Build a simulation to test glm()

```{r}

sim <- function(N = 1000 # number of simulation iterations
              , n = 189 # size of each simulation dataset
              , B = c(-1.0871, 0.7041) # true model parameters
              , p_e = 0.4 # probability of exposure (being a smoker)
              , n_vars # number of variables 1:n_vars to model
              , alpha = 0.05 
                ) {
  
  # generate N simulated data sets and run the model to be tested
  if(missing(n_vars)) {n_vars <- length(B)}
  res <- list()
  for(i in 1:N) {
    dat <- sim_lowbwt(n = n, B = B, p_e = p_e)
    vnames <- names(dat)
    res[[i]] <- glm(paste(vnames[1],'~', paste(vnames[2:n_vars], collapse = '+')), data = dat, family = binomial())
  }
  
  # performance
  bias <- list()
  cover <- list()
  power <- list()
  for(i in 1:N) {
    bias[[i]]  <- (coef(res[[i]]) - B) / B
    ci         <- confint.default(res[[i]] , level = 1-alpha)
    cover[[i]] <- ci[,1]  <= B  & B  <= ci[,2] 
    power[[i]] <- summary(res[[i]])$coefficients[,4] < alpha
  }
  bias  <- colMeans(do.call('rbind', bias))
  cover <- colMeans(do.call('rbind', cover))
  power <- colMeans(do.call('rbind', power))
   
  return(data.frame(bias = bias, cover = cover, power = power))
  
}

```

Bias and coverage probability for the effect of smoking (at simulation settings)
```{r}
set.seed(54321)
simu1 <- sim( )
simu1
 
```


an alternative to the smoking effect
```{r}
set.seed(54321)
simu1 <- sim( B = c(-1.0871, 0.9041) )
simu1
 
```


Construct the data generating mechanism, with a covariate

```{r}
sim_lowbwt2 <- function( n = 189 # size of each simulation dataset
                     , B = c(-1.0871, 0.7041, 0)# true model parameters
                     , p_e = 0.4         # probability of exposure (being a smoker)
                     , cov_mu = c(0, 0)  # means of x variables
                     , cov_var = diag(2) # covariance matrix of X variables
                     ) {
  
  # add generation of multiple, possibly correlated variables using the multivariate normal distribution (first variable still dichotomous)
  x <- MASS::mvrnorm(n=n, mu = cov_mu, Sigma = cov_var)
  x[,1] <- ifelse(
    qnorm(1-p_e) < x[,1]  
    , 1
    , 0)
  
  x_matrix <- cbind(intercept = rep(1, n), x)
  
  v = as.vector(x_matrix %*% B)
  low = sapply(v, function(x){rbinom(1,1,invlogit(x))})
  
  simdat <- as.data.frame(cbind(low, x))
  vnames <- c('low','smoke')
  if(length(B)>2) {vnames <- c(vnames,paste0('x',2:(length(B)-1)))}
  names(simdat) <- vnames
  
  return(simdat)
}

testit <- sim_lowbwt2()
```


Expand the simulation to test glm()

```{r}

sim2 <- function(N = 1000 # number of simulation iterations
              , n = 189 # size of each simulation dataset
              , B = c(-1.0871, 0.7041, 0) # true model parameters
              , p_e = 0.4 # probability of exposure (being a smoker)
              , n_vars # number of variables 1:n_vars to model
              , cov_mu = c(0, 0)  # means of x variables
              , cov_var = diag(2) # covariance matrix of X variables
              , alpha = 0.05

                ) {
  if(missing(n_vars)) {n_vars <- length(B)}
  res <- list()
  for(i in 1:N) {
    dat <- sim_lowbwt2(n = n, B = B, p_e = p_e, cov_mu = cov_mu, cov_var = cov_var)
    vnames <- names(dat)
    res[[i]] <- glm(paste(vnames[1],'~', paste(vnames[2:n_vars], collapse = '+')), data = dat, family = binomial())
  }
  
  # performance
  bias <- list()
  cover <- list()
  power <- list()
  for(i in 1:N) {
    bias[[i]]  <- (coef(res[[i]]) - B[1:n_vars])/ B[1:n_vars]
    ci         <- confint.default(res[[i]] , level = 1-alpha)
    cover[[i]] <- ci[,1]  <= B[1:n_vars]  & B[1:n_vars]  <= ci[,2] 
    power[[i]] <- summary(res[[i]])$coefficients[,4] < alpha
  }
  bias  <- colMeans(do.call('rbind', bias))
  cover <- colMeans(do.call('rbind', cover))
  power <- colMeans(do.call('rbind', power))
   
  return(data.frame(bias = bias, cover = cover, power = power))
  
}

```

```{r}

set.seed(54321)
simu2 <- sim2(n_vars = 2) 
simu2
```

 


Add correlation between smoke and x2
```{r}

set.seed(54321)
simu3 <- sim2(n_vars = 2, B = c(-1.0871, 0.7041, 0.25), cov_var = matrix(c(1,.75,.75,1), 2, 2)) 
simu3
```



QUESTION: do you think the increase in power for the effect of smoking if fair?




Add correlation between smoke and x2, but correct for x2
```{r}


simu4 <- sim2(n_vars = 3, B = c(-1.0871, 0.7041, 0.25), cov_var = matrix(c(1,.75,.75,1), 2, 2)) 
simu4
```


QUESTION: power for the effect of smoking is less than in simu2, why?

QUESTION: could this measure of power be used for study design?

