---
title: "\\large Assessing the Performance of Different Variable Selection Methods"
author: Diana Felipe
output: 
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    latex_engine: xelatex
    extra_dependencies: ["longtable", "float"]
mainfont: Arial
fontsize: 12pt
geometry: margin=1in
header-includes:
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage{float}
  - \usepackage{caption}
  - \usepackage{titling}
  - \usepackage{enumitem}
  - \setlength{\droptitle}{-4cm}
  - \preauthor{\vspace{-0.6cm}}
  - \postauthor{\vspace{-2.3cm}}
  - \setlist{nosep}
bibliography: Felipe_Proj4.bib
---


```{r, echo = F, warning=F, message = F}
knitr::opts_chunk$set(echo = TRUE)

library(tinytex)
library(readr)
library(knitr)
library(ggplot2)
library(kableExtra)
library(stats)
library(tibble)
library(labelled)
library(see)
library(here)
library(expss)
library(hdrm)
library(MASS) # AIC/BIC variable selection
library(olsrr) # Backwards variable selection
library(glmnet) # LASSO
library(tidyverse)
library(kableExtra)
library(here)
library(dplyr)
library(tidyr)

opts_chunk$set(tidy = F)
```

```{r, echo = F}
# Read in the rds file from the simulation run
simu_results <- readRDS("simulation_results_2500.rds")

# Source the functions from Project4.R
source("../Code/Project4_Functions.R")
```

```{r, echo = F}
##### MODEL PERFORMANCE
# Create data frame to assess multiple methods
sim_model_performance <- list()

# For the 2 profiles
for (p in seq_along(simu_results)) {
  # Get the pth profile summary
  profile_summary <- summarize_profiles(simu_results[[p]])
  # The pth profile info
  profile_info <- profiles[p, ]
  
  # List of all model performance for all selection methods
  model_perf_list <- list(
    Backward = profile_summary$backward$model_performance,
    AIC = profile_summary$AIC$model_performance,
    BIC = profile_summary$BIC$model_performance,
    LASSO_lambda.1se = profile_summary$lasso_lambda.1se$model_performance,
    LASSO_lambda.min = profile_summary$lasso_lambda.min$model_performance,
    Elastic_lambda.1se = profile_summary$elastic_net_lambda.1se$model_performance,
    Elastic_lambda.min = profile_summary$elastic_net_lambda.min$model_performance
  )
  
  # Create a data frame
  sim_model_performance[[p]] <- do.call(
    # Combine into rows
    rbind,
    # Create data frame with the cols of interst
    lapply(names(model_perf_list), function(m) {
      data.frame(
        n = profile_info$N,
        rho = profile_info$rho,
        method = m,
        model_perf_list[[m]]
      )
    })
  )
}

model_performance <- do.call(rbind, sim_model_performance)
row.names(model_performance) <- NULL

# Adjust the order of the variables
# Make variable a factor
model_performance$variable <- factor(model_performance$variable,
                                         levels = paste0("X", 1:20))
# Sort
model_performance <- model_performance[order(model_performance$variable,
                                                     model_performance$method), ]

### VARIABLE SELECTION METHOD PERFORMANCE
# Create data frame to assess multiple methods
sim_selection_performance <- list()

# For the 2 profiles
for (p in seq_along(simu_results)) {
  # Get the pth profile summary
  profile_summary <- summarize_profiles(simu_results[[p]])
  # The pth profile info
  profile_info <- profiles[p, ]
  
  # List of all varsel performance for all selection methods
  varsel_perf_list <- list(
    Backward = profile_summary$backward$selection_performance,
    AIC = profile_summary$AIC$selection_performance,
    BIC = profile_summary$BIC$selection_performance,
    LASSO_lambda.1se = profile_summary$lasso_lambda.1se$selection_performance,
    LASSO_lambda.min = profile_summary$lasso_lambda.min$selection_performance,
    Elastic_lambda.1se = profile_summary$elastic_net_lambda.1se$selection_performance,
    Elastic_lambda.min = profile_summary$elastic_net_lambda.min$selection_performance
  )
  
  # Create a data frame
  sim_selection_performance[[p]] <- do.call(
    # Combine into rows
    rbind,
    # Create data frame with the cols of interst
    lapply(names(varsel_perf_list), function(m) {
      data.frame(
        n = profile_info$N,
        rho = profile_info$rho,
        method = m,
        varsel_perf_list[[m]]
      )
    })
  )
}

selection_performance <- do.call(rbind, sim_selection_performance)
row.names(selection_performance) <- NULL

# Adjust the order of the variables
# Make variable a factor
selection_performance$variable <- factor(selection_performance$variable,
                                         levels = paste0("X", 1:20))
# Sort
selection_performance <- selection_performance[order(selection_performance$variable,
                                                  selection_performance$method), ]
```


## Introduction
\vspace{-6mm}
<!-- 1/2 PAGE -->
<!-- Review the problem and perhaps cite some papers posted or others -->
When building a statistical model, variable selection is important whether we are interested in prediction, explanation, or description for a given research question.
Selecting the most appropriate predictors is one of the most important and difficult parts of model building.
Variable selection is necessary when there are a large number of variables being considered and it can be a challenge to identify the most relevant predictors for the outcome of interest.
Variable selection can be done through automated selection techniques, often alongside background knowledge of the research question.
However, these automated approaches can lead to the incorrect inclusion or exclusion of predictors, which may affect the accuracy, variability, and overall robustness of the final model (Heinze et al., 2017).
The primary aim of this simulation study is to investigate five variable selection methods and compare their performance. Ultimately, we aimed to provide the recommended variable selection methods and under which specific circumstances. This report covers the findings of our simulation study. It fills in the results from the simulation plan agreed upon in the prior interim simulation plan report. 

## Model Notation
<!-- 1/2 PAGE -->
<!-- here give the notation for the model -->

## Methods
<!-- 1 PAGE -->
\vspace{-6mm}
<!-- Describe model selection approaches and provide references if necessary -->
Simulation studies involve repeatedly generated from a specified data-generating mechanism and analyzing it to assess the performance of statistical methods under controlled conditions (Morris et al., 2017).
In this study, simulation will be used to investigate the performance of different variable selection techniques.
Specifically, we will look at five model selection techniques in linear regression: Backwards Variable Selection, Akaike Information Criteria (AIC), Bayesia Alkaline Criteria (BIC), Least Absolute Shrinkage and Selection Operator (LASSO), and Elastic Net.
We will simulate data with one outcome and 20 variables, 5 of which will have varying strengths of relationships with the outcome. These 5 variables will have coefficients of approximately 0.17, 0.33, 0.5, 0.67, and 0.83, while the remaining 15 variables will have coefficients of 0.
The variables with non-zero coefficients represent the true predictors of the outcome. This setup allows us to evaluate how well each variable selection method identifies important predictors and avoids selecting variables that do not contribute to the model.
Performance will be evaluated using bias, coverage of 95% confidence intervals, and measures of variable selection performance including true and false positive rates and type I and II error.

We will simulate data with one outcome and 20 variables, 5 of which will have varying strengths of relationships with the outcome. We will specify the 5 covariates to have coefficients of $\approx$ 0.17, 0.33, 0.5, 0.67, and 0.83 and the remaining 15 variables to have coefficients of 0. 
The 5 non-zero covariates are essentially what we are setting to be as the most important predictors of the outcome and simulation will allow us to assess how the different variable selection techniques do in regard to selecting predictors that actually matter.

Because FPR was calculated separately for each null variable X6–X20, the FPR is equivalent to the variable-specific Type I error rate. In this context, a false positive means that a variable with a true coefficient of 0 was incorrectly selected into the final model. Therefore, the Type I error results followed the same pattern as the FPR results.

<!-- - MAKE SURE U MENTION THAT FOR VARIABLE PERFORMANCE YOU ARE DONG PER VAR PER METHOD!! -->
<!-- NOT FOR ANY OF X1-X5 OR X6-X20 but for ALL -->

\vspace{-6mm}
## Simulation
<!-- 1-1.5 PAGES -->
\vspace{-6mm}
<!-- Write on how you will simulate the data and how we are fitting those data. How will you summarize and present results? -->
We will run 10,000 iterations or simulations to obtain - come back here after reading that one paper.
We will be using the gen_data() function from the High-Dimensional Regression Modeling (hdrm) package in R to simulate the data for model fitting. We will specify n = 10,000 data frames, 5 non-zero predictors, 15 predictors that have a null value of 0, an exchangeable correlation structure (forces the correlation between each pair of variables to be the same), and correlation coefficients of 0, 0.35, and 0.70. The different correlation coefficients will allow us to assess how varying amounts of linear relationships or correlations between variables may affect results from model selection techniques. 
We will repeat this for a sample size of 250 and a sample size of 500 to assess the effect of sample size as well.
For each iteration of the simulation, 
we will fit a linear regression model using the lm() function from the stats R package. 
Following this, the model will be put into each of the 5 model selection approaches. We will extract the selected variables from each technique and summarize results. Specifically, we will calculate the percentage of times X1-X5 are in the final model after applying the variable selection technique (True Positives) and 
the percentage of times X15-X20 are included in the final model after variable selection (False Positives). 
For the variables that are retained in the model, 
we will calculate the deviation of the model estimates from the "true" or expected estimate defined (Bias), the long run probability a given 95% Confidence Interval (CI) will contain the true parameter (Coverage), as well as the Type I and Type II error of the variables.
In addition, for each of the five variable selection techniques, 
we will re-fit a final linear regression model with all the variables that 
a given technique identified as important predictors.
Alternatively, we may consider another appropriate approach as re-fitting models as if no variable selection was performed does introduce bias.

In the final report, we will include a table summarizing the performance of the variable selection techniques and a table summarizing the statistical/inference errors for each of the variables. We will additionally accompany these tables with visualizations.
We will use a significance level of $\alpha = 0.5$ for all analyses.  
All analyses will be conducted with R veresion 4.4.1.

\vspace{-6mm}
## Results and Discussion
\vspace{-6mm}
<!-- 2 PAGES -->
<!-- Summarize findings. Organization of findings is very important.  -->
<!-- What method do I recommend and under what cases. Cover the big picture  -->
<!-- recommendations after giving the results. -->

Bias was generally consistent across the different correlation settings, so we calculated and visualized the average bias for X1–X5 by selection method, sample size, and correlation setting (Figure 1). Across most settings, AIC, BIC, anc backwards selection had the smallest average bias. In contrast, 
the shrinkage methods showed more noticeable negative bias, indicating that these methods tended to underestimate the coefficients for X1-X5. This is expected since shrinkage methods penalize coefficient size and pull estimates toward 0. 
For both LASSO and elastic net, the 1SE lambda generally resulted in more bias than the minimum lambda, likely due to stronger penalization. These patterns were similar across correlation settings, although bias was slightly reduced with the larger sample size. 
Thus, if one is mainly interested in reducing bias, the AIC, BIC, or backwards variable selection methods are preferable.  

```{r, echo = F, warning=F,message=F, fig.cap="Average bias for each variable selection method. Variables X1–X5 are shown in the column panels, and the different predictor correlation settings are shown in the row panels. Circles represent n=250, and triangles represent n=500. AIC, BIC, and backward selection generally had smaller mean bias compared to the penalized regression methods. For both LASSO and elastic net, the 1SE lambda showed more negative bias than the minimum lambda.", position='H', include=F}
#### Visualization of bias for the different sel methods

# Look at each variable
# model_performance %>%
#   # filter((variable %in% c("X1", "X2", "X3", "X4", "X5"))) %>%
#   group_by(n, rho, method, variable) %>%
#   summarise(mean_bias = mean(bias)) %>%
#   pivot_wider(names_from = variable,
#               values_from = mean_bias)

# Table
bias_table <- model_performance %>%
  filter(variable %in% paste0("X", 1:5)) %>%
  group_by(n, rho, method, variable) %>%
  summarise(mean_bias = mean(bias, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(
    names_from = variable,
    values_from = mean_bias
  )

# Create a df to visualize bias
bias_plot_df <- model_performance %>% 
  filter(variable %in% paste0("X", 1:5)) %>% 
  group_by(n, rho, variable, method) %>% 
  summarise(
    mean_bias = mean(bias, na.rm = TRUE),
    .groups = "drop"
  )

# Rename the method labels
bias_plot_df$method <- factor(bias_plot_df$method,
                                  levels = c("AIC", "Backward",
                                             "BIC",
                                    "Elastic_lambda.1se", "Elastic_lambda.min",
                                    "LASSO_lambda.1se", "LASSO_lambda.min" ),
                                  labels = c("AIC", "Backward",
                                             "BIC",
                                             "Elastic 1se", "Elastic Min",
                                             "LASSO 1SE", "LASSO Min"))

# Plot - Mean Bias by variable selection method
ggplot(bias_plot_df, aes(y = fct_rev(method), 
                         x = mean_bias,
                         color = method,
                         shape = factor(n))) +
  theme_lucid(legend.position = "bottom") +
  guides(color = "none") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  geom_segment(aes(y = fct_rev(method), 
                   yend = fct_rev(method)),
               linewidth = 0.8) +
  geom_point(size = 2) +
  facet_grid(rho ~ variable,
             labeller = labeller(
               variable = c(
                 "X1" = "X[1]~(beta==0.17)",
                  "X2" = "X[2]~(beta==0.33)",
                  "X3" = "X[3]~(beta==0.50)",
                  "X4" = "X[4]~(beta==0.67)",
                  "X5" = "X[5]~(beta==0.83)"),
               .default = label_parsed)) +
  labs(
    x = "Mean Bias",
    y = "Method",
    shape = "sample Size"
  ) +
  scale_x_continuous(breaks = seq(-0.2, 0, by = 0.1))
   

```

For coverage, we calculated the average percentage of times the 95% interval covered the true coefficient for each variable and simulation setting, visualized in Figure ?? in the \textit{Supplemental Material} section. We observed the widest range in coverage for X1, the variable with the weakest relationship with the outcome, compared to variables X2-X20. The BIC method with $\rho = 0.7$ and sample size of 250 resulted in the lowest coverage for X1, with 20% coverage. Overall, BIC performed the least favorably for X1 across the different simulation settings. However, coverage was $\approx$ 100\% for variables X2-X5 across all simulation scenarios. 
Based on these results, if coverage is of primary concern, special caution may be needed when using more conservative selection methods such as BIC, particularly for predictors with weak relationships with the outcome and when predictors are highly correlated.

```{r, echo = F, results='hide', warning=F, message=F, fig.cap="Coverage percentage for each variable selection method. Column panels show the five variables with increasing strengths of association with the outcome, and row panels show the different correlation settings among predictors. Circles represent n=250, and triangles represent n=500. BIC had the lowest coverage for X1, the weakest predictor, particularly when predictor correlation was high", include = F}

### Display table like Dr.Sevick has to look at coverage
# Coverage for all vars for all sel methods
coverage <- model_performance %>% 
  group_by(n, rho, method, variable) %>% 
  summarise(prob_cover = mean(coverage) * 100)

# Df for now to look at
cov_df <- pivot_wider(coverage,
                      names_from = variable,
                      values_from = prob_cover)

# Create some forest plots as done in Morris et al. paper
## A separate color for each of the 5 var sel methods
## A row for bias and a row for CI coverage
## A col for each sample size


##### Visualization of coverage for the different sel methods
## This was not created in the final report but may be included in 
# Supplemental material. 
# Create a df to visualize bias
coverage_plot_df <- model_performance %>% 
  filter(variable %in% paste0("X", 1:5)) %>% 
  group_by(n, rho, method, variable) %>% 
  summarise(
    coverage_percent = mean(coverage, na.rm = TRUE) * 100,
    .groups = "drop"
  )

# Rename the method labels
coverage_plot_df$method <- factor(coverage_plot_df$method,
                                  levels = c("AIC", "Backward",
                                             "BIC",
                                    "Elastic_lambda.1se", "Elastic_lambda.min",
                                    "LASSO_lambda.1se", "LASSO_lambda.min" ),
                                  labels = c("AIC", "Backward",
                                             "BIC",
                                             "Elastic 1se", "Elastic Min",
                                             "LASSO 1SE", "LASSO Min"))

# Plot
ggplot(coverage_plot_df, aes(y = fct_rev(method), 
                         x = coverage_percent,
                         color = method,
                         shape = factor(n))) +
  theme_lucid(legend.position = "bottom") +
  geom_vline(xintercept = 95, linetype = "dashed", color = "red") +
  guides(color = "none")+
  geom_point(size = 2) +
  facet_grid(rho ~ variable,
             # Chnage the labels
  labeller = labeller(
    variable = c(
      "X1" = "X[1]~(beta==0.17)", # To get the betas to display
      "X2" = "X[2]~(beta==0.33)",
      "X3" = "X[3]~(beta==0.50)",
      "X4" = "X[4]~(beta==0.67)",
      "X5" = "X[5]~(beta==0.83)"
    ),
    .default = label_parsed)) +
  labs(
    x = "Mean Coverage (%)",
    y = "Method",
    shape = "N")
```


As FPRs across variables X6-X20 were generally consistent for each variable selection method, we calculated ahd visualized the average FP percentage across the null variables. 
From Figure 3, elastic net with $\lambda_{\text{min}}$ had the highest FP percentages, while AIC and BIC had the lowest percentages across the different correlation settings. Correlation among predictors appeared to have the largest effect on elastic net and LASSO using $\lambda_{\text{1SE}}$, where higher correlation among predictors resulted in higher FP percentages. We also note that sample sizes did not seem to influence results. These results suggest that AIC and BIC were more robust to changes in predictor correlation and produced some of the lowest FP percentages in this simulation setting.

```{r, echo = F, message = F, warning = F, fig.cap="Average false positive rate percentage by variable selection method. Column panels correspond to sample size. Circles represent rho=0, triangles represent rho=0.35, and squares represent rho=0.70. Elastic net, particularly using lambda min, produced the highest FP percentages, while BIC produced the lowest FP percentages across correlation settings. The shrinkage methods were more influenced by predictor correlation, with higher correlation generally resulting in higher FP percentages.", include = F}
# False positives == TYPE I ERROR since we are looking at individual vars for each method
fps_summary <- selection_performance %>% 
  filter(!(variable %in% c("X1", "X2", "X3", "X4", "X5"))) %>% 
  group_by(n, rho, method) %>% 
  summarise(fps_rate = mean(false_positives, na.rm = TRUE),
            fps_percent = 100 * fps_rate,
            .groups = "drop")

# Look at each variable
# selection_performance %>%
#   filter(!(variable %in% c("X1", "X2", "X3", "X4", "X5"))) %>%
#   group_by(n, rho, method, variable) %>%
#   summarise(fps_percent = mean(false_positives) * 100) %>%
#   pivot_wider(names_from = variable,
#               values_from = fps_percent)

# Looking at these results, seems that vals for X6-X20 are pretty similar
# Going to summarise with mean to digest it better yummm


# Rename the method labels
fps_summary$method <- factor(fps_summary$method,
                                  levels = c("AIC", "Backward",
                                             "BIC",
                                    "Elastic_lambda.1se", "Elastic_lambda.min",
                                    "LASSO_lambda.1se", "LASSO_lambda.min" ),
                                  labels = c("AIC", "Backward",
                                             "BIC",
                                             "Elastic 1se", "Elastic Min",
                                             "LASSO 1SE", "LASSO Min"))

# Plot
ggplot(fps_summary, aes(y = fct_rev(method), 
                         x = fps_percent,
                         color = method,
                         shape = factor(rho))) +
  theme_lucid(legend.position = "bottom") +
  guides(color = "none") +
  # geom_vline(xintercept = 0.95, linetype = "dashed", color = "red") +
  geom_point(size = 2) +
  facet_grid(~ n,
             labeller = labeller(
               variable = c(
                 "X1" = "X1 (rho=0.17)",
                 "X2" = "X2 (rho=0.33)",
                 "X3" = "X3 (rho=0.5)",
                 "X4" = "X4 (rho=0.67)",
                 "X5" = "X5 (rho=0.83)"))) +
  labs(
    x = "Mean False Positive Rate (%)",
    y = "Method",
    shape = expression(rho),
    title = "FPR by Variable Selection Method - Averaged by Method"
  )

# There does not appear to be a difference for different rhos
# There is no difference for different rhos for AIC,BIC,Backwards but there
# is a difference with shrinkage methods for different rhos, specifically
# most apparent with lamdab = 1se where rho = 0 results in lowest FPR and
# rho = 0.7 results in highest FPR aka so higher correlation with outcome tends # to result in more false positives which makes sense
# Samp size does not seem to have an effect
```

The average TPR was calculated across the different simulation settings.
Assessing TPRs for X1-X5, it was evident that X1 was the only variable were TPRs noticeably differed across the selection methods. 
In Table 1, we summarize the TPR percentages for X1. 
Overall, BIC had the lowest TPRs and within BIC, higher correlations among predictors decreased the TPR. 
We observed similar patterns at smaller scales for AIC and Backwards selection. 
The two shrinkage methods with $\lambda_{\text{min}}$ consistenly had the highest TPRs across all settings. We also note that TPRs were consistently higher for sample size of 500 compared to 250. 
There were no apparent differences in TPRs across selection methods, sample sizes, or correlation settings for variables X2-X5 as these were selected nearly all of the time. 

```{r, echo = F, message = F, warning = F, include = F}
# False postiives == TYPE I ERROR since we are looking at individual vars for each method
# Look at each variable -- Seems that X1 is most clearly different here
# selection_performance %>%
#   filter((variable %in% c("X1", "X2", "X3", "X4", "X5"))) %>%
#   group_by(n, rho, method, variable) %>%
#   summarise(tps_percent = mean(true_positives) * 100) %>%
#   pivot_wider(names_from = variable,
#               values_from = tps_percent) 

# Looking at just X1
# selection_performance %>%
#   filter(variable == "X1") %>%
#   group_by(n, rho, method, variable) %>%
#   summarise(tps_percent = mean(true_positives) * 100) %>%
#   pivot_wider(names_from = variable,
#               values_from = tps_percent) 

# Table for report
x1_tpr_long <- selection_performance %>%
  dplyr::filter(variable == "X1") %>%
  dplyr::group_by(n, rho, method) %>%
  dplyr::summarise(
    tpr_percent = mean(true_positives, na.rm = TRUE) * 100,
    .groups = "drop") 

# Create a row for each method
x1_tpr_table <- x1_tpr_long %>%
  tidyr::pivot_wider(
    id_cols = c(n, rho),
    names_from = method,
    values_from = tpr_percent) %>%
  dplyr::arrange(n, rho) %>% 
  # Change the colnames
  rename(N = n,
         Correlation = rho,
         "Elastic 1SE" = Elastic_lambda.1se,
         "Elastic Min" = Elastic_lambda.min,
         "LASSO 1SE" = LASSO_lambda.1se,
         "LASSO Min" = LASSO_lambda.min)

# Display with kable
kable(x1_tpr_table,
      # format = "latex",
      booktabs = TRUE,
      digits = 2,
      caption = "Mean True Positive Rate (\\%) for X1") %>% 
  kable_styling(latex_options = c("striped","hold_position",
                                  "scale_down"))
  
```

In regard to testing-related performance measures, we summarized Type I and Type II error percentages.
Because FPR was calculated separately for each null variable X6–X20, the FPR is equivalent to the variable-specific Type I error rate. In this context, a false positive means that a variable with a true coefficient of 0 was incorrectly selected into the final model. Therefore, the Type I error results followed the same pattern as the FPR results.

Type II error was evaluated for variables X1–X5. Mean Type II error percentages for X2–X5 were generally close to \0% across the different variable selection methods, correlation settings, and sample sizes, as shown in Figure 4 for X2 and X3. This suggests that moderate to strong relationships with the outcome were consistently detected by all methods.
However, the X1 variable had consistently lower Type II error percentages across most selection methods. We do note that increasing the sample size does appear to modestly decrease the Type II error rates. 
BIC tended to have the highest Type II error for X1. 
These results suggest that variables with weaker relationships with the outcome are more likely to be missed during variable selection methods, particularly with more conservative methods such as BIC. 

```{r, echo = F, warning = F, message=F, fig.cap="Averaged Type II error percentages for X1-X3. Type II error percentages for X2–X5 were approximately 0 across most simulation settings, indicating that moderate to strong predictors were consistently selected. X1 had the highest Type II error rates across all variable selection methods and correlation settings. Sample size slightly improved rates with a noticeable improvement for BIC when sample size was increased and under high predictor correlation.", include = F}
# Type II error summary - get the percentage
errorII_summary <- selection_performance %>% 
  filter(variable %in% c("X1", "X2", "X3")) %>% 
  group_by(n, rho, method, variable) %>% 
  summarise(
    errorII_rate = mean(typeII_error, na.rm = TRUE),
    errorII_percent = 100 * errorII_rate,
    .groups = "drop"
  )

# Plot
ggplot(errorII_summary, aes(y = fct_rev(method),
                            x = errorII_percent,
                            shape = factor(n),
                            color = method)) +
  theme_lucid(legend.position = "bottom") +
  geom_point(size = 2) +
  facet_grid(rho ~ variable) +
  guides(color = "none") +
  labs(
    x = "Type II Error (%)",
    y = "Method",
    shape = expression(n)
  )

```



\vspace{-6mm}
## Future Work 
\vspace{-6mm}
<!-- 1/2 PAGE -->
<!-- List things you learned and ideas you might want to invesitgate further that  -->
<!-- were beyond the scope of this project. -->
In the final report, we will include a table summarizing the performance of the variable selection techniques and a table summarizing the statistical/inference errors for each of the variables. We will additionally accompany these tables with visualizations.
We will use a significance level of $\alpha = 0.5$ for all analyses.  
All analyses will be conducted with R veresion 4.4.1.

<!-- TABLES OF SIMULATION PARAMETERS ARE COMMON IN METHODS PAPERS -->

\newpage
## Code and Reproducibility

\newpage
## References

\newpage 

### Supplemental Material




