---
title: Project 4 Data Analysis Plan
author: Diana Felipe
output: 
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    latex_engine: xelatex
    extra_dependencies: ["longtable", "float"]
mainfont: Arial
fontsize: 12pt
geometry: margin=1in
header-includes:
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage{float}
  - \usepackage{caption}
  - \usepackage{titling}
  - \usepackage{enumitem}
  - \setlength{\droptitle}{-4cm}
  - \preauthor{\vspace{-0.6cm}}
  - \postauthor{\vspace{-2.5cm}}
  - \setlist{nosep}
---


```{r, echo = F, warning=F, message = F}
knitr::opts_chunk$set(echo = TRUE)

library(tinytex)
library(readr)
library(knitr)
library(ggplot2)
library(kableExtra)
library(stats)
library(dplyr)
library(tibble)
library(labelled)
library(see)
library(here)
library(expss)
library(hdrm)
library(MASS) # AIC/BIC variable selection
library(olsrr) # Backwards variable selection
library(glmnet) # LASSO
library(tidyverse)
library(kableExtra)
library(here)

opts_chunk$set(tidy = F)
```

```{r, echo = F}
# Read in the rds file from the simulation run
simu_results <- readRDS("simulation_results_2500.rds")

# Source the functions from Project4.R
source("../Code/Project4_Functions.R")
```



## Introduction
<!-- 1/2 PAGE -->
<!-- Review the problem and perhaps cite some papers posted or others -->
Variable selection is important whether we are interested in prediction, explanation, or description for a given research question when building a statistical model.
Selecting the most appropriate predictors is one of the most important and difficult parts of model building.
Variable selection is necessary when there are a large number of variables being considered/available and it can be a challenge to identify the most relevant predictors for the outcome of interest.
Variable selection can be done through automated selection techniques, often alongside background knowledge of the research question.
However, these automated approaches can lead to the incorrect inclusion or exclusion of predictors, which may affect the accuracy, variability, and overall robustness of the final model (Heinze et al., 2017).
The primary aim of this simulation study is to investigate five variable selection methods and compare their performance. Ultimately, we aimed to provide the recommended variable selection methods and under which specific circumstances. This report covers the findings of our simulation study. It fills in the results from the simulation plan agreed upon in the prior interim simulation plan report. 

## Model Notation
<!-- 1/2 PAGE -->
<!-- here give the notation for the model -->

## Methods
<!-- 1 PAGE -->
\vspace{-6mm}
<!-- Describe model selection approaches and provide references if necessary -->
Simulation studies involve repeatedly generated from a specified data-generating mechanism and analyzing it to assess the performance of statistical methods under controlled conditions (Morris et al., 2017).
In this study, simulation will be used to investigate the performance of different variable selection techniques.
Specifically, we will look at five model selection techniques in linear regression: Backwards Variable Selection, Akaike Information Criteria (AIC), Bayesia Alkaline Criteria (BIC), Least Absolute Shrinkage and Selection Operator (LASSO), and Elastic Net.
We will simulate data with one outcome and 20 variables, 5 of which will have varying strengths of relationships with the outcome. These 5 variables will have coefficients of approximately 0.17, 0.33, 0.5, 0.67, and 0.83, while the remaining 15 variables will have coefficients of 0.
The variables with non-zero coefficients represent the true predictors of the outcome. This setup allows us to evaluate how well each variable selection method identifies important predictors and avoids selecting variables that do not contribute to the model.
Performance will be evaluated using bias, coverage of 95% confidence intervals, and measures of variable selection performance including true and false positive rates and type I and II error.

We will simulate data with one outcome and 20 variables, 5 of which will have varying strengths of relationships with the outcome. We will specify the 5 covariates to have coefficients of $\approx$ 0.17, 0.33, 0.5, 0.67, and 0.83 and the remaining 15 variables to have coefficients of 0. 
The 5 non-zero covariates are essentially what we are setting to be as the most important predictors of the outcome and simulation will allow us to assess how the different variable selection techniques do in regard to selecting predictors that actually matter.

<!-- - MAKE SURE U MENTION THAT FOR VARIABLE PERFORMANCE YOU ARE DONG PER VAR PER METHOD!! -->
<!-- NOT FOR ANY OF X1-X5 OR X6-X20 but for ALL -->

\vspace{-6mm}
## Simulation
<!-- 1-1.5 PAGES -->
\vspace{-6mm}
<!-- Write on how you will simulate the data and how we are fitting those data. How will you summarize and present results? -->
We will run 10,000 iterations or simulations to obtain - come back here after reading that one paper.
We will be using the gen_data() function from the High-Dimensional Regression Modeling (hdrm) package in R to simulate the data for model fitting. We will specify n = 10,000 data frames, 5 non-zero predictors, 15 predictors that have a null value of 0, an exchangeable correlation structure (forces the correlation between each pair of variables to be the same), and correlation coefficients of 0, 0.35, and 0.70. The different correlation coefficients will allow us to assess how varying amounts of linear relationships or correlations between variables may affect results from model selection techniques. 
We will repeat this for a sample size of 250 and a sample size of 500 to assess the effect of sample size as well.
For each iteration of the simulation, 
we will fit a linear regression model using the lm() function from the stats R package. 
Following this, the model will be put into each of the 5 model selection approaches. We will extract the selected variables from each technique and summarize results. Specifically, we will calculate the percentage of times X1-X5 are in the final model after applying the variable selection technique (True Positives) and 
the percentage of times X15-X20 are included in the final model after variable selection (False Positives). 
For the variables that are retained in the model, 
we will calculate the deviation of the model estimates from the "true" or expected estimate defined (Bias), the long run probability a given 95% Confidence Interval (CI) will contain the true parameter (Coverage), as well as the Type I and Type II error of the variables.
In addition, for each of the five variable selection techniques, 
we will re-fit a final linear regression model with all the variables that 
a given technique identified as important predictors.
Alternatively, we may consider another appropriate approach as re-fitting models as if no variable selection was performed does introduce bias.

In the final report, we will include a table summarizing the performance of the variable selection techniques and a table summarizing the statistical/inference errors for each of the variables. We will additionally accompany these tables with visualizations.
We will use a significance level of $\alpha = 0.5$ for all analyses.  
All analyses will be conducted with R veresion 4.4.1.

\vspace{-6mm}
## Results and Discussion
\vspace{-6mm}
<!-- 2 PAGES -->
<!-- Summarize findings. Organization of findings is very important.  -->
<!-- What method do I recommend and under what cases. Cover the big picture  -->
<!-- recommendations after giving the results. -->

```{r, echo = F}
##### MODEL PERFORMANCE
# Create data frame to assess multiple methods
sim_model_performance <- list()

# For the 2 profiles
for (p in seq_along(simu_results)) {
  # Get the pth profile summary
  profile_summary <- summarize_profiles(simu_results[[p]])
  # The pth profile info
  profile_info <- profiles[p, ]
  
  # List of all model performance for all selection methods
  model_perf_list <- list(
    Backward = profile_summary$backward$model_performance,
    AIC = profile_summary$AIC$model_performance,
    BIC = profile_summary$BIC$model_performance,
    LASSO_lambda.1se = profile_summary$lasso_lambda.1se$model_performance,
    LASSO_lambda.min = profile_summary$lasso_lambda.min$model_performance,
    Elastic_lambda.1se = profile_summary$elastic_net_lambda.1se$model_performance,
    Elastic_lambda.min = profile_summary$elastic_net_lambda.min$model_performance
  )
  
  # Create a data frame
  sim_model_performance[[p]] <- do.call(
    # Combine into rows
    rbind,
    # Create data frame with the cols of interst
    lapply(names(model_perf_list), function(m) {
      data.frame(
        n = profile_info$N,
        rho = profile_info$rho,
        method = m,
        model_perf_list[[m]]
      )
    })
  )
}

model_performance <- do.call(rbind, sim_model_performance)
row.names(model_performance) <- NULL

# Adjust the order of the variables
# Make variable a factor
model_performance$variable <- factor(model_performance$variable,
                                         levels = paste0("X", 1:20))
# Sort
model_performance <- model_performance[order(model_performance$variable,
                                                     model_performance$method), ]

### Display something Dr. S has..
# Coverage for all vars for all sel methods
coverage <- model_performance %>% 
  group_by(n, rho, method, variable) %>% 
  summarise(prob_cover = mean(coverage) * 100)

# Df for now to look at
cov_df <- pivot_wider(coverage,
                      names_from = variable,
                      values_from = prob_cover)

# Create some forest plots as done in Morris et al. paper
## A separate color for each of the 5 var sel methods
## A row for bias and a row for CI coverage
## A col for each sample size

```

```{r, echo = F, warning=F,message=F, fig.cap="The bias by variable selection method. Circle points corresond to sample size of 250 while square points correspond to 500."}
#### Visualization of bias for the different sel methods
# Create a df to visualize bias
bias_plot_df <- model_performance %>% 
  filter(variable %in% paste0("X", 1:5)) %>% 
  group_by(n, rho, method, variable) %>% 
  summarise(
    mean_bias = mean(bias, na.rm = TRUE),
    lower = quantile(bias, 0.025, na.rm = TRUE),
    upper = quantile(bias, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

# Rename the method labels
bias_plot_df$method <- factor(bias_plot_df$method,
                                  levels = c("AIC", "Backward",
                                             "BIC",
                                    "Elastic_lambda.1se", "Elastic_lambda.min",
                                    "LASSO_lambda.1se", "LASSO_lambda.min" ),
                                  labels = c("AIC", "Backward",
                                             "BIC",
                                             "Elastic 1se", "Elastic Min",
                                             "LASSO 1se", "LASSO Min"))

# Plot
ggplot(bias_plot_df, aes(y = fct_rev(method), 
                         x = mean_bias,
                         color = method,
                         shape = factor(n))) +
  theme_lucid(legend.position = "none") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  geom_segment(aes(x = lower, 
                   xend = upper,
                   y = fct_rev(method), 
                   yend = fct_rev(method)),
               linewidth = 0.8) +
  geom_point(size = 2) +
  facet_grid(rho ~ variable,
             labeller = labeller(
               variable = c(
                 "X1" = "X1 (rho=0.17)",
                 "X2" = "X2 (rho=0.33)",
                 "X3" = "X3 (rho=0.5)",
                 "X4" = "X4 (rho=0.67)",
                 "X5" = "X5 (rho=0.83)"))) +
  labs(
    x = "Bias",
    y = "Method"
  )
   
## Perhaps just use this for interpreting results


```


```{r, echo = F, results='hide', warning=F, message=F}
### Visaulization of coverage for the different sel methods
# Create a df to visualize bias
coverage_plot_df <- model_performance %>% 
  filter(variable %in% paste0("X", 1:5)) %>% 
  group_by(n, rho, method, variable) %>% 
  summarise(
    mean_coverage = mean(coverage, na.rm = TRUE),
    lower = quantile(coverage, 0.025, na.rm = TRUE),
    upper = quantile(coverage, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

# Rename the method labels
coverage_plot_df$method <- factor(coverage_plot_df$method,
                                  levels = c("AIC", "Backward",
                                             "BIC",
                                    "Elastic_lambda.1se", "Elastic_lambda.min",
                                    "LASSO_lambda.1se", "LASSO_lambda.min" ),
                                  labels = c("AIC", "Backward",
                                             "BIC",
                                             "Elastic 1se", "Elastic Min",
                                             "LASSO 1se", "LASSO Min"))

# Plot
ggplot(coverage_plot_df, aes(y = fct_rev(method), 
                         x = mean_coverage,
                         color = method,
                         shape = factor(n))) +
  theme_lucid(legend.position = "none") +
  geom_vline(xintercept = 0.95, linetype = "dashed", color = "red") +
  geom_segment(aes(x = lower, 
                   xend = upper,
                   y = fct_rev(method), 
                   yend = fct_rev(method)),
               linewidth = 0.8) +
  geom_point(size = 2) +
  facet_grid(rho ~ variable,
             labeller = labeller(
               variable = c(
                 "X1" = "X1 (rho=0.17)",
                 "X2" = "X2 (rho=0.33)",
                 "X3" = "X3 (rho=0.5)",
                 "X4" = "X4 (rho=0.67)",
                 "X5" = "X5 (rho=0.83)"))) +
  labs(
    x = "Coverage",
    y = "Method",
    title = "Coverage by Variable Selection Method"
  )
```

```{r, echo = F, warning = F, message = F}
### VARIABLE SELECTION METHOD PERFORMANCE
# Create data frame to assess multiple methods
sim_selection_performance <- list()

# For the 2 profiles
for (p in seq_along(simu_results)) {
  # Get the pth profile summary
  profile_summary <- summarize_profiles(simu_results[[p]])
  # The pth profile info
  profile_info <- profiles[p, ]
  
  # List of all varsel performance for all selection methods
  varsel_perf_list <- list(
    Backward = profile_summary$backward$selection_performance,
    AIC = profile_summary$AIC$selection_performance,
    BIC = profile_summary$BIC$selection_performance,
    LASSO_lambda.1se = profile_summary$lasso_lambda.1se$selection_performance,
    LASSO_lambda.min = profile_summary$lasso_lambda.min$selection_performance,
    Elastic_lambda.1se = profile_summary$elastic_net_lambda.1se$selection_performance,
    Elastic_lambda.min = profile_summary$elastic_net_lambda.min$selection_performance
  )
  
  # Create a data frame
  sim_selection_performance[[p]] <- do.call(
    # Combine into rows
    rbind,
    # Create data frame with the cols of interst
    lapply(names(varsel_perf_list), function(m) {
      data.frame(
        n = profile_info$N,
        rho = profile_info$rho,
        method = m,
        varsel_perf_list[[m]]
      )
    })
  )
}

selection_performance <- do.call(rbind, sim_selection_performance)
row.names(selection_performance) <- NULL

# Adjust the order of the variables
# Make variable a factor
selection_performance$variable <- factor(selection_performance$variable,
                                         levels = paste0("X", 1:20))
# Sort
selection_performance <- selection_performance[order(selection_performance$variable,
                                                     selection_performance$method), ]

# False postiives == TYPE I ERROR since we are looking at individual vars for each method
fps <- selection_performance %>% 
  filter(!(variable %in% c("X1", "X2", "X3", "X4", "X5"))) %>% 
  group_by(n, rho, method, variable) %>% 
  summarise(prob_fps = mean(false_positives) * 100)

# Df for now to look at
fps_df <- pivot_wider(fps,
                      names_from = variable,
                      values_from = prob_fps)
# fps_df

# Looking at these results, seems that vals for X6-X20 are pretty similar
# Going to summarise with mean to digest it better yummm
# 

# fps %>% 
#   group_by(n, rho, method) %>% 
#   summarise(mean_fp = mean(prob_fps))

# Create a df to visualize bias
fps_plot_df <- selection_performance %>% 
  filter(variable %in% paste0("X", 6:20)) %>% 
  group_by(n, rho, method) %>% 
  summarise(
    mean_fpr = mean(false_positives, na.rm = TRUE),
    .groups = "drop"
  )

# Rename the method labels
fps_plot_df$method <- factor(fps_plot_df$method,
                                  levels = c("AIC", "Backward",
                                             "BIC",
                                    "Elastic_lambda.1se", "Elastic_lambda.min",
                                    "LASSO_lambda.1se", "LASSO_lambda.min" ),
                                  labels = c("AIC", "Backward",
                                             "BIC",
                                             "Elastic 1se", "Elastic Min",
                                             "LASSO 1se", "LASSO Min"))

# Plot
ggplot(fps_plot_df, aes(y = fct_rev(method), 
                         x = mean_fpr,
                         color = method,
                         shape = factor(rho))) +
  theme_lucid(legend.position = "bottom") +
  guides(color = "none") +
  # geom_vline(xintercept = 0.95, linetype = "dashed", color = "red") +
  geom_point(size = 2) +
  facet_grid(~ n,
             labeller = labeller(
               variable = c(
                 "X1" = "X1 (rho=0.17)",
                 "X2" = "X2 (rho=0.33)",
                 "X3" = "X3 (rho=0.5)",
                 "X4" = "X4 (rho=0.67)",
                 "X5" = "X5 (rho=0.83)"))) +
  labs(
    x = "False Positive Rate",
    y = "Method",
    shape = expression(rho),
    title = "FPR by Variable Selection Method"
  )

# There does not appear to be a difference for different rhos
# There is no difference for different rhos for AIC,BIC,Backwards but there
# is a difference with shrinkage methods for different rhos, specifically
# most apparent with lamdab = 1se where rho = 0 results in lowest FPR and
# rho = 0.7 results in highest FPR aka so higher correlation with outcome tends # to result in more false positives which makes sense
# Samp size does not seem to have an effect
```


```{r, echo = F, warning = F, message=F}
# Type II error summary - get the percentage
errorII_summary <- selection_performance %>% 
  filter(variable %in% paste0("X", 1:5)) %>% 
  group_by(n, rho, method, variable) %>% 
  summarise(
    errorII_rate = mean(typeII_error, na.rm = TRUE),
    errorII_percent = 100 * errorII_rate,
    .groups = "drop"
  )

# A df or table to look at
errorII_df <- errorII_summary %>% 
  select(n, rho, method, variable, errorII_percent) %>% 
  pivot_wider(
    names_from = variable,
    values_from = errorII_percent
  )

# Plot
ggplot(errorII_summary, aes(y = fct_rev(method),
                            x = errorII_percent,
                            shape = factor(rho),
                            color = method)) +
  theme_lucid(legend.position = "bottom") +
  geom_point(size = 2) +
  facet_grid(~ n) +
  guides(color = "none") +
  labs(
    x = "Type II Error (%)",
    y = "Method",
    shape = expression(rho),
    title = "Type II Error by Variable Selection Method"
  )

```

```{r, echo = F, warning=F, message=F}

```


In the final report, we will include a table summarizing the performance of the variable selection techniques and a table summarizing the statistical/inference errors for each of the variables. We will additionally accompany these tables with visualizations.
We will use a significance level of $\alpha = 0.5$ for all analyses.  
All analyses will be conducted with R veresion 4.4.1.

\vspace{-6mm}
## Future Work 
\vspace{-6mm}
<!-- 1/2 PAGE -->
<!-- List things you learned and ideas you might want to invesitgate further that  -->
<!-- were beyond the scope of this project. -->
In the final report, we will include a table summarizing the performance of the variable selection techniques and a table summarizing the statistical/inference errors for each of the variables. We will additionally accompany these tables with visualizations.
We will use a significance level of $\alpha = 0.5$ for all analyses.  
All analyses will be conducted with R veresion 4.4.1.

<!-- TABLES OF SIMULATION PARAMETERS ARE COMMON IN METHODS PAPERS -->


## References




