---
title: "\\large Assessing the Performance of Different Variable Selection Methods"
author: Diana Felipe
output: 
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    latex_engine: xelatex
    extra_dependencies: ["longtable", "float"]
mainfont: Arial
fontsize: 12pt
geometry: margin=1in
header-includes:
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage{float}
  - \usepackage{caption}
  - \usepackage{titling}
  - \usepackage{enumitem}
  - \setlength{\droptitle}{-4cm}
  - \preauthor{\vspace{-0.6cm}}
  - \postauthor{\vspace{-2.3cm}}
  - \setlist{nosep}
bibliography: Felipe_Proj4.bib
csl: diana_vancouver_final.csl
---


```{r, echo = F, warning=F, message = F}
knitr::opts_chunk$set(echo = TRUE)

library(tinytex)
library(readr)
library(knitr)
library(ggplot2)
library(kableExtra)
library(stats)
library(tibble)
library(labelled)
library(see)
library(here)
library(expss)
library(hdrm)
library(MASS) # AIC/BIC variable selection
library(olsrr) # Backwards variable selection
library(glmnet) # LASSO
library(tidyverse)
library(kableExtra)
library(here)
library(dplyr)
library(tidyr)

opts_chunk$set(tidy = F)
```

```{r, echo = F}
# Read in the rds file from the simulation run
simu_results <- readRDS("simulation_results_2500.rds")

# Source the functions from Project4.R
source("../Code/Project4_Functions.R")
```

```{r, echo = F}
##### MODEL PERFORMANCE
# Create data frame to assess multiple methods
sim_model_performance <- list()

# For the 2 profiles
for (p in seq_along(simu_results)) {
  # Get the pth profile summary
  profile_summary <- summarize_profiles(simu_results[[p]])
  # The pth profile info
  profile_info <- profiles[p, ]
  
  # List of all model performance for all selection methods
  model_perf_list <- list(
    Backward = profile_summary$backward$model_performance,
    AIC = profile_summary$AIC$model_performance,
    BIC = profile_summary$BIC$model_performance,
    LASSO_lambda.1se = profile_summary$lasso_lambda.1se$model_performance,
    LASSO_lambda.min = profile_summary$lasso_lambda.min$model_performance,
    Elastic_lambda.1se = profile_summary$elastic_net_lambda.1se$model_performance,
    Elastic_lambda.min = profile_summary$elastic_net_lambda.min$model_performance
  )
  
  # Create a data frame
  sim_model_performance[[p]] <- do.call(
    # Combine into rows
    rbind,
    # Create data frame with the cols of interst
    lapply(names(model_perf_list), function(m) {
      data.frame(
        n = profile_info$N,
        rho = profile_info$rho,
        method = m,
        model_perf_list[[m]]
      )
    })
  )
}

model_performance <- do.call(rbind, sim_model_performance)
row.names(model_performance) <- NULL

# Adjust the order of the variables
# Make variable a factor
model_performance$variable <- factor(model_performance$variable,
                                         levels = paste0("X", 1:20))
# Sort
model_performance <- model_performance[order(model_performance$variable,
                                                     model_performance$method), ]

### VARIABLE SELECTION METHOD PERFORMANCE
# Create data frame to assess multiple methods
sim_selection_performance <- list()

# For the 2 profiles
for (p in seq_along(simu_results)) {
  # Get the pth profile summary
  profile_summary <- summarize_profiles(simu_results[[p]])
  # The pth profile info
  profile_info <- profiles[p, ]
  
  # List of all varsel performance for all selection methods
  varsel_perf_list <- list(
    Backward = profile_summary$backward$selection_performance,
    AIC = profile_summary$AIC$selection_performance,
    BIC = profile_summary$BIC$selection_performance,
    LASSO_lambda.1se = profile_summary$lasso_lambda.1se$selection_performance,
    LASSO_lambda.min = profile_summary$lasso_lambda.min$selection_performance,
    Elastic_lambda.1se = profile_summary$elastic_net_lambda.1se$selection_performance,
    Elastic_lambda.min = profile_summary$elastic_net_lambda.min$selection_performance
  )
  
  # Create a data frame
  sim_selection_performance[[p]] <- do.call(
    # Combine into rows
    rbind,
    # Create data frame with the cols of interst
    lapply(names(varsel_perf_list), function(m) {
      data.frame(
        n = profile_info$N,
        rho = profile_info$rho,
        method = m,
        varsel_perf_list[[m]]
      )
    })
  )
}

selection_performance <- do.call(rbind, sim_selection_performance)
row.names(selection_performance) <- NULL

# Adjust the order of the variables
# Make variable a factor
selection_performance$variable <- factor(selection_performance$variable,
                                         levels = paste0("X", 1:20))
# Sort
selection_performance <- selection_performance[order(selection_performance$variable,
                                                  selection_performance$method), ]
```


## Introduction
\vspace{-6mm}
<!-- 1/2 PAGE -->
<!-- Review the problem and perhaps cite some papers posted or others -->
When building a statistical model, variable selection is important whether we are interested in prediction, explanation, or description for a given research question.
Selecting the most appropriate predictors is one of the most important and difficult parts of model building.
Variable selection is necessary when there are a large number of variables being considered and it can be a challenge to identify the most relevant predictors for the outcome of interest.
Variable selection can be done through automated selection techniques, often alongside background knowledge of the research question.
However, these automated approaches can lead to the incorrect inclusion or exclusion of predictors, which may affect the accuracy, variability, and overall robustness of the final model [@Heinzeetal2018].
The primary aim of this simulation study is to investigate five variable selection methods and compare their performance in regard to bias, coverage, False Positive Rates (FPRs),
True Positive Rates (TPRs), as well as Type I & II error rates. 
Ultimately, we aimed to provide recommendations on variable selection methods for specific circumstances.
This report describes the simulation study in which we evaluated the five variable selection methods. 

\vspace{-6mm}
## Model Notation
\vspace{-6mm}
<!-- 1/2 PAGE -->
<!-- here give the notation for the model -->
For each simulation setting, data were generated from a linear regression model with one continuous outcome and 20 predictors. 
Let $Y_i$ be the outcome for subject $i$, where $i = 1, 2, \dots, n$. 
We specified the model as: 
\vspace{-3mm}
$$Y_i = \beta_{0} + \sum^{20}_{j = 1} \beta_{j} X_{ij} + \epsilon_{i}$$
where $X_{ij}$ represents predictor $j$ for subject $i$, $\beta_{j}$ is the true coefficient for predictor $j$, and $\epsilon_{i}$ is the random error term. 
We defined the true coefficient vector as  $\beta = (0.17, 0.33, 0.5, 0.67, 0.83, 0, \dots, 0)$. 
Variables X1-X5 were defined to be the true predictors while X6-X20 were null variables. 
The coefficients for X1-X5 were chosen to represent increasing strengths of association with the outcome. 
The predictors were generated using an exchangeable correlation structure, where every pair of predictors had the same correlation $\rho$. We evaluated three values of $\rho$: 0, 0.35, and 0.7.
This range of correlation values allowed us to assess how increasing correlation among predictors affected the performance of the variable selection methods.
We looked at two sample sizes: n = 250 and n = 500. These settings resulted in 6 simulation scenarios. 
Variable selection methods were assessed based on their ability to identify true predictors, avoid selecting null predictors, and produce estimates with low bias and appopriate confidence interval coverage. 

\vspace{-6mm}
## Methods
<!-- 1 PAGE -->
\vspace{-6mm}
<!-- Describe model selection approaches and provide references if necessary -->
Simulation studies involve repeatedly generating data from a specified data-generating mechanism and analyzing the data to assess the performance of statistical methods under controlled conditions [@Morrisetal2019]. For this simulation study, we investigate five model selection techniques in linear regression: backward variable selection, Akaike Information Criterion (AIC), Bayesian Information Criterion (BIC), Least Absolute Shrinkage and Selection Operator (LASSO), and Elastic Net. We briefly describe the variable selection methods below.

\vspace{-3mm}
1. **Backward Selection:** This method begins with all variables being considered in the full model. Variables are then deleted one at a time until the remaining variables are considered to have a meaningful contribution to the outcome [@Heinzeetal2018]. In this study, we used the \texttt{ols\_step\_backward\_p()} function from the \texttt{olsrr} package in R. This function removes predictors based on p-values in a stepwise manner until no additional variables meet the removal criterion.

2. **AIC**: AIC is a model selection criterion that allows for the comparison of different models. AIC attempts to balance selecting too few variables, which may lead to underfitting, and selecting too many variables, which may lead to overfitting. It provides a way to balance model fit and model complexity [@ChowdhuryTurin2020]. Smaller AIC values suggest a more favorable model. In this study, we used the \texttt{step()} function from the \texttt{stats} package in R with the AIC penalty.

3. **BIC**: BIC is similar to AIC in that it compares models based on fit and complexity. However, BIC uses a stronger penalty for the number of variables included in the model, so it tends to choose more parsimonious models compared to AIC [@ChowdhuryTurin2020]. In this study, we used the \texttt{step()} function from the \texttt{stats} package in R with the BIC penalty.

4. **LASSO**: LASSO is a penalized regression method that performs variable selection by shrinking coefficient estimates toward zero. The LASSO penalty can shrink some coefficients exactly to zero, meaning those variables are excluded from the final model [@Tibshirani1996]. This makes LASSO useful when many predictors are being considered. In this study, LASSO was fit using the \texttt{cv.glmnet()} function from the \texttt{glmnet} package in R, with both \(\lambda_{\min}\) and \(\lambda_{1se}\) considered.

5. **Elastic Net:** Elastic Net is also a penalized regression method and can be viewed as a combination of LASSO and ridge regression penalties. Like LASSO, it can shrink some coefficients to zero and perform variable selection. However, Elastic Net can be useful when predictors are correlated because it encourages groups of correlated predictors to be selected together [@ZouHastie2005]. In this study, Elastic Net was fit using the \texttt{cv.glmnet()} function from the \texttt{glmnet} package in R, with both \(\lambda_{\min}\) and \(\lambda_{1se}\) considered.
\vspace{-3mm}

These methods allowed us to compare traditional stepwise selection approaches with penalized regression approaches. The stepwise methods select variables by removing predictors from a full model, while LASSO and Elastic Net use penalties that shrink coefficients toward zero. We compared the performance of these methods using simulation-based measures including bias, coverage, true positive rates, false positive rates, and Type I and Type II error rates.

\vspace{-6mm}
## Simulation
<!-- 1-1.5 PAGES -->
\vspace{-6mm}
<!-- Write on how you will simulate the data and how we are fitting those data. How will you summarize and present results? -->
Because our key performance measure for this study was coverage, we used the Monte Carlo standard error approach described by Morris et al. [-@Morrisetal2019] to determine the number of simulation repetitions. Coverage is a proportion, so the Monte Carlo standard error can be approximated as $MCSE(\hat{p}) = \sqrt{\frac{p(1-p)}{n_{sim}}}$
Following the approach used in lecture, we used \(p = 0.50\), which gives the most conservative estimate because \(p(1-p)\) is maximized when \(p = 0.50\). We specified the desired Monte Carlo standard error to be 0.01. Therefore, $n_{sim} = \frac{0.50(1-0.50)}{0.01^2} = 2500$. Thus, we used 2500 simulation repetitions for each simulation scenario.
We used the `gen_data()` function from the High-Dimensional Regression Modeling (`hdrm`) package in R to simulate the data for model fitting. For each simulated dataset, we generated one continuous outcome and 20 predictor variables, described previously in the Model Notation section. We had a total of six simulation scenarios for the two sample sizes and three correlation values.
For each simulated dataset, we first fit a full linear regression model using the `lm()` function from the `stats` package in R. We then applied the five variable selection approaches. For LASSO and Elastic Net, both \(\lambda_{\min}\) and \(\lambda_{1se}\) were considered. For the penalzed regression methods, we refit a linear regression model using the selected variables in order to obtain confidence intervals.

True positiive rates were calculated as the percentage of simulations in which X1–X5 were selected in the final model. False positive rates were calculated as the percentage of simulations in which each null variable, X6–X20, was incorrectly selected. Since false positive rates were calculated separately for each null variable, these values are equivalent to variable-specific Type I error rates. Type II error was calculated for X1–X5 as the percentage of simulations in which each true predictor was not selected.
Bias was calculated as the difference between the estimated coefficient and the true coefficient. We used unconditional bias, meaning that if a variable was not selected, its coefficient estimate was set to 0 before calculating bias. Coverage was also calculated using an unconditional approach. For selected variables, coverage was defined as whether the 95% confidence interval contained the true coefficient.

The simulation was organized using several custom functions. The `model_var_performance()` function calculated bias, coverage, true positives, false positives, and Type I/II error for one selected model. The `model_var_summary()` function handled the different output types from each variable selection method and passed the selected variables and coefficients to `model_var_performance()`. The `all_var_sel_methods()` function simulated one dataset for a given sample size and correlation value, fit the full linear regression model, applied all five variable selection methods, and evaluated each method. The `sim_one_profile()` function repeated this process for one simulation profile, where a profile was defined as one combination of sample size and predictor correlation. Finally, `sim_results()` ran the full simulation across all six profiles, and `summarize_profiles()` aggregated results across simulation repetitions.

All analyses were conducted in R version 4.4.1 using a significance level of \(\alpha = 0.05\).

\vspace{-6mm}
## Results and Discussion
\vspace{-6mm}
<!-- 2 PAGES -->
<!-- Summarize findings. Organization of findings is very important.  -->
<!-- What method do I recommend and under what cases. Cover the big picture  -->
<!-- recommendations after giving the results. -->

Bias was generally consistent across the different correlation settings, so we summarized the average bias for X1–X5 by selection method, sample size, and correlation setting (Figure 1). Across most settings, AIC, BIC, and backward selection had the smallest average bias. In contrast, the shrinkage methods showed more noticeable negative bias, indicating that LASSO and elastic net tended to underestimate the coefficients for X1–X5. This is expected since shrinkage methods penalize coefficient size and pull estimates toward 0. For both LASSO and elastic net, $\lambda_{1SE}$ generally resulted in more bias than $\lambda_{min}$, likely due to stronger penalization. Bias was slightly reduced with the larger sample size. 

Coverage was calculated as the average percentage of times the 95% confidence interval contained the true coefficient. The widest range in coverage was observed for X1, the weakest predictor.

```{r, echo = F, warning=F,message=F, fig.cap="Average bias for each variable selection method. Variables X1–X5 are shown in the column panels, and the different predictor correlation settings are shown in the row panels. Circles represent n=250, and triangles represent n=500. AIC, BIC, and backward selection generally had smaller mean bias compared to the penalized regression methods. For both LASSO and elastic net, the 1SE lambda showed more negative bias than the minimum lambda."}

#### Visualization of bias for the different sel methods

# Look at each variable
# model_performance %>%
#   # filter((variable %in% c("X1", "X2", "X3", "X4", "X5"))) %>%
#   group_by(n, rho, method, variable) %>%
#   summarise(mean_bias = mean(bias)) %>%
#   pivot_wider(names_from = variable,
#               values_from = mean_bias)

# Table
bias_table <- model_performance %>%
  filter(variable %in% paste0("X", 1:5)) %>%
  group_by(n, rho, method, variable) %>%
  summarise(mean_bias = mean(bias, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(
    names_from = variable,
    values_from = mean_bias
  )

# Create a df to visualize bias
bias_plot_df <- model_performance %>% 
  filter(variable %in% paste0("X", 1:5)) %>% 
  group_by(n, rho, variable, method) %>% 
  summarise(
    mean_bias = mean(bias, na.rm = TRUE),
    .groups = "drop"
  )

# Rename the method labels
bias_plot_df$method <- factor(bias_plot_df$method,
                                  levels = c("AIC", "Backward",
                                             "BIC",
                                    "Elastic_lambda.1se", "Elastic_lambda.min",
                                    "LASSO_lambda.1se", "LASSO_lambda.min" ),
                                  labels = c("AIC", "Backward",
                                             "BIC",
                                             "Elastic 1se", "Elastic Min",
                                             "LASSO 1SE", "LASSO Min"))

# Plot - Mean Bias by variable selection method
ggplot(bias_plot_df, aes(y = fct_rev(method), 
                         x = mean_bias,
                         color = method,
                         shape = factor(n))) +
  theme_lucid(legend.position = "bottom") +
  guides(color = "none") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  geom_segment(aes(y = fct_rev(method), 
                   yend = fct_rev(method)),
               linewidth = 0.8) +
  geom_point(size = 2) +
  facet_grid(rho ~ variable,
             labeller = labeller(
               variable = c(
                 "X1" = "X[1]~(beta==0.17)",
                  "X2" = "X[2]~(beta==0.33)",
                  "X3" = "X[3]~(beta==0.50)",
                  "X4" = "X[4]~(beta==0.67)",
                  "X5" = "X[5]~(beta==0.83)"),
               .default = label_parsed)) +
  labs(
    x = "Mean Bias",
    y = "Method",
    shape = "sample Size"
  ) +
  scale_x_continuous(breaks = seq(-0.2, 0, by = 0.1))
```

BIC with ρ=0.7 and n=250 had the lowest coverage for X1, with 20% coverage. In contrast, coverage was approximately 100% for X2–X5 across most simulation scenarios. These results suggest that conservative methods such as BIC may have poor coverage for weak predictors, especially when predictors are highly correlated.

False positive rates were generaly consistent across the null variables X6–X20, so we summarized the average FP percentage across these variables. Elastic net with $\lambda_{min}$ had the highest FP percentages, while AIC and BIC had the lowest percentages across the correlation settings (Figure 3). Correlation among predictors appeared to have the largest effect on elastic net and LASSO using $\lambda_{1SE}$, where higher corelation resulted in higher FP percentages. Sample size did not appear to strongly influence FP percentages.

True positive ates were high for X2–X5 across nearly all methods and simulation settings. 
The main differences were obsered for X1, the weakest true predictor (Table 1). BIC had the lowest TPRs for X1, and higher correlation among predictors further decreased the TPR.

```{r, echo = F, warning=F, message=F, fig.cap="Coverage percentage for each variable selection method. Column panels show the five variables with increasing strengths of association with the outcome, and row panels show the different correlation settings among predictors. Circles represent n=250, and triangles represent n=500. BIC had the lowest coverage for X1, the weakest predictor, particularly when predictor correlation was high"}

### Display table like Dr.Sevick has to look at coverage
# Coverage for all vars for all sel methods
coverage <- model_performance %>% 
  group_by(n, rho, method, variable) %>% 
  summarise(prob_cover = mean(coverage) * 100)

# Df for now to look at
cov_df <- pivot_wider(coverage,
                      names_from = variable,
                      values_from = prob_cover)

# Create some forest plots as done in Morris et al. paper
## A separate color for each of the 5 var sel methods
## A row for bias and a row for CI coverage
## A col for each sample size


##### Visualization of coverage for the different sel methods
## This was not created in the final report but may be included in 
# Supplemental material. 
# Create a df to visualize bias
coverage_plot_df <- model_performance %>% 
  filter(variable %in% paste0("X", 1:5)) %>% 
  group_by(n, rho, method, variable) %>% 
  summarise(
    coverage_percent = mean(coverage, na.rm = TRUE) * 100,
    .groups = "drop"
  )

# Rename the method labels
coverage_plot_df$method <- factor(coverage_plot_df$method,
                                  levels = c("AIC", "Backward",
                                             "BIC",
                                    "Elastic_lambda.1se", "Elastic_lambda.min",
                                    "LASSO_lambda.1se", "LASSO_lambda.min" ),
                                  labels = c("AIC", "Backward",
                                             "BIC",
                                             "Elastic 1se", "Elastic Min",
                                             "LASSO 1SE", "LASSO Min"))

# Plot
ggplot(coverage_plot_df, aes(y = fct_rev(method), 
                         x = coverage_percent,
                         color = method,
                         shape = factor(n))) +
  theme_lucid(legend.position = "bottom") +
  geom_vline(xintercept = 95, linetype = "dashed", color = "red") +
  guides(color = "none")+
  geom_point(size = 2) +
  facet_grid(rho ~ variable,
             # Chnage the labels
  labeller = labeller(
    variable = c(
      "X1" = "X[1]~(beta==0.17)", # To get the betas to display
      "X2" = "X[2]~(beta==0.33)",
      "X3" = "X[3]~(beta==0.50)",
      "X4" = "X[4]~(beta==0.67)",
      "X5" = "X[5]~(beta==0.83)"
    ),
    .default = label_parsed)) +
  labs(
    x = "Mean Coverage (%)",
    y = "Method",
    shape = "N")
```

Similar but smaller patterns were observed for AIC and backward selection. The shrinkage methods using $\lambda_{min}$ consistently had the highest TPRs for X1, although this came with higher false positive rates. TPRs were generally higher for n=500 compared to n=250.

```{r, echo = F, message = F, warning = F, fig.cap="Average false positive rate percentage by variable selection method. Column panels correspond to sample size. Circles represent rho=0, triangles represent rho=0.35, and squares represent rho=0.70. Elastic net, particularly using lambda min, produced the highest FP percentages, while BIC produced the lowest FP percentages across correlation settings. The shrinkage methods were more influenced by predictor correlation, with higher correlation generally resulting in higher FP percentages."}
# False positives == TYPE I ERROR since we are looking at individual vars for each method
fps_summary <- selection_performance %>% 
  filter(!(variable %in% c("X1", "X2", "X3", "X4", "X5"))) %>% 
  group_by(n, rho, method) %>% 
  summarise(fps_rate = mean(false_positives, na.rm = TRUE),
            fps_percent = 100 * fps_rate,
            .groups = "drop")

# Look at each variable
# selection_performance %>%
#   filter(!(variable %in% c("X1", "X2", "X3", "X4", "X5"))) %>%
#   group_by(n, rho, method, variable) %>%
#   summarise(fps_percent = mean(false_positives) * 100) %>%
#   pivot_wider(names_from = variable,
#               values_from = fps_percent)

# Looking at these results, seems that vals for X6-X20 are pretty similar
# Going to summarise with mean to digest it better yummm


# Rename the method labels
fps_summary$method <- factor(fps_summary$method,
                                  levels = c("AIC", "Backward",
                                             "BIC",
                                    "Elastic_lambda.1se", "Elastic_lambda.min",
                                    "LASSO_lambda.1se", "LASSO_lambda.min" ),
                                  labels = c("AIC", "Backward",
                                             "BIC",
                                             "Elastic 1se", "Elastic Min",
                                             "LASSO 1SE", "LASSO Min"))

# Plot
ggplot(fps_summary, aes(y = fct_rev(method), 
                         x = fps_percent,
                         color = method,
                         shape = factor(rho))) +
  theme_lucid(legend.position = "bottom") +
  guides(color = "none") +
  # geom_vline(xintercept = 0.95, linetype = "dashed", color = "red") +
  geom_point(size = 2) +
  facet_grid(~ n,
             labeller = labeller(
               variable = c(
                 "X1" = "X1 (rho=0.17)",
                 "X2" = "X2 (rho=0.33)",
                 "X3" = "X3 (rho=0.5)",
                 "X4" = "X4 (rho=0.67)",
                 "X5" = "X5 (rho=0.83)"))) +
  labs(
    x = "Mean False Positive Rate (%)",
    y = "Method",
    shape = expression(rho),
    title = "FPR by Variable Selection Method - Averaged by Method"
  )
```


```{r, echo = F, message = F, warning = F}
# False postiives == TYPE I ERROR since we are looking at individual vars for each method
# Look at each variable -- Seems that X1 is most clearly different here
# selection_performance %>%
#   filter((variable %in% c("X1", "X2", "X3", "X4", "X5"))) %>%
#   group_by(n, rho, method, variable) %>%
#   summarise(tps_percent = mean(true_positives) * 100) %>%
#   pivot_wider(names_from = variable,
#               values_from = tps_percent) 

# Looking at just X1
# selection_performance %>%
#   filter(variable == "X1") %>%
#   group_by(n, rho, method, variable) %>%
#   summarise(tps_percent = mean(true_positives) * 100) %>%
#   pivot_wider(names_from = variable,
#               values_from = tps_percent) 

# Table for report
x1_tpr_long <- selection_performance %>%
  dplyr::filter(variable == "X1") %>%
  dplyr::group_by(n, rho, method) %>%
  dplyr::summarise(
    tpr_percent = mean(true_positives, na.rm = TRUE) * 100,
    .groups = "drop") 

# Create a row for each method
x1_tpr_table <- x1_tpr_long %>%
  tidyr::pivot_wider(
    id_cols = c(n, rho),
    names_from = method,
    values_from = tpr_percent) %>%
  dplyr::arrange(n, rho) %>% 
  # Change the colnames
  rename(N = n,
         Correlation = rho,
         "Elastic 1SE" = Elastic_lambda.1se,
         "Elastic Min" = Elastic_lambda.min,
         "LASSO 1SE" = LASSO_lambda.1se,
         "LASSO Min" = LASSO_lambda.min)

# Display with kable
kable(x1_tpr_table,
      # format = "latex",
      booktabs = TRUE,
      digits = 2,
      caption = "Mean True Positive Rate (\\%) for X1") %>% 
  kable_styling(latex_options = c("striped",
                                  "scale_down"))
  
```

For testing-related performance measures, the variable-specific Type I error results followed the same pattern as the false positive rates, since FPR was calculated separately for each null variable X6–X20. Type II error was evaluated for X1–X5. Mean Type II error percentages for X2–X5 were generally close to 0\% across methods, correlation settings, and sample sizes, suggesting tht moderate to strong predictors were consistently detected. X1 had the highest Type II error, especially for BIC, indicating that weaker predictors were more likly to be missed by conservative selection methods. Increasing sample size modestly reduced Type II error.

```{r, echo = F, warning = F, message=F, fig.cap="Averaged Type II error percentages for X1-X3. Type II error percentages for X2–X5 were approximately 0 across most simulation settings, indicating that moderate to strong predictors were consistently selected. X1 had the highest Type II error rates across all variable selection methods and correlation settings. Sample size slightly improved rates with a noticeable improvement for BIC when sample size was increased and under high predictor correlation."}

# Type II error summary - get the percentage
errorII_summary <- selection_performance %>% 
  filter(variable %in% c("X1", "X2", "X3")) %>% 
  group_by(n, rho, method, variable) %>% 
  summarise(
    errorII_rate = mean(typeII_error, na.rm = TRUE),
    errorII_percent = 100 * errorII_rate,
    .groups = "drop"
  )

# Plot
ggplot(errorII_summary, aes(y = fct_rev(method),
                            x = errorII_percent,
                            shape = factor(n),
                            color = method)) +
  theme_lucid(legend.position = "bottom") +
  geom_point(size = 2) +
  facet_grid(rho ~ variable) +
  guides(color = "none") +
  labs(
    x = "Type II Error (%)",
    y = "Method",
    shape = expression(n)
  )
```

Overalll, the results show a tradeoff between conservative selection and sensitivity to weaker predictors. BIC had low false positive rates but were more likely to miss X1 and had lower coverage for this weak predictor. LASSO and elastic net with $\lambda_{min}$ had higher TPRs but also higher FP percentages. AIC and backward selection generally had low bias and reasonable TPRs, but backward selection tended to include more null predictors. Therefore, the preferred method depends on the study goal! BIC may be preferable when avoiding false positives is most important, while AIC, backward selection, or penalized methods with $\lambda_{min}$ may be preferable when retaining weaker true predictors is a priority.

\vspace{-6mm}
## Future Work 
\vspace{-6mm}
<!-- 1/2 PAGE -->
<!-- List things you learned and ideas you might want to invesitgate further that  -->
<!-- were beyond the scope of this project. -->

This study focused on a limited set of conditions, including two sample sizes, three correlation settings, and five true predictors with fixed effect sizes. Future work could expand the simulation design by considering additional sample sizes, different numbers of predictors, and weaker or more variable effect sizes. This may be especially useful because the weakest predictor, X1, showed the most noticeable differences across selection methods.
Additional work could also examine other correlation structures beyond the exchangeable setting used here. For example, autoregressive correlation structures may better represent situations where only certain groups of predictors are highly correlated. It may also be useful to compare additional variable selection methods or evaluate how results change when using different tuning parameter choices for LASSO and elastic net.
Finally, this study focused on linear regression with a continuous outcome. Future studies could extend this framework to logistic regression, survival models, or other types of outcomes. These extensions would help determine whether the patterns observed here, such as the tradeoff between false positives and missing weak predictors, remain consistent in other modeling settings.

\newpage
## Code and Reproducibility
The data and statistical code used to generate the results and figures in this report are
available in the following GitHub repository:

- GitHub repository: https://github.com/DianaF65/BIOS6624

Directory structure:

- Background/ - A description of the background of the project.
- Code/
  – The “Project4.R” file contains initial data exploration as well as code for con-
struction of the final data set used for all analyses.
  - The "Project4_Functions.R" file contains all functions used for simulation and is sourced in this final report. 

- RawData/ – No data was used in this analysis. 

- Reports/ The “Project4_FinalReport.Rmd” R Markdown file was used to generate
this final report of analyses.
All analyses and figures were generated using the R statistical programming language v.
4.4.1, and the repository contains the R scripts and R Markdown files used to produce this
report

\newpage
## References

