---
title: The Effects of Hard-Drug Use at Baseline on HAART Response & Outcomes
subtitle: Frequentist Analysis
author: Diana Felipe
output: 
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    latex_engine: xelatex
    extra_dependencies: ["longtable", "float"]
mainfont: Arial
fontsize: 12pt
documentclass: article
geometry: margin=1in
header-includes:
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage{float}
  - \usepackage{caption}
  - \usepackage{titling}
  - \setlength{\droptitle}{-2cm}
  - \preauthor{\vspace{-0.6cm}}
  - \postauthor{\vspace{-3cm}}
---

```{r, echo = F, message=F, warning = F}
knitr::opts_chunk$set(echo = FALSE)
library(readr)
library(knitr)
library(ggplot2)
library(MASS)
library(emmeans)
library(multcomp)
library(kableExtra)
library(stats)
library(dplyr)
library(tibble)
library(labelled)
library(tableone)
library(lme4)
library(lmerTest)
library(nlme)
library(gtsummary)
library(gt)
library(ggeffects)
library(UpSetR)
library(visdat)
library(lubridate)
library(see)
library(data.table)
library(gridExtra)
library(kableExtra)
library(regressinator)
library(broom)
library(expss)
library(mediation)
library(bda)

# Libraries for Bayesian analysis
# Dependencies 
library(cmdstanr)
library(bayesplot)  # diagnostic plots of the MCMC chains
library(posterior)  # for summarizing posterior draws
library(bayestestR) # for calculating higest density posterior intervals
library(mcmcse)     # for calculating MCMCSE's
library(loo)        # for getting model fit statistics (WAIC and LOO-IC)
library(dplyr)
library(tibble)
library(brms)

opts_chunk$set(tidy = F)
```

```{r, echo = F}
# Read in the Data
data <- read.csv("../Data/final_hiv_data.csv")

# Examine the data and data types
# str(data)


```

```{r, echo = F}
# Assessment of the data first
# str(data)

# Create categorical labels and variables 

# Hard drugs
data$hard_drugs_cat <- factor(data$hard_drugs,
                              levels = c(0, 1),
                              labels = c("No", "Yes"))

# Move after hard drugs variable
data <- data %>% 
  relocate(hard_drugs_cat, .after = hard_drugs)

# Convert collapse race to factor
data$collapse_RACE <- factor(data$collapse_RACE)

# Convert collapse educ to factor
data$collapsed_EDUCBAS <- factor(data$collapsed_EDUCBAS)

# Convert SMOKE cat to factor
data$collapse_SMOKE <- factor(data$collapse_SMOKE)

# Relabel the variables
med_data <- apply_labels(data,
                   hard_drugs = "Hard Drug Use at BL",
                   ADH_yr2 = "HAART Adherence",
                   VLOAD_yr2 = "VLOAD - yr2")
```


A table1 separating by hard drug use at baseline: 

```{r, echo = F}
med_data %>% 
  tbl_summary(include = c(VLOAD_yr2, LEU3N_yr2,
                          AGG_MENT_yr2, AGG_PHYS_yr2, 
                          BMI, SMOKE, age, collapse_RACE,
                          collapsed_EDUCBAS,
                          ADH_yr2),
              by = hard_drugs,
              statistic = list(all_continuous() ~ "{mean}{sd}")) %>% 
  modify_header(label = "**Variable**") %>% 
  bold_labels()
```



# Frequentist Analysis

For these simple linear regression analyses, we will use the lm() function from the stats package in R. 

```{r, echo = F}
################################################################################
###                           Frequentist Analysis                           ###
################################################################################

```


## VLOAD
The number of HIV copies in a mL of blood.

Hypothesis: VLOAD is significantly higher 2 years after HART for subjects who reported hard drug use at baseline. 


### Mediation Analysis

```{r, echo = F}
### Parametric Approach
# Fit two models: With adherence and without adherence
# We will not be doing the crude model because of the 
# Very unbalanced hard drug use groups 

# Primary model - no adherence
vload_no_adh <- lm(log_VLOAD_yr2 ~ hard_drugs_cat + log_VLOAD_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS,
                     data = data)

# Fit the model with adherence
vload_adh <- lm(log_VLOAD_yr2 ~ hard_drugs_cat + log_VLOAD_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS
                    + collapsed_ADH,
                     data = data)
```

Comparison of estimates between models with and without adherence: 

```{r, echo = F}
# Compare betas 
t1 <- tbl_regression(vload_no_adh,
                     # exponentiate = TRUE,
                     include = c("hard_drugs_cat"),
                     pvalue_fun = label_style_pvalue(digits = 3),
                     estimate_fun = ~ style_number(.x, digits = 3)) 

t2 <- tbl_regression(vload_adh, 
                     # exponentiate = TRUE,
                     include = c("hard_drugs_cat"),
                     pvalue_fun = label_style_pvalue(digits = 3),
                     estimate_fun = ~ style_number(.x, digits = 3))


tbl_merge(
  tbls = list(t1, t2),
  tab_spanner = c("**Model Without ADH**", 
                  "**Model With ADH**")
)
```

We observe a difference of 0.119 units between the estimate for the two models. 
The proportion mediated by adherence is $\frac{-0.136 - (-0.017)}{-0.136} = 85\%$.
We have that 85% of the effect of hard drug use on copies of viral load is mediated through or can be explained by adherence to HAART. 
In other words, there is a difference between whether subjects adhered to HAART for those that did and did not report hard-drug use at baseline. 
The differences that we observe in the estimates can be explained by differences in adherence. 
Due to these findings, we included adherence in the VLOAD model. 


### Primary Analysis

Model diagnostics: 

```{r, echo = F, out.width="80%", fig.align='center', include=FALSE}
### Model diagnostics
# The final model for VLOAD
vload_primary <- vload_adh

# Residual plot
augment(vload_primary) |>
  ggplot(aes(x = .fitted, y = .resid)) +
  geom_point(size = 0.8) +
  # geom_smooth(se = FALSE) +
  labs(x = "Fitted value", y = "Residual") + 
  theme_lucid()

```

```{r, echo = F, out.width="80%", fig.align='center', include=FALSE}
# QQ Plot
augment(vload_primary) |>
  ggplot(aes(sample = .std.resid)) +
  geom_qq() +
  geom_qq_line() +
  labs(title = "Normal Q-Q plot of standardized residuals",
       x = "Theoretical quantiles", y = "Observed quantiles") + 
  theme_lucid()

```

Interpreting the results: 

```{r, echo = F, include=FALSE, eval = F, results = 'hide'}
summary(vload_primary)
```

After controlling for covariates, there is not significant difference (p = 0.970) in VLOAD 2 years after initiating HART between those who reported hard-drug use at baseline vs. those who did. 

<!-- REMEMBER TO INTERPRET ON LOG SCALE -->

\newpage

## LEU3N

### Mediation Analysis

The number of CD4+ T cell counts which is a measure of immunologic health.
With treatment, we expect CD4+ T Cll counts to increase.

Hypothesis: CD4+ T cell counts are significantly higher 2 years after HART for subjects who reported hard drug use at baseline. 

### Mediation Analysis

```{r, echo = F}
### Parametric Approach
# Fit two models: With adherence and without adherence
# We will not be doing the crude model because of the 
# Very unbalanced hard drug use groups 

# Primary model - no adherence
leu3n_no_adh <- lm(LEU3N_yr2 ~ hard_drugs_cat + LEU3N_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS,
                     data = data)

# Fit the model with adherence
leu3n_adh <- lm(LEU3N_yr2 ~ hard_drugs_cat + LEU3N_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS
                    + collapsed_ADH,
                     data = data)
```

Comparison of estimates between models with and without adherence: 

```{r, echo = F}
# Compare betas 
t1_leu3n <- tbl_regression(leu3n_no_adh,
                     include = c("hard_drugs_cat"),
                     pvalue_fun = label_style_pvalue(digits = 3),
                     estimate_fun = ~ style_number(.x, digits = 3)) 

t2_leu3n <- tbl_regression(leu3n_adh, 
                     include = c("hard_drugs_cat"),
                     pvalue_fun = label_style_pvalue(digits = 3),
                     estimate_fun = ~ style_number(.x, digits = 3))


tbl_merge(
  tbls = list(t1_leu3n, t2_leu3n),
  tab_spanner = c("**Model Without ADH**", 
                  "**Model With ADH**")
)
```

meow meow meow

### Primary Analysis

Model diagnostics: 

```{r, echo = F, out.width="80%", fig.align='center', include=FALSE}
### Model diagnostics
# The final model for VLOAD
leu3n_primary <- leu3n_no_adh

# Residual plot
augment(leu3n_primary) |>
  ggplot(aes(x = .fitted, y = .resid)) +
  geom_point(size = 0.8) +
  # geom_smooth(se = FALSE) +
  labs(x = "Fitted value", y = "Residual") + 
  theme_lucid()

```

```{r, echo = F, out.width="80%", fig.align='center', include=FALSE}
# QQ Plot
augment(leu3n_primary) |>
  ggplot(aes(sample = .std.resid)) +
  geom_qq() +
  geom_qq_line() +
  labs(title = "Normal Q-Q plot of standardized residuals",
       x = "Theoretical quantiles", y = "Observed quantiles") + 
  theme_lucid()

```

Interpreting the results: 

```{r, echo = F, include=FALSE, results = 'hide', eval = F}
summary(leu3n_primary)
```

After controlling for covariates, there is not significant difference (p = 0.970) in VLOAD 2 years after initiating HART between those who reported hard-drug use at baseline vs. those who did. 

\newpage

## AGG PHYS

The number of HIV copies in a mL of blood.

Hypothesis: VLOAD is significantly higher 2 years after HART for subjects who reported hard drug use at baseline. 


### Mediation Analysis

```{r, echo = F}
### Parametric Approach
# Fit two models: With adherence and without adherence
# We will not be doing the crude model because of the 
# Very unbalanced hard drug use groups 

# Primary model - no adherence
aggphys_no_adh <- lm(AGG_PHYS_yr2 ~ hard_drugs_cat + AGG_PHYS_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS,
                     data = data)

# Fit the model with adherence
aggphys_adh <- lm(AGG_PHYS_yr2 ~ hard_drugs_cat + AGG_PHYS_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS
                    + collapsed_ADH,
                     data = data)
```

Comparison of estimates between models with and without adherence: 

```{r, echo = F}
# Compare betas 
t1_aggphs <- tbl_regression(aggphys_no_adh,
                     # exponentiate = TRUE,
                     include = c("hard_drugs_cat"),
                     pvalue_fun = label_style_pvalue(digits = 3),
                     estimate_fun = ~ style_number(.x, digits = 3)) 

t2_aggphs <- tbl_regression(aggphys_adh, 
                     # exponentiate = TRUE,
                     include = c("hard_drugs_cat"),
                     pvalue_fun = label_style_pvalue(digits = 3),
                     estimate_fun = ~ style_number(.x, digits = 3))


tbl_merge(
  tbls = list(t1_aggphs, t2_aggphs),
  tab_spanner = c("**Model Without ADH**", 
                  "**Model With ADH**")
)
```

We observe a difference of 0.119 units between the estimate for the two models. 
The proportion mediated by adherence is $\frac{-0.136 - (-0.017)}{-0.136} = 85\%$.
We have that 85% of the effect of hard drug use on copies of viral load is mediated through or can be explained by adherence to HAART. 
In other words, there is a difference between whether subjects adhered to HAART for those that did and did not report hard-drug use at baseline. 
The differences that we observe in the estimates can be explained by differences in adherence. 
Due to these findings, we included adherence in the VLOAD model. 


### Primary Analysis

Model diagnostics: 

```{r, echo = F, out.width="80%", fig.align='center'}
### Model diagnostics
# The final model for VLOAD
aggphys_primary <- aggphys_no_adh

# Residual plot
augment(aggphys_primary) |>
  ggplot(aes(x = .fitted, y = .resid)) +
  geom_point(size = 0.8) +
  # geom_smooth(se = FALSE) +
  labs(x = "Fitted value", y = "Residual") + 
  theme_lucid()

```

```{r, echo = F, out.width="80%", fig.align='center'}
# QQ Plot
augment(aggphys_primary) |>
  ggplot(aes(sample = .std.resid)) +
  geom_qq() +
  geom_qq_line() +
  labs(title = "Normal Q-Q plot of standardized residuals",
       x = "Theoretical quantiles", y = "Observed quantiles") + 
  theme_lucid()

```

Interpreting the results: 

```{r, echo = F}
summary(aggphys_primary)
```

After controlling for covariates, there is not significant difference (p = 0.970) in VLOAD 2 years after initiating HART between those who reported hard-drug use at baseline vs. those who did. 

<!-- REMEMBER TO INTERPRET ON LOG SCALE -->

\newpage


## AGG MENT

The number of HIV copies in a mL of blood.

Hypothesis: VLOAD is significantly higher 2 years after HART for subjects who reported hard drug use at baseline. 


### Mediation Analysis

```{r, echo = F}
### Parametric Approach
# Fit two models: With adherence and without adherence
# We will not be doing the crude model because of the 
# Very unbalanced hard drug use groups 

# Primary model - no adherence
aggment_no_adh <- lm(AGG_PHYS_yr2 ~ hard_drugs_cat + AGG_PHYS_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS,
                     data = data)

# Fit the model with adherence
aggment_adh <- lm(AGG_PHYS_yr2 ~ hard_drugs_cat + AGG_PHYS_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS
                    + collapsed_ADH,
                     data = data)
```

Comparison of estimates between models with and without adherence: 

```{r, echo = F}
# Compare betas 
t1 <- tbl_regression(aggment_no_adh,
                     # exponentiate = TRUE,
                     include = c("hard_drugs_cat"),
                     pvalue_fun = label_style_pvalue(digits = 3),
                     estimate_fun = ~ style_number(.x, digits = 3)) 

t2 <- tbl_regression(aggment_adh, 
                     # exponentiate = TRUE,
                     include = c("hard_drugs_cat"),
                     pvalue_fun = label_style_pvalue(digits = 3),
                     estimate_fun = ~ style_number(.x, digits = 3))


tbl_merge(
  tbls = list(t1, t2),
  tab_spanner = c("**Model Without ADH**", 
                  "**Model With ADH**")
)
```

We observe a difference of 0.119 units between the estimate for the two models. 
The proportion mediated by adherence is $\frac{-0.136 - (-0.017)}{-0.136} = 85\%$.
We have that 85% of the effect of hard drug use on copies of viral load is mediated through or can be explained by adherence to HAART. 
In other words, there is a difference between whether subjects adhered to HAART for those that did and did not report hard-drug use at baseline. 
The differences that we observe in the estimates can be explained by differences in adherence. 
Due to these findings, we included adherence in the VLOAD model. 


### Primary Analysis

Model diagnostics: 

```{r, echo = F, out.width="80%", fig.align='center'}
### Model diagnostics
# The final model for VLOAD
aggment_primary <- aggment_no_adh

# Residual plot
augment(aggment_primary) |>
  ggplot(aes(x = .fitted, y = .resid)) +
  geom_point(size = 0.8) +
  # geom_smooth(se = FALSE) +
  labs(x = "Fitted value", y = "Residual") + 
  theme_lucid()

```

```{r, echo = F, out.width="80%", fig.align='center'}
# QQ Plot
augment(aggment_primary) |>
  ggplot(aes(sample = .std.resid)) +
  geom_qq() +
  geom_qq_line() +
  labs(title = "Normal Q-Q plot of standardized residuals",
       x = "Theoretical quantiles", y = "Observed quantiles") + 
  theme_lucid()

```

Interpreting the results: 

```{r, echo = F}
summary(aggment_primary)
```


\newpage

# Bayesian Analysis

```{r, echo = F}
# Read in data
data <- read.csv("../Data/final_hiv_data.csv")

# Assessment of the data first
# str(data)

# Create categorical labels and variables 

# Hard drugs
data$hard_drugs_cat <- factor(data$hard_drugs,
                              levels = c(0, 1),
                              labels = c("No", "Yes"))

# Move after hard drugs variable
data <- data %>% 
  relocate(hard_drugs_cat, .after = hard_drugs)

# Convert collapse race to factor
data$collapse_RACE <- factor(data$collapse_RACE)

# Convert collapse educ to factor
data$collapsed_EDUCBAS <- factor(data$collapsed_EDUCBAS)

# Convert SMOKE cat to factor
data$collapse_SMOKE <- factor(data$collapse_SMOKE)

# Relabel the variables
med_data <- apply_labels(data,
                   hard_drugs = "Hard Drug Use at BL",
                   ADH_yr2 = "HAART Adherence",
                   VLOAD_yr2 = "VLOAD - yr2")
```


## VLOAD

Laboratory in vitro and animal studies have shown that the use of hard drugs has increases HIV replication. 
However, since there are not any biologically plausible values that we could pick for an informative optimist prior, we will continue with a vague prior for our Bayesian models. 


For viral load, we specified a Normal $(0, 10000)$ prior for the regression coefficients and a Half-Normal $(0, 100^2)$ prior for the residual variance.

```{r, echo = F}
################################################################################
###                                 VLOAD                                    ###
################################################################################


### Linear model WITHOUT ADHERENCE
vload_linear <- lm(log_VLOAD_yr2 ~ hard_drugs_cat + log_VLOAD_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS,
                     data = data)

# sink("Output/P1_VLOAD_Freq_Output.txt")
# print(vload_linear)
# summary(vload_linear)
# confint(vload_linear)
# anova(vload_linear)
# sink()

``` 

```{r, echo = F}
### Linear model WITH ADHERENCE
vload_adh_linear <- lm(log_VLOAD_yr2 ~ hard_drugs_cat + log_VLOAD_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS
                    + collapsed_ADH,
                     data = data)

```



```{r, echo = F}
### Bayesian Model

# Linear model WITHOUT ADHERENCE

###########################################################################
# STEP 1: Define the model
# Normal priors on the regression coefficients
# Half normal priors on sigma
###########################################################################


stan_file <- write_stan_file("data {
  int<lower=0> N;                  // number of observations
  int<lower=0> P;                  // number of predictors including intercept
  matrix[N, P] X;                  // design matrix (first column = intercept)
  vector[N] y;                     // outcome

  vector[P] prior_mean;            // prior means for each beta
  vector<lower=0>[P] prior_sd;     // prior SDs for each beta

  real<lower=0> sigma_prior_sd;    // SD for half-normal prior on sigma
}

parameters {
  vector[P] beta;                  // regression coefficients
  real<lower=0> sigma;             // residual SD
}

model {
  // Vectorized priors for regression coefficients
  beta ~ normal(prior_mean, prior_sd);

  // Half-normal prior for sigma
  sigma ~ normal(0, sigma_prior_sd);

  // Likelihood
  y ~ normal(X * beta, sigma);
  

}

generated quantities {
  // log likelihood for each observation for calculating model fit stats
  vector[N] log_lik;
  for (n in 1:N) {
    log_lik[n] = normal_lpdf(y[n] | X[n] * beta, sigma);}
  }", dir="STAN", basename='VLOAD_primary'
)

```

```{r, echo = F}
# Linear Model WITH ADHERENCE

###########################################################################
# STEP 1: Define the model
# Normal priors on the regression coefficients
# Half normal priors on sigma
###########################################################################


stan_file_vload_adh <- write_stan_file("data {
  int<lower=0> N;                  // number of observations
  int<lower=0> P;                  // number of predictors including intercept
  matrix[N, P] X;                  // design matrix (first column = intercept)
  vector[N] y;                     // outcome

  vector[P] prior_mean;            // prior means for each beta
  vector<lower=0>[P] prior_sd;     // prior SDs for each beta

  real<lower=0> sigma_prior_sd;    // SD for half-normal prior on sigma
}

parameters {
  vector[P] beta;                  // regression coefficients
  real<lower=0> sigma;             // residual SD
}

model {
  // Vectorized priors for regression coefficients
  beta ~ normal(prior_mean, prior_sd);

  // Half-normal prior for sigma
  sigma ~ normal(0, sigma_prior_sd);

  // Likelihood
  y ~ normal(X * beta, sigma);
  

}

generated quantities {
  // log likelihood for each observation for calculating model fit stats
  vector[N] log_lik;
  for (n in 1:N) {
    log_lik[n] = normal_lpdf(y[n] | X[n] * beta, sigma);}
  }", dir="STAN", basename='VLOAD_adh'
)

```


```{r, echo = F}
###########################################################################
# STEP 2: Compile the Stan program
###########################################################################

# Model WITHOUT adherence
vload_mod <- cmdstan_model('STAN/VLOAD_primary.stan')

```

```{r, echo = F}
# Model WITH adherence
vload_mod_adh <- cmdstan_model('STAN/VLOAD_adh.stan')

```



```{r, echo = F}
###########################################################################
# STEP 3: Prepare your data to pass to Stan
###########################################################################

# Model WITHOUT adherence
# Outcome data
y <- data$log_VLOAD_yr2

# Design matrix in our linear regression
X <- model.matrix(~ hard_drugs_cat + log_VLOAD_yr0
   + BMI + collapse_SMOKE
   + age + collapse_RACE + collapsed_EDUCBAS,
   data = data)

# These are variables needed in Stan
# N is the number of observations in your data set 
# P is the number of columns in the design matrix
N <- nrow(X)
P <- ncol(X)

# Hyper-parameters for prior distributions
# Prior for regression coefficients is Normal(0, 10000)
# Prior for model error SD (sigma) is Half-Normal(0, 100)

# MVN mean (mean in the prior on the regression coefficients)
m <- c(1, rep(0, (P-1)))   

# # SD in the prior on regression coefficients --> variance 100^2
s <- rep(100, P) 

sigma_sd <- 100

# create data list to pass to STAN
data_list <- list(
  N = N,
  P = P,
  X = X,
  y = y,
  prior_mean = m,
  prior_sd = s,
  sigma_prior_sd = sigma_sd
)
```


```{r, echo = F}
# Model WITH adherence
# Outcome data
y_adh <- data$log_VLOAD_yr2

# Design matrix in our linear regression
X_adh <- model.matrix(~ hard_drugs_cat + log_VLOAD_yr0
   + BMI + collapse_SMOKE
   + age + collapse_RACE + collapsed_EDUCBAS
   + collapsed_ADH,
   data = data)

# These are variables needed in Stan
# N is the number of observations in your data set 
# P is the number of columns in the design matrix
N_adh <- nrow(X_adh)
P_adh <- ncol(X_adh)

# Hyper-parameters for prior distributions
# Prior for regression coefficients is Normal(0, 10000)
# Prior for model error SD (sigma) is Half-Normal(0, 100)

# MVN mean (mean in the prior on the regression coefficients)
m_adh <- c(1, rep(0, (P_adh - 1)))   

# # SD in the prior on regression coefficients --> variance 100^2
s_adh <- rep(100, P_adh) 

sigma_sd_adh <- 100

# create data list to pass to STAN
data_list_adh <- list(
  N = N_adh,
  P = P_adh,
  X = X_adh,
  y = y_adh,
  prior_mean = m_adh,
  prior_sd = s_adh,
  sigma_prior_sd = sigma_sd_adh
)
```


```{r, echo = F, message=F, warning = F, results = 'hide'}
###########################################################################
# STEP 4: Fit Model
###########################################################################

# Model WITHOUT adherence
# Specify the warmup and sampling iterations
vload_fit <- suppressMessages(
  suppressWarnings(
  vload_mod$sample(
  data = data_list,
  chains = 4,
  iter_warmup = 1000,
  iter_sampling = 2000,
  seed = 123,
  refresh = 0
 )
)
)

```

```{r, echo = F, message=F, warning = F, results = 'hide'}
# Model WITH adherence
# Specify the warmup and sampling iterations
vload_fit_adh <- vload_mod_adh$sample(
  data = data_list_adh,
  chains = 4,
  iter_warmup = 1000,
  iter_sampling = 2000,
  seed = 1234
)
```


```{r, echo = F}
###########################################################################
# STEP 5: Summarize the posterior
###########################################################################

# Model without ADH
vload_fit$summary(variables=c("beta[1]", "beta[2]", 
                              "beta[3]",
        # including extra betas to double check results with linear model
                              "beta[4]", "beta[5]",
                              "beta[6]", "beta[7]",
                              "beta[8]", 
                              "sigma"))


# Fancier Table of Results of Interest from the Posterior
# Extract posterior draws from cmdstanr fit
draws_vload <- vload_fit$draws()  

draws_mat_vload <- as_draws_matrix(draws_vload)
params_vload <- colnames(draws_mat_vload)

# Exclude non-parameter columns from summarization: lp__ and log_lik[n]
params_vload <- params_vload[!grepl("lp__|log_lik", params_vload)]

```

```{r, echo = F}
# Model WITH ADH

vload_fit_adh$summary(variables=c("beta[1]", "beta[2]", 
                              "beta[3]",
        # including extra betas to double check results with linear model
                              "beta[4]", "beta[5]",
                              "beta[6]", "beta[7]",
                              "beta[8]", 
                              "sigma"))


# Fancier Table of Results of Interest from the Posterior
# Extract posterior draws from cmdstanr fit
draws_vload_adh <- vload_fit_adh$draws()  

draws_mat_vload_adh <- as_draws_matrix(draws_vload_adh)
params_vload_adh <- colnames(draws_mat_vload_adh)

# Exclude non-parameter columns from summarization: lp__ and log_lik[n]
params_vload_adh <- params_vload_adh[!grepl("lp__|log_lik", params_vload_adh)]

```

```{r, echo = F}
# Model WITH ADH
# Build summary table
summary_table_vload <- lapply(params_vload, function(p) {
  vals <- as.numeric(draws_mat_vload[, p])
  
  # Monte Carlo standard error
  # Used to determine if we have run enough iterations
  # rule of thumb: MCSE should be less than 6.27% of the posterior standard deviation
  mcse_val <- mcmcse::mcse(vals)$se
  
  # Effective sample size
  ess_val <- ess_bulk(vals)
  
  # 95% HPDI
  hpd <- hdi(vals, ci = 0.95)
  
  tibble(
    Parameter = p,
    Estimate  = mean(vals),
    MCSE      = mcse_val,
    Std_Dev   = sd(vals),
    HPDI_2.5  = hpd$CI_low,
    HPDI_97.5 = hpd$CI_high,
    ESS       = ess_val
  )}) %>% 
  bind_rows()

print(summary_table_vload)

# Compare with linear fit
summary(vload_linear)

# check that mcmcse < 6% of posterior SD
(100*summary_table_vload$MCSE/summary_table_vload$Std_Dev) < 6
# All True


```


```{r, echo = F}
# Model WITH ADH
# Build summary table
summary_table_vload_adh <- lapply(params_vload_adh, function(p) {
  vals <- as.numeric(draws_mat_vload_adh[, p])
  
  # Monte Carlo standard error
  # Used to determine if we have run enough iterations
  # rule of thumb: MCSE should be less than 6.27% of the posterior standard deviation
  mcse_val <- mcmcse::mcse(vals)$se
  
  # Effective sample size
  ess_val <- ess_bulk(vals)
  
  # 95% HPDI
  hpd <- hdi(vals, ci = 0.95)
  
  tibble(
    Parameter = p,
    Estimate  = mean(vals),
    MCSE      = mcse_val,
    Std_Dev   = sd(vals),
    HPDI_2.5  = hpd$CI_low,
    HPDI_97.5 = hpd$CI_high,
    ESS       = ess_val
  )}) %>% 
  bind_rows()

print(summary_table_vload_adh)

# Compare with linear fit
summary(vload_adh_linear)

# check that mcmcse < 6% of posterior SD
(100*summary_table_vload_adh$MCSE/summary_table_vload_adh$Std_Dev) < 6
# All True

```


```{r, echo = F}
# Compare estimates for models with and without adherence

# Model without ADH
vload_terms <- summary_table_vload[2, ]

# Subset to Parameter, Estimate, HPDIs
vload_terms <- subset(vload_terms,
                      select = c("Parameter",
                                 "Estimate",
                                 "HPDI_2.5",
                                 "HPDI_97.5"))

# Model WITH ADH
# Pull term for hard drugs variable
vload_adh_terms <- summary_table_vload_adh[2, ]

# Subset to Parameter, Estimate, HPDIs
vload_adh_terms <- subset(vload_adh_terms,
                    select = c("Parameter", 
                               "Estimate",
                               "HPDI_2.5",
                               "HPDI_97.5"))

# Display side by side like with frequentist models
# Create data frame
comp_models_vload <- data.frame(Variable = "HD at Baseline (Yes)",
                 Estimate = round(vload_terms$Estimate, 3),
                 `95% HDI` = paste(round(vload_terms$HPDI_2.5, 3), 
                                    round(vload_terms$HPDI_97.5, 3), 
                                    sep = ", "),
                 Estimate = round(vload_adh_terms$Estimate, 3),
                 `95% HDI` = paste(round(vload_adh_terms$HPDI_2.5, 3),
                                    round(vload_adh_terms$HPDI_97.5, 3),
                                    sep = ", "),
                 check.names = FALSE)

# Display with kable
kable(
  comp_models_vload,
  # format = "latex",
  booktabs = TRUE,
  digits = 3,
  caption = "Model Summary") |>
  kable_styling(latex_options = c("striped", "hold_position")) |>
  add_header_above(c(" ", "Model Without ADH" = 2, 
                     "Model With ADH" = 2)) %>% 
  footnote(general = "HPDI - Highest Density Interval ")
  

```

In the model without taking HAART adherence into account, 
on average, viral load is 14% lower for hard drug users vs non hard drug users, 2 years after being on HAART. 
The HDI does not contain 0 which indicates that this difference is not significant.
When HAART adherence is taken into account, viral load is on average, 2.8% lower for hard-drug users compared to non-hard drug users, 2 years after HAART. 
The HDI excludes 0 and we have that this difference is not significant.
There is a noticeable difference in 



```{r, echo = F}
###########################################################################
# STEP 6: Model fit statistics (you only need 1 of these, not all 3)
###########################################################################
# Extract the log likelihood info
# For vload model without ADH
loglik_mat_vload <- as_draws_matrix(vload_fit$draws("log_lik")) 
# iterations x N

# Get the LOO-CV
loo_res <- loo(loglik_mat_vload)
# looic: 2212.6

```

```{r, echo = F}
# Model WITH ADH
loglik_mat_vload_adh <- as_draws_matrix(vload_fit_adh$draws("log_lik")) 
# iterations x N

# Get the LOO-CV
loo_res_vload_adh <- loo(loglik_mat_vload_adh)
# print(loo_res_vload_adh)
# 2208.8
```


```{r, echo = F}
###########################################################################
# STEP 6: MCMC Diagnostics
###########################################################################

# Model without ADH
# Dimensions: iterations x chains x parameters
vload_draws <- as_draws_array(vload_fit$draws())

# Tidy format for bayesplot / ggplot
vload_draws_df <- as_draws_df(vload_fit$draws())  

# parameters
vload_params <- c("beta[1]", "beta[2]", 
            "beta[3]", "beta[4]",
            "beta[5]", "beta[6]",
            "beta[7]", "beta[8]",
            "sigma")

```

```{r, echo = F}
# Dimensions: iterations x chains x parameters
vload_draws_adh <- as_draws_array(vload_fit_adh$draws())

# Tidy format for bayesplot / ggplot
vload_draws_df_adh <- as_draws_df(vload_fit_adh$draws())  

# parameters
vload_params_adh <- c("beta[1]", "beta[2]", 
            "beta[3]", "beta[4]",
            "beta[5]", "beta[6]",
            "beta[7]", "beta[8]",
            "sigma")
```


```{r, echo = F, results = 'hide'}
# Trace plots
# Trace plots show whether chains are mixing well and exploring the parameter space.
# Chains should mix well (lines overlap)
# No long trends (no “stickiness”)
# Ideally all chains overlap around the same mean

# Model without ADH
mcmc_trace(vload_draws, pars = vload_params)
# They look beautiful

```


```{r, echo = F, results = 'hide'}
# Model WITH ADH
mcmc_trace(vload_draws_adh, pars = vload_params_adh)


```



```{r, echo = F, results = 'hide'}
# R-hat and ESS diagnostics
# R-hat ≈ 1.00 → converged
# ESS should be reasonably large (e.g., >100–200 per parameter)
vload_fit$summary(variables = vload_params)
# We see that rhats are all approx 1
# ESS should be around > 200*9 = 1800 and that seems to be the case for most
```

```{r, echo = F}

vload_fit_adh$summary(variables = vload_params_adh)

```


```{r, echo = F, results='hide'}
# Posterior Density Plots
# Overlaid densities from multiple chains should match closely
# If densities differ substantially between chains → poor mixing
mcmc_dens_overlay(draws_vload, pars = vload_params)

```

```{r, echo = F, include = F, eval = F, results='hide'}
# Posterior density plots
# Model WITH ADH

mcmc_dens_overlay(draws_vload_adh, pars = vload_params_adh)
```


```{r, echo = F, results='hide'}
# Auto-correlation 
# Autocorrelation should decay quickly
# High autocorrelation → may need more iterations or thinning
mcmc_acf(draws_vload, pars = vload_params)

```

```{r, echo = F, results='hide'}
# Model WITH ADH
mcmc_acf(draws_vload_adh, pars = vload_params_adh)
```


```{r, echo = F, results='hide'}
# Diagnostics from cmdstan
vload_fit$cmdstan_diagnose()
# Need to load actual vload and not vload .rds for this

### Output
# Checking sampler transitions treedepth.
# Treedepth satisfactory for all transitions.
# 
# Checking sampler transitions for divergences.
# No divergent transitions found.
# 
# Checking E-BFMI - sampler transitions HMC potential energy.
# E-BFMI satisfactory.
# 
# Rank-normalized split effective sample size satisfactory for all parameters.
# 
# Rank-normalized split R-hat values satisfactory for all parameters.
# 
# Processing complete, no problems detected.

# Can use .rds file for future runs

```

```{r, echo = F, results='hide'}
# Model WITH ADH
# Diagnostics from cmdstan
vload_fit_adh$cmdstan_diagnose()
```


```{r, echo = F, eval = F}
## Save output if desired ###
# MCMC Output
sink("Output/P1_VLOAD_MCMCOutput.txt")
print(summary_table)
print(loo_res)
print(vload_fit$cmdstan_diagnose())
print(vload_fit$summary(variables=vload_params))
sink()

# Diagnostic plots
pdf('Output/P1_VLOAD_MCMC_DiagnosticPlots.pdf')
mcmc_trace(vload_draws, pars = vload_params)
mcmc_dens_overlay(vload_draws, pars = vload_params)
mcmc_acf(vload_draws, pars = vload_params)
dev.off()
```

We have that on average, the number of HIV copies in a mL of blood 2 years after HAART is 0.14 units lower for subjects who report hard drug use at baseline compared to subjects who report no hard drug use at baseline. However, this is not statistically significant (P=0.77). 
In other words, there is no statistically significant difference in viral load between subjects who report hard drug use and subjects who did not report hard drug use at baseline. 


\newpage

## LEU3N

For CD4+ T cell counts, we specified an vague Normal $(0, 10000)$ prior for the regression coefficients and a Half-Normal $(0, 100^2)$ prior for the residual variance.


For CD4+ T cell counts, we specified an vague Normal$(0, 10000)$ prior for the regression coefficients and a Half-Normal$(0, 100^2)$ prior for the residual variance.


```{r, echo = F}
# Read in data
data <- read.csv("../Data/final_hiv_data.csv")

# Assessment of the data first
# str(data)

# Create categorical labels and variables 

# Hard drugs
data$hard_drugs_cat <- factor(data$hard_drugs,
                              levels = c(0, 1),
                              labels = c("No", "Yes"))

# Move after hard drugs variable
data <- data %>% 
  relocate(hard_drugs_cat, .after = hard_drugs)

# Convert collapse race to factor
data$collapse_RACE <- factor(data$collapse_RACE)

# Convert collapse educ to factor
data$collapsed_EDUCBAS <- factor(data$collapsed_EDUCBAS)

# Convert SMOKE cat to factor
data$collapse_SMOKE <- factor(data$collapse_SMOKE)

# Relabel the variables
med_data <- apply_labels(data,
                   hard_drugs = "Hard Drug Use at BL",
                   ADH_yr2 = "HAART Adherence",
                   VLOAD_yr2 = "VLOAD - yr2")
```



```{r, echo = F}
################################################################################
###                                    LEU3N                                 ###
################################################################################


### Linear model WITHOUT ADHERENCE
leu3n_linear <- lm(LEU3N_yr2 ~ hard_drugs_cat + LEU3N_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS,
                     data = data)

# sink("Output/P1_leu3n_Freq_Output.txt")
# print(leu3n_linear)
# summary(leu3n_linear)
# confint(leu3n_linear)
# anova(leu3n_linear)
# sink()

``` 

```{r, echo = F}
### Linear model WITH ADHERENCE
leu3n_adh_linear <- lm(LEU3N_yr2 ~ hard_drugs_cat + LEU3N_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS
                    + collapsed_ADH,
                     data = data)

```



```{r, echo = F}
### Bayesian Model

# Linear model WITHOUT ADHERENCE

###########################################################################
# STEP 1: Define the model
# Normal priors on the regression coefficients
# Half normal priors on sigma
###########################################################################


stan_file_leu3n <- write_stan_file("data {
  int<lower=0> N;                  // number of observations
  int<lower=0> P;                  // number of predictors including intercept
  matrix[N, P] X;                  // design matrix (first column = intercept)
  vector[N] y;                     // outcome

  vector[P] prior_mean;            // prior means for each beta
  vector<lower=0>[P] prior_sd;     // prior SDs for each beta

  real<lower=0> sigma_prior_sd;    // SD for half-normal prior on sigma
}

parameters {
  vector[P] beta;                  // regression coefficients
  real<lower=0> sigma;             // residual SD
}

model {
  // Vectorized priors for regression coefficients
  beta ~ normal(prior_mean, prior_sd);

  // Half-normal prior for sigma
  sigma ~ normal(0, sigma_prior_sd);

  // Likelihood
  y ~ normal(X * beta, sigma);
  

}

generated quantities {
  // log likelihood for each observation for calculating model fit stats
  vector[N] log_lik;
  for (n in 1:N) {
    log_lik[n] = normal_lpdf(y[n] | X[n] * beta, sigma);}
  }", dir="STAN", basename='leu3n_primary'
)

```

```{r, echo = F}
# Linear Model WITH ADHERENCE

###########################################################################
# STEP 1: Define the model
# Normal priors on the regression coefficients
# Half normal priors on sigma
###########################################################################


stan_file_leu3n_adh <- write_stan_file("data {
  int<lower=0> N;                  // number of observations
  int<lower=0> P;                  // number of predictors including intercept
  matrix[N, P] X;                  // design matrix (first column = intercept)
  vector[N] y;                     // outcome

  vector[P] prior_mean;            // prior means for each beta
  vector<lower=0>[P] prior_sd;     // prior SDs for each beta

  real<lower=0> sigma_prior_sd;    // SD for half-normal prior on sigma
}

parameters {
  vector[P] beta;                  // regression coefficients
  real<lower=0> sigma;             // residual SD
}

model {
  // Vectorized priors for regression coefficients
  beta ~ normal(prior_mean, prior_sd);

  // Half-normal prior for sigma
  sigma ~ normal(0, sigma_prior_sd);

  // Likelihood
  y ~ normal(X * beta, sigma);
  

}

generated quantities {
  // log likelihood for each observation for calculating model fit stats
  vector[N] log_lik;
  for (n in 1:N) {
    log_lik[n] = normal_lpdf(y[n] | X[n] * beta, sigma);}
  }", dir="STAN", basename='leu3n_adh'
)

```


```{r, echo = F, message = F, warning = F}
###########################################################################
# STEP 2: Compile the Stan program
###########################################################################

# Model WITHOUT adherence
leu3n_mod <- cmdstan_model('STAN/leu3n_primary.stan')

```

```{r, echo = F, message = F, warning = F}
# Model WITH adherence
leu3n_mod_adh <- cmdstan_model('STAN/leu3n_adh.stan')

```


```{r, echo = F}
###########################################################################
# STEP 3: Prepare your data to pass to Stan
###########################################################################

# Model WITHOUT adherence
# Outcome data
y_leu3n <- data$LEU3N_yr2

# Design matrix in our linear regression
X_leu3n <- model.matrix(~ hard_drugs_cat + LEU3N_yr0
   + BMI + collapse_SMOKE
   + age + collapse_RACE + collapsed_EDUCBAS,
   data = data)

# These are variables needed in Stan
# N is the number of observations in your data set 
# P is the number of columns in the design matrix
N_leu3n <- nrow(X_leu3n)
P_leu3n <- ncol(X_leu3n)

# Hyper-parameters for prior distributions
# Prior for regression coefficients is Normal(0, 10000)
# Prior for model error SD (sigma) is Half-Normal(0, 100)

# MVN mean (mean in the prior on the regression coefficients)
m_leu3n <- c(1, rep(0, (P_leu3n - 1)))   

# # SD in the prior on regression coefficients --> variance 100^2
s_leu3n <- rep(1000, P_leu3n) 

sigma_sd_leu3n <- 1000

# create data list to pass to STAN
data_list <- list(
  N = N_leu3n,
  P = P_leu3n,
  X = X_leu3n,
  y = y_leu3n,
  prior_mean = m_leu3n,
  prior_sd = s_leu3n,
  sigma_prior_sd = sigma_sd_leu3n
)
```


```{r, echo = F}
# Model WITH adherence
# Outcome data
y_leu3n_adh <- data$LEU3N_yr2

# Design matrix in our linear regression
X_leu3n_adh <- model.matrix(~ hard_drugs_cat + LEU3N_yr0
   + BMI + collapse_SMOKE
   + age + collapse_RACE + collapsed_EDUCBAS
   + collapsed_ADH,
   data = data)

# These are variables needed in Stan
# N is the number of observations in your data set 
# P is the number of columns in the design matrix
N_leu3n_adh <- nrow(X_leu3n_adh)
P_leu3n_adh <- ncol(X_leu3n_adh)

# Hyper-parameters for prior distributions
# Prior for regression coefficients is Normal(0, 10000)
# Prior for model error SD (sigma) is Half-Normal(0, 100)

# MVN mean (mean in the prior on the regression coefficients)
m_leu3n_adh <- c(1, rep(0, (P_leu3n_adh - 1)))   

# # SD in the prior on regression coefficients --> variance 100^2
s_leu3n_adh <- rep(1000, P_leu3n_adh) 

sigma_sd_leu3n_adh <- 1000

# create data list to pass to STAN
data_list_adh <- list(
  N = N_leu3n_adh,
  P = P_leu3n_adh,
  X = X_leu3n_adh,
  y = y_leu3n_adh,
  prior_mean = m_leu3n_adh,
  prior_sd = s_leu3n_adh,
  sigma_prior_sd = sigma_sd_leu3n_adh
)
```


```{r, echo = F, message=F, warning = F, results = 'hide'}
###########################################################################
# STEP 4: Fit Model
###########################################################################

# Model WITHOUT adherence
# Specify the warmup and sampling iterations
leu3n_fit <- suppressMessages(
  suppressWarnings(
  leu3n_mod$sample(
  data = data_list,
  chains = 4,
  iter_warmup = 1000,
  iter_sampling = 2000,
  seed = 234,
  refresh = 0
 )
)
)

```

```{r, echo = F, message=F, warning = F, results = 'hide'}
# Model WITH adherence
# Specify the warmup and sampling iterations
leu3n_fit_adh <- leu3n_mod_adh$sample(
  data = data_list_adh,
  chains = 4,
  iter_warmup = 1000,
  iter_sampling = 2000,
  seed = 456
)
```


```{r, echo = F, results='hide'}
###########################################################################
# STEP 5: Summarize the posterior
###########################################################################

# Model without ADH
leu3n_fit$summary(variables=c("beta[1]", "beta[2]", 
                              "beta[3]",
        # including extra betas to double check results with linear model
                              "beta[4]", "beta[5]",
                              "beta[6]", "beta[7]",
                              "beta[8]", 
                              "sigma"))


# Fancier Table of Results of Interest from the Posterior
# Extract posterior draws from cmdstanr fit
draws_leu3n <- leu3n_fit$draws()  

draws_mat_leu3n <- as_draws_matrix(draws_leu3n)
params_leu3n <- colnames(draws_mat_leu3n)

# Exclude non-parameter columns from summarization: lp__ and log_lik[n]
params_leu3n <- params_leu3n[!grepl("lp__|log_lik", params_leu3n)]

```

```{r, echo = F, results = 'hide'}
# Model WITH ADH

leu3n_fit_adh$summary(variables=c("beta[1]", "beta[2]", 
                              "beta[3]",
        # including extra betas to double check results with linear model
                              "beta[4]", "beta[5]",
                              "beta[6]", "beta[7]",
                              "beta[8]", 
                              "sigma"))


# Fancier Table of Results of Interest from the Posterior
# Extract posterior draws from cmdstanr fit
draws_leu3n_adh <- leu3n_fit_adh$draws()  

draws_mat_leu3n_adh <- as_draws_matrix(draws_leu3n_adh)
params_leu3n_adh <- colnames(draws_mat_leu3n_adh)

# Exclude non-parameter columns from summarization: lp__ and log_lik[n]
params_leu3n_adh <- params_leu3n_adh[!grepl("lp__|log_lik", params_leu3n_adh)]

```

```{r, echo = F, results = 'hide'}
# Model without ADH
# Build summary table
summary_table_leu3n <- lapply(params_leu3n, function(p) {
  vals <- as.numeric(draws_mat_leu3n[, p])
  
  # Monte Carlo standard error
  # Used to determine if we have run enough iterations
  # rule of thumb: MCSE should be less than 6.27% of the posterior standard deviation
  mcse_val <- mcmcse::mcse(vals)$se
  
  # Effective sample size
  ess_val <- ess_bulk(vals)
  
  # 95% HPDI
  hpd <- hdi(vals, ci = 0.95)
  
  tibble(
    Parameter = p,
    Estimate  = mean(vals),
    MCSE      = mcse_val,
    Std_Dev   = sd(vals),
    HPDI_2.5  = hpd$CI_low,
    HPDI_97.5 = hpd$CI_high,
    ESS       = ess_val
  )}) %>% 
  bind_rows()

print(summary_table_leu3n)

# Compare with linear fit
summary(leu3n_linear)

# check that mcmcse < 6% of posterior SD
(100*summary_table_leu3n$MCSE/summary_table_leu3n$Std_Dev) < 6
# All True


```


```{r, echo = F, results = 'hide'}
# Model WITH ADH
# Build summary table
summary_table_leu3n_adh <- lapply(params_leu3n_adh, function(p) {
  vals <- as.numeric(draws_mat_leu3n_adh[, p])
  
  # Monte Carlo standard error
  # Used to determine if we have run enough iterations
  # rule of thumb: MCSE should be less than 6.27% of the posterior standard deviation
  mcse_val <- mcmcse::mcse(vals)$se
  
  # Effective sample size
  ess_val <- ess_bulk(vals)
  
  # 95% HPDI
  hpd <- hdi(vals, ci = 0.95)
  
  tibble(
    Parameter = p,
    Estimate  = mean(vals),
    MCSE      = mcse_val,
    Std_Dev   = sd(vals),
    HPDI_2.5  = hpd$CI_low,
    HPDI_97.5 = hpd$CI_high,
    ESS       = ess_val
  )}) %>% 
  bind_rows()

print(summary_table_leu3n_adh)

# Compare with linear fit
summary(leu3n_adh_linear)

# check that mcmcse < 6% of posterior SD
(100*summary_table_leu3n_adh$MCSE/summary_table_leu3n_adh$Std_Dev) < 6
# All True

```


```{r, echo = F}
# Compare estimates for models with and without adherence

# Model without ADH
leu3n_terms <- summary_table_leu3n[2, ]

# Subset to Parameter, Estimate, HPDIs
leu3n_terms <- subset(leu3n_terms,
                      select = c("Parameter",
                                 "Estimate",
                                 "HPDI_2.5",
                                 "HPDI_97.5"))

# Model WITH ADH
# Pull term for hard drugs variable
leu3n_adh_terms <- summary_table_leu3n_adh[2, ]

# Subset to Parameter, Estimate, HPDIs
leu3n_adh_terms <- subset(leu3n_adh_terms,
                    select = c("Parameter", 
                               "Estimate",
                               "HPDI_2.5",
                               "HPDI_97.5"))

# Display side by side like with frequentist models
# Create data frame
comp_models_leu3n <- data.frame(Variable = "HD at Baseline (Yes)",
                 Estimate = round(leu3n_terms$Estimate, 3),
                 `95% HDI` = paste(round(leu3n_terms$HPDI_2.5, 3), 
                                    round(leu3n_terms$HPDI_97.5, 3), 
                                    sep = ", "),
                 Estimate = round(leu3n_adh_terms$Estimate, 3),
                 `95% HDI` = paste(round(leu3n_adh_terms$HPDI_2.5, 3),
                                    round(leu3n_adh_terms$HPDI_97.5, 3),
                                    sep = ", "),
                 check.names = FALSE)

# Display with kable
kable(
  comp_models_leu3n,
  # format = "latex",
  booktabs = TRUE,
  digits = 3,
  caption = "Model Summary") |>
  kable_styling(latex_options = c("striped", "hold_position")) |>
  add_header_above(c(" ", "Model Without ADH" = 2, 
                     "Model With ADH" = 2)) %>% 
  footnote(general = "HPDI - Highest Density Interval ")

```


```{r, echo = F}
###########################################################################
# STEP 6: Model fit statistics (you only need 1 of these, not all 3)
###########################################################################
# Extract the log likelihood info
# For leu3n model without ADH
loglik_mat_leu3n <- as_draws_matrix(leu3n_fit$draws("log_lik")) 
# iterations x N

# Get the LOO-CV
loo_res_leu3n <- loo(loglik_mat_leu3n)
# looic: 6124.8 

```

```{r, echo = F}
# Model WITH ADH
loglik_mat_leu3n_adh <- as_draws_matrix(leu3n_fit_adh$draws("log_lik")) 
# iterations x N

# Get the LOO-CV
loo_res_leu3n_adh <- loo(loglik_mat_leu3n_adh)
# print(loo_res_leu3n_adh)
# 6122.8
```


```{r, echo = F}
###########################################################################
# STEP 6: MCMC Diagnostics
###########################################################################

# Model without ADH
# Dimensions: iterations x chains x parameters
leu3n_draws <- as_draws_array(leu3n_fit$draws())

# Tidy format for bayesplot / ggplot
leu3n_draws_df <- as_draws_df(leu3n_fit$draws())  

# parameters
leu3n_params <- c("beta[1]", "beta[2]", 
            "beta[3]", "beta[4]",
            "beta[5]", "beta[6]",
            "beta[7]", "beta[8]",
            "sigma")

```

```{r, echo = F}
# Dimensions: iterations x chains x parameters
leu3n_draws_adh <- as_draws_array(leu3n_fit_adh$draws())

# Tidy format for bayesplot / ggplot
leu3n_draws_df_adh <- as_draws_df(leu3n_fit_adh$draws())  

# parameters
leu3n_params_adh <- c("beta[1]", "beta[2]", 
            "beta[3]", "beta[4]",
            "beta[5]", "beta[6]",
            "beta[7]", "beta[8]",
            "sigma")
```


```{r, echo = F, results = 'hide', eval = F}
# Trace plots
# Trace plots show whether chains are mixing well and exploring the parameter space.
# Chains should mix well (lines overlap)
# No long trends (no “stickiness”)
# Ideally all chains overlap around the same mean

# Model without ADH
mcmc_trace(leu3n_draws, pars = leu3n_params)
# They look beautiful

```


```{r, echo = F, results = 'hide', eval = F}
# Model WITH ADH
mcmc_trace(leu3n_draws_adh, pars = leu3n_params_adh)
```



```{r, echo = F, results = 'hide'}
# R-hat and ESS diagnostics
# R-hat ≈ 1.00 → converged
# ESS should be reasonably large (e.g., >100–200 per parameter)
leu3n_fit$summary(variables = leu3n_params)
# We see that rhats are all approx 1
# ESS should be around > 200*9 = 1800 and that seems to be the case for most
```

```{r, echo = F, results='hide'}
# Model WITH ADH
leu3n_fit_adh$summary(variables = leu3n_params_adh)

```


```{r, echo = F, results='hide', eval = F}
# Posterior Density Plots
# Overlaid densities from multiple chains should match closely
# If densities differ substantially between chains → poor mixing
mcmc_dens_overlay(draws_leu3n, pars = leu3n_params)

```

```{r, echo = F, include = F, eval = F, results='hide'}
# Posterior density plots
# Model WITH ADH
mcmc_dens_overlay(draws_leu3n_adh, pars = leu3n_params_adh)
```


```{r, echo = F, results='hide', eval = F}
# Auto-correlation 
# Autocorrelation should decay quickly
# High autocorrelation → may need more iterations or thinning
mcmc_acf(draws_leu3n, pars = leu3n_params)

```

```{r, echo = F, results='hide', eval = F}
# Model WITH ADH
mcmc_acf(draws_leu3n_adh, pars = leu3n_params_adh)
```


```{r, echo = F, results='hide', eval = F}
# Diagnostics from cmdstan
leu3n_fit$cmdstan_diagnose()
# Need to load actual leu3n and not leu3n .rds for this

### Output
# Checking sampler transitions treedepth.
# Treedepth satisfactory for all transitions.
# 
# Checking sampler transitions for divergences.
# No divergent transitions found.
# 
# Checking E-BFMI - sampler transitions HMC potential energy.
# E-BFMI satisfactory.
# 
# Rank-normalized split effective sample size satisfactory for all parameters.
# 
# Rank-normalized split R-hat values satisfactory for all parameters.
# 
# Processing complete, no problems detected.

```

```{r, echo = F, results='hide', eval = F}
# Model WITH ADH
# Diagnostics from cmdstan
leu3n_fit_adh$cmdstan_diagnose()

### Output
# Checking sampler transitions treedepth.
# Treedepth satisfactory for all transitions.
# 
# Checking sampler transitions for divergences.
# No divergent transitions found.
# 
# Checking E-BFMI - sampler transitions HMC potential energy.
# E-BFMI satisfactory.
# 
# Rank-normalized split effective sample size satisfactory for all parameters.
# 
# Rank-normalized split R-hat values satisfactory for all parameters.
# 
# Processing complete, no problems detected.
```


```{r, echo = F, eval = F}
## Save output if desired ###
# MCMC Output
sink("Output/P1_leu3n_MCMCOutput.txt")
print(summary_table)
print(loo_res)
print(leu3n_fit$cmdstan_diagnose())
print(leu3n_fit$summary(variables=leu3n_params))
sink()

# Diagnostic plots
pdf('Output/P1_leu3n_MCMC_DiagnosticPlots.pdf')
mcmc_trace(leu3n_draws, pars = leu3n_params)
mcmc_dens_overlay(leu3n_draws, pars = leu3n_params)
mcmc_acf(leu3n_draws, pars = leu3n_params)
dev.off()

```

## AGG PHYS

For aggregate physical quality of life scores, we specified an vague Normal$(0, 10000)$ prior for the regression coefficients and a Half-Normal$(0, 100^2)$ prior for the residual variance.


```{r, echo = F}
# Read in data
data <- read.csv("../Data/final_hiv_data.csv")

# Assessment of the data first
# str(data)

# Create categorical labels and variables 

# Hard drugs
data$hard_drugs_cat <- factor(data$hard_drugs,
                              levels = c(0, 1),
                              labels = c("No", "Yes"))

# Move after hard drugs variable
data <- data %>% 
  relocate(hard_drugs_cat, .after = hard_drugs)

# Convert collapse race to factor
data$collapse_RACE <- factor(data$collapse_RACE)

# Convert collapse educ to factor
data$collapsed_EDUCBAS <- factor(data$collapsed_EDUCBAS)

# Convert SMOKE cat to factor
data$collapse_SMOKE <- factor(data$collapse_SMOKE)

# Relabel the variables
med_data <- apply_labels(data,
                   hard_drugs = "Hard Drug Use at BL",
                   ADH_yr2 = "HAART Adherence",
                   VLOAD_yr2 = "VLOAD - yr2")
```



```{r, echo = F}
################################################################################
###                                AGG PHYS                                  ###
################################################################################


### Linear model WITHOUT ADHERENCE
aggphys_linear <- lm(AGG_PHYS_yr2 ~ hard_drugs_cat + AGG_PHYS_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS,
                     data = data)

# sink("Output/P1_aggphys_Freq_Output.txt")
# print(aggphys_linear)
# summary(aggphys_linear)
# confint(aggphys_linear)
# anova(aggphys_linear)
# sink()

``` 

```{r, echo = F}
### Linear model WITH ADHERENCE
aggphys_adh_linear <- lm(AGG_PHYS_yr2 ~ hard_drugs_cat + AGG_PHYS_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS
                    + collapsed_ADH,
                     data = data)

```



```{r, echo = F}
### Bayesian Model

# Linear model WITHOUT ADHERENCE

###########################################################################
# STEP 1: Define the model
# Normal priors on the regression coefficients
# Half normal priors on sigma
###########################################################################


stan_file_aggphys <- write_stan_file("data {
  int<lower=0> N;                  // number of observations
  int<lower=0> P;                  // number of predictors including intercept
  matrix[N, P] X;                  // design matrix (first column = intercept)
  vector[N] y;                     // outcome

  vector[P] prior_mean;            // prior means for each beta
  vector<lower=0>[P] prior_sd;     // prior SDs for each beta

  real<lower=0> sigma_prior_sd;    // SD for half-normal prior on sigma
}

parameters {
  vector[P] beta;                  // regression coefficients
  real<lower=0> sigma;             // residual SD
}

model {
  // Vectorized priors for regression coefficients
  beta ~ normal(prior_mean, prior_sd);

  // Half-normal prior for sigma
  sigma ~ normal(0, sigma_prior_sd);

  // Likelihood
  y ~ normal(X * beta, sigma);
  

}

generated quantities {
  // log likelihood for each observation for calculating model fit stats
  vector[N] log_lik;
  for (n in 1:N) {
    log_lik[n] = normal_lpdf(y[n] | X[n] * beta, sigma);}
  }", dir="STAN", basename='aggphys_primary'
)

```

```{r, echo = F}
# Linear Model WITH ADHERENCE

###########################################################################
# STEP 1: Define the model
# Normal priors on the regression coefficients
# Half normal priors on sigma
###########################################################################


stan_file_aggphys_adh <- write_stan_file("data {
  int<lower=0> N;                  // number of observations
  int<lower=0> P;                  // number of predictors including intercept
  matrix[N, P] X;                  // design matrix (first column = intercept)
  vector[N] y;                     // outcome

  vector[P] prior_mean;            // prior means for each beta
  vector<lower=0>[P] prior_sd;     // prior SDs for each beta

  real<lower=0> sigma_prior_sd;    // SD for half-normal prior on sigma
}

parameters {
  vector[P] beta;                  // regression coefficients
  real<lower=0> sigma;             // residual SD
}

model {
  // Vectorized priors for regression coefficients
  beta ~ normal(prior_mean, prior_sd);

  // Half-normal prior for sigma
  sigma ~ normal(0, sigma_prior_sd);

  // Likelihood
  y ~ normal(X * beta, sigma);
  

}

generated quantities {
  // log likelihood for each observation for calculating model fit stats
  vector[N] log_lik;
  for (n in 1:N) {
    log_lik[n] = normal_lpdf(y[n] | X[n] * beta, sigma);}
  }", dir="STAN", basename='aggphys_adh'
)

```


```{r, echo = F, message = F, warning = F}
###########################################################################
# STEP 2: Compile the Stan program
###########################################################################

# Model WITHOUT adherence
aggphys_mod <- cmdstan_model('STAN/aggphys_primary.stan')

```

```{r, echo = F, message = F, warning = F}
# Model WITH adherence
aggphys_mod_adh <- cmdstan_model('STAN/aggphys_adh.stan')

```


```{r, echo = F}
###########################################################################
# STEP 3: Prepare your data to pass to Stan
###########################################################################

# Model WITHOUT adherence
# Outcome data
y_aggphys <- data$AGG_PHYS_yr2

# Design matrix in our linear regression
X_aggphys <- model.matrix(~ hard_drugs_cat + AGG_PHYS_yr0
   + BMI + collapse_SMOKE
   + age + collapse_RACE + collapsed_EDUCBAS,
   data = data)

# These are variables needed in Stan
# N is the number of observations in your data set 
# P is the number of columns in the design matrix
N_aggphys <- nrow(X_aggphys)
P_aggphys <- ncol(X_aggphys)

# Hyper-parameters for prior distributions
# Prior for regression coefficients is Normal(0, 10000)
# Prior for model error SD (sigma) is Half-Normal(0, 100)

# MVN mean (mean in the prior on the regression coefficients)
m_aggphys <- c(1, rep(0, (P_aggphys - 1)))   

# # SD in the prior on regression coefficients --> variance 100^2
s_aggphys <- rep(1000, P_aggphys) 

sigma_sd_aggphys <- 1000

# create data list to pass to STAN
data_list <- list(
  N = N_aggphys,
  P = P_aggphys,
  X = X_aggphys,
  y = y_aggphys,
  prior_mean = m_aggphys,
  prior_sd = s_aggphys,
  sigma_prior_sd = sigma_sd_aggphys
)
```


```{r, echo = F}
# Model WITH adherence
# Outcome data
y_aggphys_adh <- data$AGG_PHYS_yr2

# Design matrix in our linear regression
X_aggphys_adh <- model.matrix(~ hard_drugs_cat + AGG_PHYS_yr0
   + BMI + collapse_SMOKE
   + age + collapse_RACE + collapsed_EDUCBAS
   + collapsed_ADH,
   data = data)

# These are variables needed in Stan
# N is the number of observations in your data set 
# P is the number of columns in the design matrix
N_aggphys_adh <- nrow(X_aggphys_adh)
P_aggphys_adh <- ncol(X_aggphys_adh)

# Hyper-parameters for prior distributions
# Prior for regression coefficients is Normal(0, 10000)
# Prior for model error SD (sigma) is Half-Normal(0, 100)

# MVN mean (mean in the prior on the regression coefficients)
m_aggphys_adh <- c(1, rep(0, (P_aggphys_adh - 1)))   

# # SD in the prior on regression coefficients --> variance 100^2
s_aggphys_adh <- rep(1000, P_aggphys_adh) 

sigma_sd_aggphys_adh <- 1000

# create data list to pass to STAN
data_list_adh <- list(
  N = N_aggphys_adh,
  P = P_aggphys_adh,
  X = X_aggphys_adh,
  y = y_aggphys_adh,
  prior_mean = m_aggphys_adh,
  prior_sd = s_aggphys_adh,
  sigma_prior_sd = sigma_sd_aggphys_adh
)
```


```{r, echo = F, message=F, warning = F, results = 'hide'}
###########################################################################
# STEP 4: Fit Model
###########################################################################

# Model WITHOUT adherence
# Specify the warmup and sampling iterations
aggphys_fit <- suppressMessages(
  suppressWarnings(
  aggphys_mod$sample(
  data = data_list,
  chains = 4,
  iter_warmup = 1000,
  iter_sampling = 2000,
  seed = 235,
  refresh = 0
 )
)
)

```

```{r, echo = F, message=F, warning = F, results = 'hide'}
# Model WITH adherence
# Specify the warmup and sampling iterations
aggphys_fit_adh <- aggphys_mod_adh$sample(
  data = data_list_adh,
  chains = 4,
  iter_warmup = 1000,
  iter_sampling = 2000,
  seed = 456
)
```


```{r, echo = F, results='hide'}
###########################################################################
# STEP 5: Summarize the posterior
###########################################################################

# Model without ADH
aggphys_fit$summary(variables=c("beta[1]", "beta[2]", 
                              "beta[3]",
        # including extra betas to double check results with linear model
                              "beta[4]", "beta[5]",
                              "beta[6]", "beta[7]",
                              "beta[8]", 
                              "sigma"))


# Fancier Table of Results of Interest from the Posterior
# Extract posterior draws from cmdstanr fit
draws_aggphys <- aggphys_fit$draws()  

draws_mat_aggphys <- as_draws_matrix(draws_aggphys)
params_aggphys <- colnames(draws_mat_aggphys)

# Exclude non-parameter columns from summarization: lp__ and log_lik[n]
params_aggphys <- params_aggphys[!grepl("lp__|log_lik", params_aggphys)]

```

```{r, echo = F, results = 'hide'}
# Model WITH ADH

aggphys_fit_adh$summary(variables=c("beta[1]", "beta[2]", 
                              "beta[3]",
        # including extra betas to double check results with linear model
                              "beta[4]", "beta[5]",
                              "beta[6]", "beta[7]",
                              "beta[8]", 
                              "sigma"))


# Fancier Table of Results of Interest from the Posterior
# Extract posterior draws from cmdstanr fit
draws_aggphys_adh <- aggphys_fit_adh$draws()  

draws_mat_aggphys_adh <- as_draws_matrix(draws_aggphys_adh)
params_aggphys_adh <- colnames(draws_mat_aggphys_adh)

# Exclude non-parameter columns from summarization: lp__ and log_lik[n]
params_aggphys_adh <- params_aggphys_adh[!grepl("lp__|log_lik", params_aggphys_adh)]

```

```{r, echo = F, results = 'hide'}
# Model without ADH
# Build summary table
summary_table_aggphys <- lapply(params_aggphys, function(p) {
  vals <- as.numeric(draws_mat_aggphys[, p])
  
  # Monte Carlo standard error
  # Used to determine if we have run enough iterations
  # rule of thumb: MCSE should be less than 6.27% of the posterior standard deviation
  mcse_val <- mcmcse::mcse(vals)$se
  
  # Effective sample size
  ess_val <- ess_bulk(vals)
  
  # 95% HPDI
  hpd <- hdi(vals, ci = 0.95)
  
  tibble(
    Parameter = p,
    Estimate  = mean(vals),
    MCSE      = mcse_val,
    Std_Dev   = sd(vals),
    HPDI_2.5  = hpd$CI_low,
    HPDI_97.5 = hpd$CI_high,
    ESS       = ess_val
  )}) %>% 
  bind_rows()

print(summary_table_aggphys)

# Compare with linear fit
summary(aggphys_linear)

# check that mcmcse < 6% of posterior SD
(100*summary_table_aggphys$MCSE/summary_table_aggphys$Std_Dev) < 6
# All True


```


```{r, echo = F, results = 'hide'}
# Model WITH ADH
# Build summary table
summary_table_aggphys_adh <- lapply(params_aggphys_adh, function(p) {
  vals <- as.numeric(draws_mat_aggphys_adh[, p])
  
  # Monte Carlo standard error
  # Used to determine if we have run enough iterations
  # rule of thumb: MCSE should be less than 6.27% of the posterior standard deviation
  mcse_val <- mcmcse::mcse(vals)$se
  
  # Effective sample size
  ess_val <- ess_bulk(vals)
  
  # 95% HPDI
  hpd <- hdi(vals, ci = 0.95)
  
  tibble(
    Parameter = p,
    Estimate  = mean(vals),
    MCSE      = mcse_val,
    Std_Dev   = sd(vals),
    HPDI_2.5  = hpd$CI_low,
    HPDI_97.5 = hpd$CI_high,
    ESS       = ess_val
  )}) %>% 
  bind_rows()

print(summary_table_aggphys_adh)

# Compare with linear fit
summary(aggphys_adh_linear)

# check that mcmcse < 6% of posterior SD
(100*summary_table_aggphys_adh$MCSE/summary_table_aggphys_adh$Std_Dev) < 6
# All True

```


```{r, echo = F}
# Compare estimates for models with and without adherence

# Model without ADH
aggphys_terms <- summary_table_aggphys[2, ]

# Subset to Parameter, Estimate, HPDIs
aggphys_terms <- subset(aggphys_terms,
                      select = c("Parameter",
                                 "Estimate",
                                 "HPDI_2.5",
                                 "HPDI_97.5"))

# Model WITH ADH
# Pull term for hard drugs variable
aggphys_adh_terms <- summary_table_aggphys_adh[2, ]

# Subset to Parameter, Estimate, HPDIs
aggphys_adh_terms <- subset(aggphys_adh_terms,
                    select = c("Parameter", 
                               "Estimate",
                               "HPDI_2.5",
                               "HPDI_97.5"))

# Display side by side like with frequentist models
# Create data frame
comp_models_aggphys <- data.frame(Variable = "HD at Baseline (Yes)",
                 Estimate = round(aggphys_terms$Estimate, 3),
                 `95% HDI` = paste(round(aggphys_terms$HPDI_2.5, 3), 
                                    round(aggphys_terms$HPDI_97.5, 3), 
                                    sep = ", "),
                 Estimate = round(aggphys_adh_terms$Estimate, 3),
                 `95% HDI` = paste(round(aggphys_adh_terms$HPDI_2.5, 3),
                                    round(aggphys_adh_terms$HPDI_97.5, 3),
                                    sep = ", "),
                 check.names = FALSE)

# Display with kable
kable(
  comp_models_aggphys,
  # format = "latex",
  booktabs = TRUE,
  digits = 3,
  caption = "Model Summary") |>
  kable_styling(latex_options = c("striped", "hold_position")) |>
  add_header_above(c(" ", "Model Without ADH" = 2, 
                     "Model With ADH" = 2)) %>% 
  footnote(general = "HPDI - Highest Density Interval ")

```


```{r, echo = F}
###########################################################################
# STEP 6: Model fit statistics (you only need 1 of these, not all 3)
###########################################################################
# Extract the log likelihood info
# For aggphys model without ADH
loglik_mat_aggphys <- as_draws_matrix(aggphys_fit$draws("log_lik")) 
# iterations x N

# Get the LOO-CV
loo_res_aggphys <- loo(loglik_mat_aggphys)
# looic: 3223.6 

```

```{r, echo = F}
# Model WITH ADH
loglik_mat_aggphys_adh <- as_draws_matrix(aggphys_fit_adh$draws("log_lik")) 
# iterations x N

# Get the LOO-CV
loo_res_aggphys_adh <- loo(loglik_mat_aggphys_adh)
# print(loo_res_aggphys_adh)
# 3222.5
```


```{r, echo = F}
###########################################################################
# STEP 6: MCMC Diagnostics
###########################################################################

# Model without ADH
# Dimensions: iterations x chains x parameters
aggphys_draws <- as_draws_array(aggphys_fit$draws())

# Tidy format for bayesplot / ggplot
aggphys_draws_df <- as_draws_df(aggphys_fit$draws())  

# parameters
aggphys_params <- c("beta[1]", "beta[2]", 
            "beta[3]", "beta[4]",
            "beta[5]", "beta[6]",
            "beta[7]", "beta[8]",
            "sigma")

```

```{r, echo = F}
# Dimensions: iterations x chains x parameters
aggphys_draws_adh <- as_draws_array(aggphys_fit_adh$draws())

# Tidy format for bayesplot / ggplot
aggphys_draws_df_adh <- as_draws_df(aggphys_fit_adh$draws())  

# parameters
aggphys_params_adh <- c("beta[1]", "beta[2]", 
            "beta[3]", "beta[4]",
            "beta[5]", "beta[6]",
            "beta[7]", "beta[8]",
            "sigma")
```


```{r, echo = F, results = 'hide', eval = F}
# Trace plots
# Trace plots show whether chains are mixing well and exploring the parameter space.
# Chains should mix well (lines overlap)
# No long trends (no “stickiness”)
# Ideally all chains overlap around the same mean

# Model without ADH
mcmc_trace(aggphys_draws, pars = aggphys_params)
# They look beautiful

```


```{r, echo = F, results = 'hide', eval = F}
# Model WITH ADH
mcmc_trace(aggphys_draws_adh, pars = aggphys_params_adh)
```



```{r, echo = F, results = 'hide'}
# R-hat and ESS diagnostics
# R-hat ≈ 1.00 → converged
# ESS should be reasonably large (e.g., >100–200 per parameter)
aggphys_fit$summary(variables = aggphys_params)
# We see that rhats are all approx 1
# ESS should be around > 200*9 = 1800 and that seems to be the case for most
```

```{r, echo = F, results='hide'}
# Model WITH ADH
aggphys_fit_adh$summary(variables = aggphys_params_adh)

```


```{r, echo = F, results='hide', eval = F}
# Posterior Density Plots
# Overlaid densities from multiple chains should match closely
# If densities differ substantially between chains → poor mixing
mcmc_dens_overlay(draws_aggphys, pars = aggphys_params)

```

```{r, echo = F, include = F, eval = F, results='hide'}
# Posterior density plots
# Model WITH ADH
mcmc_dens_overlay(draws_aggphys_adh, pars = aggphys_params_adh)
```


```{r, echo = F, results='hide', eval = F}
# Auto-correlation 
# Autocorrelation should decay quickly
# High autocorrelation → may need more iterations or thinning
mcmc_acf(draws_aggphys, pars = aggphys_params)

```

```{r, echo = F, results='hide', eval = F}
# Model WITH ADH
mcmc_acf(draws_aggphys_adh, pars = aggphys_params_adh)
```


```{r, echo = F, results='hide', eval = F}
# Diagnostics from cmdstan
aggphys_fit$cmdstan_diagnose()
# Need to load actual aggphys and not aggphys .rds for this

### Output
# Checking sampler transitions treedepth.
# Treedepth satisfactory for all transitions.
# 
# Checking sampler transitions for divergences.
# No divergent transitions found.
# 
# Checking E-BFMI - sampler transitions HMC potential energy.
# E-BFMI satisfactory.
# 
# Rank-normalized split effective sample size satisfactory for all parameters.
# 
# Rank-normalized split R-hat values satisfactory for all parameters.
# 
# Processing complete, no problems detected

```

```{r, echo = F, results='hide', eval = F}
# Model WITH ADH
# Diagnostics from cmdstan
aggphys_fit_adh$cmdstan_diagnose()

### Output
# Checking sampler transitions treedepth.
# Treedepth satisfactory for all transitions.
# 
# Checking sampler transitions for divergences.
# No divergent transitions found.
# 
# Checking E-BFMI - sampler transitions HMC potential energy.
# E-BFMI satisfactory.
# 
# Rank-normalized split effective sample size satisfactory for all parameters.
# 
# Rank-normalized split R-hat values satisfactory for all parameters.
# 
# Processing complete, no problems detected.

```


```{r, echo = F, eval = F}
## Save output if desired ###
# MCMC Output
sink("Output/P1_aggphys_MCMCOutput.txt")
print(summary_table)
print(loo_res)
print(aggphys_fit$cmdstan_diagnose())
print(aggphys_fit$summary(variables=aggphys_params))
sink()

# Diagnostic plots
pdf('Output/P1_aggphys_MCMC_DiagnosticPlots.pdf')
mcmc_trace(aggphys_draws, pars = aggphys_params)
mcmc_dens_overlay(aggphys_draws, pars = aggphys_params)
mcmc_acf(aggphys_draws, pars = aggphys_params)
dev.off()

```

## AGG MENT

For aggregate mental quality of life scores, we specified an vague Normal$(0, 10000)$ prior for the regression coefficients and a Half-Normal$(0, 100^2)$ prior for the residual variance.


```{r, echo = F}
# Read in data
data <- read.csv("../Data/final_hiv_data.csv")

# Assessment of the data first
# str(data)

# Create categorical labels and variables 

# Hard drugs
data$hard_drugs_cat <- factor(data$hard_drugs,
                              levels = c(0, 1),
                              labels = c("No", "Yes"))

# Move after hard drugs variable
data <- data %>% 
  relocate(hard_drugs_cat, .after = hard_drugs)

# Convert collapse race to factor
data$collapse_RACE <- factor(data$collapse_RACE)

# Convert collapse educ to factor
data$collapsed_EDUCBAS <- factor(data$collapsed_EDUCBAS)

# Convert SMOKE cat to factor
data$collapse_SMOKE <- factor(data$collapse_SMOKE)

# Relabel the variables
med_data <- apply_labels(data,
                   hard_drugs = "Hard Drug Use at BL",
                   ADH_yr2 = "HAART Adherence",
                   VLOAD_yr2 = "VLOAD - yr2")
```



```{r, echo = F}
################################################################################
###                                AGG MENT                                  ###
################################################################################


### Linear model WITHOUT ADHERENCE
aggment_linear <- lm(AGG_PHYS_yr2 ~ hard_drugs_cat + AGG_PHYS_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS,
                     data = data)

# sink("Output/P1_aggment_Freq_Output.txt")
# print(aggment_linear)
# summary(aggment_linear)
# confint(aggment_linear)
# anova(aggment_linear)
# sink()

``` 

```{r, echo = F}
### Linear model WITH ADHERENCE
aggment_adh_linear <- lm(AGG_PHYS_yr2 ~ hard_drugs_cat + AGG_PHYS_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS
                    + collapsed_ADH,
                     data = data)

```



```{r, echo = F}
### Bayesian Model

# Linear model WITHOUT ADHERENCE

###########################################################################
# STEP 1: Define the model
# Normal priors on the regression coefficients
# Half normal priors on sigma
###########################################################################


stan_file_aggment <- write_stan_file("data {
  int<lower=0> N;                  // number of observations
  int<lower=0> P;                  // number of predictors including intercept
  matrix[N, P] X;                  // design matrix (first column = intercept)
  vector[N] y;                     // outcome

  vector[P] prior_mean;            // prior means for each beta
  vector<lower=0>[P] prior_sd;     // prior SDs for each beta

  real<lower=0> sigma_prior_sd;    // SD for half-normal prior on sigma
}

parameters {
  vector[P] beta;                  // regression coefficients
  real<lower=0> sigma;             // residual SD
}

model {
  // Vectorized priors for regression coefficients
  beta ~ normal(prior_mean, prior_sd);

  // Half-normal prior for sigma
  sigma ~ normal(0, sigma_prior_sd);

  // Likelihood
  y ~ normal(X * beta, sigma);
  

}

generated quantities {
  // log likelihood for each observation for calculating model fit stats
  vector[N] log_lik;
  for (n in 1:N) {
    log_lik[n] = normal_lpdf(y[n] | X[n] * beta, sigma);}
  }", dir="STAN", basename='aggment_primary'
)

```

```{r, echo = F}
# Linear Model WITH ADHERENCE

###########################################################################
# STEP 1: Define the model
# Normal priors on the regression coefficients
# Half normal priors on sigma
###########################################################################


stan_file_aggment_adh <- write_stan_file("data {
  int<lower=0> N;                  // number of observations
  int<lower=0> P;                  // number of predictors including intercept
  matrix[N, P] X;                  // design matrix (first column = intercept)
  vector[N] y;                     // outcome

  vector[P] prior_mean;            // prior means for each beta
  vector<lower=0>[P] prior_sd;     // prior SDs for each beta

  real<lower=0> sigma_prior_sd;    // SD for half-normal prior on sigma
}

parameters {
  vector[P] beta;                  // regression coefficients
  real<lower=0> sigma;             // residual SD
}

model {
  // Vectorized priors for regression coefficients
  beta ~ normal(prior_mean, prior_sd);

  // Half-normal prior for sigma
  sigma ~ normal(0, sigma_prior_sd);

  // Likelihood
  y ~ normal(X * beta, sigma);
  

}

generated quantities {
  // log likelihood for each observation for calculating model fit stats
  vector[N] log_lik;
  for (n in 1:N) {
    log_lik[n] = normal_lpdf(y[n] | X[n] * beta, sigma);}
  }", dir="STAN", basename='aggment_adh'
)

```


```{r, echo = F, message = F, warning = F}
###########################################################################
# STEP 2: Compile the Stan program
###########################################################################

# Model WITHOUT adherence
aggment_mod <- cmdstan_model('STAN/aggment_primary.stan')

```

```{r, echo = F, message = F, warning = F}
# Model WITH adherence
aggment_mod_adh <- cmdstan_model('STAN/aggment_adh.stan')

```


```{r, echo = F}
###########################################################################
# STEP 3: Prepare your data to pass to Stan
###########################################################################

# Model WITHOUT adherence
# Outcome data
y_aggment <- data$AGG_PHYS_yr2

# Design matrix in our linear regression
X_aggment <- model.matrix(~ hard_drugs_cat + AGG_PHYS_yr0
   + BMI + collapse_SMOKE
   + age + collapse_RACE + collapsed_EDUCBAS,
   data = data)

# These are variables needed in Stan
# N is the number of observations in your data set 
# P is the number of columns in the design matrix
N_aggment <- nrow(X_aggment)
P_aggment <- ncol(X_aggment)

# Hyper-parameters for prior distributions
# Prior for regression coefficients is Normal(0, 10000)
# Prior for model error SD (sigma) is Half-Normal(0, 100)

# MVN mean (mean in the prior on the regression coefficients)
m_aggment <- c(1, rep(0, (P_aggment - 1)))   

# # SD in the prior on regression coefficients --> variance 100^2
s_aggment <- rep(1000, P_aggment) 

sigma_sd_aggment <- 1000

# create data list to pass to STAN
data_list <- list(
  N = N_aggment,
  P = P_aggment,
  X = X_aggment,
  y = y_aggment,
  prior_mean = m_aggment,
  prior_sd = s_aggment,
  sigma_prior_sd = sigma_sd_aggment
)
```


```{r, echo = F}
# Model WITH adherence
# Outcome data
y_aggment_adh <- data$AGG_PHYS_yr2

# Design matrix in our linear regression
X_aggment_adh <- model.matrix(~ hard_drugs_cat + AGG_PHYS_yr0
   + BMI + collapse_SMOKE
   + age + collapse_RACE + collapsed_EDUCBAS
   + collapsed_ADH,
   data = data)

# These are variables needed in Stan
# N is the number of observations in your data set 
# P is the number of columns in the design matrix
N_aggment_adh <- nrow(X_aggment_adh)
P_aggment_adh <- ncol(X_aggment_adh)

# Hyper-parameters for prior distributions
# Prior for regression coefficients is Normal(0, 10000)
# Prior for model error SD (sigma) is Half-Normal(0, 100)

# MVN mean (mean in the prior on the regression coefficients)
m_aggment_adh <- c(1, rep(0, (P_aggment_adh - 1)))   

# # SD in the prior on regression coefficients --> variance 100^2
s_aggment_adh <- rep(1000, P_aggment_adh) 

sigma_sd_aggment_adh <- 1000

# create data list to pass to STAN
data_list_adh <- list(
  N = N_aggment_adh,
  P = P_aggment_adh,
  X = X_aggment_adh,
  y = y_aggment_adh,
  prior_mean = m_aggment_adh,
  prior_sd = s_aggment_adh,
  sigma_prior_sd = sigma_sd_aggment_adh
)
```


```{r, echo = F, message=F, warning = F, results = 'hide'}
###########################################################################
# STEP 4: Fit Model
###########################################################################

# Model WITHOUT adherence
# Specify the warmup and sampling iterations
aggment_fit <- suppressMessages(
  suppressWarnings(
  aggment_mod$sample(
  data = data_list,
  chains = 4,
  iter_warmup = 1000,
  iter_sampling = 2000,
  seed = 789,
  refresh = 0
 )
)
)

```

```{r, echo = F, message=F, warning = F, results = 'hide'}
# Model WITH adherence
# Specify the warmup and sampling iterations
aggment_fit_adh <- aggment_mod_adh$sample(
  data = data_list_adh,
  chains = 4,
  iter_warmup = 1000,
  iter_sampling = 2000,
  seed = 798
)
```


```{r, echo = F, results='hide'}
###########################################################################
# STEP 5: Summarize the posterior
###########################################################################

# Model without ADH
aggment_fit$summary(variables=c("beta[1]", "beta[2]", 
                              "beta[3]",
        # including extra betas to double check results with linear model
                              "beta[4]", "beta[5]",
                              "beta[6]", "beta[7]",
                              "beta[8]", 
                              "sigma"))


# Fancier Table of Results of Interest from the Posterior
# Extract posterior draws from cmdstanr fit
draws_aggment <- aggment_fit$draws()  

draws_mat_aggment <- as_draws_matrix(draws_aggment)
params_aggment <- colnames(draws_mat_aggment)

# Exclude non-parameter columns from summarization: lp__ and log_lik[n]
params_aggment <- params_aggment[!grepl("lp__|log_lik", params_aggment)]

```

```{r, echo = F, results = 'hide'}
# Model WITH ADH

aggment_fit_adh$summary(variables=c("beta[1]", "beta[2]", 
                              "beta[3]",
        # including extra betas to double check results with linear model
                              "beta[4]", "beta[5]",
                              "beta[6]", "beta[7]",
                              "beta[8]", 
                              "sigma"))


# Fancier Table of Results of Interest from the Posterior
# Extract posterior draws from cmdstanr fit
draws_aggment_adh <- aggment_fit_adh$draws()  

draws_mat_aggment_adh <- as_draws_matrix(draws_aggment_adh)
params_aggment_adh <- colnames(draws_mat_aggment_adh)

# Exclude non-parameter columns from summarization: lp__ and log_lik[n]
params_aggment_adh <- params_aggment_adh[!grepl("lp__|log_lik", params_aggment_adh)]

```

```{r, echo = F, results = 'hide'}
# Model without ADH
# Build summary table
summary_table_aggment <- lapply(params_aggment, function(p) {
  vals <- as.numeric(draws_mat_aggment[, p])
  
  # Monte Carlo standard error
  # Used to determine if we have run enough iterations
  # rule of thumb: MCSE should be less than 6.27% of the posterior standard deviation
  mcse_val <- mcmcse::mcse(vals)$se
  
  # Effective sample size
  ess_val <- ess_bulk(vals)
  
  # 95% HPDI
  hpd <- hdi(vals, ci = 0.95)
  
  tibble(
    Parameter = p,
    Estimate  = mean(vals),
    MCSE      = mcse_val,
    Std_Dev   = sd(vals),
    HPDI_2.5  = hpd$CI_low,
    HPDI_97.5 = hpd$CI_high,
    ESS       = ess_val
  )}) %>% 
  bind_rows()

print(summary_table_aggment)

# Compare with linear fit
summary(aggment_linear)

# check that mcmcse < 6% of posterior SD
(100*summary_table_aggment$MCSE/summary_table_aggment$Std_Dev) < 6
# All True


```


```{r, echo = F, results = 'hide'}
# Model WITH ADH
# Build summary table
summary_table_aggment_adh <- lapply(params_aggment_adh, function(p) {
  vals <- as.numeric(draws_mat_aggment_adh[, p])
  
  # Monte Carlo standard error
  # Used to determine if we have run enough iterations
  # rule of thumb: MCSE should be less than 6.27% of the posterior standard deviation
  mcse_val <- mcmcse::mcse(vals)$se
  
  # Effective sample size
  ess_val <- ess_bulk(vals)
  
  # 95% HPDI
  hpd <- hdi(vals, ci = 0.95)
  
  tibble(
    Parameter = p,
    Estimate  = mean(vals),
    MCSE      = mcse_val,
    Std_Dev   = sd(vals),
    HPDI_2.5  = hpd$CI_low,
    HPDI_97.5 = hpd$CI_high,
    ESS       = ess_val
  )}) %>% 
  bind_rows()

print(summary_table_aggment_adh)

# Compare with linear fit
summary(aggment_adh_linear)

# check that mcmcse < 6% of posterior SD
(100*summary_table_aggment_adh$MCSE/summary_table_aggment_adh$Std_Dev) < 6
# All True

```


```{r, echo = F}
# Compare estimates for models with and without adherence

# Model without ADH
aggment_terms <- summary_table_aggment[2, ]

# Subset to Parameter, Estimate, HPDIs
aggment_terms <- subset(aggment_terms,
                      select = c("Parameter",
                                 "Estimate",
                                 "HPDI_2.5",
                                 "HPDI_97.5"))

# Model WITH ADH
# Pull term for hard drugs variable
aggment_adh_terms <- summary_table_aggment_adh[2, ]

# Subset to Parameter, Estimate, HPDIs
aggment_adh_terms <- subset(aggment_adh_terms,
                    select = c("Parameter", 
                               "Estimate",
                               "HPDI_2.5",
                               "HPDI_97.5"))

# Display side by side like with frequentist models
# Create data frame
comp_models_aggment <- data.frame(Variable = "HD at Baseline (Yes)",
                 Estimate = round(aggment_terms$Estimate, 3),
                 `95% HDI` = paste(round(aggment_terms$HPDI_2.5, 3), 
                                    round(aggment_terms$HPDI_97.5, 3), 
                                    sep = ", "),
                 Estimate = round(aggment_adh_terms$Estimate, 3),
                 `95% HDI` = paste(round(aggment_adh_terms$HPDI_2.5, 3),
                                    round(aggment_adh_terms$HPDI_97.5, 3),
                                    sep = ", "),
                 check.names = FALSE)

# Display with kable
kable(
  comp_models_aggment,
  # format = "latex",
  booktabs = TRUE,
  digits = 3,
  caption = "Model Summary") |>
  kable_styling(latex_options = c("striped", "hold_position")) |>
  add_header_above(c(" ", "Model Without ADH" = 2, 
                     "Model With ADH" = 2)) %>% 
  footnote(general = "HPDI - Highest Density Interval ")

```


```{r, echo = F}
###########################################################################
# STEP 6: Model fit statistics (you only need 1 of these, not all 3)
###########################################################################
# Extract the log likelihood info
# For aggment model without ADH
loglik_mat_aggment <- as_draws_matrix(aggment_fit$draws("log_lik")) 
# iterations x N

# Get the LOO-CV
loo_res_aggment <- loo(loglik_mat_aggment)
# looic: 3223.8  

```

```{r, echo = F}
# Model WITH ADH
loglik_mat_aggment_adh <- as_draws_matrix(aggment_fit_adh$draws("log_lik")) 
# iterations x N

# Get the LOO-CV
loo_res_aggment_adh <- loo(loglik_mat_aggment_adh)
# print(loo_res_aggment_adh)
# 3222.1
```


```{r, echo = F}
###########################################################################
# STEP 6: MCMC Diagnostics
###########################################################################

# Model without ADH
# Dimensions: iterations x chains x parameters
aggment_draws <- as_draws_array(aggment_fit$draws())

# Tidy format for bayesplot / ggplot
aggment_draws_df <- as_draws_df(aggment_fit$draws())  

# parameters
aggment_params <- c("beta[1]", "beta[2]", 
            "beta[3]", "beta[4]",
            "beta[5]", "beta[6]",
            "beta[7]", "beta[8]",
            "sigma")

```

```{r, echo = F}
# Dimensions: iterations x chains x parameters
aggment_draws_adh <- as_draws_array(aggment_fit_adh$draws())

# Tidy format for bayesplot / ggplot
aggment_draws_df_adh <- as_draws_df(aggment_fit_adh$draws())  

# parameters
aggment_params_adh <- c("beta[1]", "beta[2]", 
            "beta[3]", "beta[4]",
            "beta[5]", "beta[6]",
            "beta[7]", "beta[8]",
            "sigma")
```


```{r, echo = F, results = 'hide', eval = F}
# Trace plots
# Trace plots show whether chains are mixing well and exploring the parameter space.
# Chains should mix well (lines overlap)
# No long trends (no “stickiness”)
# Ideally all chains overlap around the same mean

# Model without ADH
mcmc_trace(aggment_draws, pars = aggment_params)
# They look beautiful

```


```{r, echo = F, results = 'hide', eval = F}
# Model WITH ADH
mcmc_trace(aggment_draws_adh, pars = aggment_params_adh)
```



```{r, echo = F, results = 'hide'}
# R-hat and ESS diagnostics
# R-hat ≈ 1.00 → converged
# ESS should be reasonably large (e.g., >100–200 per parameter)
aggment_fit$summary(variables = aggment_params)
# We see that rhats are all approx 1
# ESS should be around > 200*9 = 1800 and that seems to be the case for most
```

```{r, echo = F, results='hide'}
# Model WITH ADH
aggment_fit_adh$summary(variables = aggment_params_adh)

```


```{r, echo = F, results='hide', eval = F}
# Posterior Density Plots
# Overlaid densities from multiple chains should match closely
# If densities differ substantially between chains → poor mixing
mcmc_dens_overlay(draws_aggment, pars = aggment_params)

```

```{r, echo = F, include = F, eval = F, results='hide'}
# Posterior density plots
# Model WITH ADH
mcmc_dens_overlay(draws_aggment_adh, pars = aggment_params_adh)
```


```{r, echo = F, results='hide', eval = F}
# Auto-correlation 
# Autocorrelation should decay quickly
# High autocorrelation → may need more iterations or thinning
mcmc_acf(draws_aggment, pars = aggment_params)

```

```{r, echo = F, results='hide', eval = F}
# Model WITH ADH
mcmc_acf(draws_aggment_adh, pars = aggment_params_adh)
```


```{r, echo = F, results='hide', eval = F}
# Diagnostics from cmdstan
aggment_fit$cmdstan_diagnose()
# Need to load actual aggment and not aggment .rds for this

### Output
# Checking sampler transitions treedepth.
# Treedepth satisfactory for all transitions.
# 
# Checking sampler transitions for divergences.
# No divergent transitions found.
# 
# Checking E-BFMI - sampler transitions HMC potential energy.
# E-BFMI satisfactory.
# 
# Rank-normalized split effective sample size satisfactory for all parameters.
# 
# Rank-normalized split R-hat values satisfactory for all parameters.
# 
# Processing complete, no problems detected.

```

```{r, echo = F, results='hide', eval = F}
# Model WITH ADH
# Diagnostics from cmdstan
aggment_fit_adh$cmdstan_diagnose()

### Output
# Checking sampler transitions treedepth.
# Treedepth satisfactory for all transitions.
# 
# Checking sampler transitions for divergences.
# No divergent transitions found.
# 
# Checking E-BFMI - sampler transitions HMC potential energy.
# E-BFMI satisfactory.
# 
# Rank-normalized split effective sample size satisfactory for all parameters.
# 
# Rank-normalized split R-hat values satisfactory for all parameters.
# 
# Processing complete, no problems detected.

```


```{r, echo = F, eval = F}
## Save output if desired ###
# MCMC Output
sink("Output/P1_aggment_MCMCOutput.txt")
print(summary_table)
print(loo_res)
print(aggment_fit$cmdstan_diagnose())
print(aggment_fit$summary(variables=aggment_params))
sink()

# Diagnostic plots
pdf('Output/P1_aggment_MCMC_DiagnosticPlots.pdf')
mcmc_trace(aggment_draws, pars = aggment_params)
mcmc_dens_overlay(aggment_draws, pars = aggment_params)
mcmc_acf(aggment_draws, pars = aggment_params)
dev.off()

```





