---
title: The Effects of Hard-Drug Use at Baseline on HAART Response & Outcomes
subtitle: Frequentist Analysis
author: Diana Felipe
output: 
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    latex_engine: xelatex
    extra_dependencies: ["longtable", "float"]
mainfont: Arial
fontsize: 12pt
documentclass: article
geometry: margin=1in
header-includes:
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage{float}
  - \usepackage{caption}
  - \usepackage{titling}
  - \setlength{\droptitle}{-2cm}
  - \preauthor{\vspace{-0.6cm}}
  - \postauthor{\vspace{-3cm}}
---

```{r, echo = F, message=F, warning = F}
knitr::opts_chunk$set(echo = FALSE)
library(readr)
library(knitr)
library(ggplot2)
library(MASS)
library(emmeans)
library(multcomp)
library(kableExtra)
library(stats)
library(dplyr)
library(tibble)
library(labelled)
library(tableone)
library(lme4)
library(lmerTest)
library(nlme)
library(gtsummary)
library(gt)
library(ggeffects)
library(UpSetR)
library(visdat)
library(lubridate)
library(see)
library(data.table)
library(gridExtra)
library(kableExtra)
library(regressinator)
library(broom)
library(expss)
library(mediation)
library(bda)

# Libraries for Bayesian analysis
# Dependencies 
library(cmdstanr)
library(bayesplot)  # diagnostic plots of the MCMC chains
library(posterior)  # for summarizing posterior draws
library(bayestestR) # for calculating higest density posterior intervals
library(mcmcse)     # for calculating MCMCSE's
library(loo)        # for getting model fit statistics (WAIC and LOO-IC)
library(dplyr)
library(tibble)

opts_chunk$set(tidy = F)
```

```{r, echo = F}
# Read in the Data
data <- read.csv("../Data/final_hiv_data.csv")

# Examine the data and data types
# str(data)


```

```{r, echo = F}
# Assessment of the data first
# str(data)

# Create categorical labels and variables 

# Hard drugs
data$hard_drugs_cat <- factor(data$hard_drugs,
                              levels = c(0, 1),
                              labels = c("No", "Yes"))

# adherence
data$ADH_yr2_cat <- factor(data$ADH_yr2,
                       levels = c(1, 2, 3, 4),
                       labels = c("100%", "95-99%",
                                  "75-94%",
                                  "<75%"))

# Move after hard drugs variable
data <- data %>% 
  relocate(hard_drugs_cat, .after = hard_drugs)

# Convert collapse race to factor
data$collapse_RACE <- factor(data$collapse_RACE)

# Convert collapse educ to factor
data$collapsed_EDUCBAS <- factor(data$collapsed_EDUCBAS)

# Convert SMOKE cat to factor
data$collapse_SMOKE <- factor(data$collapse_SMOKE)

# Relabel the variables
med_data <- apply_labels(data,
                   hard_drugs = "Hard Drug Use at BL",
                   ADH_yr2 = "HAART Adherence",
                   VLOAD_yr2 = "VLOAD - yr2")
```


A table1: 
 

```{r, echo = F}
data %>% 
  tbl_summary(include = c(VLOAD_yr2, LEU3N_yr2,
                          AGG_MENT_yr2, AGG_PHYS_yr2, 
                          BMI, SMOKE, age, hard_drugs, collapse_RACE,
                          collapsed_EDUCBAS),
              statistic = list(all_continuous() ~ "{mean}{sd}")) %>% 
  modify_header(label = "**Variable**") %>% 
  bold_labels()
```

Separating by hard drug use at baseline: 

```{r, echo = F}
data %>% 
  tbl_summary(include = c(VLOAD_yr2, LEU3N_yr2,
                          AGG_MENT_yr2, AGG_PHYS_yr2, 
                          BMI, SMOKE, age, collapse_RACE,
                          collapsed_EDUCBAS),
              by = hard_drugs,
              statistic = list(all_continuous() ~ "{mean}{sd}")) %>% 
  modify_header(label = "**Variable**") %>% 
  bold_labels()
```



# Frequentist Analysis

For these simple linear regression analyses, we will use the lm() function from the stats package in R. 

```{r, echo = F}
################################################################################
###                           Frequentist Analysis                           ###
################################################################################

```


## VLOAD
The number of HIV copies in a mL of blood.

Hypothesis: VLOAD is significantly higher 2 years after HART for subjects who reported hard drug use at baseline. 


### Mediation Analysis

```{r, echo = F}
### Parametric Approach
# Fit two models: With adherence and without adherence
# We will not be doing the crude model because of the 
# Very unbalanced hard drug use groups 

# Primary model - no adherence
vload_no_adh <- lm(log_VLOAD_yr2 ~ hard_drugs_cat + log_VLOAD_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS,
                     data = data)

# Fit the model with adherence
vload_adh <- lm(log_VLOAD_yr2 ~ hard_drugs_cat + log_VLOAD_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS
                    + ADH_yr2_cat,
                     data = data)
```

Comparison of estimates between models with and without adherence: 

```{r, echo = F}
# Compare betas 
t1 <- tbl_regression(vload_no_adh,
                     # exponentiate = TRUE,
                     include = c("hard_drugs_cat"),
                     pvalue_fun = label_style_pvalue(digits = 3),
                     estimate_fun = ~ style_number(.x, digits = 3)) 

t2 <- tbl_regression(vload_adh, 
                     # exponentiate = TRUE,
                     include = c("hard_drugs_cat"),
                     pvalue_fun = label_style_pvalue(digits = 3),
                     estimate_fun = ~ style_number(.x, digits = 3))


tbl_merge(
  tbls = list(t1, t2),
  tab_spanner = c("**Model Without ADH**", 
                  "**Model With ADH**")
)
```

We observe a difference of 0.119 units between the estimate for the two models. 
The proportion mediated by adherence is $\frac{-0.136 - (-0.017)}{-0.136} = 85\%$.
We have that 85% of the effect of hard drug use on copies of viral load is mediated through or can be explained by adherence to HAART. 
In other words, there is a difference between whether subjects adhered to HAART for those that did and did not report hard-drug use at baseline. 
The differences that we observe in the estimates can be explained by differences in adherence. 
Due to these findings, we included adherence in the VLOAD model. 


### Primary Analysis

Model diagnostics: 

```{r, echo = F, out.width="80%", fig.align='center'}
### Model diagnostics
# The final model for VLOAD
vload_primary <- vload_adh

# Residual plot
augment(vload_primary) |>
  ggplot(aes(x = .fitted, y = .resid)) +
  geom_point(size = 0.8) +
  # geom_smooth(se = FALSE) +
  labs(x = "Fitted value", y = "Residual") + 
  theme_lucid()

```

```{r, echo = F, out.width="80%", fig.align='center'}
# QQ Plot
augment(vload_primary) |>
  ggplot(aes(sample = .std.resid)) +
  geom_qq() +
  geom_qq_line() +
  labs(title = "Normal Q-Q plot of standardized residuals",
       x = "Theoretical quantiles", y = "Observed quantiles") + 
  theme_lucid()

```

Interpreting the results: 

```{r, echo = F}
summary(vload_primary)
```

After controlling for covariates, there is not significant difference (p = 0.970) in VLOAD 2 years after initiating HART between those who reported hard-drug use at baseline vs. those who did. 

<!-- REMEMBER TO INTERPRET ON LOG SCALE -->

\newpage

## LEU3N

### Mediation Analysis

The number of CD4+ T cell counts which is a measure of immunologic health.
With treatment, we expect CD4+ T Cll counts to increase.

Hypothesis: CD4+ T cell counts are significantly higher 2 years after HART for subjects who reported hard drug use at baseline. 

### Mediation Analysis

```{r, echo = F}
### Parametric Approach
# Fit two models: With adherence and without adherence
# We will not be doing the crude model because of the 
# Very unbalanced hard drug use groups 

# Primary model - no adherence
leu3n_no_adh <- lm(LEU3N_yr2 ~ hard_drugs_cat + LEU3N_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS,
                     data = data)

# Fit the model with adherence
leu3n_adh <- lm(LEU3N_yr2 ~ hard_drugs_cat + LEU3N_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS
                    + ADH_yr2_cat,
                     data = data)
```

Comparison of estimates between models with and without adherence: 

```{r, echo = F}
# Compare betas 
t1_leu3n <- tbl_regression(leu3n_no_adh,
                     include = c("hard_drugs_cat"),
                     pvalue_fun = label_style_pvalue(digits = 3),
                     estimate_fun = ~ style_number(.x, digits = 3)) 

t2_leu3n <- tbl_regression(leu3n_adh, 
                     include = c("hard_drugs_cat"),
                     pvalue_fun = label_style_pvalue(digits = 3),
                     estimate_fun = ~ style_number(.x, digits = 3))


tbl_merge(
  tbls = list(t1_leu3n, t2_leu3n),
  tab_spanner = c("**Model Without ADH**", 
                  "**Model With ADH**")
)
```

meow meow meow

### Primary Analysis

Model diagnostics: 

```{r, echo = F, out.width="80%", fig.align='center'}
### Model diagnostics
# The final model for VLOAD
leu3n_primary <- leu3n_no_adh

# Residual plot
augment(leu3n_primary) |>
  ggplot(aes(x = .fitted, y = .resid)) +
  geom_point(size = 0.8) +
  # geom_smooth(se = FALSE) +
  labs(x = "Fitted value", y = "Residual") + 
  theme_lucid()

```

```{r, echo = F, out.width="80%", fig.align='center'}
# QQ Plot
augment(leu3n_primary) |>
  ggplot(aes(sample = .std.resid)) +
  geom_qq() +
  geom_qq_line() +
  labs(title = "Normal Q-Q plot of standardized residuals",
       x = "Theoretical quantiles", y = "Observed quantiles") + 
  theme_lucid()

```

Interpreting the results: 

```{r, echo = F}
summary(leu3n_primary)
```

After controlling for covariates, there is not significant difference (p = 0.970) in VLOAD 2 years after initiating HART between those who reported hard-drug use at baseline vs. those who did. 

\newpage

## AGG PHYS

The number of HIV copies in a mL of blood.

Hypothesis: VLOAD is significantly higher 2 years after HART for subjects who reported hard drug use at baseline. 


### Mediation Analysis

```{r, echo = F}
### Parametric Approach
# Fit two models: With adherence and without adherence
# We will not be doing the crude model because of the 
# Very unbalanced hard drug use groups 

# Primary model - no adherence
aggphys_no_adh <- lm(AGG_PHYS_yr2 ~ hard_drugs_cat + AGG_PHYS_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS,
                     data = data)

# Fit the model with adherence
aggphys_adh <- lm(AGG_PHYS_yr2 ~ hard_drugs_cat + AGG_PHYS_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS
                    + ADH_yr2_cat,
                     data = data)
```

Comparison of estimates between models with and without adherence: 

```{r, echo = F}
# Compare betas 
t1_aggphs <- tbl_regression(aggphys_no_adh,
                     # exponentiate = TRUE,
                     include = c("hard_drugs_cat"),
                     pvalue_fun = label_style_pvalue(digits = 3),
                     estimate_fun = ~ style_number(.x, digits = 3)) 

t2_aggphs <- tbl_regression(aggphys_adh, 
                     # exponentiate = TRUE,
                     include = c("hard_drugs_cat"),
                     pvalue_fun = label_style_pvalue(digits = 3),
                     estimate_fun = ~ style_number(.x, digits = 3))


tbl_merge(
  tbls = list(t1_aggphs, t2_aggphs),
  tab_spanner = c("**Model Without ADH**", 
                  "**Model With ADH**")
)
```

We observe a difference of 0.119 units between the estimate for the two models. 
The proportion mediated by adherence is $\frac{-0.136 - (-0.017)}{-0.136} = 85\%$.
We have that 85% of the effect of hard drug use on copies of viral load is mediated through or can be explained by adherence to HAART. 
In other words, there is a difference between whether subjects adhered to HAART for those that did and did not report hard-drug use at baseline. 
The differences that we observe in the estimates can be explained by differences in adherence. 
Due to these findings, we included adherence in the VLOAD model. 


### Primary Analysis

Model diagnostics: 

```{r, echo = F, out.width="80%", fig.align='center'}
### Model diagnostics
# The final model for VLOAD
aggphys_primary <- aggphys_no_adh

# Residual plot
augment(aggphys_primary) |>
  ggplot(aes(x = .fitted, y = .resid)) +
  geom_point(size = 0.8) +
  # geom_smooth(se = FALSE) +
  labs(x = "Fitted value", y = "Residual") + 
  theme_lucid()

```

```{r, echo = F, out.width="80%", fig.align='center'}
# QQ Plot
augment(aggphys_primary) |>
  ggplot(aes(sample = .std.resid)) +
  geom_qq() +
  geom_qq_line() +
  labs(title = "Normal Q-Q plot of standardized residuals",
       x = "Theoretical quantiles", y = "Observed quantiles") + 
  theme_lucid()

```

Interpreting the results: 

```{r, echo = F}
summary(aggphys_primary)
```

After controlling for covariates, there is not significant difference (p = 0.970) in VLOAD 2 years after initiating HART between those who reported hard-drug use at baseline vs. those who did. 

<!-- REMEMBER TO INTERPRET ON LOG SCALE -->

\newpage


## AGG MENT

The number of HIV copies in a mL of blood.

Hypothesis: VLOAD is significantly higher 2 years after HART for subjects who reported hard drug use at baseline. 


### Mediation Analysis

```{r, echo = F}
### Parametric Approach
# Fit two models: With adherence and without adherence
# We will not be doing the crude model because of the 
# Very unbalanced hard drug use groups 

# Primary model - no adherence
aggment_no_adh <- lm(AGG_PHYS_yr2 ~ hard_drugs_cat + AGG_PHYS_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS,
                     data = data)

# Fit the model with adherence
aggment_adh <- lm(AGG_PHYS_yr2 ~ hard_drugs_cat + AGG_PHYS_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS
                    + ADH_yr2_cat,
                     data = data)
```

Comparison of estimates between models with and without adherence: 

```{r, echo = F}
# Compare betas 
t1 <- tbl_regression(aggment_no_adh,
                     # exponentiate = TRUE,
                     include = c("hard_drugs_cat"),
                     pvalue_fun = label_style_pvalue(digits = 3),
                     estimate_fun = ~ style_number(.x, digits = 3)) 

t2 <- tbl_regression(aggment_adh, 
                     # exponentiate = TRUE,
                     include = c("hard_drugs_cat"),
                     pvalue_fun = label_style_pvalue(digits = 3),
                     estimate_fun = ~ style_number(.x, digits = 3))


tbl_merge(
  tbls = list(t1, t2),
  tab_spanner = c("**Model Without ADH**", 
                  "**Model With ADH**")
)
```

We observe a difference of 0.119 units between the estimate for the two models. 
The proportion mediated by adherence is $\frac{-0.136 - (-0.017)}{-0.136} = 85\%$.
We have that 85% of the effect of hard drug use on copies of viral load is mediated through or can be explained by adherence to HAART. 
In other words, there is a difference between whether subjects adhered to HAART for those that did and did not report hard-drug use at baseline. 
The differences that we observe in the estimates can be explained by differences in adherence. 
Due to these findings, we included adherence in the VLOAD model. 


### Primary Analysis

Model diagnostics: 

```{r, echo = F, out.width="80%", fig.align='center'}
### Model diagnostics
# The final model for VLOAD
aggment_primary <- aggment_adh

# Residual plot
augment(aggment_primary) |>
  ggplot(aes(x = .fitted, y = .resid)) +
  geom_point(size = 0.8) +
  # geom_smooth(se = FALSE) +
  labs(x = "Fitted value", y = "Residual") + 
  theme_lucid()

```

```{r, echo = F, out.width="80%", fig.align='center'}
# QQ Plot
augment(aggment_primary) |>
  ggplot(aes(sample = .std.resid)) +
  geom_qq() +
  geom_qq_line() +
  labs(title = "Normal Q-Q plot of standardized residuals",
       x = "Theoretical quantiles", y = "Observed quantiles") + 
  theme_lucid()

```

Interpreting the results: 

```{r, echo = F}
summary(aggment_primary)
```


\newpage

# Bayesian Analysis

Laboratory in vitro and animal studies have shown that the use of hard drugs has increases HIV replication. 
However, since there are not any biologically plausible values that we could pickk for an informative optimist prior, we will continue with a vague prior for our Bayesian models. 

For viral load, we specified a $Normal(0, 10000)$ prior for the regression coefficients and a $Half Normal(0, 100^2)$ prior for the residual variance.

```{r, echo = F}
################################################################################
###                             Bayesian Analysis                            ###
################################################################################

### Linear model
vload_linear <- lm(VLOAD_yr2 ~ factor(hard_drugs) + BMI + factor(SMOKE)
                    + age + factor(collapse_RACE) + factor(collapsed_EDUCBAS),
                     data = data)

sink("Output/P1_VLOAD_ModelOutput.txt")
print(vload_linear)
summary(vload_linear)
confint(vload_linear)
anova(vload_linear)
sink()

``` 

```{r, echo = F}
### Bayesian Model

###########################################################################
# STEP 1: Define the model
# Normal priors on the regression coefficients
# Half normal priors on sigma
###########################################################################


stan_file <- write_stan_file("data {
  int<lower=0> N;                  // number of observations
  int<lower=0> P;                  // number of predictors including intercept
  matrix[N, P] X;                  // design matrix (first column = intercept)
  vector[N] y;                     // outcome

  vector[P] prior_mean;            // prior means for each beta
  vector<lower=0>[P] prior_sd;     // prior SDs for each beta

  real<lower=0> sigma_prior_sd;    // SD for half-normal prior on sigma
}

parameters {
  vector[P] beta;                  // regression coefficients
  real<lower=0> sigma;             // residual SD
}

model {
  // Vectorized priors for regression coefficients
  beta ~ normal(prior_mean, prior_sd);

  // Half-normal prior for sigma
  sigma ~ normal(0, sigma_prior_sd);

  // Likelihood
  y ~ normal(X * beta, sigma);
  

}

generated quantities {
  // log likelihood for each observation for calculating model fit stats
  vector[N] log_lik;
  for (n in 1:N) {
    log_lik[n] = normal_lpdf(y[n] | X[n] * beta, sigma);}
  }", dir="STAN", basename='linear_regression_half_normal'
)
```



```{r, echo = F}
###########################################################################
# STEP 2: Compile the Stan program
###########################################################################
vload_mod <- cmdstan_model('STAN/linear_regression_half_normal.stan')

```

```{r, echo = F}
###########################################################################
# STEP 3: Prepare your data to pass to Stan
###########################################################################

# Outcome data
y <- data$VLOAD_yr2

# Design matrix in our linear regression
X <- model.matrix(~ VLOAD_yr0 + hard_drugs_cat
                     + BMI + SMOKE_cat 
                     + age + collapse_RACE + collapsed_EDUCBAS,
                  data = data) 

# These are variables needed in Stan
# N is the number of observations in your data set 
# P is the number of columns in the design matrix
N <- nrow(X)
P <- ncol(X)

# Hyper-parameters for prior distributions
# Prior for regression coefficients is Normal(0, 10000)
# Prior for model error SD (sigma) is Half-Normal(0, 100)

# MVN mean (mean in the prior on the regression coefficients)
m <- c(1, rep(0, (P-1)))   

# # SD in the prior on regression coefficients --> variance 100^2
s <- rep(100, P) 

sigma_sd <- 100

# create data list to pass to STAN
data_list <- list(
  N = N,
  P = P,
  X = X,
  y = y,
  prior_mean = m,
  prior_sd = s,
  sigma_prior_sd = sigma_sd
)


```


```{r, echo = F}
###########################################################################
# STEP 4: Fit Model
###########################################################################

# Specify the warmup and sampling iterations
# Save as a .rds file
vload_fit <- vload_mod$sample(
  data = data_list,
  chains = 4,
  iter_warmup = 4000,
  iter_sampling = 5000,
  seed = 123
)

```


```{r, echo  F}
###########################################################################
# STEP 5: Summarize the posterior
###########################################################################

# For a quick posterior summary
# If you do not specify variables, this can take a log time 
# since it will summarize the log likelihoods for each observation that we saved
vload_fit$summary(variables=c("beta[1]", "beta[2]", 
                              "beta[3]", "beta[4]",
                              "beta[5]", "beta[6]",
                              "beta[7]", "beta[8]",
                              "beta[9]",
                              "sigma"))


# Fancier Table of Results of Interest from the Posterior
# Extract posterior draws from cmdstanr fit
draws <- vload_fit$draws()  

draws_mat <- as_draws_matrix(draws)
params <- colnames(draws_mat)

# Exclude non-parameter columns from summarization: lp__ and log_lik[n]
params <- params[!grepl("lp__|log_lik", params)]

```

```{r, echo = F}
# Build summary table
summary_table <- lapply(params, function(p) {
  vals <- as.numeric(draws_mat[, p])
  
  # Monte Carlo standard error
  # Used to determine if we have run enough iterations
  # rule of thumb: MCSE should be less than 6.27% of the posterior standard deviation
  mcse_val <- mcmcse::mcse(vals)$se
  
  # Effective sample size
  ess_val <- ess_bulk(vals)
  
  # 95% HPDI
  hpd <- hdi(vals, ci = 0.95)
  
  tibble(
    Parameter = p,
    Estimate  = mean(vals),
    MCSE      = mcse_val,
    Std_Dev   = sd(vals),
    HPDI_2.5  = hpd$CI_low,
    HPDI_97.5 = hpd$CI_high,
    ESS       = ess_val
  )}) %>% 
  bind_rows()

print(summary_table)

# check that mcmcse < 6% of posterior SD
(100*summary_table$MCSE/summary_table$Std_Dev) < 6
# All True

```

```{r, echo = F}
###########################################################################
# STEP 6: Model fit statistics (you only need 1 of these, not all 3)
###########################################################################
# Extract the log likelihood info
loglik_mat <- as_draws_matrix(vload_fit$draws("log_lik"))  # iterations x N

# Get the LOO-CV
loo_res <- loo(loglik_mat)
print(loo_res)

# Warning with Pareto k diagnostic values
# > pareto_k_table(loo_res)
# Pareto k diagnostic values:
#                          Count Pct.    Min. ESS
# (-Inf, 0.7]   (good)     457   98.7%   6912    
#    (0.7, 1]   (bad)        0    0.0%   <NA>    
#    (1, Inf)   (very bad)   6    1.3%   <NA> 

```

```{r, echo = F}
## DIC
# Deviance per iteration: -2 * log-likelihood summed over observations
D_theta <- -2 * rowSums(loglik_mat)
mean_D <- mean(D_theta)
# Going to stick with DIC for now..

# Deviance at posterior mean
beta_mean <- colMeans(draws_mat[, grep("^beta\\[", colnames(draws_mat))])
sigma_mean <- mean(draws_mat[, "sigma"])
D_bar_theta <- -2 * sum(dnorm(y, mean = X %*% beta_mean, sd = sigma_mean, log = TRUE))

DIC <- 2 * mean_D - D_bar_theta

print(DIC)
```

```{r, echo = F}
# Compare model fit against...

```



```{r, echo = F}
###########################################################################
# STEP 6: MCMC Diagnostics
###########################################################################

# Dimensions: iterations x chains x parameters
vload_draws <- as_draws_array(vload_fit$draws())

# Tidy format for bayesplot / ggplot
vload_draws_df <- as_draws_df(vload_fit$draws())  

# parameters
vload_params <- c("beta[1]", "beta[2]", 
            "beta[3]", "beta[4]",
            "beta[5]", "beta[6]",
            "beta[7]", "beta[8]",
            "beta[9]",
            "sigma")

```

```{r, echo = F}
# Trace plots
# Trace plots show whether chains are mixing well and exploring the parameter space.
# Chains should mix well (lines overlap)
# No long trends (no “stickiness”)
# Ideally all chains overlap around the same mean

mcmc_trace(vload_draws, pars = vload_params)
# They look beautiful

```

```{r, echo = F}
# R-hat and ESS diagnostics
# R-hat ≈ 1.00 → converged
# ESS should be reasonably large (e.g., >100–200 per parameter)
vload_fit$summary(variables = vload_params)
# We see that rhats are all approx 1
# ESS should be around > 200*9 = 1800 and that seems to be the case for most
```

```{r, echo = F}
# Posterior Density Plots
# Overlaid densities from multiple chains should match closely
# If densities differ substantially between chains → poor mixing
mcmc_dens_overlay(draws, pars = vload_params)

```

```{r, echo = F}
# Auto-correlation 
# Autocorrelation should decay quickly
# High autocorrelation → may need more iterations or thinning
mcmc_acf(draws, pars = vload_params)

```

```{r, echo = F}
# Diagnostics from cmdstan
vload_fit$cmdstan_diagnose()
```

```{r, echo = F}
## Save output if desired ###
# MCMC Output
sink("Output/P1_VLOAD_MCMCOutput.txt")
print(summary_table)
print(DIC)
print(waic_res)
print(loo_res)
print(vload_fit$cmdstan_diagnose())
print(vload_fit$summary(variables=vload_params))
sink()

# Diagnostic plots
pdf('Output/P1_VLOAD_MCMC_DiagnosticPlots.pdf')
mcmc_trace(vload_draws, pars = vload_params)
mcmc_dens_overlay(vload_draws, pars = vload_params)
mcmc_acf(vload_draws, pars = vload_params)
dev.off()
```


For CD4+ T cell counts, we specified a $Normal(0, 10000)$ prior for the regression coefficients and a $Half Normal(0, 100^2)$ prior for the residual variance.

For the aggregate mental quality of life score, we specified a $Normal(0, 10000)$ prior for the regression coefficients and a $Half Normal(0, 100^2)$ prior for the residual variance.

For the aggregate physical quality of life score, we specified a $Normal(0, 10000)$ prior for the regression coefficients and a $$Half Normal(0, 100^2) prior for the residual variance. 













