---
title: The Effects of Hard-Drug Use at Baseline on HAART Response & Outcomes
subtitle: Frequentist Analysis
author: Diana Felipe
output: 
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    latex_engine: xelatex
    extra_dependencies: ["longtable", "float"]
mainfont: Arial
fontsize: 12pt
documentclass: article
geometry: margin=1in
header-includes:
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage{float}
  - \usepackage{caption}
  - \usepackage{titling}
  - \setlength{\droptitle}{-2cm}
  - \preauthor{\vspace{-0.6cm}}
  - \postauthor{\vspace{-3cm}}
---

```{r, echo = F, message=F, warning = F}
knitr::opts_chunk$set(echo = FALSE)
library(readr)
library(knitr)
library(ggplot2)
library(MASS)
library(emmeans)
library(multcomp)
library(kableExtra)
library(stats)
library(dplyr)
library(tibble)
library(labelled)
library(tableone)
library(lme4)
library(lmerTest)
library(nlme)
library(gtsummary)
library(gt)
library(ggeffects)
library(UpSetR)
library(visdat)
library(lubridate)
library(see)
library(data.table)
library(gridExtra)
library(kableExtra)
library(regressinator)
library(broom)
opts_chunk$set(tidy = F)
```

```{r, echo = F}
# Read in the Data
data <- read.csv("../Data/final_hiv_data.csv")

# Examine the data and data types
# str(data)

# Add a logged VLOAD variable
data$log_VLOAD <- log10(data$VLOAD_yr2)
```


# Primary Outcomes

For these simple linear regression analyses, we will use the lm() function from the stats package in R. 

## VLOAD
The number of HIV copies in a mL of blood.

Hypothesis: VLOAD is significantly higher 2 years after HART for subjects who reported hard drug use at baseline. 

```{r, echo = F}
# Fit the model for VLOAD - with LEU3N
vload_mod1 <- lm(log_VLOAD ~ LEU3N_yr2 + factor(hard_drugs)
                   + BMI + age
                   + factor(SMOKE) + factor(ADH_yr2),
                  data = data)

# Fit the model for VLOAD - without LEU3N
vload_mod2 <- lm(log_VLOAD ~ factor(hard_drugs)
                   + BMI + age
                   + factor(SMOKE) + factor(ADH_yr2),
                  data = data)

# Compare AIC
AIC(vload_mod1, vload_mod2)

# Conduct a LRT 
anova(vload_mod1, vload_mod2)
# Yes, including LEU3n significantly improves model fit

# Final model 
vload_model <- lm(log_VLOAD ~ LEU3N_yr2 + factor(hard_drugs)
                   + BMI + age
                   + factor(SMOKE) + factor(ADH_yr2),
                  data = data)

# Summary of the model
# summary(vload_model)
```

```{r, echo = F, out.width="80%", fig.align='center'}
### Model diagnostics
# From: 

# Residual plot
augment(vload_model) |>
  ggplot(aes(x = .fitted, y = .resid)) +
  geom_point(size = 0.8) +
  # geom_smooth(se = FALSE) +
  labs(x = "Fitted value", y = "Residual") + 
  theme_lucid()

```

```{r, echo = F, out.width="80%", fig.align='center'}
# QQ Plot
augment(vload_model) |>
  ggplot(aes(sample = .std.resid)) +
  geom_qq() +
  geom_qq_line() +
  labs(title = "Normal Q-Q plot of standardized residuals",
       x = "Theoretical quantiles", y = "Observed quantiles") + 
  theme_lucid()

```

Interpreting the results: 

```{r, echo = F}
summary(vload_model)
```

After controlling for covariates, there is not significant difference (p = 0.22) in VLOAD 2 years after initiating HART between those who reported hard-drug use at baseline vs. those who did. 

<!-- REMEMBER TO INTERPRET ON LOG SCALE -->


\newpage

# LEU3N

```{r, echo = F}
# Fit the model for LEU3N - with VLOAD
leu3n_mod1 <- lm(LEU3N_yr2 ~ VLOAD_yr2 + factor(hard_drugs)
                   + BMI + age
                   + factor(SMOKE) + factor(ADH_yr2),
                  data = data)

# Fit the model for LEU3N - without VLOAD
leu3n_mod2 <- lm(LEU3N_yr2 ~ factor(hard_drugs)
                   + BMI + age
                   + factor(SMOKE) + factor(ADH_yr2),
                  data = data)

# Compare AIC
AIC(leu3n_mod1, leu3n_mod2)

# Conduct a LRT 
anova(leu3n_mod1, leu3n_mod2)
# Yes, including LEU3n significantly improves model fit

# Final model 
LEU3N_model <- lm(log_LEU3N ~ LEU3N_yr2 + factor(hard_drugs)
                   + BMI + age
                   + factor(SMOKE) + factor(ADH_yr2),
                  data = data)

# Summary of the model
# summary(LEU3N_model)
```




