---
title: The Effects of Hard-Drug Use at Baseline on HAART Response & Outcomes
subtitle: Frequentist Analysis
author: Diana Felipe
output: 
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    latex_engine: xelatex
    extra_dependencies: ["longtable", "float"]
mainfont: Arial
fontsize: 12pt
documentclass: article
geometry: margin=1in
header-includes:
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage{float}
  - \usepackage{caption}
  - \usepackage{titling}
  - \setlength{\droptitle}{-2cm}
  - \preauthor{\vspace{-0.6cm}}
  - \postauthor{\vspace{-3cm}}
---

```{r, echo = F, message=F, warning = F}
knitr::opts_chunk$set(echo = FALSE)
library(readr)
library(knitr)
library(ggplot2)
library(MASS)
library(emmeans)
library(multcomp)
library(kableExtra)
library(stats)
library(dplyr)
library(tibble)
library(labelled)
library(tableone)
library(lme4)
library(lmerTest)
library(nlme)
library(gtsummary)
library(gt)
library(ggeffects)
library(UpSetR)
library(visdat)
library(lubridate)
library(see)
library(data.table)
library(gridExtra)
library(kableExtra)
library(regressinator)
library(broom)
opts_chunk$set(tidy = F)
```

```{r, echo = F}
# Read in the Data
data <- read.csv("../Data/final_hiv_data.csv")

# Examine the data and data types
# str(data)

# Add a logged VLOAD variable
data$log_VLOAD <- log10(data$VLOAD_yr2)
data$log_VLOAD_yr0 <- log10(data$VLOAD_yr0)
```

# Median Analysis

The investigator is interested in knowing if differences in adherence to HAART can explain differences in the outcomes between subjects who did and did not report hard hard drug use at baseline.
In other words, we aim to assess if Adherence to HAART is a mediator - does it explain differences in the outcomes between drug use group for all 4 primary outcomes?

A mediator is defined as an intermediate variable on the causal path between an independent variable and a dependent variable.

In ou particular case, the possible mediator of interest is adherence to HAART, the independent variable is hard-drug use at basleine, and the dependent variable is one of: VLOAD, LEU3N, AGG MENT, or AGG PHYS. 

The indirect effect is the effect of hard drug use on the outcome that works through HAART adherence. 
The direct effect is the effect of hard drug use on the outcome that does NOT work through HAART adherence. In other words, the remaining effect after adjusting for the mediator. 

We will use this RPubs tutorial for Mediation analyses: 
[Mediation Analysis Using R](https://rpubs.com/mbounthavong/mediation_analysis_r) which uses the [mediation](https://cran.r-project.org/web/packages/mediation/index.html) R package. 

# VLOAD 





# Primary Outcomes

For these simple linear regression analyses, we will use the lm() function from the stats package in R. 

## VLOAD
The number of HIV copies in a mL of blood.

Hypothesis: VLOAD is significantly higher 2 years after HART for subjects who reported hard drug use at baseline. 

```{r, echo = F}
# Fit the model for VLOAD - with LEU3N
vload_mod1 <- lm(log_VLOAD ~ log_VLOAD_yr0 + LEU3N_yr0 + factor(hard_drugs)
                   + BMI + age
                   + factor(SMOKE) + factor(ADH_yr2),
                  data = data)

# Fit the model for VLOAD - without LEU3N
vload_mod2 <- lm(log_VLOAD ~ log_VLOAD_yr0 + factor(hard_drugs)
                   + BMI + age
                   + factor(SMOKE) + factor(ADH_yr2),
                  data = data)

# Compare AIC
AIC(vload_mod1, vload_mod2)
# Proceed with LEU3N in model

# How about the QOL Scores
# Fit the model for VLOAD with AGG MENT
vload_mod3 <- lm(log_VLOAD ~ log_VLOAD_yr0 + LEU3N_yr0 + factor(hard_drugs)
                   + BMI + age
                   + factor(SMOKE) + factor(ADH_yr2)
                   + AGG_MENT_yr0,
                  data = data)

# Compare AIC
AIC(vload_mod1, vload_mod3)
# Basically equivalent - not contuing with AGG MENT

# Model with AGG PHYS
vload_mod5 <- lm(log_VLOAD ~ log_VLOAD_yr0 + LEU3N_yr0 + factor(hard_drugs)
                   + BMI + age
                   + factor(SMOKE) + factor(ADH_yr2)
                   + AGG_PHYS_yr0,
                  data = data)
# Compare AGG PHYS and AGG MENT models
AIC(vload_mod1, vload_mod5)
# They are basically equivalent - with the simpler models, no QOL scores

# Final model 
VLOAD_model <- lm(log_VLOAD ~ log_VLOAD_yr0 + LEU3N_yr0 + factor(hard_drugs)
                   + BMI + age
                   + factor(SMOKE) + factor(ADH_yr2),
                  data = data)

# Summary of the model
# summary(VLOAD_model)
```

Model diagnostics: 

```{r, echo = F, out.width="80%", fig.align='center'}
### Model diagnostics
# From: 

# Residual plot
augment(VLOAD_model) |>
  ggplot(aes(x = .fitted, y = .resid)) +
  geom_point(size = 0.8) +
  # geom_smooth(se = FALSE) +
  labs(x = "Fitted value", y = "Residual") + 
  theme_lucid()

```

```{r, echo = F, out.width="80%", fig.align='center'}
# QQ Plot
augment(VLOAD_model) |>
  ggplot(aes(sample = .std.resid)) +
  geom_qq() +
  geom_qq_line() +
  labs(title = "Normal Q-Q plot of standardized residuals",
       x = "Theoretical quantiles", y = "Observed quantiles") + 
  theme_lucid()

```

Interpreting the results: 

```{r, echo = F}
summary(VLOAD_model)
```

After controlling for covariates, there is not significant difference (p = 0.96) in VLOAD 2 years after initiating HART between those who reported hard-drug use at baseline vs. those who did. 

<!-- REMEMBER TO INTERPRET ON LOG SCALE -->


\newpage

# LEU3N

```{r, echo = F}
# Fit the model for LEU3N - with VLOAD
leu3n_mod1 <- lm(LEU3N_yr2 ~ LEU3N_yr0 + VLOAD_yr0 + factor(hard_drugs)
                   + BMI + age
                   + factor(SMOKE) + factor(ADH_yr2),
                  data = data)

# Fit the model for LEU3N - without VLOAD
leu3n_mod2 <- lm(LEU3N_yr2 ~ LEU3N_yr0 + factor(hard_drugs)
                   + BMI + age
                   + factor(SMOKE) + factor(ADH_yr2),
                  data = data)

# Compare AIC
AIC(leu3n_mod1, leu3n_mod2)
# These models are equivalent

# Conduct a LRT 
anova(leu3n_mod1, leu3n_mod2)
# No, including VLOAD does not significantly improve model fit

## QOL Scores
# AGG MENT
leu3n_mod3 <- lm(LEU3N_yr2 ~ LEU3N_yr0 + factor(hard_drugs)
                   + BMI + age
                   + factor(SMOKE) + factor(ADH_yr2)
                   + AGG_MENT_yr0,
                  data = data)

# Compare with AIC
AIC(leu3n_mod2, leu3n_mod3)
# No difference again - proceeding with simpler model 

# AGG PHYS
leu3n_mod4 <- lm(LEU3N_yr2 ~ LEU3N_yr0 + factor(hard_drugs)
                   + BMI + age
                   + factor(SMOKE) + factor(ADH_yr2)
                   + AGG_PHYS_yr0,
                  data = data)

# Compare with AIC
AIC(leu3n_mod2, leu3n_mod4)
# No difference again - proceeding with simpler model 

# Final model 
LEU3N_model <- lm(LEU3N_yr2 ~ LEU3N_yr0 + factor(hard_drugs)
                   + BMI + age
                   + factor(SMOKE) + factor(ADH_yr2),
                  data = data)

# Summary of the model
# summary(LEU3N_model)
```

Model diagnostics: 

```{r, echo = F, out.width="80%", fig.align='center'}
### Model diagnostics
# From: 

# Residual plot
augment(LEU3N_model) |>
  ggplot(aes(x = .fitted, y = .resid)) +
  geom_point(size = 0.8) +
  # geom_smooth(se = FALSE) +
  labs(x = "Fitted value", y = "Residual") + 
  theme_lucid()

```

```{r, echo = F, out.width="80%", fig.align='center'}
# QQ Plot
augment(LEU3N_model) |>
  ggplot(aes(sample = .std.resid)) +
  geom_qq() +
  geom_qq_line() +
  labs(title = "Normal Q-Q plot of standardized residuals",
       x = "Theoretical quantiles", y = "Observed quantiles") + 
  theme_lucid()

```


Interpreting the results: 

```{r, echo = F}
summary(LEU3N_model)

# Get the CI
# confint(LEU3N_model)
```
After adjusting for covariates, there is a significant difference (p < 0.0001)
in CD4+ T cell count 2 years after HART between subjects who did and did not report hard drug use at baseline. 

On average, CD4+ T cell count was about 176.59 units lower (95% CI: (-239.62, -113.58)) for subjects that reported hard drug use at baseline compared to non-hard drug use at baseline, 2 years after HAART. This is a clinically significant meaningful difference and is what we hypothesized for differences between hard-drug users and non-hard drug users. 

\newpage








