---
title: The Effects of Hard-Drug Use at Baseline on HAART Response & Outcomes
subtitle: Frequentist Analysis
author: Diana Felipe
output: 
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    latex_engine: xelatex
    extra_dependencies: ["longtable", "float"]
mainfont: Arial
fontsize: 12pt
documentclass: article
geometry: margin=1in
header-includes:
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage{float}
  - \usepackage{caption}
  - \usepackage{titling}
  - \setlength{\droptitle}{-2cm}
  - \preauthor{\vspace{-0.6cm}}
  - \postauthor{\vspace{-3cm}}
---

```{r, echo = F, message=F, warning = F}
knitr::opts_chunk$set(echo = FALSE)
library(readr)
library(knitr)
library(ggplot2)
library(MASS)
library(emmeans)
library(multcomp)
library(kableExtra)
library(stats)
library(dplyr)
library(tibble)
library(labelled)
library(tableone)
library(lme4)
library(lmerTest)
library(nlme)
library(gtsummary)
library(gt)
library(ggeffects)
library(UpSetR)
library(visdat)
library(lubridate)
library(see)
library(data.table)
library(gridExtra)
library(kableExtra)
library(regressinator)
library(broom)
library(expss)
library(mediation)

# Libraries for Bayesian analysis
# Dependencies 
library(cmdstanr)
library(bayesplot)  # diagnostic plots of the MCMC chains
library(posterior)  # for summarizing posterior draws
library(bayestestR) # for calculating higest density posterior intervals
library(mcmcse)     # for calculating MCMCSE's
library(loo)        # for getting model fit statistics (WAIC and LOO-IC)
library(dplyr)
library(tibble)

opts_chunk$set(tidy = F)
```

```{r, echo = F}
# Read in the Data
data <- read.csv("../Data/final_hiv_data.csv")

# Examine the data and data types
# str(data)

# Add a logged VLOAD variable
data$log_VLOAD <- log10(data$VLOAD_yr2)
data$log_VLOAD_yr0 <- log10(data$VLOAD_yr0)
```

```{r, echo = F}
# Assessment of the data first
# str(data)

# Create categorical labels and variables 
# SMOKE
data$SMOKE_cat <- factor(data$SMOKE,
                         levels = c(1, 2, 3),
                         labels = c("Never", "Former", "Current"))
# Move after SMOKE variable
data <- data %>% 
  relocate(SMOKE_cat, .after = SMOKE)

# Hard drugs
data$hard_drugs_cat <- factor(data$hard_drugs,
                              levels = c(0, 1),
                              "No", "Yes")
# Move after hard drugs variable
data <- data %>% 
  relocate(hard_drugs_cat, .after = hard_drugs)

# Convert collapse race to factor
data$collapse_RACE <- factor(data$collapse_RACE)

# Convert collapse educ to factor
data$collapsed_EDUCBAS <- factor(data$collapsed_EDUCBAS)

# Relabel the variables
med_data <- apply_labels(data,
                   hard_drugs = "Hard Drug Use at BL",
                   ADH_yr2 = "HAART Adherence",
                   VLOAD_yr2 = "VLOAD - yr2")
```


A table1: 
 

```{r, echo = F}
data %>% 
  tbl_summary(include = c(VLOAD_yr2, LEU3N_yr2,
                          AGG_MENT_yr2, AGG_PHYS_yr2, 
                          BMI, SMOKE, age, hard_drugs, collapse_RACE,
                          collapsed_EDUCBAS),
              statistic = list(all_continuous() ~ "{mean}{sd}")) %>% 
  modify_header(label = "**Variable**") %>% 
  bold_labels()
```

Separating by hard drug use at baseline: 

```{r, echo = F}
data %>% 
  tbl_summary(include = c(VLOAD_yr2, LEU3N_yr2,
                          AGG_MENT_yr2, AGG_PHYS_yr2, 
                          BMI, SMOKE, age, collapse_RACE,
                          collapsed_EDUCBAS),
              by = hard_drugs,
              statistic = list(all_continuous() ~ "{mean}{sd}")) %>% 
  modify_header(label = "**Variable**") %>% 
  bold_labels()
```



# Frequentist Analysis

For these simple linear regression analyses, we will use the lm() function from the stats package in R. 

```{r, echo = F}
################################################################################
###                             Bayesian Analysis                            ###
################################################################################

```


## VLOAD
The number of HIV copies in a mL of blood.

Hypothesis: VLOAD is significantly higher 2 years after HART for subjects who reported hard drug use at baseline. 

```{r, echo = F}
### Parametric Approach
# Fit two models: With adherence and without adherence
# We will not be doing the crude model because of the 
# Very unbalanced hard drug use groups 

# Primary model - no adherence
vload_primary <- lm(VLOAD_yr2 ~ hard_drugs + BMI + SMOKE 
                    + age + collapse_RACE + collapsed_EDUCBAS,
                     data = data)
```

Model diagnostics: 

```{r, echo = F, out.width="80%", fig.align='center'}
### Model diagnostics
# From: 

# Residual plot
augment(VLOAD_model) |>
  ggplot(aes(x = .fitted, y = .resid)) +
  geom_point(size = 0.8) +
  # geom_smooth(se = FALSE) +
  labs(x = "Fitted value", y = "Residual") + 
  theme_lucid()

```

```{r, echo = F, out.width="80%", fig.align='center'}
# QQ Plot
augment(VLOAD_model) |>
  ggplot(aes(sample = .std.resid)) +
  geom_qq() +
  geom_qq_line() +
  labs(title = "Normal Q-Q plot of standardized residuals",
       x = "Theoretical quantiles", y = "Observed quantiles") + 
  theme_lucid()

```

Interpreting the results: 

```{r, echo = F}
summary(VLOAD_model)
```

After controlling for covariates, there is not significant difference (p = 0.96) in VLOAD 2 years after initiating HART between those who reported hard-drug use at baseline vs. those who did. 

<!-- REMEMBER TO INTERPRET ON LOG SCALE -->

Comparison of estimates between models with and without adherence: 

```{r, echo = F}
# Model WITH adherence
vload_adh <- lm(VLOAD_yr2 ~ hard_drugs + ADH_yr2
                     + BMI + SMOKE 
                     + age + collapse_RACE + collapsed_EDUCBAS,
                     data = data)

# Compare betas 
t1 <- tbl_regression(vload_primary, 
                     estimate_fun = ~ style_number(.x, digits = 3)) %>%
  modify_column_hide(columns = p.value) %>% 
  modify_column_hide(c(conf.low, conf.high))

t2 <- tbl_regression(vload_adh, 
                     estimate_fun = ~ style_number(.x, digits = 3)) %>%
  modify_column_hide(columns = p.value) %>%
  modify_column_hide(c(conf.low, conf.high))

tbl_merge(
  tbls = list(t1, t2),
  tab_spanner = c("**Model Without ADH**", 
                  "**Model With ADH**")
)
```

\newpage

## LEU3N


\newpage

# Bayesian Analysis

Laboratory in vitro and animal studies have shown that the use of hard drugs has increases HIV replication. 
However, since there are not any biologically plausible values that we could pickk for an informative optimist prior, we will continue with a vague prior for our Bayesian models. 

For viral load, we specified a $Normal(0, 10000)$ prior for the regression coefficients and a $Half Normal(0, 100^2)$ prior for the residual variance.

```{r, echo = F}
################################################################################
###                             Bayesian Analysis                            ###
################################################################################

# Linear model
vload_linear <- lm(VLOAD_yr2 ~ factor(hard_drugs) + BMI + factor(SMOKE)
                    + age + factor(collapse_RACE) + factor(collapsed_EDUCBAS),
                     data = data)

sink("Output/P1_VLOAD_ModelOutput.txt")
print(vload_linear)
summary(vload_linear)
confint(vload_linear)
anova(vload_linear)
sink()

# Bayesian Model



``` 


For CD4+ T cell counts, we specified a $Normal(0, 10000)$ prior for the regression coefficients and a $Half Normal(0, 100^2)$ prior for the residual variance.

For the aggregate mental quality of life score, we specified a $Normal(0, 10000)$ prior for the regression coefficients and a $Half Normal(0, 100^2)$ prior for the residual variance.

For the aggregate physical quality of life score, we specified a $Normal(0, 10000)$ prior for the regression coefficients and a $$Half Normal(0, 100^2) prior for the residual variance. 













