---
title: The Effectiveness of HART
author: Diana Felipe
output: 
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    latex_engine: xelatex
    extra_dependencies: ["longtable", "float"]
mainfont: Arial
fontsize: 12pt
documentclass: article
geometry: margin=1in
header-includes:
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage{float}
  - \usepackage{caption}
  - \usepackage{titling}
  - \setlength{\droptitle}{-3cm}
  - \preauthor{\vspace{-0.6cm}}
  - \postauthor{\vspace{-3cm}}
---

```{r, echo = F, warning=F, message = F}
knitr::opts_chunk$set(echo = FALSE)
library(readr)
library(knitr)
library(tidyr)
library(ggplot2)
library(MASS)
library(emmeans)
library(multcomp)
library(kableExtra)
library(stats)
library(dplyr)
library(tibble)
library(labelled)
library(tableone)
library(lme4)
library(lmerTest)
library(nlme)
library(gtsummary)
library(gt)
library(ggeffects)
library(UpSetR)
library(visdat)
library(lubridate)
library(see)
library(data.table)
library(gridExtra)
library(kableExtra)
library(regressinator)
library(broom)
library(expss)
library(mediation)
library(bda)
library(patchwork)
library(purrr)
library(stringr)

# Libraries for Bayesian analysis
# Dependencies 
library(cmdstanr)
library(rstan)
library(bayesplot)  # diagnostic plots of the MCMC chains
library(posterior)  # for summarizing posterior draws
library(bayestestR) # for calculating higest density posterior intervals
library(mcmcse)     # for calculating MCMCSE's
library(loo)        # for getting model fit statistics (WAIC and LOO-IC)
library(dplyr)
library(tibble)
library(brms)

opts_chunk$set(tidy = F)
```

```{r, echo = F}
# Read in the Data
data <- read.csv("../Data/final_hiv_data.csv")

# Examine the data and data types
# str(data)

# Create categorical labels and variables 

# Hard drugs
data$hard_drugs_cat <- factor(data$hard_drugs,
                              levels = c(0, 1),
                              labels = c("No", "Yes"))

# Move after hard drugs variable
data <- data %>% 
  relocate(hard_drugs_cat, .after = hard_drugs)

# Convert collapse race to factor
data$collapse_RACE <- factor(data$collapse_RACE)

# Convert collapse educ to factor
data$collapsed_EDUCBAS <- factor(data$collapsed_EDUCBAS)

# Convert SMOKE cat to factor
data$collapse_SMOKE <- factor(data$collapse_SMOKE)

# Relabel the variables
med_data <- apply_labels(data,
                   hard_drugs = "Hard Drug Use at BL",
                   ADH_yr2 = "HAART Adherence",
                   VLOAD_yr2 = "VLOAD - yr2")
```


\vspace{-8mm}
# Introduction
\vspace{-8mm}

The Multicenter AIDS Cohort Study is a prospective, observational, and longitudinal cohort of homosexual and bisexual men with Human Immunodeficiency Virus (HIV-1). 
This study examines the effect of Highly Active Antiretroviral Treatment (HAART), a standard treatment given to HIV-1 patients. Data was collected once annually for eight years and subjects were started on HART years 1 - 8.
This project aims to determine whether the use of hard drugs at baseline affects a patient's treatment response 2 years after HAART.
There is minimal evidence on the relationship between hard drugs and treatment response. These analyses aims to investigate the following hypotheses: 
Hypothesis 1: Viral load is higher for hard-drug users compared to non-hard drug users after 2 years of HAART.
Hypothesis 2: CD4+ T-Cell Counts are higher for hard-drug users compared to non-hard drug users after 2 years of HAART.
Hypothesis 3: Aggregate and physical quality of life scores are lower for hard-drug users compared to non-hard drug users after 2 years of HAART.
Hypothesis 4: Differences in the four measures of treatment response between hard drug and non-hard drug users can be explained by differences in adherence to medication. 
<!-- If hard drugs negatively or positively affect a patient's response to HART, clinicians may be able to better assist patients by providing treatment that is better suited to patients based on past hard drug history. -->

\vspace{-10mm}
# Methods
\vspace{-6mm}

\vspace{-6mm}
## Data Cleaning
\vspace{-6mm}

The original 8-year longitudinal data set contained 3935 observations for 715 subjects. 
Before filtering the data to the years of interest for the analysis, we investigated missingness in the whole data set to ensure concerning missingness patterns were not present. There were not any relationships or missingness patterns between any of the four primary outcomes and other variables in the data set. 
We then subset the data to information for years 0 and 2. Investigating this particular subset more closely, we filtered out individuals whose BMI did not fall in the range of (10, 60) as anything outside of this range is highly biologically implausible BMI values for adults in the population of interest. Following this, we only kept subjects that contained information for both years 0 and 2. 
In this step, we also created the variables necessary for analysis including viral load at year 2 and viral load at year 0, and so on for the three other outcomes. 
With the exception of adherence to medication, the baseline values were retained for hard drug use, BMI, Age, Race, Education, and Smoking Status.
At this stage, missingness was highest for viral load and CD4+ T-Cell counts at ~3%. 
We removed any missingness for any variables at this step ($\approx$ 10% missingness). 
We assessed distributions for all variables and from this determined that a log transformation of viral load was necessary.
Lastly, we collapsed categories for specific variables. 
The original 8 race categories were collapsed to 'White' and 'Non-White',
the original 3 smoking categories were reduced to 'Never/Former' and 'Current',
Education was collapsed to 'Less than College' and 'Completed College or Higher',
and adherence to HAART was collapsed to '<95%' and '>95%'.
Our final data set for analysis contained 463 subjects with complete data for years 0 and 2.

<!-- Extras -->
<!-- With the exception of adherence, all other variables in the data set had no more than ~7% missingness. -->

\vspace{-8mm}
## Data Analysis
\vspace{-6mm}

To investigate differences in treatment response between subjects who did and did not report hard drug use at baseline, we used linear regression and conducted analyses using both a Frequentist and a Bayesian framework.
For each of the four measures of treatment response, we used the measurements in the second year as the outcome for the four models. 
We adjusted for hard drug use, age, BMI, race, education at baseline, and measurements for the outcome of interest at year 0. As an example, the model for viral load:
$\text{Viral Load (Yr 2)} \sim \text{Viral Load (Yr 0)} + \text{Hard Drug Use} + \text{Age} + \ldots$ 
Assumptions of linearity, homoscedasticity, and normality were confirmed with diagnostic plots.
For the bayesian analyses, we specified prior distributions for each outcome.
Because prior information on effect sizes was limited, we used weakly informative priors for regression coefficients and the residual standard deviation. 
For viral load, we specified a vague Normal$(0, 100)$ prior for the regression coefficients and a Half-Normal$(0, 100)$ prior for the residual variance.
For the other three outcomes, we specified a vague Normal$(0, 1000)$ prior for the regression coefficients and a Half-Normal$(0, 1000)$ prior for the residual variance.
We used 4 chains with 1000 warmup iterations and 2000 sampling iterations for each model.
We used the 95% Highest Density Interval (HDI) and the 95% Confidence Interval (CI) to determine significance. Variables were significant if the intervals \textit{excluded} 0. 

We additionally assessed the impact subject's adherence to HAART had on the outcomes due to the concern that subjects who report hard drug use have a higher probability of lower adherence to medication which in turn affects the measures of treatment response. 
For each of the models described above, we created an additional model with adherence to HAART as an additional covariate. As an example: 
$\text{Viral Load (Yr 2)} \sim \text{Viral Load (Yr 0)} + \text{Adherence} + \text{Hard Drug Use} + ...$
This allowed us to quantify differences in the estimates when adherence was and was not taken into account to determine whether differences in any of the four outcomes could be explained by differences in adherence. 
A complete case analysis was used for each model. 
The R Statistical Programming Language v. 4.4.1 was used for all analyses. 
<!-- Extras  -->
<!-- For the frequentist analyses, estimates were significant if the 95% confidence intervals excluded 0. -->
<!-- For the bayesian analyses, estimates were significant if the 95% highest density intervals (HDIs) did not include 0. -->

\vspace{-10mm}
# Results
\vspace{-6mm}

Table 1 contains a description of the data used for all analyses. 
The mean and SD are displayed for the primary outcomes while the number and percentage of patients are displayed for the 5 covariates.
We note the highly unbalanced distribution of hard drug and non-hard drug users.
We observe higher viral load for both hard drug use groups, lower CD4+ T-Cell counts, and overall, similar (describe QOL scores).

<!------------------------------------ Table 1 -------------------------------->

```{r, echo = F}
# Fix levels for categorical variables for Table1
# Race
data$RACE_tab1 <- factor(data$collapse_RACE,
                         levels = c("Other", "White_NH"),
                         labels = c("Other", "White Non-Hispanic"))
# Education
data$EDU_tab1 <- factor(data$collapsed_EDUCBAS,
                        levels = c("Comp_Coll_Higher", "Less_Than_College"),
                        labels = c("Complete College Or Higher",
                                   "Less Than College"))
# Adherence


# Relabel the variables for the Table 1

tab1_data <- apply_labels(data,
                     VLOAD_yr0 = "Viral Load - (Yr 0)",
                     VLOAD_yr2 = "Viral Load - (Yr 2)",
                     LEU3N_yr0 = "CD4+ T-Cell Count (Yr 0)",
                     LEU3N_yr2 = "CD4+ T-Cell Count (Yr 2)",
                     AGG_MENT_yr0 = "Agg. Mental Quality of Life Score (Yr 0)",
                     AGG_MENT_yr2 = "Agg. Mental Quality of Life Score (Yr 2)",
                     AGG_PHYS_yr0 = "Agg. Physical Quality of Life Score (Yr 0)",
                     AGG_PHYS_yr2 = "Agg. Physical Quality of Life Score (Yr 2)",
                     collapse_SMOKE = "Smoking Status (Baseline)",
                     EDU_tab1 = "Education (Baseline)",
                     RACE_tab1 = "Race",
                     age = "Age (Baseline)",
                     hard_drugs_cat = "Hard Drug Use (Baseline)",
                     collapsed_ADH = "HAART Adherence")

# Display the modified Table1
tab1_data %>% 
  tbl_summary(include = c(VLOAD_yr2, VLOAD_yr0,
                          LEU3N_yr2, LEU3N_yr0,
                          AGG_MENT_yr2, AGG_MENT_yr0,
                          AGG_PHYS_yr2, AGG_PHYS_yr0,
                          BMI, collapse_SMOKE, 
                          age, RACE_tab1,
                          EDU_tab1,
                          collapsed_ADH),
              by = hard_drugs_cat,
              statistic = list(all_continuous() ~ "{mean}{sd}")) %>% 
  modify_header(label = "**Hard Drug Use at Baseline**") %>% 
  bold_labels() %>% 
  modify_caption(caption = "**Table 1 - Patient Characteristics by Hard Drug Use at Baseline**")
```


<!------------------------------------- VLOAD --------------------------------->


```{r, echo = F}
################################################################################
###                                   VLOAD                                  ###
################################################################################


### Linear model WITHOUT ADHERENCE
vload_linear <- lm(log_VLOAD_yr2 ~ hard_drugs_cat + log_VLOAD_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS,
                     data = data)

# sink("Output/P1_VLOAD_Freq_Output.txt")
# print(vload_linear)
# summary(vload_linear)
# confint(vload_linear)
# anova(vload_linear)
# sink()

``` 

```{r, echo = F}
### Linear model WITH ADHERENCE
vload_adh_linear <- lm(log_VLOAD_yr2 ~ hard_drugs_cat + log_VLOAD_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS
                    + collapsed_ADH,
                     data = data)

```



```{r, echo = F, cache = TRUE}
### Bayesian Model

# Linear model WITHOUT ADHERENCE

###########################################################################
# STEP 1: Define the model
# Normal priors on the regression coefficients
# Half normal priors on sigma
###########################################################################


stan_file <- write_stan_file("data {
  int<lower=0> N;                  // number of observations
  int<lower=0> P;                  // number of predictors including intercept
  matrix[N, P] X;                  // design matrix (first column = intercept)
  vector[N] y;                     // outcome

  vector[P] prior_mean;            // prior means for each beta
  vector<lower=0>[P] prior_sd;     // prior SDs for each beta

  real<lower=0> sigma_prior_sd;    // SD for half-normal prior on sigma
}

parameters {
  vector[P] beta;                  // regression coefficients
  real<lower=0> sigma;             // residual SD
}

model {
  // Vectorized priors for regression coefficients
  beta ~ normal(prior_mean, prior_sd);

  // Half-normal prior for sigma
  sigma ~ normal(0, sigma_prior_sd);

  // Likelihood
  y ~ normal(X * beta, sigma);
  

}

generated quantities {
  // log likelihood for each observation for calculating model fit stats
  vector[N] log_lik;
  for (n in 1:N) {
    log_lik[n] = normal_lpdf(y[n] | X[n] * beta, sigma);}
  }", dir="STAN", basename='VLOAD_primary'
)

```

```{r, echo = F, cache = TRUE}
# Linear Model WITH ADHERENCE

###########################################################################
# STEP 1: Define the model
# Normal priors on the regression coefficients
# Half normal priors on sigma
###########################################################################


stan_file_vload_adh <- write_stan_file("data {
  int<lower=0> N;                  // number of observations
  int<lower=0> P;                  // number of predictors including intercept
  matrix[N, P] X;                  // design matrix (first column = intercept)
  vector[N] y;                     // outcome

  vector[P] prior_mean;            // prior means for each beta
  vector<lower=0>[P] prior_sd;     // prior SDs for each beta

  real<lower=0> sigma_prior_sd;    // SD for half-normal prior on sigma
}

parameters {
  vector[P] beta;                  // regression coefficients
  real<lower=0> sigma;             // residual SD
}

model {
  // Vectorized priors for regression coefficients
  beta ~ normal(prior_mean, prior_sd);

  // Half-normal prior for sigma
  sigma ~ normal(0, sigma_prior_sd);

  // Likelihood
  y ~ normal(X * beta, sigma);
  

}

generated quantities {
  // log likelihood for each observation for calculating model fit stats
  vector[N] log_lik;
  for (n in 1:N) {
    log_lik[n] = normal_lpdf(y[n] | X[n] * beta, sigma);}
  }", dir="STAN", basename='VLOAD_adh'
)

```


```{r, echo = F, cache = TRUE}
###########################################################################
# STEP 2: Compile the Stan program
###########################################################################

# Model WITHOUT adherence
vload_mod <- cmdstan_model('STAN/VLOAD_primary.stan')

```

```{r, echo = F, cache = TRUE}
# Model WITH adherence
vload_mod_adh <- cmdstan_model('STAN/VLOAD_adh.stan')

```



```{r, echo = F}
###########################################################################
# STEP 3: Prepare your data to pass to Stan
###########################################################################

# Model WITHOUT adherence
# Outcome data
y <- data$log_VLOAD_yr2

# Design matrix in our linear regression
X <- model.matrix(~ hard_drugs_cat + log_VLOAD_yr0
   + BMI + collapse_SMOKE
   + age + collapse_RACE + collapsed_EDUCBAS,
   data = data)

# These are variables needed in Stan
# N is the number of observations in your data set 
# P is the number of columns in the design matrix
N <- nrow(X)
P <- ncol(X)

# Hyper-parameters for prior distributions
# Prior for regression coefficients is Normal(0, 10000)
# Prior for model error SD (sigma) is Half-Normal(0, 100)

# MVN mean (mean in the prior on the regression coefficients)
m <- c(1, rep(0, (P-1)))   

# # SD in the prior on regression coefficients --> variance 100^2
s <- rep(100, P) 

sigma_sd <- 100

# create data list to pass to STAN
data_list <- list(
  N = N,
  P = P,
  X = X,
  y = y,
  prior_mean = m,
  prior_sd = s,
  sigma_prior_sd = sigma_sd
)
```


```{r, echo = F}
# Model WITH adherence
# Outcome data
y_adh <- data$log_VLOAD_yr2

# Design matrix in our linear regression
X_adh <- model.matrix(~ hard_drugs_cat + log_VLOAD_yr0
   + BMI + collapse_SMOKE
   + age + collapse_RACE + collapsed_EDUCBAS
   + collapsed_ADH,
   data = data)

# These are variables needed in Stan
# N is the number of observations in your data set 
# P is the number of columns in the design matrix
N_adh <- nrow(X_adh)
P_adh <- ncol(X_adh)

# Hyper-parameters for prior distributions
# Prior for regression coefficients is Normal(0, 10000)
# Prior for model error SD (sigma) is Half-Normal(0, 100)

# MVN mean (mean in the prior on the regression coefficients)
m_adh <- c(1, rep(0, (P_adh - 1)))   

# # SD in the prior on regression coefficients --> variance 100^2
s_adh <- rep(100, P_adh) 

sigma_sd_adh <- 100

# create data list to pass to STAN
data_list_adh <- list(
  N = N_adh,
  P = P_adh,
  X = X_adh,
  y = y_adh,
  prior_mean = m_adh,
  prior_sd = s_adh,
  sigma_prior_sd = sigma_sd_adh
)
```


```{r, echo = F, message=F, warning = F, results='hide'}
###########################################################################
# STEP 4: Fit Model
###########################################################################

# Model WITHOUT adherence
# Specify the warmup and sampling iterations
vload_fit <- suppressMessages(
  suppressWarnings(
  vload_mod$sample(
  data = data_list,
  chains = 4,
  iter_warmup = 1000,
  iter_sampling = 2000,
  seed = 123,
  refresh = 0
 )
)
)

```

```{r, echo = F, message=F, warning = F, results = 'hide'}
# Model WITH adherence
# Specify the warmup and sampling iterations
vload_fit_adh <- vload_mod_adh$sample(
  data = data_list_adh,
  chains = 4,
  iter_warmup = 1000,
  iter_sampling = 2000,
  seed = 1234
)
```


```{r, echo = F, results='hide'}
###########################################################################
# STEP 5: Summarize the posterior
###########################################################################

# Model without ADH
vload_fit$summary(variables=c("beta[1]", "beta[2]", 
                              "beta[3]",
        # including extra betas to double check results with linear model
                              "beta[4]", "beta[5]",
                              "beta[6]", "beta[7]",
                              "beta[8]", 
                              "sigma"))


# Fancier Table of Results of Interest from the Posterior
# Extract posterior draws from cmdstanr fit
draws_vload <- vload_fit$draws()  

draws_mat_vload <- as_draws_matrix(draws_vload)
params_vload <- colnames(draws_mat_vload)

# Exclude non-parameter columns from summarization: lp__ and log_lik[n]
params_vload <- params_vload[!grepl("lp__|log_lik", params_vload)]

```

```{r, echo = F, cache = TRUE, results='hide'}
# Model WITH ADH

vload_fit_adh$summary(variables=c("beta[1]", "beta[2]", 
                              "beta[3]",
        # including extra betas to double check results with linear model
                              "beta[4]", "beta[5]",
                              "beta[6]", "beta[7]",
                              "beta[8]", 
                              "sigma"))


# Fancier Table of Results of Interest from the Posterior
# Extract posterior draws from cmdstanr fit
draws_vload_adh <- vload_fit_adh$draws()  

draws_mat_vload_adh <- as_draws_matrix(draws_vload_adh)
params_vload_adh <- colnames(draws_mat_vload_adh)

# Exclude non-parameter columns from summarization: lp__ and log_lik[n]
params_vload_adh <- params_vload_adh[!grepl("lp__|log_lik", params_vload_adh)]

```

```{r, echo = F, results='hide', cache = TRUE}
# Model WITH ADH
# Build summary table
summary_table_vload <- lapply(params_vload, function(p) {
  vals <- as.numeric(draws_mat_vload[, p])
  
  # Monte Carlo standard error
  # Used to determine if we have run enough iterations
  # rule of thumb: MCSE should be less than 6.27% of the posterior standard deviation
  mcse_val <- mcmcse::mcse(vals)$se
  
  # Effective sample size
  ess_val <- ess_bulk(vals)
  
  # 95% HPDI
  hpd <- hdi(vals, ci = 0.95)
  
  tibble(
    Parameter = p,
    Estimate  = mean(vals),
    MCSE      = mcse_val,
    Std_Dev   = sd(vals),
    HPDI_2.5  = hpd$CI_low,
    HPDI_97.5 = hpd$CI_high,
    ESS       = ess_val
  )}) %>% 
  bind_rows()

print(summary_table_vload)

# Compare with linear fit
summary(vload_linear)

# check that mcmcse < 6% of posterior SD
(100*summary_table_vload$MCSE/summary_table_vload$Std_Dev) < 6
# All True


```


```{r, echo = F, results='hide', cache = TRUE}
# Model WITH ADH
# Build summary table
summary_table_vload_adh <- lapply(params_vload_adh, function(p) {
  vals <- as.numeric(draws_mat_vload_adh[, p])
  
  # Monte Carlo standard error
  # Used to determine if we have run enough iterations
  # rule of thumb: MCSE should be less than 6.27% of the posterior standard deviation
  mcse_val <- mcmcse::mcse(vals)$se
  
  # Effective sample size
  ess_val <- ess_bulk(vals)
  
  # 95% HPDI
  hpd <- hdi(vals, ci = 0.95)
  
  tibble(
    Parameter = p,
    Estimate  = mean(vals),
    MCSE      = mcse_val,
    Std_Dev   = sd(vals),
    HPDI_2.5  = hpd$CI_low,
    HPDI_97.5 = hpd$CI_high,
    ESS       = ess_val
  )}) %>% 
  bind_rows()

print(summary_table_vload_adh)

# Compare with linear fit
summary(vload_adh_linear)

# check that mcmcse < 6% of posterior SD
(100*summary_table_vload_adh$MCSE/summary_table_vload_adh$Std_Dev) < 6
# All True

```



```{r, echo = F, cache = TRUE}
###########################################################################
# STEP 6: Model fit statistics (you only need 1 of these, not all 3)
###########################################################################
# Extract the log likelihood info
# For vload model without ADH
loglik_mat_vload <- as_draws_matrix(vload_fit$draws("log_lik")) 
# iterations x N

# Get the LOO-CV
loo_res <- loo(loglik_mat_vload)
# looic: 2212.6

```

```{r, echo = F, cache = TRUE}
# Model WITH ADH
loglik_mat_vload_adh <- as_draws_matrix(vload_fit_adh$draws("log_lik")) 
# iterations x N

# Get the LOO-CV
loo_res_vload_adh <- loo(loglik_mat_vload_adh)
# print(loo_res_vload_adh)
# 2208.8
```


```{r, echo = F, cache = TRUE}
###########################################################################
# STEP 6: MCMC Diagnostics
###########################################################################

# Model without ADH
# Dimensions: iterations x chains x parameters
vload_draws <- as_draws_array(vload_fit$draws())

# Tidy format for bayesplot / ggplot
vload_draws_df <- as_draws_df(vload_fit$draws())  

# parameters
vload_params <- c("beta[1]", "beta[2]", 
            "beta[3]", "beta[4]",
            "beta[5]", "beta[6]",
            "beta[7]", "beta[8]",
            "sigma")

```

```{r, echo = F, cache = TRUE}
# Dimensions: iterations x chains x parameters
vload_draws_adh <- as_draws_array(vload_fit_adh$draws())

# Tidy format for bayesplot / ggplot
vload_draws_df_adh <- as_draws_df(vload_fit_adh$draws())  

# parameters
vload_params_adh <- c("beta[1]", "beta[2]", 
            "beta[3]", "beta[4]",
            "beta[5]", "beta[6]",
            "beta[7]", "beta[8]",
            "sigma")
```


```{r, echo = F, results = 'hide', eval = F}
# Trace plots
# Trace plots show whether chains are mixing well and exploring the parameter space.
# Chains should mix well (lines overlap)
# No long trends (no “stickiness”)
# Ideally all chains overlap around the same mean

# Model without ADH
mcmc_trace(vload_draws, pars = vload_params)
# They look beautiful

```


```{r, echo = F, results = 'hide', eval = F}
# Model WITH ADH
mcmc_trace(vload_draws_adh, pars = vload_params_adh)
```



```{r, echo = F, results = 'hide', cache = TRUE}
# R-hat and ESS diagnostics
# R-hat ≈ 1.00 → converged
# ESS should be reasonably large (e.g., >100–200 per parameter)
vload_fit$summary(variables = vload_params)
# We see that rhats are all approx 1
# ESS should be around > 200*9 = 1800 and that seems to be the case for most
```

```{r, echo = F, cache = TRUE, results='hide'}

vload_fit_adh$summary(variables = vload_params_adh)

```


```{r, echo = F, results='hide', eval = F}
# Posterior Density Plots
# Overlaid densities from multiple chains should match closely
# If densities differ substantially between chains → poor mixing
mcmc_dens_overlay(draws_vload, pars = vload_params)

```

```{r, echo = F, eval = F, results='hide'}
# Posterior density plots
# Model WITH ADH

mcmc_dens_overlay(draws_vload_adh, pars = vload_params_adh)
```


```{r, echo = F, results='hide', eval = F}
# Auto-correlation 
# Autocorrelation should decay quickly
# High autocorrelation → may need more iterations or thinning
mcmc_acf(draws_vload, pars = vload_params)

```

```{r, echo = F, results='hide', eval = F}
# Model WITH ADH
mcmc_acf(draws_vload_adh, pars = vload_params_adh)
```


```{r, echo = F, results='hide', eval = F}
# Diagnostics from cmdstan
vload_fit$cmdstan_diagnose()
# Need to load actual vload and not vload .rds for this

### Output
# Checking sampler transitions treedepth.
# Treedepth satisfactory for all transitions.
# 
# Checking sampler transitions for divergences.
# No divergent transitions found.
# 
# Checking E-BFMI - sampler transitions HMC potential energy.
# E-BFMI satisfactory.
# 
# Rank-normalized split effective sample size satisfactory for all parameters.
# 
# Rank-normalized split R-hat values satisfactory for all parameters.
# 
# Processing complete, no problems detected.

# Can use .rds file for future runs

```

```{r, echo = F, results='hide', eval = F}
# Model WITH ADH
# Diagnostics from cmdstan
vload_fit_adh$cmdstan_diagnose()
```


```{r, echo = F, eval = F}
## Save output if desired ###
# MCMC Output
sink("Output/P1_VLOAD_MCMCOutput.txt")
print(summary_table)
print(loo_res)
print(vload_fit$cmdstan_diagnose())
print(vload_fit$summary(variables=vload_params))
sink()

# Diagnostic plots
pdf('Output/P1_VLOAD_MCMC_DiagnosticPlots.pdf')
mcmc_trace(vload_draws, pars = vload_params)
mcmc_dens_overlay(vload_draws, pars = vload_params)
mcmc_acf(vload_draws, pars = vload_params)
dev.off()
```



```{r, echo = F, eval = F}
# Read in data
data <- read.csv("../Data/final_hiv_data.csv")

# Assessment of the data first
# str(data)

# Create categorical labels and variables 

# Hard drugs
data$hard_drugs_cat <- factor(data$hard_drugs,
                              levels = c(0, 1),
                              labels = c("No", "Yes"))

# Move after hard drugs variable
data <- data %>% 
  relocate(hard_drugs_cat, .after = hard_drugs)

# Convert collapse race to factor
data$collapse_RACE <- factor(data$collapse_RACE)

# Convert collapse educ to factor
data$collapsed_EDUCBAS <- factor(data$collapsed_EDUCBAS)

# Convert SMOKE cat to factor
data$collapse_SMOKE <- factor(data$collapse_SMOKE)

# Relabel the variables
med_data <- apply_labels(data,
                   hard_drugs = "Hard Drug Use at BL",
                   ADH_yr2 = "HAART Adherence",
                   VLOAD_yr2 = "VLOAD - yr2")
```



```{r, echo = F}
################################################################################
###                                    LEU3N                                 ###
################################################################################


### Linear model WITHOUT ADHERENCE
leu3n_linear <- lm(LEU3N_yr2 ~ hard_drugs_cat + LEU3N_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS,
                     data = data)

# sink("Output/P1_leu3n_Freq_Output.txt")
# print(leu3n_linear)
# summary(leu3n_linear)
# confint(leu3n_linear)
# anova(leu3n_linear)
# sink()

``` 

```{r, echo = F}
### Linear model WITH ADHERENCE
leu3n_adh_linear <- lm(LEU3N_yr2 ~ hard_drugs_cat + LEU3N_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS
                    + collapsed_ADH,
                     data = data)

```



```{r, echo = F}
### Bayesian Model

# Linear model WITHOUT ADHERENCE

###########################################################################
# STEP 1: Define the model
# Normal priors on the regression coefficients
# Half normal priors on sigma
###########################################################################


stan_file_leu3n <- write_stan_file("data {
  int<lower=0> N;                  // number of observations
  int<lower=0> P;                  // number of predictors including intercept
  matrix[N, P] X;                  // design matrix (first column = intercept)
  vector[N] y;                     // outcome

  vector[P] prior_mean;            // prior means for each beta
  vector<lower=0>[P] prior_sd;     // prior SDs for each beta

  real<lower=0> sigma_prior_sd;    // SD for half-normal prior on sigma
}

parameters {
  vector[P] beta;                  // regression coefficients
  real<lower=0> sigma;             // residual SD
}

model {
  // Vectorized priors for regression coefficients
  beta ~ normal(prior_mean, prior_sd);

  // Half-normal prior for sigma
  sigma ~ normal(0, sigma_prior_sd);

  // Likelihood
  y ~ normal(X * beta, sigma);
  

}

generated quantities {
  // log likelihood for each observation for calculating model fit stats
  vector[N] log_lik;
  for (n in 1:N) {
    log_lik[n] = normal_lpdf(y[n] | X[n] * beta, sigma);}
  }", dir="STAN", basename='leu3n_primary'
)

```

```{r, echo = F}
# Linear Model WITH ADHERENCE

###########################################################################
# STEP 1: Define the model
# Normal priors on the regression coefficients
# Half normal priors on sigma
###########################################################################


stan_file_leu3n_adh <- write_stan_file("data {
  int<lower=0> N;                  // number of observations
  int<lower=0> P;                  // number of predictors including intercept
  matrix[N, P] X;                  // design matrix (first column = intercept)
  vector[N] y;                     // outcome

  vector[P] prior_mean;            // prior means for each beta
  vector<lower=0>[P] prior_sd;     // prior SDs for each beta

  real<lower=0> sigma_prior_sd;    // SD for half-normal prior on sigma
}

parameters {
  vector[P] beta;                  // regression coefficients
  real<lower=0> sigma;             // residual SD
}

model {
  // Vectorized priors for regression coefficients
  beta ~ normal(prior_mean, prior_sd);

  // Half-normal prior for sigma
  sigma ~ normal(0, sigma_prior_sd);

  // Likelihood
  y ~ normal(X * beta, sigma);
  

}

generated quantities {
  // log likelihood for each observation for calculating model fit stats
  vector[N] log_lik;
  for (n in 1:N) {
    log_lik[n] = normal_lpdf(y[n] | X[n] * beta, sigma);}
  }", dir="STAN", basename='leu3n_adh'
)

```


```{r, echo = F, message = F, warning = F}
###########################################################################
# STEP 2: Compile the Stan program
###########################################################################

# Model WITHOUT adherence
leu3n_mod <- cmdstan_model('STAN/leu3n_primary.stan')

```

```{r, echo = F, message = F, warning = F}
# Model WITH adherence
leu3n_mod_adh <- cmdstan_model('STAN/leu3n_adh.stan')

```


```{r, echo = F}
###########################################################################
# STEP 3: Prepare your data to pass to Stan
###########################################################################

# Model WITHOUT adherence
# Outcome data
y_leu3n <- data$LEU3N_yr2

# Design matrix in our linear regression
X_leu3n <- model.matrix(~ hard_drugs_cat + LEU3N_yr0
   + BMI + collapse_SMOKE
   + age + collapse_RACE + collapsed_EDUCBAS,
   data = data)

# These are variables needed in Stan
# N is the number of observations in your data set 
# P is the number of columns in the design matrix
N_leu3n <- nrow(X_leu3n)
P_leu3n <- ncol(X_leu3n)

# Hyper-parameters for prior distributions
# Prior for regression coefficients is Normal(0, 10000)
# Prior for model error SD (sigma) is Half-Normal(0, 100)

# MVN mean (mean in the prior on the regression coefficients)
m_leu3n <- c(1, rep(0, (P_leu3n - 1)))   

# # SD in the prior on regression coefficients --> variance 100^2
s_leu3n <- rep(1000, P_leu3n) 

sigma_sd_leu3n <- 1000

# create data list to pass to STAN
data_list <- list(
  N = N_leu3n,
  P = P_leu3n,
  X = X_leu3n,
  y = y_leu3n,
  prior_mean = m_leu3n,
  prior_sd = s_leu3n,
  sigma_prior_sd = sigma_sd_leu3n
)
```


```{r, echo = F}
# Model WITH adherence
# Outcome data
y_leu3n_adh <- data$LEU3N_yr2

# Design matrix in our linear regression
X_leu3n_adh <- model.matrix(~ hard_drugs_cat + LEU3N_yr0
   + BMI + collapse_SMOKE
   + age + collapse_RACE + collapsed_EDUCBAS
   + collapsed_ADH,
   data = data)

# These are variables needed in Stan
# N is the number of observations in your data set 
# P is the number of columns in the design matrix
N_leu3n_adh <- nrow(X_leu3n_adh)
P_leu3n_adh <- ncol(X_leu3n_adh)

# Hyper-parameters for prior distributions
# Prior for regression coefficients is Normal(0, 10000)
# Prior for model error SD (sigma) is Half-Normal(0, 100)

# MVN mean (mean in the prior on the regression coefficients)
m_leu3n_adh <- c(1, rep(0, (P_leu3n_adh - 1)))   

# # SD in the prior on regression coefficients --> variance 100^2
s_leu3n_adh <- rep(1000, P_leu3n_adh) 

sigma_sd_leu3n_adh <- 1000

# create data list to pass to STAN
data_list_adh <- list(
  N = N_leu3n_adh,
  P = P_leu3n_adh,
  X = X_leu3n_adh,
  y = y_leu3n_adh,
  prior_mean = m_leu3n_adh,
  prior_sd = s_leu3n_adh,
  sigma_prior_sd = sigma_sd_leu3n_adh
)
```


```{r, echo = F, message=F, warning = F, results = 'hide'}
###########################################################################
# STEP 4: Fit Model
###########################################################################

# Model WITHOUT adherence
# Specify the warmup and sampling iterations
leu3n_fit <- suppressMessages(
  suppressWarnings(
  leu3n_mod$sample(
  data = data_list,
  chains = 4,
  iter_warmup = 1000,
  iter_sampling = 2000,
  seed = 234,
  refresh = 0
 )
)
)

```

```{r, echo = F, message=F, warning = F, results = 'hide'}
# Model WITH adherence
# Specify the warmup and sampling iterations
leu3n_fit_adh <- leu3n_mod_adh$sample(
  data = data_list_adh,
  chains = 4,
  iter_warmup = 1000,
  iter_sampling = 2000,
  seed = 456
)
```


```{r, echo = F, results='hide'}
###########################################################################
# STEP 5: Summarize the posterior
###########################################################################

# Model without ADH
leu3n_fit$summary(variables=c("beta[1]", "beta[2]", 
                              "beta[3]",
        # including extra betas to double check results with linear model
                              "beta[4]", "beta[5]",
                              "beta[6]", "beta[7]",
                              "beta[8]", 
                              "sigma"))


# Fancier Table of Results of Interest from the Posterior
# Extract posterior draws from cmdstanr fit
draws_leu3n <- leu3n_fit$draws()  

draws_mat_leu3n <- as_draws_matrix(draws_leu3n)
params_leu3n <- colnames(draws_mat_leu3n)

# Exclude non-parameter columns from summarization: lp__ and log_lik[n]
params_leu3n <- params_leu3n[!grepl("lp__|log_lik", params_leu3n)]

```

```{r, echo = F, results = 'hide'}
# Model WITH ADH

leu3n_fit_adh$summary(variables=c("beta[1]", "beta[2]", 
                              "beta[3]",
        # including extra betas to double check results with linear model
                              "beta[4]", "beta[5]",
                              "beta[6]", "beta[7]",
                              "beta[8]", 
                              "sigma"))


# Fancier Table of Results of Interest from the Posterior
# Extract posterior draws from cmdstanr fit
draws_leu3n_adh <- leu3n_fit_adh$draws()  

draws_mat_leu3n_adh <- as_draws_matrix(draws_leu3n_adh)
params_leu3n_adh <- colnames(draws_mat_leu3n_adh)

# Exclude non-parameter columns from summarization: lp__ and log_lik[n]
params_leu3n_adh <- params_leu3n_adh[!grepl("lp__|log_lik", params_leu3n_adh)]

```

```{r, echo = F, results = 'hide'}
# Model without ADH
# Build summary table
summary_table_leu3n <- lapply(params_leu3n, function(p) {
  vals <- as.numeric(draws_mat_leu3n[, p])
  
  # Monte Carlo standard error
  # Used to determine if we have run enough iterations
  # rule of thumb: MCSE should be less than 6.27% of the posterior standard deviation
  mcse_val <- mcmcse::mcse(vals)$se
  
  # Effective sample size
  ess_val <- ess_bulk(vals)
  
  # 95% HPDI
  hpd <- hdi(vals, ci = 0.95)
  
  tibble(
    Parameter = p,
    Estimate  = mean(vals),
    MCSE      = mcse_val,
    Std_Dev   = sd(vals),
    HPDI_2.5  = hpd$CI_low,
    HPDI_97.5 = hpd$CI_high,
    ESS       = ess_val
  )}) %>% 
  bind_rows()

print(summary_table_leu3n)

# Compare with linear fit
summary(leu3n_linear)

# check that mcmcse < 6% of posterior SD
(100*summary_table_leu3n$MCSE/summary_table_leu3n$Std_Dev) < 6
# All True


```


```{r, echo = F, results = 'hide'}
# Model WITH ADH
# Build summary table
summary_table_leu3n_adh <- lapply(params_leu3n_adh, function(p) {
  vals <- as.numeric(draws_mat_leu3n_adh[, p])
  
  # Monte Carlo standard error
  # Used to determine if we have run enough iterations
  # rule of thumb: MCSE should be less than 6.27% of the posterior standard deviation
  mcse_val <- mcmcse::mcse(vals)$se
  
  # Effective sample size
  ess_val <- ess_bulk(vals)
  
  # 95% HPDI
  hpd <- hdi(vals, ci = 0.95)
  
  tibble(
    Parameter = p,
    Estimate  = mean(vals),
    MCSE      = mcse_val,
    Std_Dev   = sd(vals),
    HPDI_2.5  = hpd$CI_low,
    HPDI_97.5 = hpd$CI_high,
    ESS       = ess_val
  )}) %>% 
  bind_rows()

print(summary_table_leu3n_adh)

# Compare with linear fit
summary(leu3n_adh_linear)

# check that mcmcse < 6% of posterior SD
(100*summary_table_leu3n_adh$MCSE/summary_table_leu3n_adh$Std_Dev) < 6
# All True

```


```{r, echo = F, eval = F, results = 'hide'}
# Compare estimates for models with and without adherence

# Model without ADH
leu3n_terms <- summary_table_leu3n[2, ]

# Subset to Parameter, Estimate, HPDIs
leu3n_terms <- subset(leu3n_terms,
                      select = c("Parameter",
                                 "Estimate",
                                 "HPDI_2.5",
                                 "HPDI_97.5"))

# Model WITH ADH
# Pull term for hard drugs variable
leu3n_adh_terms <- summary_table_leu3n_adh[2, ]

# Subset to Parameter, Estimate, HPDIs
leu3n_adh_terms <- subset(leu3n_adh_terms,
                    select = c("Parameter", 
                               "Estimate",
                               "HPDI_2.5",
                               "HPDI_97.5"))

# Display side by side like with frequentist models
# Create data frame
comp_models_leu3n <- data.frame(Variable = "HD at Baseline (Yes)",
                 Estimate = round(leu3n_terms$Estimate, 3),
                 `95% HDI` = paste(round(leu3n_terms$HPDI_2.5, 3), 
                                    round(leu3n_terms$HPDI_97.5, 3), 
                                    sep = ", "),
                 Estimate = round(leu3n_adh_terms$Estimate, 3),
                 `95% HDI` = paste(round(leu3n_adh_terms$HPDI_2.5, 3),
                                    round(leu3n_adh_terms$HPDI_97.5, 3),
                                    sep = ", "),
                 check.names = FALSE)

# Display with kable
kable(
  comp_models_leu3n,
  # format = "latex",
  booktabs = TRUE,
  digits = 3,
  caption = "Model Summary") |>
  kable_styling(latex_options = c("striped", "hold_position")) |>
  add_header_above(c(" ", "Model Without ADH" = 2, 
                     "Model With ADH" = 2)) %>% 
  footnote(general = "HPDI - Highest Density Interval ")

```


```{r, echo = F}
###########################################################################
# STEP 6: Model fit statistics (you only need 1 of these, not all 3)
###########################################################################
# Extract the log likelihood info
# For leu3n model without ADH
loglik_mat_leu3n <- as_draws_matrix(leu3n_fit$draws("log_lik")) 
# iterations x N

# Get the LOO-CV
loo_res_leu3n <- loo(loglik_mat_leu3n)
# looic: 6124.8 

```

```{r, echo = F}
# Model WITH ADH
loglik_mat_leu3n_adh <- as_draws_matrix(leu3n_fit_adh$draws("log_lik")) 
# iterations x N

# Get the LOO-CV
loo_res_leu3n_adh <- loo(loglik_mat_leu3n_adh)
# print(loo_res_leu3n_adh)
# 6122.8
```


```{r, echo = F}
###########################################################################
# STEP 6: MCMC Diagnostics
###########################################################################

# Model without ADH
# Dimensions: iterations x chains x parameters
leu3n_draws <- as_draws_array(leu3n_fit$draws())

# Tidy format for bayesplot / ggplot
leu3n_draws_df <- as_draws_df(leu3n_fit$draws())  

# parameters
leu3n_params <- c("beta[1]", "beta[2]", 
            "beta[3]", "beta[4]",
            "beta[5]", "beta[6]",
            "beta[7]", "beta[8]",
            "sigma")

```

```{r, echo = F}
# Dimensions: iterations x chains x parameters
leu3n_draws_adh <- as_draws_array(leu3n_fit_adh$draws())

# Tidy format for bayesplot / ggplot
leu3n_draws_df_adh <- as_draws_df(leu3n_fit_adh$draws())  

# parameters
leu3n_params_adh <- c("beta[1]", "beta[2]", 
            "beta[3]", "beta[4]",
            "beta[5]", "beta[6]",
            "beta[7]", "beta[8]",
            "sigma")
```


```{r, echo = F, results = 'hide', eval = F}
# Trace plots
# Trace plots show whether chains are mixing well and exploring the parameter space.
# Chains should mix well (lines overlap)
# No long trends (no “stickiness”)
# Ideally all chains overlap around the same mean

# Model without ADH
mcmc_trace(leu3n_draws, pars = leu3n_params)
# They look beautiful

```


```{r, echo = F, results = 'hide', eval = F}
# Model WITH ADH
mcmc_trace(leu3n_draws_adh, pars = leu3n_params_adh)
```



```{r, echo = F, results = 'hide'}
# R-hat and ESS diagnostics
# R-hat ≈ 1.00 → converged
# ESS should be reasonably large (e.g., >100–200 per parameter)
leu3n_fit$summary(variables = leu3n_params)
# We see that rhats are all approx 1
# ESS should be around > 200*9 = 1800 and that seems to be the case for most
```

```{r, echo = F, results='hide'}
# Model WITH ADH
leu3n_fit_adh$summary(variables = leu3n_params_adh)

```


```{r, echo = F, results='hide', eval = F}
# Posterior Density Plots
# Overlaid densities from multiple chains should match closely
# If densities differ substantially between chains → poor mixing
mcmc_dens_overlay(draws_leu3n, pars = leu3n_params)

```

```{r, echo = F, include = F, eval = F, results='hide'}
# Posterior density plots
# Model WITH ADH
mcmc_dens_overlay(draws_leu3n_adh, pars = leu3n_params_adh)
```


```{r, echo = F, results='hide', eval = F}
# Auto-correlation 
# Autocorrelation should decay quickly
# High autocorrelation → may need more iterations or thinning
mcmc_acf(draws_leu3n, pars = leu3n_params)

```

```{r, echo = F, results='hide', eval = F}
# Model WITH ADH
mcmc_acf(draws_leu3n_adh, pars = leu3n_params_adh)
```


```{r, echo = F, results='hide', eval = F}
# Diagnostics from cmdstan
leu3n_fit$cmdstan_diagnose()
# Need to load actual leu3n and not leu3n .rds for this

### Output
# Checking sampler transitions treedepth.
# Treedepth satisfactory for all transitions.
# 
# Checking sampler transitions for divergences.
# No divergent transitions found.
# 
# Checking E-BFMI - sampler transitions HMC potential energy.
# E-BFMI satisfactory.
# 
# Rank-normalized split effective sample size satisfactory for all parameters.
# 
# Rank-normalized split R-hat values satisfactory for all parameters.
# 
# Processing complete, no problems detected.

```

```{r, echo = F, results='hide', eval = F}
# Model WITH ADH
# Diagnostics from cmdstan
leu3n_fit_adh$cmdstan_diagnose()

### Output
# Checking sampler transitions treedepth.
# Treedepth satisfactory for all transitions.
# 
# Checking sampler transitions for divergences.
# No divergent transitions found.
# 
# Checking E-BFMI - sampler transitions HMC potential energy.
# E-BFMI satisfactory.
# 
# Rank-normalized split effective sample size satisfactory for all parameters.
# 
# Rank-normalized split R-hat values satisfactory for all parameters.
# 
# Processing complete, no problems detected.
```


```{r, echo = F, eval = F}
## Save output if desired ###
# MCMC Output
sink("Output/P1_leu3n_MCMCOutput.txt")
print(summary_table)
print(loo_res)
print(leu3n_fit$cmdstan_diagnose())
print(leu3n_fit$summary(variables=leu3n_params))
sink()

# Diagnostic plots
pdf('Output/P1_leu3n_MCMC_DiagnosticPlots.pdf')
mcmc_trace(leu3n_draws, pars = leu3n_params)
mcmc_dens_overlay(leu3n_draws, pars = leu3n_params)
mcmc_acf(leu3n_draws, pars = leu3n_params)
dev.off()

```


```{r, echo = F, eval = F}
# Read in data
data <- read.csv("../Data/final_hiv_data.csv")

# Assessment of the data first
# str(data)

# Create categorical labels and variables 

# Hard drugs
data$hard_drugs_cat <- factor(data$hard_drugs,
                              levels = c(0, 1),
                              labels = c("No", "Yes"))

# Move after hard drugs variable
data <- data %>% 
  relocate(hard_drugs_cat, .after = hard_drugs)

# Convert collapse race to factor
data$collapse_RACE <- factor(data$collapse_RACE)

# Convert collapse educ to factor
data$collapsed_EDUCBAS <- factor(data$collapsed_EDUCBAS)

# Convert SMOKE cat to factor
data$collapse_SMOKE <- factor(data$collapse_SMOKE)

# Relabel the variables
med_data <- apply_labels(data,
                   hard_drugs = "Hard Drug Use at BL",
                   ADH_yr2 = "HAART Adherence",
                   VLOAD_yr2 = "VLOAD - yr2")
```



```{r, echo = F}
################################################################################
###                                AGG PHYS                                  ###
################################################################################


### Linear model WITHOUT ADHERENCE
aggphys_linear <- lm(AGG_PHYS_yr2 ~ hard_drugs_cat + AGG_PHYS_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS,
                     data = data)

# sink("Output/P1_aggphys_Freq_Output.txt")
# print(aggphys_linear)
# summary(aggphys_linear)
# confint(aggphys_linear)
# anova(aggphys_linear)
# sink()

``` 

```{r, echo = F}
### Linear model WITH ADHERENCE
aggphys_adh_linear <- lm(AGG_PHYS_yr2 ~ hard_drugs_cat + AGG_PHYS_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS
                    + collapsed_ADH,
                     data = data)

```


```{r, echo = F}
### Bayesian Model

# Linear model WITHOUT ADHERENCE

###########################################################################
# STEP 1: Define the model
# Normal priors on the regression coefficients
# Half normal priors on sigma
###########################################################################


stan_file_aggphys <- write_stan_file("data {
  int<lower=0> N;                  // number of observations
  int<lower=0> P;                  // number of predictors including intercept
  matrix[N, P] X;                  // design matrix (first column = intercept)
  vector[N] y;                     // outcome

  vector[P] prior_mean;            // prior means for each beta
  vector<lower=0>[P] prior_sd;     // prior SDs for each beta

  real<lower=0> sigma_prior_sd;    // SD for half-normal prior on sigma
}

parameters {
  vector[P] beta;                  // regression coefficients
  real<lower=0> sigma;             // residual SD
}

model {
  // Vectorized priors for regression coefficients
  beta ~ normal(prior_mean, prior_sd);

  // Half-normal prior for sigma
  sigma ~ normal(0, sigma_prior_sd);

  // Likelihood
  y ~ normal(X * beta, sigma);
  

}

generated quantities {
  // log likelihood for each observation for calculating model fit stats
  vector[N] log_lik;
  for (n in 1:N) {
    log_lik[n] = normal_lpdf(y[n] | X[n] * beta, sigma);}
  }", dir="STAN", basename='aggphys_primary'
)

```

```{r, echo = F}
# Linear Model WITH ADHERENCE

###########################################################################
# STEP 1: Define the model
# Normal priors on the regression coefficients
# Half normal priors on sigma
###########################################################################


stan_file_aggphys_adh <- write_stan_file("data {
  int<lower=0> N;                  // number of observations
  int<lower=0> P;                  // number of predictors including intercept
  matrix[N, P] X;                  // design matrix (first column = intercept)
  vector[N] y;                     // outcome

  vector[P] prior_mean;            // prior means for each beta
  vector<lower=0>[P] prior_sd;     // prior SDs for each beta

  real<lower=0> sigma_prior_sd;    // SD for half-normal prior on sigma
}

parameters {
  vector[P] beta;                  // regression coefficients
  real<lower=0> sigma;             // residual SD
}

model {
  // Vectorized priors for regression coefficients
  beta ~ normal(prior_mean, prior_sd);

  // Half-normal prior for sigma
  sigma ~ normal(0, sigma_prior_sd);

  // Likelihood
  y ~ normal(X * beta, sigma);
  

}

generated quantities {
  // log likelihood for each observation for calculating model fit stats
  vector[N] log_lik;
  for (n in 1:N) {
    log_lik[n] = normal_lpdf(y[n] | X[n] * beta, sigma);}
  }", dir="STAN", basename='aggphys_adh'
)

```


```{r, echo = F, message = F, warning = F}
###########################################################################
# STEP 2: Compile the Stan program
###########################################################################

# Model WITHOUT adherence
aggphys_mod <- cmdstan_model('STAN/aggphys_primary.stan')

```

```{r, echo = F, message = F, warning = F}
# Model WITH adherence
aggphys_mod_adh <- cmdstan_model('STAN/aggphys_adh.stan')

```


```{r, echo = F}
###########################################################################
# STEP 3: Prepare your data to pass to Stan
###########################################################################

# Model WITHOUT adherence
# Outcome data
y_aggphys <- data$AGG_PHYS_yr2

# Design matrix in our linear regression
X_aggphys <- model.matrix(~ hard_drugs_cat + AGG_PHYS_yr0
   + BMI + collapse_SMOKE
   + age + collapse_RACE + collapsed_EDUCBAS,
   data = data)

# These are variables needed in Stan
# N is the number of observations in your data set 
# P is the number of columns in the design matrix
N_aggphys <- nrow(X_aggphys)
P_aggphys <- ncol(X_aggphys)

# Hyper-parameters for prior distributions
# Prior for regression coefficients is Normal(0, 10000)
# Prior for model error SD (sigma) is Half-Normal(0, 100)

# MVN mean (mean in the prior on the regression coefficients)
m_aggphys <- c(1, rep(0, (P_aggphys - 1)))   

# # SD in the prior on regression coefficients --> variance 100^2
s_aggphys <- rep(1000, P_aggphys) 

sigma_sd_aggphys <- 1000

# create data list to pass to STAN
data_list <- list(
  N = N_aggphys,
  P = P_aggphys,
  X = X_aggphys,
  y = y_aggphys,
  prior_mean = m_aggphys,
  prior_sd = s_aggphys,
  sigma_prior_sd = sigma_sd_aggphys
)
```


```{r, echo = F}
# Model WITH adherence
# Outcome data
y_aggphys_adh <- data$AGG_PHYS_yr2

# Design matrix in our linear regression
X_aggphys_adh <- model.matrix(~ hard_drugs_cat + AGG_PHYS_yr0
   + BMI + collapse_SMOKE
   + age + collapse_RACE + collapsed_EDUCBAS
   + collapsed_ADH,
   data = data)

# These are variables needed in Stan
# N is the number of observations in your data set 
# P is the number of columns in the design matrix
N_aggphys_adh <- nrow(X_aggphys_adh)
P_aggphys_adh <- ncol(X_aggphys_adh)

# Hyper-parameters for prior distributions
# Prior for regression coefficients is Normal(0, 10000)
# Prior for model error SD (sigma) is Half-Normal(0, 100)

# MVN mean (mean in the prior on the regression coefficients)
m_aggphys_adh <- c(1, rep(0, (P_aggphys_adh - 1)))   

# # SD in the prior on regression coefficients --> variance 100^2
s_aggphys_adh <- rep(1000, P_aggphys_adh) 

sigma_sd_aggphys_adh <- 1000

# create data list to pass to STAN
data_list_adh <- list(
  N = N_aggphys_adh,
  P = P_aggphys_adh,
  X = X_aggphys_adh,
  y = y_aggphys_adh,
  prior_mean = m_aggphys_adh,
  prior_sd = s_aggphys_adh,
  sigma_prior_sd = sigma_sd_aggphys_adh
)
```


```{r, echo = F, message=F, warning = F, results = 'hide'}
###########################################################################
# STEP 4: Fit Model
###########################################################################

# Model WITHOUT adherence
# Specify the warmup and sampling iterations
aggphys_fit <- suppressMessages(
  suppressWarnings(
  aggphys_mod$sample(
  data = data_list,
  chains = 4,
  iter_warmup = 1000,
  iter_sampling = 2000,
  seed = 235,
  refresh = 0
 )
)
)

```

```{r, echo = F, message=F, warning = F, results = 'hide'}
# Model WITH adherence
# Specify the warmup and sampling iterations
aggphys_fit_adh <- aggphys_mod_adh$sample(
  data = data_list_adh,
  chains = 4,
  iter_warmup = 1000,
  iter_sampling = 2000,
  seed = 456
)
```


```{r, echo = F, results='hide'}
###########################################################################
# STEP 5: Summarize the posterior
###########################################################################

# Model without ADH
aggphys_fit$summary(variables=c("beta[1]", "beta[2]", 
                              "beta[3]",
        # including extra betas to double check results with linear model
                              "beta[4]", "beta[5]",
                              "beta[6]", "beta[7]",
                              "beta[8]", 
                              "sigma"))


# Fancier Table of Results of Interest from the Posterior
# Extract posterior draws from cmdstanr fit
draws_aggphys <- aggphys_fit$draws()  

draws_mat_aggphys <- as_draws_matrix(draws_aggphys)
params_aggphys <- colnames(draws_mat_aggphys)

# Exclude non-parameter columns from summarization: lp__ and log_lik[n]
params_aggphys <- params_aggphys[!grepl("lp__|log_lik", params_aggphys)]

```

```{r, echo = F, results = 'hide'}
# Model WITH ADH

aggphys_fit_adh$summary(variables=c("beta[1]", "beta[2]", 
                              "beta[3]",
        # including extra betas to double check results with linear model
                              "beta[4]", "beta[5]",
                              "beta[6]", "beta[7]",
                              "beta[8]", 
                              "sigma"))


# Fancier Table of Results of Interest from the Posterior
# Extract posterior draws from cmdstanr fit
draws_aggphys_adh <- aggphys_fit_adh$draws()  

draws_mat_aggphys_adh <- as_draws_matrix(draws_aggphys_adh)
params_aggphys_adh <- colnames(draws_mat_aggphys_adh)

# Exclude non-parameter columns from summarization: lp__ and log_lik[n]
params_aggphys_adh <- params_aggphys_adh[!grepl("lp__|log_lik", params_aggphys_adh)]

```

```{r, echo = F, results = 'hide'}
# Model without ADH
# Build summary table
summary_table_aggphys <- lapply(params_aggphys, function(p) {
  vals <- as.numeric(draws_mat_aggphys[, p])
  
  # Monte Carlo standard error
  # Used to determine if we have run enough iterations
  # rule of thumb: MCSE should be less than 6.27% of the posterior standard deviation
  mcse_val <- mcmcse::mcse(vals)$se
  
  # Effective sample size
  ess_val <- ess_bulk(vals)
  
  # 95% HPDI
  hpd <- hdi(vals, ci = 0.95)
  
  tibble(
    Parameter = p,
    Estimate  = mean(vals),
    MCSE      = mcse_val,
    Std_Dev   = sd(vals),
    HPDI_2.5  = hpd$CI_low,
    HPDI_97.5 = hpd$CI_high,
    ESS       = ess_val
  )}) %>% 
  bind_rows()

print(summary_table_aggphys)

# Compare with linear fit
summary(aggphys_linear)

# check that mcmcse < 6% of posterior SD
(100*summary_table_aggphys$MCSE/summary_table_aggphys$Std_Dev) < 6
# All True
```


```{r, echo = F, results = 'hide'}
# Model WITH ADH
# Build summary table
summary_table_aggphys_adh <- lapply(params_aggphys_adh, function(p) {
  vals <- as.numeric(draws_mat_aggphys_adh[, p])
  
  # Monte Carlo standard error
  # Used to determine if we have run enough iterations
  # rule of thumb: MCSE should be less than 6.27% of the posterior standard deviation
  mcse_val <- mcmcse::mcse(vals)$se
  
  # Effective sample size
  ess_val <- ess_bulk(vals)
  
  # 95% HPDI
  hpd <- hdi(vals, ci = 0.95)
  
  tibble(
    Parameter = p,
    Estimate  = mean(vals),
    MCSE      = mcse_val,
    Std_Dev   = sd(vals),
    HPDI_2.5  = hpd$CI_low,
    HPDI_97.5 = hpd$CI_high,
    ESS       = ess_val
  )}) %>% 
  bind_rows()

print(summary_table_aggphys_adh)

# Compare with linear fit
summary(aggphys_adh_linear)

# check that mcmcse < 6% of posterior SD
(100*summary_table_aggphys_adh$MCSE/summary_table_aggphys_adh$Std_Dev) < 6
# All True

```


```{r, echo = F, eval = F, results = 'hide'}
# Compare estimates for models with and without adherence

# Model without ADH
aggphys_terms <- summary_table_aggphys[2, ]

# Subset to Parameter, Estimate, HPDIs
aggphys_terms <- subset(aggphys_terms,
                      select = c("Parameter",
                                 "Estimate",
                                 "HPDI_2.5",
                                 "HPDI_97.5"))

# Model WITH ADH
# Pull term for hard drugs variable
aggphys_adh_terms <- summary_table_aggphys_adh[2, ]

# Subset to Parameter, Estimate, HPDIs
aggphys_adh_terms <- subset(aggphys_adh_terms,
                    select = c("Parameter", 
                               "Estimate",
                               "HPDI_2.5",
                               "HPDI_97.5"))

# Display side by side like with frequentist models
# Create data frame
comp_models_aggphys <- data.frame(Variable = "HD at Baseline (Yes)",
                 Estimate = round(aggphys_terms$Estimate, 3),
                 `95% HDI` = paste(round(aggphys_terms$HPDI_2.5, 3), 
                                    round(aggphys_terms$HPDI_97.5, 3), 
                                    sep = ", "),
                 Estimate = round(aggphys_adh_terms$Estimate, 3),
                 `95% HDI` = paste(round(aggphys_adh_terms$HPDI_2.5, 3),
                                    round(aggphys_adh_terms$HPDI_97.5, 3),
                                    sep = ", "),
                 check.names = FALSE)

# Display with kable
# kable(
#   comp_models_aggphys,
#   # format = "latex",
#   booktabs = TRUE,
#   digits = 3,
#   caption = "Model Summary") |>
#   kable_styling(latex_options = c("striped", "hold_position")) |>
#   add_header_above(c(" ", "Model Without ADH" = 2, 
#                      "Model With ADH" = 2)) %>% 
#   footnote(general = "HPDI - Highest Density Interval ")

```


```{r, echo = F}
###########################################################################
# STEP 6: Model fit statistics (you only need 1 of these, not all 3)
###########################################################################
# Extract the log likelihood info
# For aggphys model without ADH
loglik_mat_aggphys <- as_draws_matrix(aggphys_fit$draws("log_lik")) 
# iterations x N

# Get the LOO-CV
loo_res_aggphys <- loo(loglik_mat_aggphys)
# looic: 3223.6 

```

```{r, echo = F}
# Model WITH ADH
loglik_mat_aggphys_adh <- as_draws_matrix(aggphys_fit_adh$draws("log_lik")) 
# iterations x N

# Get the LOO-CV
loo_res_aggphys_adh <- loo(loglik_mat_aggphys_adh)
# print(loo_res_aggphys_adh)
# 3222.5
```


```{r, echo = F}
###########################################################################
# STEP 6: MCMC Diagnostics
###########################################################################

# Model without ADH
# Dimensions: iterations x chains x parameters
aggphys_draws <- as_draws_array(aggphys_fit$draws())

# Tidy format for bayesplot / ggplot
aggphys_draws_df <- as_draws_df(aggphys_fit$draws())  

# parameters
aggphys_params <- c("beta[1]", "beta[2]", 
            "beta[3]", "beta[4]",
            "beta[5]", "beta[6]",
            "beta[7]", "beta[8]",
            "sigma")

```

```{r, echo = F}
# Dimensions: iterations x chains x parameters
aggphys_draws_adh <- as_draws_array(aggphys_fit_adh$draws())

# Tidy format for bayesplot / ggplot
aggphys_draws_df_adh <- as_draws_df(aggphys_fit_adh$draws())  

# parameters
aggphys_params_adh <- c("beta[1]", "beta[2]", 
            "beta[3]", "beta[4]",
            "beta[5]", "beta[6]",
            "beta[7]", "beta[8]",
            "sigma")
```


```{r, echo = F, results = 'hide', eval = F}
# Trace plots
# Trace plots show whether chains are mixing well and exploring the parameter space.
# Chains should mix well (lines overlap)
# No long trends (no “stickiness”)
# Ideally all chains overlap around the same mean

# Model without ADH
mcmc_trace(aggphys_draws, pars = aggphys_params)
# They look beautiful

```


```{r, echo = F, results = 'hide', eval = F}
# Model WITH ADH
mcmc_trace(aggphys_draws_adh, pars = aggphys_params_adh)
```



```{r, echo = F, results = 'hide'}
# R-hat and ESS diagnostics
# R-hat ≈ 1.00 → converged
# ESS should be reasonably large (e.g., >100–200 per parameter)
aggphys_fit$summary(variables = aggphys_params)
# We see that rhats are all approx 1
# ESS should be around > 200*9 = 1800 and that seems to be the case for most
```

```{r, echo = F, results='hide'}
# Model WITH ADH
aggphys_fit_adh$summary(variables = aggphys_params_adh)

```


```{r, echo = F, results='hide', eval = F}
# Posterior Density Plots
# Overlaid densities from multiple chains should match closely
# If densities differ substantially between chains → poor mixing
mcmc_dens_overlay(draws_aggphys, pars = aggphys_params)

```

```{r, echo = F, include = F, eval = F, results='hide'}
# Posterior density plots
# Model WITH ADH
mcmc_dens_overlay(draws_aggphys_adh, pars = aggphys_params_adh)
```


```{r, echo = F, results='hide', eval = F}
# Auto-correlation 
# Autocorrelation should decay quickly
# High autocorrelation → may need more iterations or thinning
mcmc_acf(draws_aggphys, pars = aggphys_params)

```

```{r, echo = F, results='hide', eval = F}
# Model WITH ADH
mcmc_acf(draws_aggphys_adh, pars = aggphys_params_adh)
```


```{r, echo = F, results='hide', eval = F}
# Diagnostics from cmdstan
aggphys_fit$cmdstan_diagnose()
# Need to load actual aggphys and not aggphys .rds for this

### Output
# Checking sampler transitions treedepth.
# Treedepth satisfactory for all transitions.
# 
# Checking sampler transitions for divergences.
# No divergent transitions found.
# 
# Checking E-BFMI - sampler transitions HMC potential energy.
# E-BFMI satisfactory.
# 
# Rank-normalized split effective sample size satisfactory for all parameters.
# 
# Rank-normalized split R-hat values satisfactory for all parameters.
# 
# Processing complete, no problems detected

```

```{r, echo = F, results='hide', eval = F}
# Model WITH ADH
# Diagnostics from cmdstan
aggphys_fit_adh$cmdstan_diagnose()

### Output
# Checking sampler transitions treedepth.
# Treedepth satisfactory for all transitions.
# 
# Checking sampler transitions for divergences.
# No divergent transitions found.
# 
# Checking E-BFMI - sampler transitions HMC potential energy.
# E-BFMI satisfactory.
# 
# Rank-normalized split effective sample size satisfactory for all parameters.
# 
# Rank-normalized split R-hat values satisfactory for all parameters.
# 
# Processing complete, no problems detected.

```


```{r, echo = F, eval = F}
## Save output if desired ###
# MCMC Output
sink("Output/P1_aggphys_MCMCOutput.txt")
print(summary_table)
print(loo_res)
print(aggphys_fit$cmdstan_diagnose())
print(aggphys_fit$summary(variables=aggphys_params))
sink()

# Diagnostic plots
pdf('Output/P1_aggphys_MCMC_DiagnosticPlots.pdf')
mcmc_trace(aggphys_draws, pars = aggphys_params)
mcmc_dens_overlay(aggphys_draws, pars = aggphys_params)
mcmc_acf(aggphys_draws, pars = aggphys_params)
dev.off()

```


```{r, echo = F}
# Read in data
data <- read.csv("../Data/final_hiv_data.csv")

# Assessment of the data first
# str(data)

# Create categorical labels and variables 

# Hard drugs
data$hard_drugs_cat <- factor(data$hard_drugs,
                              levels = c(0, 1),
                              labels = c("No", "Yes"))

# Move after hard drugs variable
data <- data %>% 
  relocate(hard_drugs_cat, .after = hard_drugs)

# Convert collapse race to factor
data$collapse_RACE <- factor(data$collapse_RACE)

# Convert collapse educ to factor
data$collapsed_EDUCBAS <- factor(data$collapsed_EDUCBAS)

# Convert SMOKE cat to factor
data$collapse_SMOKE <- factor(data$collapse_SMOKE)

# Relabel the variables
med_data <- apply_labels(data,
                   hard_drugs = "Hard Drug Use at BL",
                   ADH_yr2 = "HAART Adherence",
                   VLOAD_yr2 = "VLOAD - yr2")
```



```{r, echo = F}
################################################################################
###                                AGG MENT                                  ###
################################################################################


### Linear model WITHOUT ADHERENCE
aggment_linear <- lm(AGG_MENT_yr2 ~ hard_drugs_cat + AGG_MENT_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS,
                     data = data)

# sink("Output/P1_aggment_Freq_Output.txt")
# print(aggment_linear)
# summary(aggment_linear)
# confint(aggment_linear)
# anova(aggment_linear)
# sink()

``` 

```{r, echo = F}
### Linear model WITH ADHERENCE
aggment_adh_linear <- lm(AGG_MENT_yr2 ~ hard_drugs_cat + AGG_MENT_yr0
                    + BMI + collapse_SMOKE
                    + age + collapse_RACE + collapsed_EDUCBAS
                    + collapsed_ADH,
                     data = data)

```



```{r, echo = F}
### Bayesian Model

# Linear model WITHOUT ADHERENCE

###########################################################################
# STEP 1: Define the model
# Normal priors on the regression coefficients
# Half normal priors on sigma
###########################################################################


stan_file_aggment <- write_stan_file("data {
  int<lower=0> N;                  // number of observations
  int<lower=0> P;                  // number of predictors including intercept
  matrix[N, P] X;                  // design matrix (first column = intercept)
  vector[N] y;                     // outcome

  vector[P] prior_mean;            // prior means for each beta
  vector<lower=0>[P] prior_sd;     // prior SDs for each beta

  real<lower=0> sigma_prior_sd;    // SD for half-normal prior on sigma
}

parameters {
  vector[P] beta;                  // regression coefficients
  real<lower=0> sigma;             // residual SD
}

model {
  // Vectorized priors for regression coefficients
  beta ~ normal(prior_mean, prior_sd);

  // Half-normal prior for sigma
  sigma ~ normal(0, sigma_prior_sd);

  // Likelihood
  y ~ normal(X * beta, sigma);
  

}

generated quantities {
  // log likelihood for each observation for calculating model fit stats
  vector[N] log_lik;
  for (n in 1:N) {
    log_lik[n] = normal_lpdf(y[n] | X[n] * beta, sigma);}
  }", dir="STAN", basename='aggment_primary'
)

```

```{r, echo = F}
# Linear Model WITH ADHERENCE

###########################################################################
# STEP 1: Define the model
# Normal priors on the regression coefficients
# Half normal priors on sigma
###########################################################################


stan_file_aggment_adh <- write_stan_file("data {
  int<lower=0> N;                  // number of observations
  int<lower=0> P;                  // number of predictors including intercept
  matrix[N, P] X;                  // design matrix (first column = intercept)
  vector[N] y;                     // outcome

  vector[P] prior_mean;            // prior means for each beta
  vector<lower=0>[P] prior_sd;     // prior SDs for each beta

  real<lower=0> sigma_prior_sd;    // SD for half-normal prior on sigma
}

parameters {
  vector[P] beta;                  // regression coefficients
  real<lower=0> sigma;             // residual SD
}

model {
  // Vectorized priors for regression coefficients
  beta ~ normal(prior_mean, prior_sd);

  // Half-normal prior for sigma
  sigma ~ normal(0, sigma_prior_sd);

  // Likelihood
  y ~ normal(X * beta, sigma);
  

}

generated quantities {
  // log likelihood for each observation for calculating model fit stats
  vector[N] log_lik;
  for (n in 1:N) {
    log_lik[n] = normal_lpdf(y[n] | X[n] * beta, sigma);}
  }", dir="STAN", basename='aggment_adh'
)

```


```{r, echo = F, message = F, warning = F}
###########################################################################
# STEP 2: Compile the Stan program
###########################################################################

# Model WITHOUT adherence
aggment_mod <- cmdstan_model('STAN/aggment_primary.stan')

```

```{r, echo = F, message = F, warning = F}
# Model WITH adherence
aggment_mod_adh <- cmdstan_model('STAN/aggment_adh.stan')

```


```{r, echo = F}
###########################################################################
# STEP 3: Prepare your data to pass to Stan
###########################################################################

# Model WITHOUT adherence
# Outcome data
y_aggment <- data$AGG_MENT_yr2

# Design matrix in our linear regression
X_aggment <- model.matrix(~ hard_drugs_cat + AGG_MENT_yr0
   + BMI + collapse_SMOKE
   + age + collapse_RACE + collapsed_EDUCBAS,
   data = data)

# These are variables needed in Stan
# N is the number of observations in your data set 
# P is the number of columns in the design matrix
N_aggment <- nrow(X_aggment)
P_aggment <- ncol(X_aggment)

# Hyper-parameters for prior distributions
# Prior for regression coefficients is Normal(0, 10000)
# Prior for model error SD (sigma) is Half-Normal(0, 100)

# MVN mean (mean in the prior on the regression coefficients)
m_aggment <- c(1, rep(0, (P_aggment - 1)))   

# # SD in the prior on regression coefficients --> variance 100^2
s_aggment <- rep(1000, P_aggment) 

sigma_sd_aggment <- 1000

# create data list to pass to STAN
data_list <- list(
  N = N_aggment,
  P = P_aggment,
  X = X_aggment,
  y = y_aggment,
  prior_mean = m_aggment,
  prior_sd = s_aggment,
  sigma_prior_sd = sigma_sd_aggment
)
```


```{r, echo = F}
# Model WITH adherence
# Outcome data
y_aggment_adh <- data$AGG_MENT_yr2

# Design matrix in our linear regression
X_aggment_adh <- model.matrix(~ hard_drugs_cat + AGG_MENT_yr0
   + BMI + collapse_SMOKE
   + age + collapse_RACE + collapsed_EDUCBAS
   + collapsed_ADH,
   data = data)

# These are variables needed in Stan
# N is the number of observations in your data set 
# P is the number of columns in the design matrix
N_aggment_adh <- nrow(X_aggment_adh)
P_aggment_adh <- ncol(X_aggment_adh)

# Hyper-parameters for prior distributions
# Prior for regression coefficients is Normal(0, 10000)
# Prior for model error SD (sigma) is Half-Normal(0, 100)

# MVN mean (mean in the prior on the regression coefficients)
m_aggment_adh <- c(1, rep(0, (P_aggment_adh - 1)))   

# # SD in the prior on regression coefficients --> variance 100^2
s_aggment_adh <- rep(1000, P_aggment_adh) 

sigma_sd_aggment_adh <- 1000

# create data list to pass to STAN
data_list_adh <- list(
  N = N_aggment_adh,
  P = P_aggment_adh,
  X = X_aggment_adh,
  y = y_aggment_adh,
  prior_mean = m_aggment_adh,
  prior_sd = s_aggment_adh,
  sigma_prior_sd = sigma_sd_aggment_adh
)
```


```{r, echo = F, message=F, warning = F, results = 'hide'}
###########################################################################
# STEP 4: Fit Model
###########################################################################

# Model WITHOUT adherence
# Specify the warmup and sampling iterations
aggment_fit <- suppressMessages(
  suppressWarnings(
  aggment_mod$sample(
  data = data_list,
  chains = 4,
  iter_warmup = 1000,
  iter_sampling = 2000,
  seed = 789,
  refresh = 0
 )
)
)

```

```{r, echo = F, message=F, warning = F, results = 'hide'}
# Model WITH adherence
# Specify the warmup and sampling iterations
aggment_fit_adh <- aggment_mod_adh$sample(
  data = data_list_adh,
  chains = 4,
  iter_warmup = 1000,
  iter_sampling = 2000,
  seed = 798
)
```


```{r, echo = F, results='hide'}
###########################################################################
# STEP 5: Summarize the posterior
###########################################################################

# Model without ADH
aggment_fit$summary(variables=c("beta[1]", "beta[2]", 
                              "beta[3]",
        # including extra betas to double check results with linear model
                              "beta[4]", "beta[5]",
                              "beta[6]", "beta[7]",
                              "beta[8]", 
                              "sigma"))


# Fancier Table of Results of Interest from the Posterior
# Extract posterior draws from cmdstanr fit
draws_aggment <- aggment_fit$draws()  

draws_mat_aggment <- as_draws_matrix(draws_aggment)
params_aggment <- colnames(draws_mat_aggment)

# Exclude non-parameter columns from summarization: lp__ and log_lik[n]
params_aggment <- params_aggment[!grepl("lp__|log_lik", params_aggment)]

```

```{r, echo = F, results = 'hide'}
# Model WITH ADH

aggment_fit_adh$summary(variables=c("beta[1]", "beta[2]", 
                              "beta[3]",
        # including extra betas to double check results with linear model
                              "beta[4]", "beta[5]",
                              "beta[6]", "beta[7]",
                              "beta[8]", 
                              "sigma"))


# Fancier Table of Results of Interest from the Posterior
# Extract posterior draws from cmdstanr fit
draws_aggment_adh <- aggment_fit_adh$draws()  

draws_mat_aggment_adh <- as_draws_matrix(draws_aggment_adh)
params_aggment_adh <- colnames(draws_mat_aggment_adh)

# Exclude non-parameter columns from summarization: lp__ and log_lik[n]
params_aggment_adh <- params_aggment_adh[!grepl("lp__|log_lik", params_aggment_adh)]

```

```{r, echo = F, results = 'hide'}
# Model without ADH
# Build summary table
summary_table_aggment <- lapply(params_aggment, function(p) {
  vals <- as.numeric(draws_mat_aggment[, p])
  
  # Monte Carlo standard error
  # Used to determine if we have run enough iterations
  # rule of thumb: MCSE should be less than 6.27% of the posterior standard deviation
  mcse_val <- mcmcse::mcse(vals)$se
  
  # Effective sample size
  ess_val <- ess_bulk(vals)
  
  # 95% HPDI
  hpd <- hdi(vals, ci = 0.95)
  
  tibble(
    Parameter = p,
    Estimate  = mean(vals),
    MCSE      = mcse_val,
    Std_Dev   = sd(vals),
    HPDI_2.5  = hpd$CI_low,
    HPDI_97.5 = hpd$CI_high,
    ESS       = ess_val
  )}) %>% 
  bind_rows()

print(summary_table_aggment)

# Compare with linear fit
summary(aggment_linear)

# check that mcmcse < 6% of posterior SD
(100*summary_table_aggment$MCSE/summary_table_aggment$Std_Dev) < 6
# All True


```


```{r, echo = F, results = 'hide'}
# Model WITH ADH
# Build summary table
summary_table_aggment_adh <- lapply(params_aggment_adh, function(p) {
  vals <- as.numeric(draws_mat_aggment_adh[, p])
  
  # Monte Carlo standard error
  # Used to determine if we have run enough iterations
  # rule of thumb: MCSE should be less than 6.27% of the posterior standard deviation
  mcse_val <- mcmcse::mcse(vals)$se
  
  # Effective sample size
  ess_val <- ess_bulk(vals)
  
  # 95% HPDI
  hpd <- hdi(vals, ci = 0.95)
  
  tibble(
    Parameter = p,
    Estimate  = mean(vals),
    MCSE      = mcse_val,
    Std_Dev   = sd(vals),
    HPDI_2.5  = hpd$CI_low,
    HPDI_97.5 = hpd$CI_high,
    ESS       = ess_val
  )}) %>% 
  bind_rows()

print(summary_table_aggment_adh)

# Compare with linear fit
summary(aggment_adh_linear)

# check that mcmcse < 6% of posterior SD
(100*summary_table_aggment_adh$MCSE/summary_table_aggment_adh$Std_Dev) < 6
# All True

```


```{r, echo = F, results = 'hide'}
# Compare estimates for models with and without adherence

# Model without ADH
aggment_terms <- summary_table_aggment[2, ]

# Subset to Parameter, Estimate, HPDIs
aggment_terms <- subset(aggment_terms,
                      select = c("Parameter",
                                 "Estimate",
                                 "HPDI_2.5",
                                 "HPDI_97.5"))

# Model WITH ADH
# Pull term for hard drugs variable
aggment_adh_terms <- summary_table_aggment_adh[2, ]

# Subset to Parameter, Estimate, HPDIs
aggment_adh_terms <- subset(aggment_adh_terms,
                    select = c("Parameter", 
                               "Estimate",
                               "HPDI_2.5",
                               "HPDI_97.5"))

# Display side by side like with frequentist models
# Create data frame
comp_models_aggment <- data.frame(Variable = "HD at Baseline (Yes)",
                 Estimate = round(aggment_terms$Estimate, 3),
                 `95% HDI` = paste(round(aggment_terms$HPDI_2.5, 3), 
                                    round(aggment_terms$HPDI_97.5, 3), 
                                    sep = ", "),
                 Estimate = round(aggment_adh_terms$Estimate, 3),
                 `95% HDI` = paste(round(aggment_adh_terms$HPDI_2.5, 3),
                                    round(aggment_adh_terms$HPDI_97.5, 3),
                                    sep = ", "),
                 check.names = FALSE)

# Display with kable
# kable(
#   comp_models_aggment,
#   # format = "latex",
#   booktabs = TRUE,
#   digits = 3,
#   caption = "Model Summary") |>
#   kable_styling(latex_options = c("striped", "hold_position")) |>
#   add_header_above(c(" ", "Model Without ADH" = 2, 
#                      "Model With ADH" = 2)) %>% 
#   footnote(general = "HPDI - Highest Density Interval ")

```


```{r, echo = F}
###########################################################################
# STEP 6: Model fit statistics (you only need 1 of these, not all 3)
###########################################################################
# Extract the log likelihood info
# For aggment model without ADH
loglik_mat_aggment <- as_draws_matrix(aggment_fit$draws("log_lik")) 
# iterations x N

# Get the LOO-CV
loo_res_aggment <- loo(loglik_mat_aggment)
# looic: 3223.8  

```

```{r, echo = F}
# Model WITH ADH
loglik_mat_aggment_adh <- as_draws_matrix(aggment_fit_adh$draws("log_lik")) 
# iterations x N

# Get the LOO-CV
loo_res_aggment_adh <- loo(loglik_mat_aggment_adh)
# print(loo_res_aggment_adh)
# 3222.1
```


```{r, echo = F}
###########################################################################
# STEP 6: MCMC Diagnostics
###########################################################################

# Model without ADH
# Dimensions: iterations x chains x parameters
aggment_draws <- as_draws_array(aggment_fit$draws())

# Tidy format for bayesplot / ggplot
aggment_draws_df <- as_draws_df(aggment_fit$draws())  

# parameters
aggment_params <- c("beta[1]", "beta[2]", 
            "beta[3]", "beta[4]",
            "beta[5]", "beta[6]",
            "beta[7]", "beta[8]",
            "sigma")

```

```{r, echo = F}
# Dimensions: iterations x chains x parameters
aggment_draws_adh <- as_draws_array(aggment_fit_adh$draws())

# Tidy format for bayesplot / ggplot
aggment_draws_df_adh <- as_draws_df(aggment_fit_adh$draws())  

# parameters
aggment_params_adh <- c("beta[1]", "beta[2]", 
            "beta[3]", "beta[4]",
            "beta[5]", "beta[6]",
            "beta[7]", "beta[8]",
            "sigma")
```


```{r, echo = F, results = 'hide', eval = F}
# Trace plots
# Trace plots show whether chains are mixing well and exploring the parameter space.
# Chains should mix well (lines overlap)
# No long trends (no “stickiness”)
# Ideally all chains overlap around the same mean

# Model without ADH
mcmc_trace(aggment_draws, pars = aggment_params)
# They look beautiful

```


```{r, echo = F, results = 'hide', eval = F}
# Model WITH ADH
mcmc_trace(aggment_draws_adh, pars = aggment_params_adh)
```



```{r, echo = F, results = 'hide'}
# R-hat and ESS diagnostics
# R-hat ≈ 1.00 → converged
# ESS should be reasonably large (e.g., >100–200 per parameter)
aggment_fit$summary(variables = aggment_params)
# We see that rhats are all approx 1
# ESS should be around > 200*9 = 1800 and that seems to be the case for most
```

```{r, echo = F, results='hide'}
# Model WITH ADH
aggment_fit_adh$summary(variables = aggment_params_adh)

```


```{r, echo = F, results='hide', eval = F}
# Posterior Density Plots
# Overlaid densities from multiple chains should match closely
# If densities differ substantially between chains → poor mixing
mcmc_dens_overlay(draws_aggment, pars = aggment_params)

```

```{r, echo = F, include = F, eval = F, results='hide'}
# Posterior density plots
# Model WITH ADH
mcmc_dens_overlay(draws_aggment_adh, pars = aggment_params_adh)
```


```{r, echo = F, results='hide', eval = F}
# Auto-correlation 
# Autocorrelation should decay quickly
# High autocorrelation → may need more iterations or thinning
mcmc_acf(draws_aggment, pars = aggment_params)

```

```{r, echo = F, results='hide', eval = F}
# Model WITH ADH
mcmc_acf(draws_aggment_adh, pars = aggment_params_adh)
```


```{r, echo = F, results='hide', eval = F}
# Diagnostics from cmdstan
aggment_fit$cmdstan_diagnose()
# Need to load actual aggment and not aggment .rds for this

### Output
# Checking sampler transitions treedepth.
# Treedepth satisfactory for all transitions.
# 
# Checking sampler transitions for divergences.
# No divergent transitions found.
# 
# Checking E-BFMI - sampler transitions HMC potential energy.
# E-BFMI satisfactory.
# 
# Rank-normalized split effective sample size satisfactory for all parameters.
# 
# Rank-normalized split R-hat values satisfactory for all parameters.
# 
# Processing complete, no problems detected.

```

```{r, echo = F, results='hide', eval = F}
# Model WITH ADH
# Diagnostics from cmdstan
aggment_fit_adh$cmdstan_diagnose()

### Output
# Checking sampler transitions treedepth.
# Treedepth satisfactory for all transitions.
# 
# Checking sampler transitions for divergences.
# No divergent transitions found.
# 
# Checking E-BFMI - sampler transitions HMC potential energy.
# E-BFMI satisfactory.
# 
# Rank-normalized split effective sample size satisfactory for all parameters.
# 
# Rank-normalized split R-hat values satisfactory for all parameters.
# 
# Processing complete, no problems detected.

```



```{r, echo = F, eval = F}
## Save output if desired ###
# MCMC Output
sink("Output/P1_aggment_MCMCOutput.txt")
print(summary_table)
print(loo_res)
print(aggment_fit$cmdstan_diagnose())
print(aggment_fit$summary(variables=aggment_params))
sink()

# Diagnostic plots
pdf('Output/P1_aggment_MCMC_DiagnosticPlots.pdf')
mcmc_trace(aggment_draws, pars = aggment_params)
mcmc_dens_overlay(aggment_draws, pars = aggment_params)
mcmc_acf(aggment_draws, pars = aggment_params)
dev.off()

```



<!----------------------------------------------------------------------------->
<!---                             Primary Outcomes                         ---->
<!----------------------------------------------------------------------------->

<!----------------------------------------------------------------------------->
<!---               Comparing Frequentist and Bayesian Estimates           ---->
<!----------------------------------------------------------------------------->

```{r, echo = F}
# Compare estimates for models with and without adherence

### VLOAD
## Bayesian 
# Model without ADH
vload_terms <- summary_table_vload[2, ] # Hard drugs beta

# Subset to Parameter, Estimate, HPDIs
vload_terms <- subset(vload_terms,
                      select = c("Parameter",
                                 "Estimate",
                                 "HPDI_2.5",
                                 "HPDI_97.5"))

## Frequentist
vload_lin_terms <- tidy(vload_linear, conf.int = TRUE)
# Subset to hard drugs beta
vload_lin_terms <- vload_lin_terms[2, ]
# Subse to estimate and conf values
vload_lin_terms <- subset(vload_lin_terms,
             select = c("estimate", "conf.low", "conf.high"))


### LEU3N
## Bayesian
# Model without ADH
leu3n_terms <- summary_table_leu3n[2, ]

# Subset to Parameter, Estimate, HPDIs
leu3n_terms <- subset(leu3n_terms,
                      select = c("Parameter",
                                 "Estimate",
                                 "HPDI_2.5",
                                 "HPDI_97.5"))
## Frequentist
leu3n_lin_terms <- tidy(leu3n_linear, conf.int = TRUE)
# Subset to hard drugs beta
leu3n_lin_terms <- leu3n_lin_terms[2, ]
# Subse to estimate and conf values
leu3n_lin_terms <- subset(leu3n_lin_terms,
             select = c("estimate", "conf.low", "conf.high"))


### AGG PHYS
## Bayesian
# Model without ADH
aggphys_terms <- summary_table_aggphys[2, ]

# Subset to Parameter, Estimate, HPDIs
aggphys_terms <- subset(aggphys_terms,
                      select = c("Parameter",
                                 "Estimate",
                                 "HPDI_2.5",
                                 "HPDI_97.5"))
aggphys_lin_terms <- tidy(aggphys_linear, conf.int = TRUE)
# Subset to hard drugs beta
aggphys_lin_terms <- aggphys_lin_terms[2, ]
# Subse to estimate and conf values
aggphys_lin_terms <- subset(aggphys_lin_terms,
             select = c("estimate", "conf.low", "conf.high"))


### AGG MENT
## Bayesian
# Model without ADH
aggment_terms <- summary_table_aggment[2, ]

# Subset to Parameter, Estimate, HPDIs
aggment_terms <- subset(aggment_terms,
                      select = c("Parameter",
                                 "Estimate",
                                 "HPDI_2.5",
                                 "HPDI_97.5"))

## Frequentist
aggment_lin_terms <- tidy(aggment_linear, conf.int = TRUE)
# Subset to hard drugs beta
aggment_lin_terms <- aggment_lin_terms[2, ]
# Subse to estimate and conf values
aggment_lin_terms <- subset(aggment_lin_terms,
             select = c("estimate", "conf.low", "conf.high"))



# Display side by side like with frequentist models
# Create data frame
all_outcomes <- data.frame(Outcome = 
                          # VLOAD
                          c("Log(Viral Load)",
                          # LEU3N
                          "CD4+ T-Cell Count",
                          # AGG PHYS
                          "Agg. Mental QOL Score",
                          # AGG MENT
                          "Agg. Physical QOL Score"),
                  # Bayesian Estimates
                 Estimate = 
                   # VLOAD
                   c(round(exp(vload_terms$Estimate), 3),
                   # LEU3N
                    round(leu3n_terms$Estimate, 3),
                   # AGG PHYS
                    round(aggphys_terms$Estimate, 3),
                   # AGG MENT
                   round(aggment_terms$Estimate, 3)),
                 # Bayesian HDIs
                 `95% HDI` = c(
                   # VLOAD
                   paste(round(exp(vload_terms$HPDI_2.5), 3), 
                                    round(exp(vload_terms$HPDI_97.5), 3), 
                                    sep = ", "),
                   # LEU3N
                   paste(round(leu3n_terms$HPDI_2.5, 3), 
                                    round(leu3n_terms$HPDI_97.5, 3), 
                                    sep = ", "),
                   # AGG PHYS
                   paste(round(aggphys_terms$HPDI_2.5, 3), 
                                    round(aggphys_terms$HPDI_97.5, 3), 
                                    sep = ", "),
                   # AGG MENT
                   paste(round(aggment_terms$HPDI_2.5, 3), 
                                    round(aggment_terms$HPDI_97.5, 3), 
                                    sep = ", ")),
                 # Frequentist estimates
                 Estimate = 
                   # VLOAD
                   c(round(exp(vload_lin_terms$estimate), 3),
                   # LEU3N
                    round(leu3n_lin_terms$estimate, 3),
                   # AGG PHYS
                    round(aggphys_lin_terms$estimate, 3),
                   # AGG MENT
                   round(aggment_lin_terms$estimate, 3)),
                 # Frequentist CIs
                 `95% CI` = c(
                   # VLOAD
                   paste(round(exp(vload_lin_terms$conf.low), 3), 
                              round(exp(vload_lin_terms$conf.high), 3), 
                                    sep = ", "),
                   # LEU3N
                   paste(round(leu3n_lin_terms$conf.low, 3), 
                                round(leu3n_lin_terms$conf.high, 3), 
                                    sep = ", "),
                   # AGG PHYS
                   paste(round(aggphys_lin_terms$conf.low, 3), 
                               round(aggphys_lin_terms$conf.high, 3), 
                                    sep = ", "),
                   # AGG MENT
                   paste(round(aggment_lin_terms$conf.low, 3), 
                               round(aggment_lin_terms$conf.high, 3), 
                                    sep = ", ")),
                 check.names = FALSE
)

# Display with kable
kbl(
  all_outcomes,
  booktabs = TRUE,
  digits = 3,
  caption = "Model Results for Bayesian and Frequentist Analyses") |>
  # subcaption = "Meow" |>
  kable_styling(latex_options = c("striped", "hold_position")) |> 
  add_header_above(c(" ", "Bayesian Model" = 2,
                     "Frequentist Model" = 2)) |>
  footnote(number = c("HDI = Highest Density Interval & CI = Confidence Interval",
                      "Estimates interpreted as changes in outcome for hard drug users compared to non-hard drug users",
                      "Estimates for Viral Load are exponentiated"))
```
<!-- Table 2 -->
<!-- VLOAD: 13% lower for 95% HDI: 65% lower to 109% higher -->
In Table 2 we compare estimates and measures of uncertainty for the estimates. Estimates with the frequentist and bayesian models yielded very similar results.
We confirmed that with weak priors, Bayesian estimates were consistent with frequentist results.
Both the estimates and the 95% confidence interval and 95% highest density interval are quite similar.
The frequentist analyses validates our findings with the Bayesian models. 
Interpreting the findings with the Bayesian models, we found that
after adjustment, the estimated viral load ratio comparing baseline hard drug users to non-users at 2 years was 0.87 (95% HDI: 0.35, 2.09). This corresponds to an estimated 13% lower viral load among drug users. We note that the interval was wide and included both substantial decreases and increases, indicating considerable uncertainty and no clear evidence of a difference. As HDI excluded 0, these findings are not significant.
CD4+ T-Cell Counts were ~ -163 units lower (-223.956, -102.143) for hard-drug users compared to non-hard drug users at 2 years, after adjusting for covariates The difference between hard drug users and non-hard drug users is statistically significant as the HDI does not include 0. 
The estimated difference in aggregate mental quality of life score was ~ -3 units lower (-6.147, -0.711) for hard-drug users compared to non-hard drug users after 2 years of HAART. Since the HDI excludes 0, the difference between the two hard drug use groups is statistically significant. 
Lastly, aggregate mental quality of life score was estimated to be around -0.3 units lower (-3.565, 3.035) for non-hard drug users compared to hard-drug users after 2 years of HAART and after adjustment. However, this difference is not statistically significant as the HDI includes 0. 

<!-- Table 3  -->
In Table 3 we compare estimates between the Bayesian models that did and did not include adherence for each outcome. 
In Figure 1, we visualize the differences in the estimates between the models with and without ADH. Since CD4+ T-Cell counts was on a different scale compared to the other three outcomes, we created two separate panels to better illustrate differences.
Estimates for exponentiated viral load and aggregate mental quality of life score differed by ~0.1 or so and HDIs shared overlap. 
As estimates and HDIs for the models with and without adherence yielded similar results, it does not appear to be that adherence to HAART explains differences between hard drug and non-hard drug users. 
For CD4+ T-Cell Count, estimates differed by around 6 points and HDIs were about 10 points higher for the low and high interval points for the model with ADH.
Estimates varied slightly when including and exluding ADH, however, the small difference in the model estimates suggests that differences in medication adherence does not explain differences in CD4+ T-Cell Counts between hard drug and non-hard drug users. 
Models with and without ADH were noticeably different for aggregate physical quality of life score. This suggests that HAART adherence may influence differences in aggregate physical quality of life scores between the two hard drug use groups. However, we mentioned above that differences between the hard drug use groups for this quality of life score was not significant. 


<!----------------------------------------------------------------------------->
<!---                     Estimates With & Without ADH                     ---->
<!----------------------------------------------------------------------------->


```{r, echo = F}
#### Compare estimates for models with and without adherence
### VLOAD
# Pull term for hard drugs variable
vload_adh_terms <- summary_table_vload_adh[2, ]

# Subset to Parameter, Estimate, HPDIs
vload_adh_terms <- subset(vload_adh_terms,
                    select = c("Parameter", 
                               "Estimate",
                               "HPDI_2.5",
                               "HPDI_97.5"))


### LEU3N
# Model WITH ADH
# Pull term for hard drugs variable
leu3n_adh_terms <- summary_table_leu3n_adh[2, ]

# Subset to Parameter, Estimate, HPDIs
leu3n_adh_terms <- subset(leu3n_adh_terms,
                    select = c("Parameter", 
                               "Estimate",
                               "HPDI_2.5",
                               "HPDI_97.5"))

### AGG PHYS
# Model WITH ADH
# Pull term for hard drugs variable
aggphys_adh_terms <- summary_table_aggphys_adh[2, ]

# Subset to Parameter, Estimate, HPDIs
aggphys_adh_terms <- subset(aggphys_adh_terms,
                    select = c("Parameter", 
                               "Estimate",
                               "HPDI_2.5",
                               "HPDI_97.5"))

### AGG MENT
# Model WITH ADH
# Pull term for hard drugs variable
aggment_adh_terms <- summary_table_aggment_adh[2, ]

# Subset to Parameter, Estimate, HPDIs
aggment_adh_terms <- subset(aggment_adh_terms,
                    select = c("Parameter", 
                               "Estimate",
                               "HPDI_2.5",
                               "HPDI_97.5"))

# Display side by side like with frequentist models
# Create data frame
# Display side by side like with frequentist models
# Create data frame
adh_comparison <- data.frame(Outcome = 
                          # VLOAD
                          c("Log(Viral Load)",
                          # LEU3N
                          "CD4+ T-Cell Count",
                          # AGG PHYS
                          "Agg. Mental QOL Score",
                          # AGG MENT
                          "Agg. Physical QOL Score"),
                  ## WITHOUT ADH
                 Estimate = 
                   # VLOAD
                   c(round(exp(vload_terms$Estimate), 3),
                   # LEU3N
                    round(leu3n_terms$Estimate, 3),
                   # AGG PHYS
                    round(aggphys_terms$Estimate, 3),
                   # AGG MENT
                   round(aggment_terms$Estimate, 3)),
                 ## WITHOUT ADH
                 `95% HDI` = c(
                   # VLOAD
                   paste(round(exp(vload_terms$HPDI_2.5), 3), 
                                    round(exp(vload_terms$HPDI_97.5), 3), 
                                    sep = ", "),
                   # LEU3N
                   paste(round(leu3n_terms$HPDI_2.5, 3), 
                                    round(leu3n_terms$HPDI_97.5, 3), 
                                    sep = ", "),
                   # AGG PHYS
                   paste(round(aggphys_terms$HPDI_2.5, 3), 
                                    round(aggphys_terms$HPDI_97.5, 3), 
                                    sep = ", "),
                   # AGG MENT
                   paste(round(aggment_terms$HPDI_2.5, 3), 
                                    round(aggment_terms$HPDI_97.5, 3), 
                                    sep = ", ")),
                 ## WITH ADH
                 Estimate = 
                   # VLOAD
                   c(round(exp(vload_adh_terms$Estimate), 3),
                   # LEU3N
                    round(leu3n_adh_terms$Estimate, 3),
                   # AGG PHYS
                    round(aggphys_adh_terms$Estimate, 3),
                   # AGG MENT
                   round(aggment_adh_terms$Estimate, 3)),
                 # WITH ADH
                 `95% HDI` = c(
                   # VLOAD
                   paste(round(exp(vload_adh_terms$HPDI_2.5), 3), 
                                    round(exp(vload_adh_terms$HPDI_97.5), 3), 
                                    sep = ", "),
                   # LEU3N
                   paste(round(leu3n_adh_terms$HPDI_2.5, 3), 
                                    round(leu3n_adh_terms$HPDI_97.5, 3), 
                                    sep = ", "),
                   # AGG PHYS
                   paste(round(aggphys_adh_terms$HPDI_2.5, 3), 
                                    round(aggphys_adh_terms$HPDI_97.5, 3), 
                                    sep = ", "),
                   # AGG MENT
                   paste(round(aggment_adh_terms$HPDI_2.5, 3), 
                                    round(aggment_adh_terms$HPDI_97.5, 3), 
                                    sep = ", ")),
                 check.names = FALSE
)

# Display with kable
kable(
  adh_comparison,
  # format = "latex",
  booktabs = TRUE,
  digits = 3,
  caption = "Model Summaries Without and With Medication Adherence Taken into Account") |>
  kable_styling(latex_options = c("striped", "hold_position",
                                  "scale_down")) |>
  add_header_above(c(" ", "Model Without ADH" = 2, 
                     "Model With ADH" = 2)) %>% 
  footnote(number = c("HDI = Highest Density Interval & CI = Confidence Interval",
                      "Estimates interpreted as changes in outcome for hard drug users compared to non-hard drug users",
                      "Estimates for viral load are exponentiated"))
```

```{r, echo = F, warning = F, message = F}
# Create data frame 
adh_comparison2 <- data.frame(Outcome = 
                          # VLOAD
                          c("Log(Viral Load)",
                          # LEU3N
                          "CD4+ T-Cell Count",
                          # AGG PHYS
                          "Agg. Mental QOL Score",
                          # AGG MENT
                          "Agg. Physical QOL Score"),
                  ## WITHOUT ADH
                 Estimate_No_ADH = 
                   # VLOAD
                   c(round(exp(vload_terms$Estimate), 3),
                   # LEU3N
                    round(leu3n_terms$Estimate, 3),
                   # AGG PHYS
                    round(aggphys_terms$Estimate, 3),
                   # AGG MENT
                   round(aggment_terms$Estimate, 3)),
                 ## WITHOUT ADH
                 `95% HDI (No ADH)` = c(
                   # VLOAD
                   paste(round(exp(vload_terms$HPDI_2.5), 3), 
                                    round(exp(vload_terms$HPDI_97.5), 3), 
                                    sep = ", "),
                   # LEU3N
                   paste(round(leu3n_terms$HPDI_2.5, 3), 
                                    round(leu3n_terms$HPDI_97.5, 3), 
                                    sep = ", "),
                   # AGG PHYS
                   paste(round(aggphys_terms$HPDI_2.5, 3), 
                                    round(aggphys_terms$HPDI_97.5, 3), 
                                    sep = ", "),
                   # AGG MENT
                   paste(round(aggment_terms$HPDI_2.5, 3), 
                                    round(aggment_terms$HPDI_97.5, 3), 
                                    sep = ", ")),
                 ## WITH ADH
                 Estimate_ADH = 
                   # VLOAD
                   c(round(exp(vload_adh_terms$Estimate), 3),
                   # LEU3N
                    round(leu3n_adh_terms$Estimate, 3),
                   # AGG PHYS
                    round(aggphys_adh_terms$Estimate, 3),
                   # AGG MENT
                   round(aggment_adh_terms$Estimate, 3)),
                 # WITH ADH
                 `95% HDI (ADH)` = c(
                   # VLOAD
                   paste(round(exp(vload_adh_terms$HPDI_2.5), 3), 
                                    round(exp(vload_adh_terms$HPDI_97.5), 3), 
                                    sep = ", "),
                   # LEU3N
                   paste(round(leu3n_adh_terms$HPDI_2.5, 3), 
                                    round(leu3n_adh_terms$HPDI_97.5, 3), 
                                    sep = ", "),
                   # AGG PHYS
                   paste(round(aggphys_adh_terms$HPDI_2.5, 3), 
                                    round(aggphys_adh_terms$HPDI_97.5, 3), 
                                    sep = ", "),
                   # AGG MENT
                   paste(round(aggment_adh_terms$HPDI_2.5, 3), 
                                    round(aggment_adh_terms$HPDI_97.5, 3), 
                                    sep = ", ")),
                 check.names = FALSE
)


forest_df <- adh_comparison2 %>%
  # Create a longer version of df to plot
  pivot_longer(
    cols = c(Estimate_No_ADH, Estimate_ADH),
    names_to = "Model",
    values_to = "Estimate") %>%
  mutate(
    # Relabel the HDI cols
    HDI = case_when(
      Model == "Estimate_No_ADH" ~ `95% HDI (No ADH)`,
      Model == "Estimate_ADH"    ~ `95% HDI (ADH)`
    ),
    # Relabel the Model cols
    Model = case_when(
      Model == "Estimate_No_ADH" ~ "Without ADH",
      Model == "Estimate_ADH"    ~ "With ADH"
    ),
    # Use stringr package to get the lower bounds of the HIDs
    lower = as.numeric(str_trim(str_extract(HDI, "^[^,]+"))),
    # Get the upper bounds of the HDIs
    upper = as.numeric(str_trim(str_extract(HDI, "[^,]+$"))),
    # Group the outcomes for scaling
    OutcomeGroup = case_when(
  str_detect(Outcome, "CD4") ~ "CD4+ T-Cell Counts",
  TRUE                         ~ "Other outcomes"),
    Outcome = factor(Outcome, 
                     levels = rev(unique(adh_comparison2$Outcome)))) %>%
  select(Outcome, OutcomeGroup, Model, Estimate, lower, upper)

# Reference lines for plot - 0 for significance
ref_lines <- tibble(
  OutcomeGroup = c("CD4+ T-Cell Cou ts", 
                   "Viral Load (Ratio) & Quality of Life Scores"),
  ref = c(1, 0)
)

```

```{r, echo = F, warning = F, message = F, caption = "Figure 1 - Differences in estimates between models with (red) and without (blue) taking medication adherence into account. The points are the estimates and the bars are the HDIs. HDIs that cross the dashed line at 1 (for exponentiated viral load) or 0 (for the other three outcomes) indicate that the differences in the outcome between hard drug and non-hard drug users is statistically significant."}
# Plot
ggplot(forest_df, aes(x = Estimate, 
                      y = Outcome, 
                      # shape = Model,
                      col = Model)) +
  geom_vline(data = ref_lines,
             aes(xintercept = ref),
             linetype = "dashed",
             inherit.aes = FALSE) +
  geom_errorbarh(aes(xmin = lower, xmax = upper),
                 width = 0.1,
                 position = position_dodge(width = 0.5),
                 orientation = "y") +
  geom_point(size = 2,
             position = position_dodge(width = 0.5),
             show.legend = FALSE) +
  facet_wrap(~OutcomeGroup, 
             scales = "free_x",
             ncol = 1) +
  labs(
    x = "Estimate (95% HDI)",
    y = NULL,
    shape = NULL,
    title = "Effects of Baseline Hard Drug Use On Treatment Response at Year 2\nModels With vs Without ADH") +
  theme_minimal()
```

 

\vspace{-6mm}
# Discussion
\vspace{-6mm}



meow meow meow


