---
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    latex_engine: xelatex
mainfont: Arial
fontsize: 11pt
documentclass: article
geometry: margin=0.9in
---

```{r, echo = F, warning=F, message = F}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(knitr)
library(ggplot2)
library(MASS)
library(emmeans)
library(multcomp)
library(kableExtra)
library(stats)
library(dplyr)
library(tibble)
library(labelled)
library(tableone)
library(lme4)
library(lmerTest)
library(nlme)
library(gtsummary)
library(gt)
library(ggeffects)
library(UpSetR)
library(visdat)
library(lubridate)
library(see)
opts_chunk$set(tidy = F)
```

```{r, echo = F}
# Read in the data
proj0_dat <- read.csv("../RawData/Project0_Clean_v2.csv",
                      header = TRUE,
                      check.names = FALSE)
```

# Missingness

Below we ummarize and visuaize missingness in the dataset. 

The number and percentage of missing values for each variable: 

```{r, echo = F}
# Cumulative sum of missingness for each variable
naniar::miss_var_summary(proj0_dat)
```

Investigating the NAs in MEMs: Sample Interval Decimal Time, it appears that a majority of the NAs correspond to the first time point of Collect Sample = 1. We will replace these with 0.. 

```{r, echo = F}
proj0_dat %>% 
  group_by(SubjectID) %>% 
  reframe(SubjectID, 
          `Collection Sample`,
          `MEMs: Sample interval Decimal Time (mins)`)
```

when you when you

```{r, echo = F}
# How many variables 0 - 5 missing values
naniar::miss_case_table(proj0_dat)
```

Visualizing the missingness: 


```{r, echo = F}
vis_dat(proj0_dat)
```
\newpage

# Descriptive Statistics

A summary of the data - including ranges and data types: 

```{r, echo = F}
summary(proj0_dat)
```

```{r, echo = F}
## Make a subset of the data with the relevant cols for Q1 - Q3

# Vector of the cols of interest
cols <- c("SubjectID",
          "Collection Date", "DAYNUMB",
          "Collection Sample",
          "Booket: Clock Time", "MEMs: Clock Time",
          "Sleep Diary reported wake time",
          "Cortisol (nmol/L)", "DHEA (nmol/L)")

# Subset the original data frame
wake_data <- proj0_dat[, colnames(proj0_dat) %in% cols ]

```


```{r, echo = F}
# Convert the Booklet Time to minutes
wake_data$Booklet_Minutes <-
  # Convert to date objects and convert to minutes
  hour(hm(wake_data$`Booket: Clock Time`)) * 60 +
  minute(hm(wake_data$`Booket: Clock Time`))

# Move this to after Booklet: Clock Time
wake_data <- wake_data %>% 
  relocate(Booklet_Minutes, .after = `Booket: Clock Time`)

# Convert the MEM Time to minutes
wake_data$MEM_C <-
  # Convert to date objects and convert to minutes
  hour(hm(wake_data$`MEMs: Clock Time`)) * 60 +
  minute(hm(wake_data$`MEMs: Clock Time`))

# Move this to after MEMs: Clock Time
wake_data <- wake_data %>% 
  relocate(MEM_C, .after = `MEMs: Clock Time`)

```


## Booklet Times

The range of Booklet Times: 

```{r, echo = F}
summary(wake_data$Booklet_Minutes)
```
There are 35 NAs.


There are 35 NAs

```{r, echo = F}
# A plot of the distribution of the outcome
ggplot(na.omit(wake_data), aes(x = Booklet_Minutes,)) +
         geom_histogram(bins = 10, fill = "palevioletred2",
                        col = "black") + 
  theme_lucid() + 
  facet_grid(~ DAYNUMB)
``` 

A log transformation: 

```{r, echo = F}
# A plot of the distribution of the outcome
ggplot(na.omit(wake_data), aes(x = log(Booklet_Minutes))) +
         geom_histogram(bins = 10, fill = "palevioletred2",
                        col = "black") + 
  theme_lucid() + 
  facet_grid(~ DAYNUMB)
```


## MEM Times

A summary of the MEM times

```{r, echo = F}
# Summary of the MEM times
summary(wake_data$MEM_C)
```

We do see that There are 61 NAs.

```{r, echo = F}
ggplot(na.omit(wake_data), aes(x = MEM_C,)) +
         geom_histogram(bins = 10, fill = "seagreen3",
                        col = "black") + 
  theme_lucid() + 
  facet_grid(~ DAYNUMB)
```
Visualizing the outcome by day: 

```{r, echo = F}
ggplot(na.omit(wake_data), aes(x = log(MEM_C))) +
         geom_histogram(bins = 10, fill = "seagreen3",
                        col = "black") + 
  theme_lucid() + 
  facet_grid(~ DAYNUMB)
# Log or square root transformation does not appear to change much
```


The Booklet and MEM distributions appear to agree for the most part. 

```{r, echo = F}
# Create a subset of the data frame that only has Wake times
# AKA Collection Sample == 1
wakes <- wake_data[wake_data$`Collection Sample` == 1, ]
```


## Question 1

What is the agreement between the subject's recording of sampling times compared to the times recorded by an electronic monitoring cap?

For this question, we are only interested in the minutes since Waking Up.

```{r, echo = F, warning = F, message = F}
################################################################################
###                               Question 1                                 ### 
################################################################################

# Visualize the Wake Times from Booklet data
ggplot(wakes, aes(x = DAYNUMB,
                  y = Booklet_Minutes,
                  group = SubjectID,
                  color = factor(SubjectID))) +
  geom_line() + 
  theme_lucid() + 
  labs(title = "Wake Up Times Across Days - Booklet",
       x = "Day", y = "Booklet Time (Minutes)") + 
  scale_x_continuous(breaks = seq(1, 3, by = 1))

```



```{r, echo = F, message = F, warning = F}
# Visualize the Wake Times from MEMs data
ggplot(wakes, aes(x = DAYNUMB,
                  y = MEM_C,
                  group = factor(SubjectID),
                  color = factor(SubjectID))) +
  geom_line() + 
  theme_lucid() + 
  labs(title = "Wake Up Times Across Days - MEMs",
       x = "Day", y = "MEM Time (Minutes)") + 
  scale_x_continuous(breaks = seq(1, 3, by = 1))
```

These appear to be very similar but we can see some instances where the MEM failed to record the time. 

```{r, echo = F}
### Model
# Center MEM time for interpretation relative to Mean MEM time
wakes$MEM_C <- wakes$MEM_C - mean(wakes$MEM_Minutes, na.rm = TRUE)

# LMM with random intercept for subject - Day added
model_day <- lme(Booklet_Minutes ~ MEM_C + factor(DAYNUMB), 
                random = ~ 1 | SubjectID,
                na.action = na.omit,
                method = "ML",
                data = wakes)

# No day variable in the model
model_noday <- lme(Booklet_Minutes ~ MEM_C, 
                random = ~ 1 | SubjectID,
                na.action = na.omit,
                method = "ML",
                data = wakes)

# Random intercept and slope


# AIC
# > AIC(model_q1, model_noday)
#             df      AIC
# model_q1     6 495.4461
# model_noday  4 496.3718
# Day slightly better

# ANOVA
# > anova(model_q1, model_noday)
#             Model df      AIC      BIC    logLik   Test  L.Ratio p-value
# model_q1        1  6 495.4461 509.1061 -241.7230                        
# model_noday     2  4 496.3718 505.4785 -244.1859 1 vs 2 4.925757  0.085
# Day does not significanlty improve model fit

# summary(model_q1)

# Obtain 95% CI
# intervals(model_q1)
# (0.9694074, 1.007572)
```

### Diagnostics

```{r,  echo = F, fig.cap="Model Diagnostics", fig.show='hold', out.width="70%", fig.align='center'}
# Residual plot formodel
plot(model_noday, abline = 0)
# 
# QQ plot
qqnorm(resid(model_noday), main = NULL)
qqline(resid(model_noday))
# Adding DAYNUMB improved residual diagnostics
```


```{r,  echo = F, fig.cap="Model Diagnostics", fig.show='hold', out.width="70%", fig.align='center'}
# Residual plot formodel
plot(model_day, abline = 0)
# 
# QQ plot
qqnorm(resid(model_day), main = NULL)
qqline(resid(model_day))
# Adding DAYNUMB improved residual diagnostics
```
For the standardized residual plot, there does not appear to be a violation of heteroscedasticity. 

In the qqplot, we do see some slight deviations in the tails, but not anything concerning. 

Residuals another way: 

```{r, echo = F}
# Obtain data used to fit model
model_data <- getData(model_q1)

# Obtain residuals of model
model_data$resids <- resid(model_q1)
# Obtain fitted values
model_data$fitted <- fitted(model_q1)
```

```{r, echo = F}
# Residual histogram
ggplot(model_data, aes(x = resids)) +
  geom_histogram(bins = 25,
                 fill = "cadetblue2",
                 color = "black") +
  theme_lucid()

```
The residual hist looks symmetric enough given the small sample size. 

```{r, echo = F}
ggplot(model_data, aes(x = fitted, y = resids)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = 2) +
  theme_classic()

```
These look evenly scattered.

```{r, echo = F}
# LMM with random intercept for subject - Day added
model_q1 <- lme(Booklet_Minutes ~ MEM_C + factor(DAYNUMB), 
                random = ~ 1 | SubjectID,
                na.action = na.omit,
                method = "REML",
                data = wakes)

summary(model_q1)
```

Intercept: The average wake up time recorded by the MEM was ~829 minutes on the first day. 

Holding day constant, a one-minute increase in MEM-recorded wake time is associated with an approximately one-minute increase in Booklet-recorded wake time ($\beta$ ≈ 0.99; 95% CI: (0.97–1.01); p < 0.0001). This near one-to-one association indicates strong concordance between the two wake-time measures within a given day.
In other words, the subject's recorded wake up time and the MEM's recorded wake up time track each other quite closely. Changes in MEM time are mirrored almost exactly by changes in Booklet time.  

Assessing bias: 

```{r, echo = F}
fixef(model_q1)["(Intercept)"] -
mean(wakes$MEM_Minutes, na.rm = TRUE)

```
We have bias of around 2.21, indicating that there is small bias. The mean wake up time of MEM is 417.58 whereas the intercept representing the average MEM wake up time was 415.36.

```{r, echo = F}
# Look at the variance covariance mat
VarCorr(model_q1)

```

The variability is split between the between subject variability (random intercept) and the within-subject variability (residual). 

Intercept (between subject heterogeneity): After accounting for MEM time and day, subjects differ in their mean Booklet wake time with a sd of about 6.6 minutes. 

Residual variance (within subject heterogeneity): Day to day fluctuations within a subject. Given a subject, a day, and a MEM time, individual Booklet observations vary around the subject-specific mean with a sd of about 5 minutes. 

Scatterplot: 

```{r, echo = F}
ggplot(na.omit(wakes), aes(x = MEM_Minutes, y = Booklet_Minutes,
                           color = factor(DAYNUMB))) + 
  geom_point() + 
  theme_lucid() + 
  geom_abline(intercept = 0, slope = 1, linetype = "solid") +
  labs(color = "Day") +
  labs(title = "Relationship Between Booklet and MEM Wake Up Times",
       x = "MEM Wake Up Time (Minutes)",
       y = "Booklet Wake Up Time (Minutes)")
```

```{r, echo = F}
ggplot(na.omit(wakes), aes(x = MEM_Minutes, y = Booklet_Minutes,
                           color = factor(DAYNUMB))) + 
  geom_point() + 
  theme_lucid() + 
  labs(color = "Day") + 
  facet_grid(~ DAYNUMB) +
  labs(title = "Relationship Between Booklet and MEM Wake Up Times",
       subtitle = "Separated By Day",
       x = "MEM Wake Up Time (Minutes)",
       y = "Booklet Wake Up Time (Minutes)")
```

We can see that there appears to be a strong linear association. There are not any noticeable differences across days or obvious non-linear patterns.
The scatterplot shows a strong linear association between subject recorded time and electronic recording cap wake up time. 
Since each subject has one measure across days, we tested this association with a linear mixed effects model with a random intercept for subject and including fixed day effects. The estimated slope was approximately 1 and highly significant indicating that the two measurements follow each other very closely wihin day.


## Question 2

Are subject's adhering accurately to the +30 min (Collection Sample 2) and +10-hour (600min) (Collection Sample 4) sampling times required by a study protocol?

Good adherence - Percent of samples taken within 7.5 minutes of the scheduled time

Adequate adherence - Percent of samples taken within 15 minutes of scheduled time

```{r, echo = F}
################################################################################
###                               Question 2                                 ### 
################################################################################

###### With Booklet Times

##### Good adherence - within 7.5 minutes

### 30 minute Collection Sample 2
# Filter wake data to Collection Sample 2
sample2 <- wake_data[wake_data$`Collection Sample` == 2, ]

## Good adherence - within 7.5 minutes of scheduled time
library(data.table)

# Df with Wake time and 30 min later times
wake_check <- wake_data %>%
  # filter(`Collection Sample` %in% c(1, 2)) %>%
  group_by(SubjectID, DAYNUMB) %>%
  reframe(
    wake_time = Booklet_Minutes[`Collection Sample` == 1],
    S2_time = Booklet_Minutes[`Collection Sample` == 2],
    S4_time = Booklet_Minutes[`Collection Sample` == 4])

 
# Define adherence
wake_check <- wake_check %>%
  mutate(
    # Subjects should be recording 30 min after waking
    target_time = wake_time + 30,
    # Difference between 30 min later and the recorded time
    diff_min    = S2_time - target_time,
    # Define good adherence - within 7.5 minutes
    good_adherent = abs(diff_min) <= 7.5,
    # Define adequate adherence - within 15 minutes
    adequate_adherent = abs(diff_min) <= 15
  )

```

### Good Adherence 

```{r, echo = F}

# What is the proportion of adherence to 30 min later
mean(wake_check$adherent, na.rm = TRUE)
```

Adherence by subject: 

```{r, echo = F}
subject_good_adherence <- wake_check %>%
  group_by(SubjectID) %>%
  summarise(
    # How many days out of 3 were they adherent?
    days_good_adherent = sum(good_adherent),
    # What is the proportion
    prop_good_adherent = mean(good_adherent),
    .groups = "drop"
  )

subject_good_adherence

```

```{r, echo = F}
# How many subjects adhered all 3 days
sum(subject_good_adherence$prop_good_adherent == 1, na.rm = T)

# Proportion and Percentage of subjects
# 13/31 = 0.42 or 42%

```

```{r, echo = F}
# Visualize
ggplot(na.omit(wake_check), aes(x = diff_min)) +
  geom_histogram(binwidth = 2,
                 col = "black",
                 fill = "turquoise") +
  geom_vline(xintercept = c(-7.5, 7.5), linetype = "dashed",
             col = "red") +
  theme_lucid() +
  labs(
    x = "Minutes from target (+30)",
    y = "Count"
  ) + 
  scale_x_continuous(breaks = seq(0, 50, by = 10)) + 
  scale_y_continuous(breaks = seq(0, 60, by = 10),
                     limits = c(0, 60)) + 
  labs(title = "Distribution of 30 Minutes Post Wake Up Time")

```

### Adequate Adherence



## Question 3

### DHEA

A summary of DHEA: 

```{r, echo = F}
summary(proj0_dat$`DHEA (nmol/L)`)
```
How many values equal 5.205?

```{r, echo = F}
proj0_dat %>% 
  filter(`DHEA (nmol/L)` == 5.205) %>% 
  reframe(SubjectID, `Collection Sample`, 
          DAYNUMB,`DHEA (nmol/L)`)
```
There appears to be one subject that has 4 values of 5.205 for DHEA (nmol/L).

Visualizing DHEA over time: 

```{r, echo = F, message = F, warning=F}
ggplot(proj0_dat, aes(x = `Collection Sample`,
                      y = `DHEA (nmol/L)`,
                      color = factor(SubjectID))) +
  geom_point() + 
  geom_line() +
  facet_wrap(~DAYNUMB)
  # theme(legend.position = "none")
```

Visualizing a random subset of 5 patients: 

```{r, echo = F}
# Random sample of 5 subjects
```

A plot of the mean trajectories: 

```{r, echo = F}
# Add a column of the mean trajectory...
proj0_dat %>% 
  filter(!is.na(`DHEA (nmol/L)`)) %>% 
  group_by(SubjectID) %>% 
  summarise(mean_DHEA = mean(`DHEA (nmol/L)`, na.rm = T),
            .groups = "drop") + 
  
# Plot  
ggplot(proj0_dat,
    aes(x = `Collection Sample`, 
             y = `DHEA (nmol/L)`,
             color = factor(SubjectID))) + 
  geom_point() + 
  geom_line() + 
  facet_wrap(~DAYNUMB)
```


### Cortisol

A summary of Cortisol: 

```{r, echo = F}
summary(proj0_dat$`Cortisol (nmol/L)`)
```

How many measurements are above 26?

```{r, echo = F}
proj0_dat %>% 
  filter(`Cortisol (nmol/L)` >= 26) %>% 
  reframe(SubjectID, `Collection Sample`, 
          DAYNUMB,`Cortisol (nmol/L)`)
```


How many values equal exceed 80?

```{r, echo = F}
proj0_dat %>% 
  filter(`Cortisol (nmol/L)` >= 80) %>% 
  reframe(SubjectID, `Collection Sample`, 
          DAYNUMB,`Cortisol (nmol/L)`)
```
There appears to be one subject that has a value greater than 80. We wll remove this observation for this subject as this is most likely a lab error. 

Visualizing Cortisol over time: 

```{r, echo = F, message = F, warning=F}
ggplot(proj0_dat, aes(x = `Collection Sample`,
                      y = `Cortisol (nmol/L)`,
                      color = factor(SubjectID))) +
  geom_point() + 
  geom_line() +
  facet_wrap(~DAYNUMB)
  # theme(legend.position = "none")
```

Visualizing a random subset of 5 patients: 

```{r, echo = F}
# Random sample of 5 subjects
```


