---
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    latex_engine: xelatex
mainfont: Arial
fontsize: 11pt
documentclass: article
geometry: margin=0.9in
---

```{r, echo = F, warning=F, message = F}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(knitr)
library(ggplot2)
library(MASS)
library(emmeans)
library(multcomp)
library(kableExtra)
library(stats)
library(dplyr)
library(tibble)
library(labelled)
library(tableone)
library(lme4)
library(lmerTest)
library(nlme)
library(gtsummary)
library(gt)
library(ggeffects)
library(UpSetR)
library(visdat)
opts_chunk$set(tidy = F)
```

```{r, echo = F}
# Read in the data
proj0_dat <- read.csv("../RawData/Project0_Clean_v2.csv",
                      header = TRUE,
                      check.names = FALSE)
```

# Missingness

Below we ummarize and visuaize missingness in the dataset. 

The number and percentage of missing values for each variable: 

```{r, echo = F}
# Cumulative sum of missingness for each variable
naniar::miss_var_summary(proj0_dat)
```

Investigating the NAs in MEMs: Sample Interval Decimal Time, it appears that a majority of the NAs correspond to the first time point of Collect Sample = 1. We will replace these with 0.. 

```{r, echo = F}
proj0_dat %>% 
  group_by(SubjectID) %>% 
  reframe(SubjectID, 
          `Collection Sample`,
          `MEMs: Sample interval Decimal Time (mins)`)
```

when you when you

```{r, echo = F}
# How many variables 0 - 5 missing values
naniar::miss_case_table(proj0_dat)
```

Visualizing the missingness: 


```{r, echo = F}
vis_dat(proj0_dat)
```
\newpage

# Descriptive Statistics

A summary of the data - including ranges and data types: 

```{r, echo = F}
summary(proj0_dat)
```

## Question 1

## Question 2

## Question 3

### DHEA

A summary of DHEA: 

```{r, echo = F}
summary(proj0_dat$`DHEA (nmol/L)`)
```
How many values equal 5.205?

```{r, echo = F}
proj0_dat %>% 
  filter(`DHEA (nmol/L)` == 5.205) %>% 
  reframe(SubjectID, `Collection Sample`, 
          DAYNUMB,`DHEA (nmol/L)`)
```
There appears to be one subject that has 4 values of 5.205 for DHEA (nmol/L).

Visualizing DHEA over time: 

```{r, echo = F, message = F, warning=F}
ggplot(proj0_dat, aes(x = `Collection Sample`,
                      y = `DHEA (nmol/L)`,
                      color = factor(SubjectID))) +
  geom_point() + 
  geom_line() +
  facet_wrap(~DAYNUMB)
  # theme(legend.position = "none")
```

Visualizing a random subset of 5 patients: 

```{r, echo = F}
# Random sample of 5 subjects
```

A plot of the mean trajectories: 

```{r, echo = F}
# Add a column of the mean trajectory...
proj0_dat %>% 
  filter(!is.na(`DHEA (nmol/L)`)) %>% 
  group_by(SubjectID) %>% 
  summarise(mean_DHEA = mean(`DHEA (nmol/L)`, na.rm = T),
            .groups = "drop") + 
  
# Plot  
ggplot(proj0_dat,
    aes(x = `Collection Sample`, 
             y = `DHEA (nmol/L)`,
             color = factor(SubjectID))) + 
  geom_point() + 
  geom_line() + 
  facet_wrap(~DAYNUMB)
```


### Cortisol

A summary of Cortisol: 

```{r, echo = F}
summary(proj0_dat$`Cortisol (nmol/L)`)
```

How many measurements are above 26?

```{r, echo = F}
proj0_dat %>% 
  filter(`Cortisol (nmol/L)` >= 26) %>% 
  reframe(SubjectID, `Collection Sample`, 
          DAYNUMB,`Cortisol (nmol/L)`)
```


How many values equal exceed 80?

```{r, echo = F}
proj0_dat %>% 
  filter(`Cortisol (nmol/L)` >= 80) %>% 
  reframe(SubjectID, `Collection Sample`, 
          DAYNUMB,`Cortisol (nmol/L)`)
```
There appears to be one subject that has a value greater than 80. We wll remove this observation for this subject as this is most likely a lab error. 

Visualizing Cortisol over time: 

```{r, echo = F, message = F, warning=F}
ggplot(proj0_dat, aes(x = `Collection Sample`,
                      y = `Cortisol (nmol/L)`,
                      color = factor(SubjectID))) +
  geom_point() + 
  geom_line() +
  facet_wrap(~DAYNUMB)
  # theme(legend.position = "none")
```

Visualizing a random subset of 5 patients: 

```{r, echo = F}
# Random sample of 5 subjects
```


