---
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    latex_engine: xelatex
mainfont: Arial
fontsize: 12pt
documentclass: article
geometry: margin=1in
header-includes:
  - \usepackage{setspace}
  - \doublespacing
---

```{r, echo = F, warning=F, message = F}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(knitr)
library(ggplot2)
library(MASS)
library(emmeans)
library(multcomp)
library(kableExtra)
library(stats)
library(dplyr)
library(tibble)
library(labelled)
library(tableone)
library(lme4)
library(lmerTest)
library(nlme)
library(gtsummary)
library(gt)
library(ggeffects)
library(UpSetR)
library(visdat)
library(lubridate)
library(see)
library(data.table)
library(sjPlot)
opts_chunk$set(tidy = F)
```

```{r, echo = F, warning = F, message = F}
# Read in the data
proj0_dat <- read.csv("../RawData/Project0_Clean_v2.csv",
                      header = TRUE,
                      check.names = FALSE)


```


```{r, echo = F, warning = F, message = F}
########### Question 1 Data

## Make a subset of the data with the relevant cols for Q1 - Q3

# Vector of the cols of interest
cols <- c("SubjectID",
          "Collection Date", "DAYNUMB",
          "Collection Sample",
          "Booket: Clock Time", "MEMs: Clock Time",
          "Sleep Diary reported wake time",
          "Cortisol (nmol/L)", "DHEA (nmol/L)")

# Subset the original data frame
wake_data <- proj0_dat[, colnames(proj0_dat) %in% cols ]
 
# Convert the Sleep diary Wake time to minutes
wake_data$`Sleep Diary Wake Minutes` <-
  # Convert to date objects and convert to minutes
  lubridate::hour(hm(wake_data$`Sleep Diary reported wake time`)) * 60 +
  lubridate::minute(hm(wake_data$`Sleep Diary reported wake time`))

# Move this to after Collection Sample
wake_data <- wake_data %>% 
  relocate(`Sleep Diary Wake Minutes`, 
           .after = `Sleep Diary reported wake time`)

# Convert the Booklet Time to minutes
wake_data$Booklet_Minutes <-
  # Convert to date objects and convert to minutes
  lubridate::hour(hm(wake_data$`Booket: Clock Time`)) * 60 +
  lubridate::minute(hm(wake_data$`Booket: Clock Time`))

# Move this to after Booklet: Clock Time
wake_data <- wake_data %>% 
  relocate(Booklet_Minutes, .after = `Booket: Clock Time`)

# Convert the MEM Time to minutes
wake_data$MEM_Minutes <-
  # Convert to date objects and convert to minutes
  lubridate::hour(hm(wake_data$`MEMs: Clock Time`)) * 60 +
  lubridate::minute(hm(wake_data$`MEMs: Clock Time`))

# Move this to after MEMs: Clock Time
wake_data <- wake_data %>% 
  relocate(MEM_Minutes, .after = `MEMs: Clock Time`)

# Center MEM time
wake_data$MEM_C <-
  # Convert to date objects and convert to minutes
  lubridate::hour(hm(wake_data$`MEMs: Clock Time`)) * 60 +
  lubridate::minute(hm(wake_data$`MEMs: Clock Time`))

# Move this to after MEMs: Clock Time
wake_data <- wake_data %>% 
  relocate(MEM_C, .after = `MEMs: Clock Time`)

# Create a Minutes SINCE WAKE variable
wake_data <- wake_data %>%
  mutate(
    clock_min = lubridate::hour(hm(`Booket: Clock Time`)) * 60 +
                lubridate::minute(hm(`Booket: Clock Time`))) %>%
  group_by(SubjectID, DAYNUMB) %>%
  mutate(wake_min = clock_min[`Collection Sample` == 1][1],
         minutes_since_wake = clock_min - wake_min) %>%
  ungroup() %>% 
  relocate(minutes_since_wake, .after = `Sleep Diary Wake Minutes`)


### Modify data for Person 3029
wake_data <- wake_data %>% 
  mutate(DAYNUMB = case_when(
    SubjectID == 3029 & `Collection Date` == "10/7/2018" ~ 1,
    SubjectID == 3029 & `Collection Date` == "10/10/2018" ~ 2,
    SubjectID == 3029 & `Collection Date` == "10/12/2018" ~ 3,
    TRUE ~ DAYNUMB
  )) %>% 
  ungroup()

# Create a subset of the data frame that only has Wake times
# AKA Collection Sample == 1
wakes <- wake_data[wake_data$`Collection Sample` == 1, ]

```


# Introduction

Research suggests that there is a relationship between hormonal patterns across the span of a day and conditions including sleep disorders and depression.
In this study, the focus is on the hormones, DHEA and Cortisol, which play a role in regulating stress. 
Cortisol is an important steroid hormone that significantly affects metabolism which in turn affects the central nervous system. 
DHEA is 

In this study, researchers followed the cortisol and DHEA patterns over the course of a day for a totstudal y period of three days to measure the the patterns of the two hormones in a natural environment.  
Researchers utilized two methods to collect saliva samples; a Saliva Procurement and Integrated Testing (SPIT) booklet and an automatic Electronic Monitoring Cap. Saliva samples were collected at waking, 30 minutes after waking, before lunch, and 600 minutes after waking. Subjects did not have to adhere to a specific wake up time for 31 healthy control subjects. 

This report addresses the following hypotheses of interest: 

\textbf{Hypothesis 1: } 
There was satisfactory agreement between the subject's recording of sampling times compared to the recording times by the electronic monitoring cap. More specifically, there is a one-to-one relationship between the subject's recorded times and the booklet times. 

\textbf{Hypothesis 2: } 
Subjects adhered appropriately to the study protocol of the +30 minute and +600 minute sampling times. 

\textbf{Hypothesis 3: DHEA and Cortisol levels show an initial increase from waking to 30 minutes followed by a gradual decline from 30 minutes and on. } 

We will further refer to the electronic monitoring cap as MEM time and the the SPIT Booklet as Booklet time for the remainder of the report.

# Methods

## Preliminary Data Analysis

As missingness can present serious issues for the data analysis if not explored and addressed, we investigated missingness through calculating the number and percentage of missing values for each variable in the dataset. We further visualized missingness for each variable.  

Below is a Table1 of the primary variables of interest for the analysis:

For the variables we employed in our analyses, we assessed and visualized distributions to determine if there were any outliers or considerations that needed to be taken for data analysis. 

## Data Cleaning




## Data Analysis

\textbf{1:} To address research question 1, we employed a longitudinal linear mixed effects model with Booklet time in Minutes for the outcome, MEM in Minutes and the day number (1-3) as covariates We included a random intercept for subject that allowed us to take variability between subjects into account in our model. 
Adding Day as a covariate noticeably improved diagnostic plots for the assumptions of heteroscedasticity and normality. We additionally assessed bias by.... COME BACK TO THIS

\textbf{2: }

\textbf{2: }


# Results

From the missingness investigation, we determined the MEMs had the highest amount of missing: 177 missing values which is equivalent to 47.6%. 
Booklet had the second highest amount of missing with 55 missing values or 14.8%.
Cortisol and DHEA measurements had 5 missing values or 1.34% missingness.

The graphic below allows us to visually see the quantity of missingness in the MEM times compared to the other variables in the dataset: 

## Hypothesis 1

Below is a table of the estimates, 95% Confidence Intervals and P-values for the model: 

```{r, echo = F}
# LMM with random intercept for subject - Day added
model_q1 <- lme(Booklet_Minutes ~ MEM_C + factor(DAYNUMB), 
                random = ~ 1 | SubjectID,
                na.action = na.omit,
                method = "REML",
                data = wakes)


# Display model summary with tab_model
# Do not show random effects, ICC
tab_model(model_q1, show.re.var = FALSE, show.icc = FALSE,
          show.intercept = FALSE)

```

```{r, echo = F, eval = F}
intervals(model_q1)

#  Fixed effects:
#                       lower       est.      upper
# (Intercept)      -5.0891734  3.6513329 12.3918392
# MEM_C             0.9667782  0.9859424  1.0051067
# factor(DAYNUMB)2 -6.6349646 -3.4589031 -0.2828415
# factor(DAYNUMB)3 -5.0521491 -1.8810020  1.2901450
```

Holding day constant, a one-minute increase in MEM-recorded wake time is associated with an approximately one-minute increase in Booklet-recorded wake time ($\beta$ ≈ 0.99; 95% CI: (0.97–1.01); p < 0.0001). This near one-to-one association indicates strong concordance between the two wake-time measures within a given day.
In other words, the subject's recorded wake up time and the MEM's recorded wake up time track each other quite closely. Changes in MEM time are mirrored almost exactly by changes in Booklet time and our initial hypothesis is correct. 



# Discussion


