---
title: Assessing DHEA and Cortisol Diurnal Patterns
author: Diana Felipe
output: 
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    latex_engine: xelatex
    extra_dependencies: ["longtable", "float"]
mainfont: Arial
fontsize: 12pt
documentclass: article
geometry: margin=1in
header-includes:
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage{float}
  - \usepackage{caption}
  - \usepackage{titling}
  - \setlength{\droptitle}{-2cm}
  - \preauthor{\vspace{-0.6cm}}
  - \postauthor{\vspace{-3cm}}
---

```{r, echo = F, warning=F, message = F}
knitr::opts_chunk$set(echo = TRUE)


library(readr)
library(knitr)
library(ggplot2)
library(MASS)
library(emmeans)
library(multcomp)
library(kableExtra)
library(stats)
library(dplyr)
library(tibble)
library(labelled)
library(tableone)
library(lme4)
library(lmerTest)
library(nlme)
library(gtsummary)
library(gt)
library(ggeffects)
library(UpSetR)
library(visdat)
library(lubridate)
library(see)
library(data.table)
library(gridExtra)
library(kableExtra)
opts_chunk$set(tidy = F)
```

```{r, echo = F, warning = F, message = F}
### Read in the data
# The data is located in folder called RawData
proj0_dat <- read.csv("../RawData/Project0_Clean_v2.csv",
                      header = TRUE,
                      check.names = FALSE)

### Modify data for Person 3029
proj0_dat <- proj0_dat %>% 
  mutate(DAYNUMB = case_when(
    SubjectID == 3029 & `Collection Date` == "10/7/2018" ~ 1,
    SubjectID == 3029 & `Collection Date` == "10/10/2018" ~ 2,
    SubjectID == 3029 & `Collection Date` == "10/12/2018" ~ 3,
    TRUE ~ DAYNUMB
  )) %>% 
  ungroup()

```


```{r, echo = F, warning = F, message = F}
############################## Question 1 Data #################################

## Make a subset of the data with the relevant cols for Q1 - Q3

# Vector of the cols of interest
cols <- c("SubjectID",
          "Collection Date", "DAYNUMB",
          "Collection Sample",
          "Booket: Clock Time", "MEMs: Clock Time",
          "Sleep Diary reported wake time",
          "Cortisol (nmol/L)", "DHEA (nmol/L)")

# Subset the original data frame
wake_data <- proj0_dat[, colnames(proj0_dat) %in% cols ]
 
# Convert the Sleep diary Wake time to minutes
wake_data$`Sleep Diary Wake Minutes` <-
  # Convert to date objects and convert to minutes
  lubridate::hour(hm(wake_data$`Sleep Diary reported wake time`)) * 60 +
  lubridate::minute(hm(wake_data$`Sleep Diary reported wake time`))

# Move this to after Collection Sample
wake_data <- wake_data %>% 
  relocate(`Sleep Diary Wake Minutes`, 
           .after = `Sleep Diary reported wake time`)

# Convert the Booklet Time to minutes
wake_data$Booklet_Minutes <-
  # Convert to date objects and convert to minutes
  lubridate::hour(hm(wake_data$`Booket: Clock Time`)) * 60 +
  lubridate::minute(hm(wake_data$`Booket: Clock Time`))

# Move this to after Booklet: Clock Time
wake_data <- wake_data %>% 
  relocate(Booklet_Minutes, .after = `Booket: Clock Time`)

# Convert the MEM Time to minutes
wake_data$MEM_Minutes <-
  # Convert to date objects and convert to minutes
  lubridate::hour(hm(wake_data$`MEMs: Clock Time`)) * 60 +
  lubridate::minute(hm(wake_data$`MEMs: Clock Time`))

# Move this to after MEMs: Clock Time
wake_data <- wake_data %>% 
  relocate(MEM_Minutes, .after = `MEMs: Clock Time`)

# Center MEM time
wake_data$MEM_C <-
  # Convert to date objects and convert to minutes
  lubridate::hour(hm(wake_data$`MEMs: Clock Time`)) * 60 +
  lubridate::minute(hm(wake_data$`MEMs: Clock Time`))

# Move this to after MEMs: Clock Time
wake_data <- wake_data %>% 
  relocate(MEM_C, .after = `MEMs: Clock Time`)

# Create a Minutes SINCE WAKE variable
wake_data <- wake_data %>%
  mutate(
    clock_min = lubridate::hour(hm(`Booket: Clock Time`)) * 60 +
                lubridate::minute(hm(`Booket: Clock Time`))) %>%
  group_by(SubjectID, DAYNUMB) %>%
  mutate(wake_min = clock_min[`Collection Sample` == 1][1],
         minutes_since_wake = clock_min - wake_min) %>%
  ungroup() %>% 
  relocate(minutes_since_wake, .after = `Sleep Diary Wake Minutes`)


### Modify data for Person 3029
wake_data <- wake_data %>% 
  mutate(DAYNUMB = case_when(
    SubjectID == 3029 & `Collection Date` == "10/7/2018" ~ 1,
    SubjectID == 3029 & `Collection Date` == "10/10/2018" ~ 2,
    SubjectID == 3029 & `Collection Date` == "10/12/2018" ~ 3,
    TRUE ~ DAYNUMB
  )) %>% 
  ungroup()

# Create a subset of the data frame that only has Wake times
# AKA Collection Sample == 1
wakes <- wake_data[wake_data$`Collection Sample` == 1, ]

```

# Introduction

Research suggests that there is a relationship between hormonal patterns across the span of a day and conditions including sleep disorders and depression. In this study, the focus is on the hormones DHEA and cortisol, which play an important role in regulating the body’s stress response. Cortisol is an important steroid hormone that significantly affects metabolism, which in turn influences the central nervous system. Cortisol typically follows a diurnal pattern, with levels peaking shortly after waking and then gradually declining throughout the day.
DHEA is another steroid hormone involved in the stress response and is often considered to have effects that counterbalance cortisol. DHEA has been associated with immune function, mood regulation, and overall physiological resilience, and its daily pattern is less well characterized compared to cortisol. Because both hormones are involved in stress regulation, understanding their patterns across the day may provide insight into physiological processes related to health and disease.

In this study, researchers examined cortisol and DHEA patterns over the course of a day across a total study period of three days. Saliva samples were collected using two methods: a Saliva Procurement and Integrated Testing (SPIT) booklet and an automatic electronic monitoring cap. Samples were collected at waking, 30 minutes after waking, before lunch, and 600 minutes after waking. Subjects were allowed to follow their normal daily schedules and were not required to adhere to a specific wake-up time. A total of 31 healthy control subjects participated in the study. 

This report addresses the following hypotheses of interest: 

\textbf{Hypothesis 1: } 
There was satisfactory agreement between the subject's recording of sampling times compared to the recording times by the electronic monitoring cap. More specifically, there is a one-to-one relationship between the subject's recorded times and the booklet times. 

\textbf{Hypothesis 2: } 
Subjects adhered appropriately to the study protocol of the +30 minute and +600 minute sampling times. 

\textbf{Hypothesis 3: } 
DHEA and Cortisol levels show an initial increase from waking to 30 minutes followed by a gradual decline from 30 minutes and on. 

\textit{Note: We will further refer to the electronic monitoring cap as MEM time and the the SPIT Booklet as Booklet time for the remainder of the report.}

\vspace{-6mm}
# Methods
\vspace{-6mm}

## Preliminary Data Analysis

Below is a Table1 of the primary variables of interest for the analysis. The minutes are all minutes relative to midnight.


```{r, echo = F, warning = F, message = F}

# Adjust labels for collection sample
data <- wake_data %>%
  mutate(
    `Collection Sample` = factor(
      `Collection Sample`,
      levels = c(1, 2, 3, 4),
      labels = c("Wake Time", "+30 Min", "Pre Lunch", "+600 Min")
    )
  )

# Create subject level summaries - NOT observation level
data <- data %>%
  group_by(SubjectID, `Collection Sample`) %>%
  summarise(
    mean_booklet   = mean(Booklet_Minutes, na.rm = TRUE),
    mean_mem = mean(MEM_Minutes, na.rm = TRUE),
    mean_wake = mean(minutes_since_wake, na.rm = TRUE),
    mean_dhea = mean(`DHEA (nmol/L)`, na.rm = TRUE),
    mean_cort = mean(`Cortisol (nmol/L)`, na.rm = TRUE),
    .groups = "drop")

# Define variables for the table one
variables <- c("mean_booklet", "mean_mem", 
               "mean_wake", "mean_dhea", "mean_cort")

# Separate by collection time
strata <- "Collection Sample"

# modify labels
var_label(data$mean_booklet) <- 'Booklet Minutes'
var_label(data$mean_mem) <- 'MEM Minutes'
var_label(data$mean_wake) <- 'Minutes Since Wake' 
var_label(data$mean_dhea) <- 'DHEA (nmol/L)'
var_label(data$mean_cort) <- 'Cortisol (nmol/L)'

# Create the table one
studypop_tab1 <- CreateTableOne(
  vars = variables,
  strata = strata,
  test = FALSE,
  data = data
)

studypop_tab_forkbl <- print(
  studypop_tab1,
  varLabels = TRUE,
  printToggle = FALSE
)

# Display with kbl
kbl(
  studypop_tab_forkbl,
  caption = "Summary Per Subject Within Collection Sample",
  booktabs = TRUE,
  align = "c"
) %>%
  kable_styling(latex_options = "hold_position", font_size = 10)

```


```{r, echo = F, eval = F, message=F, warning=F}
# Convert some of these back to hh:mm to talk about it
# Create function to convert minutes to hh:mm
min_to_clock <- function(x) {
  h <- floor(x / 60)
  m <- round(x %% 60)
  sprintf("%02d:%02d", h, m)
}


### Get the means
# Mean minutes since wake
mean_minwake <- mean(data$`Sleep Diary Wake Minutes`, 
                     na.rm = TRUE)
# min_to_clock(mean_minwake) 
# 06:52

# Mean booklet minutes
mean_bookmin <- mean(data$Booklet_Minutes, 
                     na.rm = TRUE) # 654.3472
# min_to_clock(mean_bookmin) 
# 10:54

# Mean MEM minutes
mean_memmin <- mean(data$MEM_Minutes, 
                     na.rm = TRUE) # 684.2379
# min_to_clock(mean_memmin) 
# 11:24
```
For analysis, we exclusively worked with the time variables in minutes. 
Looking at Minutes Since Waking, we can see that the mean minutes closely follow the study protocol times. The average DHEA and Cortisol levels appear to be similar across the 4 time points throughout the day.

As missingness can present serious issues for the data analysis if not addressed, we investigated missingness through calculating the number and percentage of missing values for each variable in the dataset. We further visualized missingness for each variable.  

For the variables we employed in our analyses, we assessed and visualized distributions to determine if there were any outliers or considerations that needed to be taken for data analysis. 

\vspace{-6mm}
## Data Cleaning
\vspace{-6mm}

For DHEA, we removed observations that were $\geq 5.205$ (nmol/L) as these are considered extremely high DHEA levels that could interfere with analysis. There were 6 observations (4 of which corresponded to Subject 3037) that had DHEA levels of 5.205 (nmol/L). We removed all observations for Subject 3037 due to multiple high DHEA values that suggest underlying health problems as well as the 2 other observations with DHEA of 5.205 (nmol/L).
For cortisol, we observed 5 observations from 4 patients that had Cortisol levels $\geq 26$. Although these values are considered high, they are biologically plausible and they were left in the analysis. One observation had a cortisol level of ~89 which was removed for the analysis. 

\vspace{-6mm}
## Data Analysis
\vspace{-6mm}

To address the first hypothesis, we employed a longitudinal linear mixed effects model with Booklet Time in Minutes as the outcome with covariates - MEM in Minutes and the Day Number (1-3). Adding Day as a covariate noticeably improved diagnostic plots for the assumptions of heteroscedasticity and normality.
We included a random intercept for subject that allowed us to take variability between subjects into account in our model. 
To assess whether booklet times tended to be recorded earlier or later than MEM time, we used the difference between MEM and Booklet times as the outcome in a longitudinal linear mixed effects model. This model had Day Number as a covariate and a random intercept for subject. 

To determine whether subjects were adhering accurately to the +30 minute and +600 minute sampling times required by study protocol, intervals of time were defined to calculate percentages for which there was good or adequate adherence.
Good adherence was defined by a collection of the saliva sample within a 7.5 minute window interval. Patients had adequate adherence if samples were within a 15 minute window. 
To calculate the percentage of individuals that had good adherence and adequate adherence, the difference between the "correct" sampling time and the observed time was calculated and this difference was used to assess whether observed times were within the good or adequate adherence windows. 
We visualized and summarized the distribution of the MEM and Booklet times to guide us in our decision to proceed with either MEM or Booklet times for our calculations.
We observed that MEM times had a considerable amount of missingness and determined that reporting both would give a better overall sense of adherence.
We summarized adherence for the study population with tables.

For the third hypothesis, we used a piecewise longitudinal linear mixed effects model to assess whether DHEA and Cortisol levels followed the expected diurnal pattern.
Due to a right-skew in the distribution of DHEA, a log transformed DHEA variable was used as the outcome to assess DHEA changes over the day. 
No transformations was made to Cortisol. 
Covariates for both models included manually calculated pre and post 30 minute variables (variables for piecewise slopes) which would give us the slopes for the two time intervals of interest. 

All analyses were conducted in R version 4.4.1 (2024-06-14), using the nlme  package for modeling.

\vspace{-9mm}

# Results
\vspace{-6mm}

From the missingness investigation, we determined the MEMs had the highest amount of missing: 177 missing values which is equivalent to 47.6%. 
Booklet had the second highest amount of missingness with 55 missing values or 14.8%.
Cortisol and DHEA measurements had 5 missing values each or 1.34% missingness.

<!-- The graphic below allowed us to visually see the quantity of missingness in the MEM times compared to the other variables in the data set, we followed certain thresholds for good and adequate adherence. -->

\vspace{-6mm}
## Hypothesis 1
\vspace{-6mm}

Below is a table of the estimates, 95% Confidence Intervals and P-values: 

```{r, echo = F, warning = F, message = F}
# LMM with random intercept for subject - Day added
model_q1 <- lme(Booklet_Minutes ~ MEM_C + factor(DAYNUMB), 
                random = ~ 1 | SubjectID,
                na.action = na.omit,
                method = "REML",
                data = wakes)


# Display model summary with kable
# Extract t and p-values from model for data frame
summary_df1 <- as.data.frame(summary(model_q1)$tTable)

# kable
kable(
  summary_df1,
  format = "latex",
  booktabs = TRUE,
  digits = 3,
  caption = "Model Summary"
) |>
  kable_styling(latex_options = c("striped", "hold_position"))


```

```{r, echo = F, eval = F}
# intervals(model_q1)

#  Fixed effects:
#                       lower       est.      upper
# (Intercept)      -5.0891734  3.6513329 12.3918392
# MEM_C             0.9667782  0.9859424  1.0051067
# factor(DAYNUMB)2 -6.6349646 -3.4589031 -0.2828415
# factor(DAYNUMB)3 -5.0521491 -1.8810020  1.2901450
```

```{r, echo = F, message = F, warning = F}
######## Assess bias
# Create a variable that is the difference between MEM and Booklet times
wake_data <- wake_data %>%
  mutate(diff_minutes = Booklet_Minutes - MEM_Minutes)

# Model to assess bias were difference is the outcome
bias_model <- lme(
  diff_minutes ~ 1 + factor(DAYNUMB),
  random = ~ 1 | SubjectID,
  data = wake_data,
  na.action = na.omit,
  method = "REML"
)

# summary(bias_model)
# intervals(bias_model)
```


Holding day constant, a one-minute increase in MEM-recorded wake time is associated with an approximately one-minute increase in Booklet-recorded wake time ($\beta$ ≈ 0.99; 95% CI: (0.97–1.01); p < 0.0001). This near one-to-one association indicates strong concordance between the two wake-time measures within a given day.
In other words, the subject's recorded wake up time and the MEM's recorded wake up time track each other quite closely. This is visualized in Figure 1 below.
Changes in MEM time are mirrored almost exactly by changes in Booklet time and our initial hypothesis is correct. 
In regard to bias, booklet times tended to be recorded approximately 7-8 minutes earlier (95% CI: (-14.28, -0.62)) than MEM times on average. This difference is statistically significant with p-val $\approx$ 3.


```{r, echo = F, warning=F, message = F, fig.cap="Booklet Vs. MEM Wake Up Times", out.width="80%", fig.align='center'}
ggplot(wakes, aes(x = MEM_Minutes, y = Booklet_Minutes,
                           color = factor(DAYNUMB))) + 
  geom_point(size = 3, alpha = 0.7) + 
  theme_lucid() + 
  geom_abline(intercept = 0, slope = 1, linetype = "solid") +
  labs(color = "Day") +
  labs(title = "Relationship Between Booklet and MEM Wake Up Times",
       x = "MEM Wake Up Time (Minutes)",
       y = "Booklet Wake Up Time (Minutes)")
```

\vspace{-6mm}
## Hypothesis 2
\vspace{-6mm}

```{r, echo = F}
################################################################################
###                               Question 2                                 ### 
################################################################################


# Df with Wake time, 30 min post wake, and 600 min post wake
wake_check <- wake_data %>%
  # filter(`Collection Sample` %in% c(1, 2)) %>%
  group_by(SubjectID, DAYNUMB) %>%
  reframe(
    # Wake time
    wake_time = `Sleep Diary Wake Minutes`[`Collection Sample` == 1],
    # The 30 minute time with Booklet
    Book_30time = Booklet_Minutes[`Collection Sample` == 2],
    # The 600 miniute time with booklet
    Book_600time = Booklet_Minutes[`Collection Sample` == 4],
    # The 30 minute time with MEM
    MEM_30time = MEM_Minutes[`Collection Sample` == 2],
    # The 600 minute time with MEM
    MEM_600time = MEM_Minutes[`Collection Sample` == 4]
    )

 
# Define adherence for BOOK
book_wake_check <- wake_check %>%
  mutate(
    # 30 min after waking target time
    target_30time = wake_time + 30,
    # 600 min after waking target time
    target_600time = wake_time + 600,
    # Difference between 30 min later and the recorded time
    diff_min30  = Book_30time - target_30time,
    # Difference between 600 min later and the recorded time
    diff_min600 = Book_600time - target_600time,
    # For 30 min: Define good adherence - within 7.5 minutes
    good_adherent30 = abs(diff_min30) <= 7.5,
    # For 600 min: Define good adherence
    good_adherent600 = abs(diff_min600) <= 7.5,
    # For 30 min: Define adequate adherence - within 15 minutes
    adequate_ad30 = abs(diff_min30) <= 15,
    # For 600 min: Define adequate adherence
    adequate_ad600 = abs(diff_min600) <= 15
  ) %>% 
  relocate(c(target_30time, diff_min30,
             good_adherent30, adequate_ad30), 
           .after = Book_30time) %>% 
  relocate(c(target_600time, diff_min600,
             good_adherent600, adequate_ad600), 
           .after = Book_600time)

# Define adherence for MEM
MEM_wake_check <- wake_check %>%
  mutate(
    # 30 min after waking target time
    target_30time = wake_time + 30,
    # 600 min after waking target time
    target_600time = wake_time + 600,
    # Difference between 30 min later and the recorded time
    diff_min30  = MEM_30time - target_30time,
    # Difference between 600 min later and the recorded time
    diff_min600 = MEM_600time - target_600time,
    # For 30 min: Define good adherence - within 7.5 minutes
    good_adherent30 = abs(diff_min30) <= 7.5,
    # For 600 min: Define good adherence
    good_adherent600 = abs(diff_min600) <= 7.5,
    # For 30 min: Define adequate adherence - within 15 minutes
    adequate_ad30 = abs(diff_min30) <= 15,
    # For 600 min: Define adequate adherence
    adequate_ad600 = abs(diff_min600) <= 15
  ) %>% 
  relocate(c(target_30time, diff_min30,
             good_adherent30, adequate_ad30), 
           .after = MEM_30time) %>% 
  relocate(c(target_600time, diff_min600,
             good_adherent600, adequate_ad600), 
           .after = MEM_600time)

```

```{r, echo = F}
################################## Good Adherence ##############################

############################## Adherence by Subject ############################

################################### +30 Minutes ################################

######## Booklet
subject_good_30adherence <- book_wake_check %>%
  group_by(SubjectID) %>%
  summarise(
    # How many days out of 3 were they adherent?
    days_good_adherent = sum(good_adherent30),
    # What is the proportion
    prop_good_adherent = mean(good_adherent30),
    .groups = "drop"
  )
# subject_good_30adherence

# How many subjects adhered all 3 days
# sum(subject_good_30adherence$prop_good_adherent == 1, na.rm = T)
# Proportion and Percentage of subjects
# 14/31 = 0.45 or 45%

######## MEM
MEM_subject_good_30adherence <- MEM_wake_check %>%
  group_by(SubjectID) %>%
  summarise(
    # How many days out of 3 were they adherent?
    days_good_adherent = sum(good_adherent30),
    # What is the proportion
    prop_good_adherent = mean(good_adherent30),
    .groups = "drop"
  )
# MEM_subject_good_30adherence

# How many subjects adhered all 3 days
# sum(MEM_subject_good_30adherence$prop_good_adherent == 1, na.rm = T)
# Proportion and Percentage of subjects
# 6/31 




################################# Adequate Adherence ###########################

############################## Adherence by Subject ############################

################################### +30 Minutes ################################

###### Book

subject_ad_30adherence <- book_wake_check %>%
  group_by(SubjectID) %>%
  summarise(
    # How many days out of 3 were they adherent?
    days_ad_adherent = sum(adequate_ad30),
    # What is the proportion
    prop_ad_adherent = mean(adequate_ad30),
    .groups = "drop"
  )

# subject_ad_30adherence

# How many subjects adhered all 3 days
# sum(subject_ad_30adherence$prop_ad_adherent == 1, na.rm = T)
# Proportion and Percentage of subjects
# 20/31 = 0.65 or 65%

###### MEM
MEM_subject_ad_30adherence <- MEM_wake_check %>%
  group_by(SubjectID) %>%
  summarise(
    # How many days out of 3 were they adherent?
    days_ad_adherent = sum(adequate_ad30),
    # What is the proportion
    prop_ad_adherent = mean(adequate_ad30),
    .groups = "drop"
  )

# MEM_subject_ad_30adherence

# How many subjects adhered all 3 days
# sum(MEM_subject_ad_30adherence$prop_ad_adherent == 1, na.rm = T)
# Proportion and Percentage of subjects
# 10/31 = 0.32 or 32%


```

Below we summarize the percentage of subjects that adhered to the study sampling times, for both MEM and Booklet times. 

```{r, echo = F}
################################## Good Adherence ##############################

################################ Adherence Overall #############################

################################### +30 Minutes ################################

######### Book
# Adherence overall
good_book <- book_wake_check %>%
  # group_by(SubjectID) %>%
  summarise(
    # What is the proportion
    prop_good_adherent = round(mean(good_adherent30, na.rm = TRUE), 4),
    # Percentage
    per_good_adherence = (prop_good_adherent * 100)
  )
# good_book
# A tibble: 1 × 2
#   prop_good_adherent per_good_adherence
#                <dbl>              <dbl>
# 1              0.782               78.2


######### MEM
# Adherence overall
good_MEM <- MEM_wake_check %>%
  # group_by(SubjectID) %>%
  summarise(
    # What is the proportion
    prop_good_adherent = round(mean(good_adherent30, na.rm = TRUE), 4),
    # Percentage
    per_good_adherence = (prop_good_adherent * 100)
  )
# good_MEM
# A tibble: 1 × 2
#   prop_good_adherent per_good_adherence
#                <dbl>              <dbl>
# 1              0.529               52.9

################################# Adequate Adherence ###########################

################################ Adherence Overall #############################

################################### +30 Minutes ################################

######## Book

# Adherence overall
adequate_book <- book_wake_check %>%
  summarise(
    # What is the proportion
    prop_adeq_adherent = round(mean(adequate_ad30, na.rm = TRUE), 4),
    # Percentage
    per_adeq_adherence = (prop_adeq_adherent * 100)
  )
# adequate_book
#   prop_adeq_adherent per_adeq_adherence
#                <dbl>              <dbl>
# 1              0.897               89.7

######## MEM
# Adherence overall
adequate_MEM <- MEM_wake_check %>%
  summarise(
    # What is the proportion
    prop_adeq_adherent = round(mean(adequate_ad30, na.rm = TRUE), 4),
    # Percentage
    per_adeq_adherence = (prop_adeq_adherent * 100)
  )
# adequate_MEM
#   prop_adeq_adherent per_adeq_adherence
#                <dbl>              <dbl>
# 1              0.714               71.4

###### Create Data frame to display with this
# Combine all the proportions into a table to display nicely
props <- data.frame(Good = c(good_book$per_good_adherence, 
                             good_MEM$per_good_adherence),
                    Adequate = c(adequate_book$per_adeq_adherence,
                                 adequate_MEM$per_adeq_adherence))

# Add rownames
rownames(props) <- c("Booklet", "MEM")

```


```{r, echo = F}
################################## Good Adherence ##############################

################################ Adherence Overall #############################

#################################### +600 Minutes ##############################

##### Book
# Adherence overall
good_book600 <- book_wake_check %>%
  # group_by(SubjectID) %>%
  summarise(
    # What is the proportion
    prop_good_adherent = round(mean(good_adherent600, na.rm = TRUE), 4),
    # Percentage
    per_good_adherence = (prop_good_adherent * 100)
  )
#   prop_good_adherent per_good_adherence
#                <dbl>              <dbl>
# 1              0.462               46.2

##### MEM
# Adherence overall
good_MEM600 <- MEM_wake_check %>%
  # group_by(SubjectID) %>%
  summarise(
    # What is the proportion
    prop_good_adherent = round(mean(good_adherent600, na.rm = TRUE), 4),
    # Percentage
    per_good_adherence = (prop_good_adherent * 100)
  )
#   prop_good_adherent per_good_adherence
#                <dbl>              <dbl>
# 1              0.325               32.5

############################### Adequate Adherence #############################

################################ Adherence Overall #############################

#################################### +600 Minutes ##############################

# Adherence overall
adequate_book600 <- book_wake_check %>%
  summarise(
    # What is the proportion
    prop_adeq_adherent = round(mean(adequate_ad600, na.rm = TRUE), 4),
    # Percentage
    per_adeq_adherence = (prop_adeq_adherent * 100)
  )
#   prop_adeq_adherent per_adeq_adherence
#                <dbl>              <dbl>
# 1              0.562               56.2

# Adherence overall
adequate_MEM600 <- MEM_wake_check %>%
  summarise(
    # What is the proportion
    prop_adeq_adherent = round(mean(adequate_ad600, na.rm = TRUE), 4),
    # Percentage
    per_adeq_adherence = (prop_adeq_adherent * 100)
  )

#   prop_adeq_adherent per_adeq_adherence
#                <dbl>              <dbl>
# 1              0.398               39.8
  
###### Create Data frame to display with this
# Combine all the proportions into a table to display nicely
props600 <- data.frame(Good = c(good_book600$per_good_adherence, 
                             good_MEM600$per_good_adherence),
                    Adequate = c(adequate_book600$per_adeq_adherence,
                                 adequate_MEM600$per_adeq_adherence))

# Add rownames
rownames(props600) <- c("Booklet", "MEM")
  

```


```{r results='asis', echo=FALSE}
# Combine information in a more cohesive way

props30_long <- props %>%
  as.data.frame() %>%
  rownames_to_column("Method") %>%
  mutate(Window = "+30 minutes")

props60_long <- props600 %>%
  as.data.frame() %>%
  rownames_to_column("Method") %>%
  mutate(Window = "+60 minutes")

final_props <- bind_rows(props30_long, props60_long) %>%
  relocate(Window, Method)

kable(
  final_props,
  format = "latex",
  booktabs = TRUE,
  digits = 2,
  caption = "Percentages for Good and Adequate Adherence by Collection Window"
) |>
  kable_styling(latex_options = c("striped", "hold_position")) |>
  collapse_rows(columns = 1, latex_hline = "major")


```


Overall, more subjects are meeting the adequate adherence thresholds compared to the good adherence thresholds for both the +30 minute and +600 minute sampling times. In addition, the MEM percentages are lower for both sampling times and for both good and adequaate thresholds.

In Figure 2 we visualize the minutes from the +30 minute sampling time (left column) and the +600 minute sampling time (right column). 
Times for Booklet are in the top row and Times for MEM are in the bottom row. 
The red dashed lines indicate the $\pm$ 7.5 window for good adherence while the pink dashed lines are the $\pm$ 15 minute window for adequate adherence. 


```{r, echo = F, warning = F, message=F, fig.align='center', fig.cap="Adherence for +30 and +600 Sampling Times", fig.pos="H"}

#################################### Visualize #################################

# Visualize the distribution of minutes from the sampling times

# Book +30 min
p1 <- ggplot(book_wake_check, aes(x = diff_min30)) +
  geom_histogram(binwidth = 3, col = "black", fill = "turquoise") +
  geom_vline(xintercept = c(-7.5, 7.5), col = "red",
             linetype = "dashed") +
  geom_vline(xintercept = c(-15, 15), color = "hotpink",
             linetype = "dashed") +
  theme_lucid() + 
  scale_y_continuous(breaks = seq(0, 40, by = 10)) + 
  labs(title = "Minutes from +30 Minute Sampling Time",
       subtitle = "Booklet",
       x = "Minutes from target (+30)",
       y = "Count") 

# MEM + 30 min
p2 <- ggplot(MEM_wake_check, aes(x = diff_min30)) +
  geom_histogram(binwidth = 10, col = "black", fill = "skyblue") +
  geom_vline(xintercept = c(-7.5, 7.5), col = "red",
             linetype = "dashed") +
  geom_vline(xintercept = c(-15, 15), color = "hotpink",
             linetype = "dashed") +
  theme_lucid() + 
  scale_y_continuous(breaks = seq(0, 40, by = 10),
                     limits = c(0, 40)) +
  labs(title = "Minutes from +30 Minute Sampling Time",
       subtitle = "MEM",
       x = "Minutes from target (+30)",
       y = "Count") 

# MEM +30 Min
p3 <- ggplot(book_wake_check, aes(x = diff_min600)) +
  geom_histogram(binwidth = 20, col = "black", fill = "turquoise") +
  geom_vline(xintercept = c(-7.5, 7.5), col = "red",
             linetype = "dashed") +
  geom_vline(xintercept = c(-15, 15), color = "hotpink",
             linetype = "dashed") +
  theme_lucid() + 
  scale_x_continuous(breaks = seq(-50, 360, by = 50)) + 
  scale_y_continuous(breaks = seq(0, 40, by = 10),
                     limits = c(0, 40)) +
  labs(title = "Minutes from +600 Sampling Time",
       subtitle = "Booklet",
       x = "Minutes from target (+600)",
       y = "Count") 

# MEM +600 min
p4 <- ggplot(MEM_wake_check, aes(x = diff_min600)) +
  geom_histogram(binwidth = 20, col = "black", fill = "skyblue") +
  geom_vline(xintercept = c(-7.5, 7.5), col = "red",
             linetype = "dashed") +
  geom_vline(xintercept = c(-15, 15), color = "hotpink",
             linetype = "dashed") +
  theme_lucid() + 
  scale_x_continuous(breaks = seq(-50, 360, by = 50)) + 
  scale_y_continuous(breaks = seq(0, 40, by = 10),
                     limits = c(0, 40)) +
  labs(title = "Minutes from +600 Minute Sampling Time",
       subtitle = "MEM",
       x = "Minutes from target (+600)",
       y = "Count") 


# Arrange 
grid.arrange(p1, p3, p2, p4, nrow = 2)

```

## Hypothesis 3

### DHEA

```{r, echo = F, warning = F, message = F}

################################################################################
###                               Question 3                                 ### 
################################################################################

##### Create data frame for question 3
## Make a subset of the data with the relevant cols for Q3

# Vector of the cols of interest
cols <- c("SubjectID",
          "Collection Date", "DAYNUMB",
          "Collection Sample",
          "Booket: Clock Time", "MEMs: Clock Time",
          "Sleep Diary reported wake time",
          "Cortisol (nmol/L)", "DHEA (nmol/L)")

# Subset the original data frame
wake_data_q3 <- proj0_dat[, colnames(proj0_dat) %in% cols ]



### Modify data for Person 3029
wake_data_q3 <- wake_data_q3 %>% 
  mutate(DAYNUMB = case_when(
    SubjectID == 3029 & `Collection Date` == "10/7/2018" ~ 1,
    SubjectID == 3029 & `Collection Date` == "10/10/2018" ~ 2,
    SubjectID == 3029 & `Collection Date` == "10/12/2018" ~ 3,
    TRUE ~ DAYNUMB
  )) %>% 
  ungroup()

# Convert the Sleep diary Wake time to minutes
wake_data_q3$`Sleep Diary Wake Minutes` <-
  # Convert to date objects and convert to minutes
  lubridate::hour(hm(wake_data_q3$`Sleep Diary reported wake time`)) * 60 +
  lubridate::minute(hm(wake_data_q3$`Sleep Diary reported wake time`))

# Move this to after Collection Sample
wake_data_q3 <- wake_data_q3 %>% 
  relocate(`Sleep Diary Wake Minutes`, 
           .after = `Sleep Diary reported wake time`)

# Convert the Booklet Time to minutes
wake_data_q3$Booklet_Minutes <-
  # Convert to date objects and convert to minutes
  lubridate::hour(hm(wake_data_q3$`Booket: Clock Time`)) * 60 +
  lubridate::minute(hm(wake_data_q3$`Booket: Clock Time`))

# Move this to after Booklet: Clock Time
wake_data_q3 <- wake_data_q3 %>% 
  relocate(Booklet_Minutes, .after = `Booket: Clock Time`)

# Convert the MEM Time to minutes
wake_data_q3$MEM_Minutes <-
  # Convert to date objects and convert to minutes
  lubridate::hour(hm(wake_data_q3$`MEMs: Clock Time`)) * 60 +
  lubridate::minute(hm(wake_data_q3$`MEMs: Clock Time`))

# Move this to after MEMs: Clock Time
wake_data_q3 <- wake_data_q3 %>% 
  relocate(MEM_Minutes, .after = `MEMs: Clock Time`)

# Create a Minutes SINCE WAKE variable
wake_data_q3 <- wake_data_q3 %>%
  group_by(SubjectID, DAYNUMB) %>%
  mutate(wake_min = Booklet_Minutes[`Collection Sample` == 1][1],
         minutes_since_wake = Booklet_Minutes - wake_min) %>%
  ungroup() %>% 
  relocate(minutes_since_wake, .after = `Sleep Diary Wake Minutes`)

# Rename cortisol col
wake_data_q3 <- wake_data_q3 %>%
  rename(Cortisol = `Cortisol (nmol/L)`)

```

```{r, echo = F}
#################################### DHEA ######################################

### Remove DHEA values that equaled 5.205
wake_data_q3 <- wake_data_q3 %>%
  filter(
    is.na(`DHEA (nmol/L)`) |
    !(dplyr::near(`DHEA (nmol/L)`, 5.205) | SubjectID == "3037"))

# 14 observations removed - 358 observations left

```

```{r, echo = F, include = F}
ggplot(wake_data_q3, aes(x = `Collection Sample`,
                      y = `DHEA (nmol/L)`,
                      color = factor(SubjectID))) +
  geom_point() + 
  geom_line() +
  facet_wrap(~DAYNUMB) + 
  theme_lucid()
```

```{r, echo = F, include = F}
##### Data checks

# A plot of the distribution of the outcome
ggplot(wake_data_q3, aes(x = `DHEA (nmol/L)`)) +
         geom_histogram(bins = 20, fill = "palevioletred2",
                        col = "black") + 
  theme_lucid() 

# A plot of the distribution of the outcome
ggplot(wake_data_q3, aes(x = log(`DHEA (nmol/L)`))) +
         geom_histogram(bins = 20, fill = "palevioletred2",
                        col = "black") + 
  theme_lucid() 

# We will log transform DHEA

```

In Table 3 we display the estimates, 95% Confidence Intervals and P-values for the DHEA model:

```{r, echo = F}

#### Fit the LMM

# Create a logged DHEA variable
wake_data_q3$log_DHEA <- log(wake_data_q3$`DHEA (nmol/L)`)

# Knot at 30 minutes - Collection Sample 2
wake_data_q3 <- wake_data_q3 %>%
  mutate(
    # Slope from waking to 30 miutes
    time_pre30  = pmin(minutes_since_wake, 30),
    # Slope from after 30 minutes
    time_post30 = pmax(minutes_since_wake - 30,
                       0)
  )

# Center Sleep Diary wake minutes
wake_data_q3 <- wake_data_q3 %>% 
  mutate(Sleep_Centered = `Sleep Diary Wake Minutes` - 
           mean(`Sleep Diary Wake Minutes`, na.rm = TRUE),
         Book_MinC = Booklet_Minutes - mean(Booklet_Minutes, 
                                            na.rm = TRUE)) %>% 
  relocate(Sleep_Centered, .after = `Sleep Diary Wake Minutes`)
# Mean wake up time is - 409.4545 minutes


##### Model
model_q3 <- lme(log_DHEA ~ time_pre30 + time_post30,
                 random = ~ 1 | SubjectID,
                 method = "REML",
                na.action = na.omit,
                 data = wake_data_q3)

# summary(model_q3)
# confint(model_q3)

# Display model summary with kable
# Extract t and p-values from model for data frame
summary_df2 <- as.data.frame(summary(model_q3)$tTable)

# kable
kable(
  summary_df2,
  format = "latex",
  booktabs = TRUE,
  digits = 5,
  caption = "Model Summary"
) |>
  kable_styling(latex_options = c("striped", "hold_position"))

```

```{r, echo = F}
#### Back transform betas 
# Pre 30
a <- -0.01724 * 30 
a_orig <- (exp(a) - 1) * 100 # -38.52871

# Confidence interval
a1 <- (exp(-0.02313 * 30) - 1) * 100 # -71.6346
a2 <- (exp(-0.0113 * 30) - 1) * 100 # -25.92

# Post 30
b <- -0.00153 * 600
b_org <- (exp(b) - 1) * 100 # -37.07332
b1 <- (exp(-0.00183* 600) - 1) * 100 
b2 <- (exp(-0.00122* 600) - 1) * 100 


```

For the time period of wake time to 30 minutes, there is approximately a 40% decrease in DHEA (95% CI: (-50.04%, -28.75%)).
This is statistically significant with p < 0.001.
DHEA shows a statistically significant decrease, not the expected increase, from waking to 30 minutes after waking. 

As hypothesized, from 30 minutes to 600 minutes, there is approimately a 60% decrease in DHEA levels (95% CI: (-66.65%, -51.91%)). This is statistically significant with p-val < 0.001.
We visualize these slopes in Figure 3 below. 


```{r, echo = F, warning = F, message = F, out.width="90%", fig.align='center', fig.cap="Diurnal DHEA Pattern"}
# For plotting purposes here, we use a model with lmer()!!
model_plot_dhea <- lmer(log_DHEA ~ time_pre30 + time_post30
                 + (1 | SubjectID),
                 REML = TRUE,
                 data = wake_data_q3)

t_max <- quantile(wake_data_q3$minutes_since_wake, 0.95, na.rm = TRUE)

pred_grid <- tibble(
  minutes_since_wake = seq(0, as.numeric(t_max), by = 5)) %>%
  mutate(
    time_pre30  = pmin(minutes_since_wake, 30),
    time_post30 = pmax(minutes_since_wake - 30, 0)
  )

# Get fixed effects predictions
pred_grid$fit <- predict(model_plot_dhea, newdata = pred_grid, re.form = NA)

# Add a 95% CI
X <- model.matrix(~ time_pre30 + time_post30, data = pred_grid)
V <- vcov(model_plot_dhea)
se_fit <- sqrt(diag(X %*% V %*% t(X)))

pred_grid <- pred_grid %>%
  mutate(
    lwr = fit - 1.96 * se_fit,
    upr = fit + 1.96 * se_fit
  )

# Plot
ggplot() +
  geom_point(data = wake_data_q3,
             aes(x = minutes_since_wake, y = log_DHEA),
             alpha = 0.25, size = 1) +
  geom_ribbon(data = pred_grid,
              aes(x = minutes_since_wake, 
                  ymin = lwr,
                  ymax = upr),
              alpha = 0.2,
              fill = "blue") +
  geom_line(data = pred_grid,
            aes(x = minutes_since_wake, y = fit),
            linewidth = 1,
            col = "blue") +
  geom_vline(xintercept = 30, linetype = 2,
             col = "red") +
  labs(x = "Minutes since waking", y = "log(DHEA)",
       title = "DHEA Pattern Over Time",
       subtitle = "Knot at 30 minutes") + 
  theme_lucid()

```


```{r, echo = F, include = F, eval = F}
###### Model diagnostics
# Obtain data used to fit model
model_dataq3 <- model_q3@frame

# Obtain residuals of model
model_dataq3$resids <- resid(model_q3)

# Obtain fitted values
model_dataq3$fitted <- fitted(model_q3)

# Residual histogram
ggplot(model_dataq3, aes(x = resids)) +
  geom_histogram(bins = 25,
                 fill = "cadetblue2",
                 color = "black") +
  theme_lucid()

# residual vs. fitted
ggplot(model_dataq3, aes(x = fitted, y = resids)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = 2) +
  theme_classic()
```

### Cortisol

```{r, echo = F}
################################## Cortisol ##################################

# Determine how many measurements are above 80
# wake_data_q3 %>% 
#   filter(`Cortisol (nmol/L)` >= 26) %>% 
#   reframe(SubjectID, `Collection Sample`, 
#           DAYNUMB,`Cortisol (nmol/L)`)

# Exclude the observation with 89 cortisol
wake_data_q3 <- wake_data_q3 %>%
  filter(!dplyr::near(Cortisol, 89.55714))

```

```{r, echo = F, message = F, warning=F, out.width="80%", eval = F, fig.align='center'}
# Visualize cortisol over time
ggplot(wake_data_q3, aes(x = `Collection Sample`,
                      y = Cortisol,
                      color = factor(SubjectID))) +
  geom_point() + 
  geom_line() +
  facet_wrap(~DAYNUMB) + 
  theme_lucid()
```


```{r, echo = F}
######### Model

# Model
model_cortisol <- lme(Cortisol ~ time_pre30 + time_post30,
                  random = ~ 1 | SubjectID,
                 method = "REML",
                 na.action = na.omit,
                 data = wake_data_q3)

# summary(model_cortisol)
# intervals(model_cortisol)

# Extract t and p-values from model for data frame
summary_df3 <- as.data.frame(summary(model_cortisol)$tTable)

# kable
kable(
  summary_df3,
  format = "latex",
  booktabs = TRUE,
  digits = 5,
  caption = "Model Summary"
) |>
  kable_styling(latex_options = c("striped", ""))

```

```{r, echo = F, eval = F}
#### Back transform betas 
# Pre 30
c <- 0.025216 * 30 
c_orig <- (exp(c) - 1) * 100 # 

# Confidence interval
c1 <- (exp(-0.0199 * 30) - 1) * 100 # 
c2 <- (exp( 0.07038 * 30) - 1) * 100 # 

# Post 30
d <- -0.009 * 600
d_org <- (exp(d) - 1) * 100 # -37.07332
d1 <- (exp(-0.0117* 600) - 1) * 100 
d2 <- (exp(-0.0070* 600) - 1) * 100 


```


For the time period of wake time to 30 minutes, there is approximately a 54% increase in Cortisol (95% CI: (-60.19%, 503.15%)).
This is not statistically significant with p = 0.53.
Although there is an increase from waking to 30 minutes after waking, it is not a statistically significant increase.  

From 30 minutes to 600 minutes, there is approimately a 99% decrease in DHEA levels (95% CI: (-99.86%, -97.27%)). This is statistically significant with p-val < 0.001. 
As expected, Cortisol showed a statistically significant decline over 10 hours. 
Cortisol follows the expected pattern of an increase (although not significant) from waking to 30 minutes followed by a decline from 30 minutes and on. 


```{r, echo = F, warning = F, message = F, out.width="90%", fig.align='center', fig.cap="Diurnal Cortisol Pattern"}
# For plotting purposes here, we use a model with lmer()!!
model_plot_cort <- lmer(Cortisol ~ time_pre30 + time_post30
                 + (1 | SubjectID),
                 REML = TRUE,
                 data = wake_data_q3)

t_max <- quantile(wake_data_q3$minutes_since_wake, 0.95, na.rm = TRUE)

pred_grid <- tibble(
  minutes_since_wake = seq(0, as.numeric(t_max), by = 5)) %>%
  mutate(
    time_pre30  = pmin(minutes_since_wake, 30),
    time_post30 = pmax(minutes_since_wake - 30, 0))

# Get fixed effects predictions
pred_grid$fit_cort <- predict(model_plot_cort, 
                              newdata = pred_grid,
                              re.form = NA)

# Add a 95% CI
X_cortisol <- model.matrix(~ time_pre30 + time_post30, 
                           data = pred_grid)
V_cortisol <- vcov(model_plot_cort)

se_fit_cort <- sqrt(diag(X_cortisol %*% V_cortisol %*% t(X_cortisol)))

pred_grid <- pred_grid %>%
  mutate(
    lwr_cort = fit_cort - 1.96 * se_fit_cort,
    upr_cort = fit_cort + 1.96 * se_fit_cort
  )

# Plot
ggplot() +
  geom_point(data = wake_data_q3,
             aes(x = minutes_since_wake, 
                 y = Cortisol),
             alpha = 0.25, size = 1) +
  geom_ribbon(data = pred_grid,
              aes(x = minutes_since_wake, 
                  ymin = lwr_cort,
                  ymax = upr_cort),
              alpha = 0.2,
              fill = "blue") +
  geom_line(data = pred_grid,
            aes(x = minutes_since_wake, 
                y = fit_cort),
            linewidth = 1,
            col = "blue") +
  geom_vline(xintercept = 30, linetype = 2,
             col = "red") +
  labs(x = "Minutes since waking", y = "Cortisol",
       title = "Cortisol Pattern Over Time",
       subtitle = "Knot at 30 minutes") + 
  theme_lucid()


```

```{r, echo = F, include = F, eval = F}
###### Model diagnostics
# Obtain data used to fit model
model_data_cort <- model_cortisol@frame

# Obtain residuals of model
model_data_cort$resids <- resid(model_cortisol)

# Obtain fitted values
model_data_cort$fitted <- fitted(model_cortisol)

# Residual histogram
ggplot(model_data_cort, aes(x = resids)) +
  geom_histogram(bins = 25,
                 fill = "cadetblue2",
                 color = "black")

# residual vs. fitted
ggplot(model_data_cort, aes(x = fitted, y = resids)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = 2) +
  theme_classic()
```



# Discussion and Summary

<!-- In our missingness investigation, we determined that the variable with the highest missingness was MEM time. This was expected due to potential cap malfunction, such as low battery, or user error. In contrast, booklet-recorded time had much lower missingness. Measurements of DHEA and cortisol did not have substantial missingness, indicating that missingness was largely limited to the timing variables instead of the biological outcomes. -->

In regard to the comparison of MEM and booklet times, we found a statistically significant and strong linear relationship between the two measures, indicating that booklet-recorded times tracked closely with MEM times. However, we also observed that booklet times tended to be recorded slightly earlier than MEM times, on average. This suggests the presence of a small systematic bias. Possible explanations include delayed electronic cap recordings due to technical issues or perhaps inaccuracies in self-reported booklet times. This highlights a known limitation of studies relying on self-reported sampling times.

For the +30 minute sampling window, the percentages for both good and adequate adherence were reasonably high, at approximately 80%.
On the other hand for the +600 minute sampling window, we saw a notable decrease in adherence. Both good and adequate adherence percentages for booklet and MEM times fell below 50%. This decrease is expected, since longer intervals from waking are more susceptible to schedule variability which undoubtedly can affect accurate timing.
Across both sampling windows and adherence thresholds, the impact of missing MEM time was apparent, with lower adherence percentages compared to booklet time. 
<!-- In Figure 2, we observed clear differences between the distributions of minutes from the target sampling times for the +30 minute and +600 minute windows, as well as between MEM and booklet times. All distributions showed long right tails, indicating larger deviations from the target sampling times for some observations. However, the distributions were centered near zero, suggesting that the majority of subjects collected samples within the good or adequate adherence windows. -->

In the assessment of DHEA and cortisol patterns over time, we found that cortisol followed the expected diurnal pattern, with higher levels earlier in the day and a decline over time. In contrast, DHEA did not exhibit a clear or consistent pattern over the sampling period. These differences may be due to several factors, including biological differences in hormone regulation, variability within this particular cohort, or limitations in sampling frequency and timing.

In summary, our findings suggest that booklet-based saliva collection using the SPIT device demonstrates strong agreement with electronic monitoring, though a small systematic bias toward earlier recorded booklet times was observed. Adherence to short-term sampling requirements was generally high, while adherence to longer-term sampling windows was more variable. Cortisol exhibited the expected diurnal decline, whereas DHEA did not show a clear temporal pattern in this sample. These results highlight the feasibility and limitations of at-home saliva collection protocols and emphasize the importance of accurate timing and careful handling of missing data in studies of diurnal hormone patterns.

# Data and Code Availability

The data and statistical code used to generate the results and figures in this report are available in the following GitHub repository:

- GitHub repository: https://github.com/DianaF65/BIOS6624

Directory structure:

\begin{itemize}
  \item Background/ - A description of the background of the project.
  \item Code/ - A R Markdown file consisting of code for exploratory and an initial report draft as well as as a .R file of code for the complete analysis.
  \item RawData/ – Raw data files.
  \item Reports/ R Markdown file used to generate this report
\end{itemize}

All analyses and figures were generated using R, and the repository contains the R scripts and R Markdown files used to produce this report



