---
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    latex_engine: xelatex
    extra_dependencies: ["longtable", "float"]
mainfont: Arial
fontsize: 12pt
documentclass: article
geometry: margin=1in
header-includes:
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage{float}
  - \usepackage{caption}
  - \usepackage{subcaption}
---

```{r, echo = F, warning=F, message = F}
knitr::opts_chunk$set(echo = TRUE)


library(readr)
library(knitr)
library(ggplot2)
library(MASS)
library(emmeans)
library(multcomp)
library(kableExtra)
library(stats)
library(dplyr)
library(tibble)
library(labelled)
library(tableone)
library(lme4)
library(lmerTest)
library(nlme)
library(gtsummary)
library(gt)
library(ggeffects)
library(UpSetR)
library(visdat)
library(lubridate)
library(see)
library(data.table)
library(gridExtra)
library(kableExtra)
opts_chunk$set(tidy = F)
```

```{r, echo = F, warning = F, message = F}
# Read in the data
proj0_dat <- read.csv("../RawData/Project0_Clean_v2.csv",
                      header = TRUE,
                      check.names = FALSE)

### Modify data for Person 3029
proj0_dat <- proj0_dat %>% 
  mutate(DAYNUMB = case_when(
    SubjectID == 3029 & `Collection Date` == "10/7/2018" ~ 1,
    SubjectID == 3029 & `Collection Date` == "10/10/2018" ~ 2,
    SubjectID == 3029 & `Collection Date` == "10/12/2018" ~ 3,
    TRUE ~ DAYNUMB
  )) %>% 
  ungroup()

```


```{r, echo = F, warning = F, message = F}
############################## Question 1 Data #################################

## Make a subset of the data with the relevant cols for Q1 - Q3

# Vector of the cols of interest
cols <- c("SubjectID",
          "Collection Date", "DAYNUMB",
          "Collection Sample",
          "Booket: Clock Time", "MEMs: Clock Time",
          "Sleep Diary reported wake time",
          "Cortisol (nmol/L)", "DHEA (nmol/L)")

# Subset the original data frame
wake_data <- proj0_dat[, colnames(proj0_dat) %in% cols ]
 
# Convert the Sleep diary Wake time to minutes
wake_data$`Sleep Diary Wake Minutes` <-
  # Convert to date objects and convert to minutes
  lubridate::hour(hm(wake_data$`Sleep Diary reported wake time`)) * 60 +
  lubridate::minute(hm(wake_data$`Sleep Diary reported wake time`))

# Move this to after Collection Sample
wake_data <- wake_data %>% 
  relocate(`Sleep Diary Wake Minutes`, 
           .after = `Sleep Diary reported wake time`)

# Convert the Booklet Time to minutes
wake_data$Booklet_Minutes <-
  # Convert to date objects and convert to minutes
  lubridate::hour(hm(wake_data$`Booket: Clock Time`)) * 60 +
  lubridate::minute(hm(wake_data$`Booket: Clock Time`))

# Move this to after Booklet: Clock Time
wake_data <- wake_data %>% 
  relocate(Booklet_Minutes, .after = `Booket: Clock Time`)

# Convert the MEM Time to minutes
wake_data$MEM_Minutes <-
  # Convert to date objects and convert to minutes
  lubridate::hour(hm(wake_data$`MEMs: Clock Time`)) * 60 +
  lubridate::minute(hm(wake_data$`MEMs: Clock Time`))

# Move this to after MEMs: Clock Time
wake_data <- wake_data %>% 
  relocate(MEM_Minutes, .after = `MEMs: Clock Time`)

# Center MEM time
wake_data$MEM_C <-
  # Convert to date objects and convert to minutes
  lubridate::hour(hm(wake_data$`MEMs: Clock Time`)) * 60 +
  lubridate::minute(hm(wake_data$`MEMs: Clock Time`))

# Move this to after MEMs: Clock Time
wake_data <- wake_data %>% 
  relocate(MEM_C, .after = `MEMs: Clock Time`)

# Create a Minutes SINCE WAKE variable
wake_data <- wake_data %>%
  mutate(
    clock_min = lubridate::hour(hm(`Booket: Clock Time`)) * 60 +
                lubridate::minute(hm(`Booket: Clock Time`))) %>%
  group_by(SubjectID, DAYNUMB) %>%
  mutate(wake_min = clock_min[`Collection Sample` == 1][1],
         minutes_since_wake = clock_min - wake_min) %>%
  ungroup() %>% 
  relocate(minutes_since_wake, .after = `Sleep Diary Wake Minutes`)


### Modify data for Person 3029
wake_data <- wake_data %>% 
  mutate(DAYNUMB = case_when(
    SubjectID == 3029 & `Collection Date` == "10/7/2018" ~ 1,
    SubjectID == 3029 & `Collection Date` == "10/10/2018" ~ 2,
    SubjectID == 3029 & `Collection Date` == "10/12/2018" ~ 3,
    TRUE ~ DAYNUMB
  )) %>% 
  ungroup()

# Create a subset of the data frame that only has Wake times
# AKA Collection Sample == 1
wakes <- wake_data[wake_data$`Collection Sample` == 1, ]

```


# Introduction

Research suggests that there is a relationship between hormonal patterns across the span of a day and conditions including sleep disorders and depression.
In this study, the focus is on the hormones, DHEA and Cortisol, which play a role in regulating stress. 
Cortisol is an important steroid hormone that significantly affects metabolism which in turn affects the central nervous system. 
DHEA is 

In this study, researchers followed the cortisol and DHEA patterns over the course of a day for a totstudal y period of three days to measure the the patterns of the two hormones in a natural environment.  
Researchers utilized two methods to collect saliva samples; a Saliva Procurement and Integrated Testing (SPIT) booklet and an automatic Electronic Monitoring Cap. Saliva samples were collected at waking, 30 minutes after waking, before lunch, and 600 minutes after waking. Subjects did not have to adhere to a specific wake up time for 31 healthy control subjects. 

This report addresses the following hypotheses of interest: 

\textbf{Hypothesis 1: } 
There was satisfactory agreement between the subject's recording of sampling times compared to the recording times by the electronic monitoring cap. More specifically, there is a one-to-one relationship between the subject's recorded times and the booklet times. 

\textbf{Hypothesis 2: } 
Subjects adhered appropriately to the study protocol of the +30 minute and +600 minute sampling times. 

\textbf{Hypothesis 3: DHEA and Cortisol levels show an initial increase from waking to 30 minutes followed by a gradual decline from 30 minutes and on. } 

We will further refer to the electronic monitoring cap as MEM time and the the SPIT Booklet as Booklet time for the remainder of the report.

# Methods

## Preliminary Data Analysis

As missingness can present serious issues for the data analysis if not explored and addressed, we investigated missingness through calculating the number and percentage of missing values for each variable in the dataset. We further visualized missingness for each variable.  

Below is a Table1 of the primary variables of interest for the analysis:

For the variables we employed in our analyses, we assessed and visualized distributions to determine if there were any outliers or considerations that needed to be taken for data analysis. 

## Data Cleaning




## Data Analysis

\textbf{1:} To address research question 1, we employed a longitudinal linear mixed effects model with Booklet time in Minutes for the outcome, MEM in Minutes and the day number (1-3) as covariates We included a random intercept for subject that allowed us to take variability between subjects into account in our model. 
Adding Day as a covariate noticeably improved diagnostic plots for the assumptions of heteroscedasticity and normality. We additionally assessed bias by.... COME BACK TO THIS

\textbf{2: } To determine whether subjects were adhering accurately to the +30 minute and +600 minute sampling times required by study protocol, intervals of time were defined to determine the number and percentage of patients that had good or adequate adherence.
Good adherence was defined by a collection of the saliva sample within a 7.5 minute window interval of the +30 minute and +600 minute times whereas adequate adherence if samples were within a 15 minute window. 
To calculate the percentage and proportion of individuals that had good adherence and adequate adherence, the difference between the "correct" sampling time and the observed time was calculated and this difference was used to assess whether observed times were within the good or adequate adherence windows. 

We visualized and summarized the distribution of the MEM and Booklet times to guide us in our decision to proceed with either MEM or Booklet times for our calculations.
We observed that MEM times had a considerable amount of missingness and determined that reporting both would give a better overall sense of adherence.
We summarized adherence by subjects as well as adherence for the whole study population.

\textbf{2: } 


# Results

From the missingness investigation, we determined the MEMs had the highest amount of missing: 177 missing values which is equivalent to 47.6%. 
Booklet had the second highest amount of missing with 55 missing values or 14.8%.
Cortisol and DHEA measurements had 5 missing values or 1.34% missingness.

The graphic below allowed us to visually see the quantity of missingness in the MEM times compared to the other variables in the data set, we followed certain thresholds for good and adequate adherence.


## Hypothesis 1

Below is a table of the estimates, 95% Confidence Intervals and P-values for the model: 

```{r, echo = F}

```


```{r, echo = F}
# LMM with random intercept for subject - Day added
model_q1 <- lme(Booklet_Minutes ~ MEM_C + factor(DAYNUMB), 
                random = ~ 1 | SubjectID,
                na.action = na.omit,
                method = "REML",
                data = wakes)


# Display model summary with kable
# Extract t and p-values from model for data frame
summary_df1 <- as.data.frame(summary(model_q1)$tTable)

# kable
kable(
  summary_df1,
  format = "latex",
  booktabs = TRUE,
  digits = 5,
  caption = "Model Summary"
) |>
  kable_styling(latex_options = c("striped"))


```

```{r, echo = F, eval = F}
intervals(model_q1)

#  Fixed effects:
#                       lower       est.      upper
# (Intercept)      -5.0891734  3.6513329 12.3918392
# MEM_C             0.9667782  0.9859424  1.0051067
# factor(DAYNUMB)2 -6.6349646 -3.4589031 -0.2828415
# factor(DAYNUMB)3 -5.0521491 -1.8810020  1.2901450
```

Holding day constant, a one-minute increase in MEM-recorded wake time is associated with an approximately one-minute increase in Booklet-recorded wake time ($\beta$ ≈ 0.99; 95% CI: (0.97–1.01); p < 0.0001). This near one-to-one association indicates strong concordance between the two wake-time measures within a given day.
In other words, the subject's recorded wake up time and the MEM's recorded wake up time track each other quite closely. Changes in MEM time are mirrored almost exactly by changes in Booklet time and our initial hypothesis is correct. 

## Hypothesis 2

```{r, echo = F}
################################################################################
###                               Question 2                                 ### 
################################################################################


# Df with Wake time, 30 min post wake, and 600 min post wake
wake_check <- wake_data %>%
  # filter(`Collection Sample` %in% c(1, 2)) %>%
  group_by(SubjectID, DAYNUMB) %>%
  reframe(
    # Wake time
    wake_time = `Sleep Diary Wake Minutes`[`Collection Sample` == 1],
    # The 30 minute time with Booklet
    Book_30time = Booklet_Minutes[`Collection Sample` == 2],
    # The 600 miniute time with booklet
    Book_600time = Booklet_Minutes[`Collection Sample` == 4],
    # The 30 minute time with MEM
    MEM_30time = MEM_Minutes[`Collection Sample` == 2],
    # The 600 minute time with MEM
    MEM_600time = MEM_Minutes[`Collection Sample` == 4]
    )

 
# Define adherence for BOOK
book_wake_check <- wake_check %>%
  mutate(
    # 30 min after waking target time
    target_30time = wake_time + 30,
    # 600 min after waking target time
    target_600time = wake_time + 600,
    # Difference between 30 min later and the recorded time
    diff_min30  = Book_30time - target_30time,
    # Difference between 600 min later and the recorded time
    diff_min600 = Book_600time - target_600time,
    # For 30 min: Define good adherence - within 7.5 minutes
    good_adherent30 = abs(diff_min30) <= 7.5,
    # For 600 min: Define good adherence
    good_adherent600 = abs(diff_min600) <= 7.5,
    # For 30 min: Define adequate adherence - within 15 minutes
    adequate_ad30 = abs(diff_min30) <= 15,
    # For 600 min: Define adequate adherence
    adequate_ad600 = abs(diff_min600) <= 15
  ) %>% 
  relocate(c(target_30time, diff_min30,
             good_adherent30, adequate_ad30), 
           .after = Book_30time) %>% 
  relocate(c(target_600time, diff_min600,
             good_adherent600, adequate_ad600), 
           .after = Book_600time)

# Define adherence for MEM
MEM_wake_check <- wake_check %>%
  mutate(
    # 30 min after waking target time
    target_30time = wake_time + 30,
    # 600 min after waking target time
    target_600time = wake_time + 600,
    # Difference between 30 min later and the recorded time
    diff_min30  = MEM_30time - target_30time,
    # Difference between 600 min later and the recorded time
    diff_min600 = MEM_600time - target_600time,
    # For 30 min: Define good adherence - within 7.5 minutes
    good_adherent30 = abs(diff_min30) <= 7.5,
    # For 600 min: Define good adherence
    good_adherent600 = abs(diff_min600) <= 7.5,
    # For 30 min: Define adequate adherence - within 15 minutes
    adequate_ad30 = abs(diff_min30) <= 15,
    # For 600 min: Define adequate adherence
    adequate_ad600 = abs(diff_min600) <= 15
  ) %>% 
  relocate(c(target_30time, diff_min30,
             good_adherent30, adequate_ad30), 
           .after = MEM_30time) %>% 
  relocate(c(target_600time, diff_min600,
             good_adherent600, adequate_ad600), 
           .after = MEM_600time)

```

```{r, echo = F}
################################## Good Adherence ##############################

############################## Adherence by Subject ############################

################################### +30 Minutes ################################

######## Booklet
subject_good_30adherence <- book_wake_check %>%
  group_by(SubjectID) %>%
  summarise(
    # How many days out of 3 were they adherent?
    days_good_adherent = sum(good_adherent30),
    # What is the proportion
    prop_good_adherent = mean(good_adherent30),
    .groups = "drop"
  )
# subject_good_30adherence

# How many subjects adhered all 3 days
# sum(subject_good_30adherence$prop_good_adherent == 1, na.rm = T)
# Proportion and Percentage of subjects
# 14/31 = 0.45 or 45%

######## MEM
MEM_subject_good_30adherence <- MEM_wake_check %>%
  group_by(SubjectID) %>%
  summarise(
    # How many days out of 3 were they adherent?
    days_good_adherent = sum(good_adherent30),
    # What is the proportion
    prop_good_adherent = mean(good_adherent30),
    .groups = "drop"
  )
# MEM_subject_good_30adherence

# How many subjects adhered all 3 days
# sum(MEM_subject_good_30adherence$prop_good_adherent == 1, na.rm = T)
# Proportion and Percentage of subjects
# 6/31 




################################# Adequate Adherence ###########################

############################## Adherence by Subject ############################

################################### +30 Minutes ################################

###### Book

subject_ad_30adherence <- book_wake_check %>%
  group_by(SubjectID) %>%
  summarise(
    # How many days out of 3 were they adherent?
    days_ad_adherent = sum(adequate_ad30),
    # What is the proportion
    prop_ad_adherent = mean(adequate_ad30),
    .groups = "drop"
  )

# subject_ad_30adherence

# How many subjects adhered all 3 days
# sum(subject_ad_30adherence$prop_ad_adherent == 1, na.rm = T)
# Proportion and Percentage of subjects
# 20/31 = 0.65 or 65%

###### MEM
MEM_subject_ad_30adherence <- MEM_wake_check %>%
  group_by(SubjectID) %>%
  summarise(
    # How many days out of 3 were they adherent?
    days_ad_adherent = sum(adequate_ad30),
    # What is the proportion
    prop_ad_adherent = mean(adequate_ad30),
    .groups = "drop"
  )

# MEM_subject_ad_30adherence

# How many subjects adhered all 3 days
# sum(MEM_subject_ad_30adherence$prop_ad_adherent == 1, na.rm = T)
# Proportion and Percentage of subjects
# 10/31 = 0.32 or 32%


################################### +60 Minutes ################################



```

Below we summarize the percentage of subjects that adhered to the study sampling times. Percentages for the +30 Minute sampling time is in Table 2 and percentages for the +600 Minute sampling time is in Table 3, for both MEM and Booklet times. 

```{r, echo = F}
################################## Good Adherence ##############################

################################ Adherence Overall #############################

################################### +30 Minutes ################################

######### Book
# Adherence overall
good_book <- book_wake_check %>%
  # group_by(SubjectID) %>%
  summarise(
    # What is the proportion
    prop_good_adherent = round(mean(good_adherent30, na.rm = TRUE), 4),
    # Percentage
    per_good_adherence = (prop_good_adherent * 100)
  )
# good_book
# A tibble: 1 × 2
#   prop_good_adherent per_good_adherence
#                <dbl>              <dbl>
# 1              0.782               78.2


######### MEM
# Adherence overall
good_MEM <- MEM_wake_check %>%
  # group_by(SubjectID) %>%
  summarise(
    # What is the proportion
    prop_good_adherent = round(mean(good_adherent30, na.rm = TRUE), 4),
    # Percentage
    per_good_adherence = (prop_good_adherent * 100)
  )
# good_MEM
# A tibble: 1 × 2
#   prop_good_adherent per_good_adherence
#                <dbl>              <dbl>
# 1              0.529               52.9

################################# Adequate Adherence ###########################

################################ Adherence Overall #############################

################################### +30 Minutes ################################

######## Book

# Adherence overall
adequate_book <- book_wake_check %>%
  summarise(
    # What is the proportion
    prop_adeq_adherent = round(mean(adequate_ad30, na.rm = TRUE), 4),
    # Percentage
    per_adeq_adherence = (prop_adeq_adherent * 100)
  )
# adequate_book
#   prop_adeq_adherent per_adeq_adherence
#                <dbl>              <dbl>
# 1              0.897               89.7

######## MEM
# Adherence overall
adequate_MEM <- MEM_wake_check %>%
  summarise(
    # What is the proportion
    prop_adeq_adherent = round(mean(adequate_ad30, na.rm = TRUE), 4),
    # Percentage
    per_adeq_adherence = (prop_adeq_adherent * 100)
  )
# adequate_MEM
#   prop_adeq_adherent per_adeq_adherence
#                <dbl>              <dbl>
# 1              0.714               71.4

###### Create Data frame to display with this
# Combine all the proportions into a table to display nicely
props <- data.frame(Good = c(good_book$per_good_adherence, 
                             good_MEM$per_good_adherence),
                    Adequate = c(adequate_book$per_adeq_adherence,
                                 adequate_MEM$per_adeq_adherence))

# Add rownames
rownames(props) <- c("Booklet", "MEM")

```


```{r, echo = F}
################################## Good Adherence ##############################

################################ Adherence Overall #############################

#################################### +600 Minutes ##############################

##### Book
# Adherence overall
good_book600 <- book_wake_check %>%
  # group_by(SubjectID) %>%
  summarise(
    # What is the proportion
    prop_good_adherent = round(mean(good_adherent600, na.rm = TRUE), 4),
    # Percentage
    per_good_adherence = (prop_good_adherent * 100)
  )
#   prop_good_adherent per_good_adherence
#                <dbl>              <dbl>
# 1              0.462               46.2

##### MEM
# Adherence overall
good_MEM600 <- MEM_wake_check %>%
  # group_by(SubjectID) %>%
  summarise(
    # What is the proportion
    prop_good_adherent = round(mean(good_adherent600, na.rm = TRUE), 4),
    # Percentage
    per_good_adherence = (prop_good_adherent * 100)
  )
#   prop_good_adherent per_good_adherence
#                <dbl>              <dbl>
# 1              0.325               32.5

############################### Adequate Adherence #############################

################################ Adherence Overall #############################

#################################### +600 Minutes ##############################

# Adherence overall
adequate_book600 <- book_wake_check %>%
  summarise(
    # What is the proportion
    prop_adeq_adherent = round(mean(adequate_ad600, na.rm = TRUE), 4),
    # Percentage
    per_adeq_adherence = (prop_adeq_adherent * 100)
  )
#   prop_adeq_adherent per_adeq_adherence
#                <dbl>              <dbl>
# 1              0.562               56.2

# Adherence overall
adequate_MEM600 <- MEM_wake_check %>%
  summarise(
    # What is the proportion
    prop_adeq_adherent = round(mean(adequate_ad600, na.rm = TRUE), 4),
    # Percentage
    per_adeq_adherence = (prop_adeq_adherent * 100)
  )

#   prop_adeq_adherent per_adeq_adherence
#                <dbl>              <dbl>
# 1              0.398               39.8
  
###### Create Data frame to display with this
# Combine all the proportions into a table to display nicely
props600 <- data.frame(Good = c(good_book600$per_good_adherence, 
                             good_MEM600$per_good_adherence),
                    Adequate = c(adequate_book600$per_adeq_adherence,
                                 adequate_MEM600$per_adeq_adherence))

# Add rownames
rownames(props600) <- c("Booklet", "MEM")
  

```


```{r results='asis', echo=FALSE}
# Combine information in a more cohesive way

props30_long <- props %>%
  as.data.frame() %>%
  rownames_to_column("Method") %>%
  mutate(Window = "+30 minutes")

props60_long <- props600 %>%
  as.data.frame() %>%
  rownames_to_column("Method") %>%
  mutate(Window = "+60 minutes")

final_props <- bind_rows(props30_long, props60_long) %>%
  relocate(Window, Method)

kable(
  final_props,
  format = "latex",
  booktabs = TRUE,
  digits = 2,
  caption = "Percentages for Good and Adequate Adherence by Collection Window"
) |>
  kable_styling(latex_options = c("striped", "hold_position")) |>
  collapse_rows(columns = 1, latex_hline = "major")


```


Overall, more subjects are meeting the adequate adherence thresholds compared to the good adherence thresholds for both the +30 minute and +600 minute sampling times. In addition, the MEM percentages are lower for both sampling times and for both good and adequaate thresholds.


Below we visualize the minutes from the +30 minute sampling time (left column) and the +600 minute sampling time (right column). 
Times for Booklet are in the top row and time for MEM are in the bottom row. 
The red dashed lines indicate the $\pm$ 7.5 window for good adherence while the pink dashed lines are the $\pm$ 15 minute window for adequate adherence. 


```{r, echo = F, warning = F, message=F, fig.align='center', fig.cap="Adherence for +30 and +600 Sampling Times", fig.pos="H"}

#################################### Visualize #################################

# Visualize the distribution of minutes from the sampling times

# Book +30 min
p1 <- ggplot(book_wake_check, aes(x = diff_min30)) +
  geom_histogram(binwidth = 3, col = "black", fill = "turquoise") +
  geom_vline(xintercept = c(-7.5, 7.5), col = "red",
             linetype = "dashed") +
  geom_vline(xintercept = c(-15, 15), color = "hotpink",
             linetype = "dashed") +
  theme_lucid() + 
  scale_y_continuous(breaks = seq(0, 40, by = 10)) + 
  labs(title = "Minutes from +30 Minute Sampling Time",
       subtitle = "Booklet",
       x = "Minutes from target (+30)",
       y = "Count") 

# MEM + 30 min
p2 <- ggplot(MEM_wake_check, aes(x = diff_min30)) +
  geom_histogram(binwidth = 10, col = "black", fill = "turquoise") +
  geom_vline(xintercept = c(-7.5, 7.5), col = "red",
             linetype = "dashed") +
  geom_vline(xintercept = c(-15, 15), color = "hotpink",
             linetype = "dashed") +
  theme_lucid() + 
  scale_y_continuous(breaks = seq(0, 40, by = 10),
                     limits = c(0, 40)) +
  labs(title = "Minutes from +30 Minute Sampling Time",
       subtitle = "MEM",
       x = "Minutes from target (+30)",
       y = "Count") 

# MEM +30 Min
p3 <- ggplot(book_wake_check, aes(x = diff_min600)) +
  geom_histogram(binwidth = 20, col = "black", fill = "turquoise") +
  geom_vline(xintercept = c(-7.5, 7.5), col = "red",
             linetype = "dashed") +
  geom_vline(xintercept = c(-15, 15), color = "hotpink",
             linetype = "dashed") +
  theme_lucid() + 
  scale_x_continuous(breaks = seq(-50, 360, by = 50)) + 
  scale_y_continuous(breaks = seq(0, 40, by = 10),
                     limits = c(0, 40)) +
  labs(title = "Minutes from +600 Sampling Time",
       subtitle = "Booklet",
       x = "Minutes from target (+600)",
       y = "Count") 

# MEM +600 min
p4 <- ggplot(MEM_wake_check, aes(x = diff_min600)) +
  geom_histogram(binwidth = 20, col = "black", fill = "turquoise") +
  geom_vline(xintercept = c(-7.5, 7.5), col = "red",
             linetype = "dashed") +
  geom_vline(xintercept = c(-15, 15), color = "hotpink",
             linetype = "dashed") +
  theme_lucid() + 
  scale_x_continuous(breaks = seq(-50, 360, by = 50)) + 
  scale_y_continuous(breaks = seq(0, 40, by = 10),
                     limits = c(0, 40)) +
  labs(title = "Minutes from +600 Minute Sampling Time",
       subtitle = "MEM",
       x = "Minutes from target (+600)",
       y = "Count") 


# Arrange 
grid.arrange(p1, p3, p2, p4, nrow = 2)

```

## Hypothesis 3



# Discussion

















