---
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    latex_engine: xelatex
mainfont: Arial
fontsize: 11pt
documentclass: article
geometry: margin=0.9in
---

```{r, echo = F, warning=F, message = F}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(knitr)
library(ggplot2)
library(MASS)
library(emmeans)
library(multcomp)
library(kableExtra)
library(stats)
library(dplyr)
library(tibble)
library(labelled)
library(tableone)
library(lme4)
library(lmerTest)
library(nlme)
library(gtsummary)
library(gt)
library(ggeffects)
library(UpSetR)
library(visdat)
library(lubridate)
library(see)
library(data.table)
opts_chunk$set(tidy = F)
```

```{r, echo = F}
# Read in the data
proj0_dat <- read.csv("../RawData/Project0_Clean_v2.csv",
                      header = TRUE,
                      check.names = FALSE)
```

# Missingness

Below we ummarize and visuaize missingness in the dataset. 

The number and percentage of missing values for each variable: 

```{r, echo = F}
# Cumulative sum of missingness for each variable
naniar::miss_var_summary(proj0_dat)
```

Investigating the NAs in MEMs: Sample Interval Decimal Time, it appears that a majority of the NAs correspond to the first time point of Collect Sample = 1. We will replace these with 0.. 

```{r, echo = F}
proj0_dat %>% 
  group_by(SubjectID) %>% 
  reframe(SubjectID, 
          `Collection Sample`,
          `MEMs: Sample interval Decimal Time (mins)`)
```



```{r, echo = F}
# How many variables 0 - 5 missing values
naniar::miss_case_table(proj0_dat)
```

Visualizing the missingness: 


```{r, echo = F}
# Remove variables for this plot
plot_dat <- proj0_dat
# Remove variables we do not use in analysis, etc.
vars <- c("Booklet: Sample interval", 
          "Booklet: Sample interval Decimal Time (mins)",
          "MEMs: Sample interval",
          "MEMs: Sample interval Decimal Time (mins)",
          "DHEA (pg/dl)",
          "Cortisol (ug/dl)")
 
plot_dat <- plot_dat[ , !(names(plot_dat) %in% vars)]


colnames(plot_dat) <- c("SubjectID", "Collection_Date",
                        "Collection_Sample",
                        "Booklet Time", "MEMs Time",
                        "Wake Time", "Booklet Sample Interval",
                        "Booklet")

vis_dat(plot_dat)
```
\newpage

# Descriptive Statistics

A summary of the data - including ranges and data types: 

```{r, echo = F}
summary(proj0_dat)
```


```{r, echo = F}
## Make a subset of the data with the relevant cols for Q1 - Q3

# Vector of the cols of interest
cols <- c("SubjectID",
          "Collection Date", "DAYNUMB",
          "Collection Sample",
          "Booket: Clock Time", "MEMs: Clock Time",
          "Sleep Diary reported wake time",
          "Cortisol (nmol/L)", "DHEA (nmol/L)")

# Subset the original data frame
wake_data <- proj0_dat[, colnames(proj0_dat) %in% cols ]


# Convert the Booklet Time to minutes
wake_data$Booklet_Minutes <-
  # Convert to date objects and convert to minutes
  lubridate::hour(hm(wake_data$`Booket: Clock Time`)) * 60 +
  lubridate::minute(hm(wake_data$`Booket: Clock Time`))

# Move this to after Booklet: Clock Time
wake_data <- wake_data %>% 
  relocate(Booklet_Minutes, .after = `Booket: Clock Time`)

# Convert MEM time to minutes
wake_data$MEM_Minutes <-
  # Convert to date objects and convert to minutes
  lubridate::hour(hm(wake_data$`MEMs: Clock Time`)) * 60 +
  lubridate::minute(hm(wake_data$`MEMs: Clock Time`))

# Move this to after Booklet: Clock Time
wake_data <- wake_data %>% 
  relocate(MEM_Minutes, .after = `MEMs: Clock Time`)

# Convert the MEM Time to minutes
wake_data$MEM_C <-
  # Convert to date objects and convert to minutes
  lubridate::hour(hm(wake_data$`MEMs: Clock Time`)) * 60 +
  lubridate::minute(hm(wake_data$`MEMs: Clock Time`))

# Move this to after MEMs: Clock Time
wake_data <- wake_data %>% 
  relocate(MEM_C, .after = `MEMs: Clock Time`)

# Create a subset of the data frame that only has Wake times
# AKA Collection Sample == 1
# wakes <- wake_data[wake_data$`Collection Sample` == 1, ]

```


## Booklet Times

The range of Booklet Times: 

```{r, echo = F}
summary(wake_data$Booklet_Minutes)
```
There are 35 NAs.


There are 35 NAs

```{r, echo = F}
# A plot of the distribution of the outcome
ggplot(na.omit(wake_data), aes(x = Booklet_Minutes,)) +
         geom_histogram(bins = 10, fill = "palevioletred2",
                        col = "black") + 
  theme_lucid() + 
  facet_grid(~ DAYNUMB)
``` 

A log transformation: 

```{r, echo = F}
# A plot of the distribution of the outcome
ggplot(na.omit(wake_data), aes(x = log(Booklet_Minutes))) +
         geom_histogram(bins = 10, fill = "palevioletred2",
                        col = "black") + 
  theme_lucid() + 
  facet_grid(~ DAYNUMB)
```


## MEM Times

A summary of the MEM times

```{r, echo = F}
# Summary of the MEM times
summary(wake_data$MEM_C)
```

We do see that There are 61 NAs.

```{r, echo = F}
ggplot(na.omit(wake_data), aes(x = MEM_C,)) +
         geom_histogram(bins = 10, fill = "seagreen3",
                        col = "black") + 
  theme_lucid() + 
  facet_grid(~ DAYNUMB)
```
Visualizing the outcome by day: 

```{r, echo = F}
ggplot(na.omit(wake_data), aes(x = log(MEM_C))) +
         geom_histogram(bins = 10, fill = "seagreen3",
                        col = "black") + 
  theme_lucid() + 
  facet_grid(~ DAYNUMB)
# Log or square root transformation does not appear to change much
```


The Booklet and MEM distributions appear to agree for the most part. 


## Question 1

What is the agreement between the subject's recording of sampling times compared to the times recorded by an electronic monitoring cap?

For this question, we are only interested in the minutes since Waking Up.

```{r, echo = F, warning = F, message = F}
################################################################################
###                               Question 1                                 ### 
################################################################################

# Visualize the Wake Times from Booklet data
ggplot(wakedata, aes(x = DAYNUMB,
                  y = Booklet_Minutes,
                  group = SubjectID,
                  color = factor(SubjectID))) +
  geom_line() + 
  theme_lucid() + 
  labs(title = "Wake Up Times Across Days - Booklet",
       x = "Day", y = "Booklet Time (Minutes)") + 
  scale_x_continuous(breaks = seq(1, 3, by = 1))

```



```{r, echo = F, message = F, warning = F}
# Visualize the Wake Times from MEMs data
ggplot(wake_data, aes(x = DAYNUMB,
                  y = MEM_C,
                  group = factor(SubjectID),
                  color = factor(SubjectID))) +
  geom_line() + 
  theme_lucid() + 
  labs(title = "Wake Up Times Across Days - MEMs",
       x = "Day", y = "MEM Time (Minutes)") + 
  scale_x_continuous(breaks = seq(1, 3, by = 1))
```

These appear to be very similar but we can see some instances where the MEM failed to record the time. 

```{r, echo = F}
### Model
# Center MEM time for interpretation relative to Mean MEM time
# wakes$MEM_C <- wakes$MEM_Minutes - mean(wakes$MEM_Minutes, na.rm = TRUE)

# LMM with random intercept for subject - Day added
model_day <- lme(Booklet_Minutes ~ MEM_C + factor(DAYNUMB), 
                random = ~ 1 | SubjectID,
                na.action = na.omit,
                method = "ML",
                data = wake_data)

# No day variable in the model
model_noday <- lme(Booklet_Minutes ~ MEM_C, 
                random = ~ 1 | SubjectID,
                na.action = na.omit,
                method = "ML",
                data = wake_data)

# Random intercept and slope


# AIC
# > AIC(model_q1, model_noday)
#             df      AIC
# model_q1     6 495.4461
# model_noday  4 496.3718
# Day slightly better

# ANOVA
# > anova(model_q1, model_noday)
#             Model df      AIC      BIC    logLik   Test  L.Ratio p-value
# model_q1        1  6 495.4461 509.1061 -241.7230                        
# model_noday     2  4 496.3718 505.4785 -244.1859 1 vs 2 4.925757  0.085
# Day does not significanlty improve model fit

# summary(model_q1)

# Obtain 95% CI
# intervals(model_q1)
# (0.9694074, 1.007572)
```

### Diagnostics

```{r,  echo = F, fig.cap="Model Diagnostics", fig.show='hold', out.width="70%", fig.align='center'}
# Residual plot formodel
plot(model_noday, abline = 0)
# 
# QQ plot
qqnorm(resid(model_noday), main = NULL)
qqline(resid(model_noday))
# Adding DAYNUMB improved residual diagnostics
```


```{r,  echo = F, fig.cap="Model Diagnostics", fig.show='hold', out.width="70%", fig.align='center'}
# Residual plot formodel
plot(model_day, abline = 0)
# 
# QQ plot
qqnorm(resid(model_day), main = NULL)
qqline(resid(model_day))
# Adding DAYNUMB improved residual diagnostics
```
For the standardized residual plot, there does not appear to be a violation of heteroscedasticity. 

In the qqplot, we do see some slight deviations in the tails, but not anything concerning. 

Residuals another way: 

```{r, echo = F}
# Obtain data used to fit model
model_data <- getData(model_q1)

# Obtain residuals of model
model_data$resids <- resid(model_q1)
# Obtain fitted values
model_data$fitted <- fitted(model_q1)
```

```{r, echo = F}
# Residual histogram
ggplot(model_data, aes(x = resids)) +
  geom_histogram(bins = 25,
                 fill = "cadetblue2",
                 color = "black") +
  theme_lucid()

```
The residual hist looks symmetric enough given the small sample size. 

```{r, echo = F}
ggplot(model_data, aes(x = fitted, y = resids)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = 2) +
  theme_classic()

```
These look evenly scattered.

```{r, echo = F}
# LMM with random intercept for subject - Day added
model_q1 <- lme(Booklet_Minutes ~ MEM_C + factor(DAYNUMB), 
                random = ~ 1 | SubjectID,
                na.action = na.omit,
                method = "REML",
                data = wake_data)

summary(model_q1)
intervals(model_q1)
```

Intercept: The average wake up time recorded by the MEM was ~829 minutes on the first day. 

Holding day constant, a one-minute increase in MEM-recorded wake time is associated with an approximately one-minute increase in Booklet-recorded wake time ($\beta$ ≈ 0.99; 95% CI: (0.97–1.01); p < 0.0001). This near one-to-one association indicates strong concordance between the two wake-time measures within a given day.
In other words, the subject's recorded wake up time and the MEM's recorded wake up time track each other quite closely. Changes in MEM time are mirrored almost exactly by changes in Booklet time.  

In regard to bias, booklet times tended to be recorded approximately 7-8 minutes earlier (95% CI: (-14.28, -0.62)) than MEM times on average. This difference is statistically significant with p-val $\approx$ 3.

```{r, echo = F}
######## Assess bias
# Create a variable that is the difference between MEM and Booklet times
wake_data <- wake_data %>%
  mutate(diff_minutes = Booklet_Minutes - MEM_Minutes)

# Model to assess bias were difference is the outcome
bias_model <- lme(
  diff_minutes ~ 1 + factor(DAYNUMB),
  random = ~ 1 | SubjectID,
  data = wake_data,
  na.action = na.omit,
  method = "REML"
)

summary(bias_model)
intervals(bias_model)
```

```{r, echo = F, warning = F}

ggplot(wake_data, aes(x = MEM_Minutes, y = diff_minutes)) +
  geom_point(alpha = 0.4) +
  geom_hline(yintercept = 0, col = "red") +
  labs(
    x = "MEM time (minutes since midnight)",
    y = "Booklet − MEM (minutes)",
    title = "Agreement Between Booklet and MEM Times"
  ) + 
  theme_lucid()

```


```{r, echo = F}
# Look at the variance covariance mat
VarCorr(model_q1)

```

The variability is split between the between subject variability (random intercept) and the within-subject variability (residual). 

Intercept (between subject heterogeneity): After accounting for MEM time and day, subjects differ in their mean Booklet wake time with a sd of about 6.6 minutes. 

Residual variance (within subject heterogeneity): Day to day fluctuations within a subject. Given a subject, a day, and a MEM time, individual Booklet observations vary around the subject-specific mean with a sd of about 5 minutes. 

Scatterplot: 

```{r, echo = F}
ggplot(wakes, aes(x = MEM_Minutes, y = Booklet_Minutes,
                           color = factor(DAYNUMB))) + 
  geom_point() + 
  theme_lucid() + 
  geom_abline(intercept = 0, slope = 1, linetype = "solid") +
  labs(color = "Day") +
  labs(title = "Relationship Between Booklet and MEM Wake Up Times",
       x = "MEM Wake Up Time (Minutes)",
       y = "Booklet Wake Up Time (Minutes)")
```

```{r, echo = F}
ggplot(na.omit(wakes), aes(x = MEM_Minutes, y = Booklet_Minutes,
                           color = factor(DAYNUMB))) + 
  geom_point() + 
  theme_lucid() + 
  labs(color = "Day") + 
  facet_grid(~ DAYNUMB) +
  labs(title = "Relationship Between Booklet and MEM Wake Up Times",
       subtitle = "Separated By Day",
       x = "MEM Wake Up Time (Minutes)",
       y = "Booklet Wake Up Time (Minutes)")
```

We can see that there appears to be a strong linear association. There are not any noticeable differences across days or obvious non-linear patterns.
The scatterplot shows a strong linear association between subject recorded time and electronic recording cap wake up time. 
Since each subject has one measure across days, we tested this association with a linear mixed effects model with a random intercept for subject and including fixed day effects. The estimated slope was approximately 1 and highly significant indicating that the two measurements follow each other very closely wihin day.


# Question 2

Are subject's adhering accurately to the +30 min (Collection Sample 2) and +10-hour (600min) (Collection Sample 4) sampling times required by a study protocol?

Good adherence - Percent of samples taken within 7.5 minutes of the scheduled time

Adequate adherence - Percent of samples taken within 15 minutes of scheduled time

```{r, echo = F}
################################################################################
###                               Question 2                                 ### 
################################################################################

## Make a subset of the data with the relevant cols for Q1 - Q3

# Vector of the cols of interest
cols <- c("SubjectID",
          "Collection Date", "DAYNUMB",
          "Collection Sample",
          "Booket: Clock Time", "MEMs: Clock Time",
          "Sleep Diary reported wake time",
          "Cortisol (nmol/L)", "DHEA (nmol/L)")

# Subset the original data frame
wake_data_q2 <- proj0_dat[, colnames(proj0_dat) %in% cols ]

# Convert the Sleep diary Wake time to minutes
wake_data_q2$`Sleep Diary Wake Minutes` <-
  # Convert to date objects and convert to minutes
  lubridate::hour(hm(wake_data_q2$`Sleep Diary reported wake time`)) * 60 +
  lubridate::minute(hm(wake_data_q2$`Sleep Diary reported wake time`))

# Move this to after Collection Sample
wake_data_q2 <- wake_data_q2 %>% 
  relocate(`Sleep Diary Wake Minutes`, 
           .after = `Sleep Diary reported wake time`)

# Convert the Booklet Time to minutes
wake_data_q2$Booklet_Minutes <-
  # Convert to date objects and convert to minutes
  lubridate::hour(hm(wake_data_q2$`Booket: Clock Time`)) * 60 +
  lubridate::minute(hm(wake_data_q2$`Booket: Clock Time`))

# Move this to after Booklet: Clock Time
wake_data_q2 <- wake_data_q2 %>% 
  relocate(Booklet_Minutes, .after = `Booket: Clock Time`)

# Convert the MEM Time to minutes
wake_data_q2$MEM_Minutes <-
  # Convert to date objects and convert to minutes
  lubridate::hour(hm(wake_data_q2$`MEMs: Clock Time`)) * 60 +
  lubridate::minute(hm(wake_data_q2$`MEMs: Clock Time`))

# Move this to after MEMs: Clock Time
wake_data_q2 <- wake_data_q2 %>% 
  relocate(MEM_Minutes, .after = `MEMs: Clock Time`)
```

First we investigate this question with the booklet times. 

```{r, echo = F}
###### With Booklet Times

##### Good adherence - within 7.5 minutes

## Good adherence - within 7.5 minutes of scheduled time

# Df with Wake time, 30 min post wake, and 600 min post wake
wake_check <- wake_data_q2 %>%
  # filter(`Collection Sample` %in% c(1, 2)) %>%
  group_by(SubjectID, DAYNUMB) %>%
  reframe(
    # Wake time
    wake_time = `Sleep Diary Wake Minutes`[`Collection Sample` == 1],
    # The 30 minute time with Booklet
    Book_30time = Booklet_Minutes[`Collection Sample` == 2],
    # The 600 miniute time with booklet
    Book_600time = Booklet_Minutes[`Collection Sample` == 4],
    # The 30 minute time with MEM
    MEM_30time = MEM_Minutes[`Collection Sample` == 2],
    # The 600 minute time with MEM
    MEM_600time = MEM_Minutes[`Collection Sample` == 4]
    )

 
# Define adherence for BOOK
book_wake_check <- wake_check %>%
  mutate(
    # 30 min after waking target time
    target_30time = wake_time + 30,
    # 600 min after waking target time
    target_600time = wake_time + 600,
    # Difference between 30 min later and the recorded time
    diff_min30  = Book_30time - target_30time,
    # Difference between 600 min later and the recorded time
    diff_min600 = Book_600time - target_600time,
    # For 30 min: Define good adherence - within 7.5 minutes
    good_adherent30 = abs(diff_min30) <= 7.5,
    # For 600 min: Define good adherence
    good_adherent600 = abs(diff_min600) <= 7.5,
    # For 30 min: Define adequate adherence - within 15 minutes
    adequate_ad30 = abs(diff_min30) <= 15,
    # For 600 min: Define adequate adherence
    adequate_ad600 = abs(diff_min600) <= 15
  ) %>% 
  relocate(c(target_30time, diff_min30,
             good_adherent30, adequate_ad30), 
           .after = Book_30time) %>% 
  relocate(c(target_600time, diff_min600,
             good_adherent600, adequate_ad600), 
           .after = Book_600time)

# Define adherence for MEM
MEM_wake_check <- wake_check %>%
  mutate(
    # 30 min after waking target time
    target_30time = wake_time + 30,
    # 600 min after waking target time
    target_600time = wake_time + 600,
    # Difference between 30 min later and the recorded time
    diff_min30  = MEM_30time - target_30time,
    # Difference between 600 min later and the recorded time
    diff_min600 = MEM_600time - target_600time,
    # For 30 min: Define good adherence - within 7.5 minutes
    good_adherent30 = abs(diff_min30) <= 7.5,
    # For 600 min: Define good adherence
    good_adherent600 = abs(diff_min600) <= 7.5,
    # For 30 min: Define adequate adherence - within 15 minutes
    adequate_ad30 = abs(diff_min30) <= 15,
    # For 600 min: Define adequate adherence
    adequate_ad600 = abs(diff_min600) <= 15
  ) %>% 
  relocate(c(target_30time, diff_min30,
             good_adherent30, adequate_ad30), 
           .after = MEM_30time) %>% 
  relocate(c(target_600time, diff_min600,
             good_adherent600, adequate_ad600), 
           .after = MEM_600time)

```

## Differences between MEM and Book Time

```{r, echo = F}
# Book time
ggplot(book_wake_check, aes(x = Book_30time)) + 
  geom_histogram(bins = 40, col = "black", fill = "pink") + 
  theme_lucid()

# MEM time
ggplot(MEM_wake_check, aes(x = MEM_30time)) + 
  geom_histogram(bins = 40, col = "black", fill = "hotpink") + 
  theme_lucid()

```

30 Min with Book: 

```{r, echo = F}
# Summary of Book 30 min time
summary(book_wake_check$Book_30time)
```
30 Min with MEMs: 

```{r, echo = F}
# Summary of MEM 30 min time
summary(MEM_wake_check$MEM_30time)
```


```{r, echo = F}
# Book time
ggplot(book_wake_check, aes(x = Book_600time)) + 
  geom_histogram(bins = 40, col = "black", fill = "pink") + 
  theme_lucid()

# MEM time
ggplot(MEM_wake_check, aes(x = MEM_600time)) + 
  geom_histogram(bins = 40, col = "black", fill = "hotpink") + 
  theme_lucid()

```



600 Min with Book: 

```{r, echo = F}
# Summary of Book 30 min time
summary(book_wake_check$Book_600time)
```

600 Min with MEM: 

```{r, echo = F}
# Summary of MEM 30 min time
summary(MEM_wake_check$MEM_600time)
```


## 30 Min Post Wake Time

### Good Adherence 

Booklet: 

Adherence by subject: 

```{r, echo = F}
subject_good_30adherence <- book_wake_check %>%
  group_by(SubjectID) %>%
  summarise(
    # How many days out of 3 were they adherent?
    days_good_adherent = sum(good_adherent30),
    # What is the proportion
    prop_good_adherent = mean(good_adherent30),
    .groups = "drop"
  )
subject_good_30adherence

# How many subjects adhered all 3 days
# sum(subject_good_30adherence$prop_good_adherent == 1, na.rm = T)

# Proportion and Percentage of subjects
# 14/31 = 0.45 or 45%

```



MEM: 

```{r, echo = F}
MEM_subject_good_30adherence <- MEM_wake_check %>%
  group_by(SubjectID) %>%
  summarise(
    # How many days out of 3 were they adherent?
    days_good_adherent = sum(good_adherent30),
    # What is the proportion
    prop_good_adherent = mean(good_adherent30),
    .groups = "drop"
  )
MEM_subject_good_30adherence

# How many subjects adhered all 3 days
# sum(MEM_subject_good_30adherence$prop_good_adherent == 1, na.rm = T)

# Proportion and Percentage of subjects
# 6/31 

```


Adherence overall: 

Booklet

```{r, echo = F}
# Adherence overall
good_book <- book_wake_check %>%
  # group_by(SubjectID) %>%
  summarise(
    # What is the proportion
    prop_good_adherent = round(mean(good_adherent30, na.rm = TRUE), 4),
    # Percentage
    per_good_adherence = (prop_good_adherent * 100)
  )
good_book

```

MEM


```{r, echo = F}
# Adherence overall
good_MEM <- MEM_wake_check %>%
  # group_by(SubjectID) %>%
  summarise(
    # What is the proportion
    prop_good_adherent = round(mean(good_adherent30, na.rm = TRUE), 4),
    # Percentage
    per_good_adherence = (prop_good_adherent * 100)
  )
good_MEM
```


### Adequate Adherence

Adherence by subject with Booket: 

```{r, echo = F}
subject_ad_30adherence <- book_wake_check %>%
  group_by(SubjectID) %>%
  summarise(
    # How many days out of 3 were they adherent?
    days_ad_adherent = sum(adequate_ad30),
    # What is the proportion
    prop_ad_adherent = mean(adequate_ad30),
    .groups = "drop"
  )

subject_ad_30adherence

# How many subjects adhered all 3 days
sum(subject_ad_30adherence$prop_ad_adherent == 1, na.rm = T)

# Proportion and Percentage of subjects
# 20/31 = 0.65 or 65%
```

Adherence by subject with MEM: 

```{r, echo = F}
MEM_subject_ad_30adherence <- MEM_wake_check %>%
  group_by(SubjectID) %>%
  summarise(
    # How many days out of 3 were they adherent?
    days_ad_adherent = sum(adequate_ad30),
    # What is the proportion
    prop_ad_adherent = mean(adequate_ad30),
    .groups = "drop"
  )

MEM_subject_ad_30adherence

# How many subjects adhered all 3 days
sum(MEM_subject_ad_30adherence$prop_ad_adherent == 1, na.rm = T)

# Proportion and Percentage of subjects
# 20/31 = 0.65 or 65%
```



Adherence overall with Booklet: 

```{r, echo = F}
# Adherence overall
adequate_book <- book_wake_check %>%
  summarise(
    # What is the proportion
    prop_adeq_adherent = round(mean(adequate_ad30, na.rm = TRUE), 4),
    # Percentage
    per_adeq_adherence = (prop_adeq_adherent * 100)
  )
adequate_book

```


Adherence overall with MEM: 

```{r, echo = F}
# Adherence overall
adequate_MEM <- MEM_wake_check %>%
  summarise(
    # What is the proportion
    prop_adeq_adherent = round(mean(adequate_ad30, na.rm = TRUE), 4),
    # Percentage
    per_adeq_adherence = (prop_adeq_adherent * 100)
  )
adequate_MEM

```
```{r, echo = F}
# Combine all the proportions into a table to display nicely
props <- data.frame(Good = c(good_book$prop_good_adherent, 
                             good_MEM$prop_good_adherent),
                    Adequate = c(adequate_book$prop_adeq_adherent,
                                 good_MEM$prop_good_adherent))

# Add rownames
rownames(props) <- c("Booklet", "MEM")
  
# Display with Kable
# Display with Kable
kable(
  props,
  format = "latex",
  booktabs = TRUE,
  digits = 5,
  caption = "Statistics for Good and Adequate Adherence"
) |>
  kable_styling(latex_options = c("striped"))
  
```



### Visualization

### Booklet

```{r, echo = F}
# Visualize
ggplot(na.omit(book_wake_check), aes(x = diff_min30)) +
  geom_histogram(binwidth = 2, col = "black", fill = "turquoise") +
  geom_vline(xintercept = c(-7.5, 7.5), col = "red",
             linetype = "dashed") +
  geom_vline(xintercept = c(-15, 15), color = "hotpink",
             linetype = "dashed") +
  theme_lucid() + 
  scale_x_continuous(breaks = seq(0, 50, by = 10)) + 
  scale_y_continuous(breaks = seq(0, 60, by = 10), limits = c(0, 60)) + 
  labs(title = "Distribution of 30 Minutes Post Wake Up Time",
       subtitle = "Red - Good Adherence and Pink - Adequate Adherence",
       x = "Minutes from target (+30)",
       y = "Count") 


```

```{r, echo = F}
# What is the range of diff times
summary(book_wake_check$diff_min30)
```

```{r, echo = F}
# How many observations were greater than 5
sum(book_wake_check$diff_min30 >= 5, na.rm = TRUE)
```

#### MEM

```{r, echo = F}
# Visualize
ggplot(na.omit(MEM_wake_check), aes(x = diff_min30)) +
  geom_histogram(binwidth = 2, col = "black", fill = "turquoise") +
  geom_vline(xintercept = c(-7.5, 7.5), col = "red",
             linetype = "dashed") +
  geom_vline(xintercept = c(-15, 15), color = "hotpink",
             linetype = "dashed") +
  theme_lucid() + 
  scale_x_continuous(breaks = seq(-10, 180, by = 20)) + 
  scale_y_continuous(breaks = seq(0, 10, by = 1), limits = c(0, 10)) + 
  labs(title = "Distribution of 30 Minutes Post Wake Up Time",
       subtitle = "Red - Good Adherence and Pink - Adequate Adherence",
       x = "Minutes from target (+30)",
       y = "Count") 


```

```{r, echo = F}
# What is the range of diff times
summary(MEM_wake_check$diff_min30)
```

```{r, echo = F}
# How many observations were greater than 5
sum(MEM_wake_check$diff_min30 >= 5, na.rm = TRUE)
```


## 600 Min Post Wake Time 

### Good Adherence 

Adherence by subject with Booklet: 

```{r, echo = F}
subject_good_600adherence <- book_wake_check %>%
  group_by(SubjectID) %>%
  summarise(
    # How many days out of 3 were they adherent?
    days_good_adherent = sum(good_adherent600),
    # What is the proportion
    prop_good_adherent = mean(good_adherent600),
    .groups = "drop"
  )
subject_good_600adherence

# How many subjects adhered all 3 days
sum(subject_good_600adherence$prop_good_adherent == 1, na.rm = T)

```

Adherence by subject with MEM: 

```{r, echo = F}
MEM_subject_good_600adherence <- MEM_wake_check %>%
  group_by(SubjectID) %>%
  summarise(
    # How many days out of 3 were they adherent?
    days_good_adherent = sum(good_adherent600),
    # What is the proportion
    prop_good_adherent = mean(good_adherent600),
    .groups = "drop"
  )
MEM_subject_good_600adherence

# How many subjects adhered all 3 days
sum(MEM_subject_good_600adherence$prop_good_adherent == 1, na.rm = T)

```

Adherence overall with booklet: 

```{r, echo = F}
# Adherence overall
book_wake_check %>%
  # group_by(SubjectID) %>%
  summarise(
    # What is the proportion
    prop_good_adherent = round(mean(good_adherent600, na.rm = TRUE), 4),
    # Percentage
    per_good_adherence = (prop_good_adherent * 100)
  )

```

Adherence overall with MEMs: 

```{r, echo = F}
# Adherence overall
MEM_wake_check %>%
  # group_by(SubjectID) %>%
  summarise(
    # What is the proportion
    prop_good_adherent = round(mean(good_adherent600, na.rm = TRUE), 4),
    # Percentage
    per_good_adherence = (prop_good_adherent * 100)
  )

```


### Adequate Adherence


Adherence by subject with Booklet: 

```{r, echo = F}
subject_ad_600adherence <- book_wake_check %>%
  group_by(SubjectID) %>%
  summarise(
    # How many days out of 3 were they adherent?
    days_ad_adherent = sum(adequate_ad600),
    # What is the proportion
    prop_ad_adherent = mean(adequate_ad600),
    .groups = "drop"
  )
subject_ad_600adherence

# How many subjects adhered all 3 days
sum(subject_ad_600adherence$prop_ad_adherent == 1, na.rm = T)

```

Adherence by subject with MEM: 

```{r, echo = F}
MEM_subject_ad_600adherence <- MEM_wake_check %>%
  group_by(SubjectID) %>%
  summarise(
    # How many days out of 3 were they adherent?
    days_ad_adherent = sum(adequate_ad600),
    # What is the proportion
    prop_ad_adherent = mean(adequate_ad600),
    .groups = "drop"
  )
MEM_subject_ad_600adherence

# How many subjects adhered all 3 days
sum(MEM_subject_ad_600adherence$prop_ad_adherent == 1, na.rm = T)

```


Adherence overall with booklet: 

```{r, echo = F}
# Adherence overall
book_wake_check %>%
  summarise(
    # What is the proportion
    prop_adeq_adherent = round(mean(adequate_ad600, na.rm = TRUE), 4),
    # Percentage
    per_adeq_adherence = (prop_adeq_adherent * 100)
  )

```


Adherence overall with MEM: 

```{r, echo = F}
# Adherence overall
MEM_wake_check %>%
  summarise(
    # What is the proportion
    prop_adeq_adherent = round(mean(adequate_ad600, na.rm = TRUE), 4),
    # Percentage
    per_adeq_adherence = (prop_adeq_adherent * 100)
  )

```


### Visualization

#### Booklet

```{r, echo = F}
# Visualize
ggplot(na.omit(book_wake_check), aes(x = diff_min600)) +
  geom_histogram(binwidth = 10, col = "black", fill = "turquoise") +
  geom_vline(xintercept = c(-7.5, 7.5), col = "red",
             linetype = "dashed") +
  geom_vline(xintercept = c(-15, 15), color = "hotpink",
             linetype = "dashed") +
  theme_lucid() + 
  scale_x_continuous(breaks = seq(-50, 360, by = 50)) + 
  # scale_y_continuous(breaks = seq(0, 600, by = 10), limits = c(0, 600)) + 
  labs(title = "Distribution of 600 Minutes Post Wake Up Time",
       subtitle = "Red - Good Adherence and Pink - Adequate Adherence",
       x = "Minutes from target (+600)",
       y = "Count") 


```

```{r, echo = F}
# What is the range of diff times
summary(book_wake_check$diff_min600)
```

```{r, echo = F}
# How many observations were greater than 44
sum(book_wake_check$diff_min600 >= 45, na.rm = TRUE)
```

#### MEM

```{r, echo = F}
# Visualize
ggplot(na.omit(MEM_wake_check), aes(x = diff_min600)) +
  geom_histogram(binwidth = 10, col = "black", fill = "turquoise") +
  geom_vline(xintercept = c(-7.5, 7.5), col = "red",
             linetype = "dashed") +
  geom_vline(xintercept = c(-15, 15), color = "hotpink",
             linetype = "dashed") +
  theme_lucid() + 
  scale_x_continuous(breaks = seq(-50, 360, by = 50)) + 
  # scale_y_continuous(breaks = seq(0, 600, by = 10), limits = c(0, 600)) + 
  labs(title = "Distribution of 600 Minutes Post Wake Up Time",
       subtitle = "Red - Good Adherence and Pink - Adequate Adherence",
       x = "Minutes from target (+600)",
       y = "Count") 


```

```{r, echo = F}
# What is the range of diff times
summary(MEM_wake_check$diff_min600)
```

```{r, echo = F}
# How many observations were greater than 44
sum(MEM_wake_check$diff_min600 >= 45, na.rm = TRUE)
```



## Question 3

```{r, echo = F}
################################################################################
###                               Question 3                                 ### 
################################################################################

##### Create data frame for question 3
## Make a subset of the data with the relevant cols for Q3

# Vector of the cols of interest
cols <- c("SubjectID",
          "Collection Date", "DAYNUMB",
          "Collection Sample",
          "Booket: Clock Time", "MEMs: Clock Time",
          "Sleep Diary reported wake time",
          "Cortisol (nmol/L)", "DHEA (nmol/L)")

# Subset the original data frame
wake_data_q3 <- proj0_dat[, colnames(proj0_dat) %in% cols ]

### Modify data for Person 3029
wake_data_q3 <- wake_data_q3 %>% 
  mutate(DAYNUMB = case_when(
    SubjectID == 3029 & `Collection Date` == "10/7/2018" ~ 1,
    SubjectID == 3029 & `Collection Date` == "10/10/2018" ~ 2,
    SubjectID == 3029 & `Collection Date` == "10/12/2018" ~ 3,
    TRUE ~ DAYNUMB
  )) %>% 
  ungroup()

# Convert the Sleep diary Wake time to minutes
wake_data_q3$`Sleep Diary Wake Minutes` <-
  # Convert to date objects and convert to minutes
  lubridate::hour(hm(wake_data_q3$`Sleep Diary reported wake time`)) * 60 +
  lubridate::minute(hm(wake_data_q3$`Sleep Diary reported wake time`))

# Move this to after Collection Sample
wake_data_q3 <- wake_data_q3 %>% 
  relocate(`Sleep Diary Wake Minutes`, 
           .after = `Sleep Diary reported wake time`)

# Convert the Booklet Time to minutes
wake_data_q3$Booklet_Minutes <-
  # Convert to date objects and convert to minutes
  lubridate::hour(hm(wake_data_q3$`Booket: Clock Time`)) * 60 +
  lubridate::minute(hm(wake_data_q3$`Booket: Clock Time`))

# Move this to after Booklet: Clock Time
wake_data_q3 <- wake_data_q3 %>% 
  relocate(Booklet_Minutes, .after = `Booket: Clock Time`)

# Convert the MEM Time to minutes
wake_data_q3$MEM_Minutes <-
  # Convert to date objects and convert to minutes
  lubridate::hour(hm(wake_data_q3$`MEMs: Clock Time`)) * 60 +
  lubridate::minute(hm(wake_data_q3$`MEMs: Clock Time`))

# Move this to after MEMs: Clock Time
wake_data_q3 <- wake_data_q3 %>% 
  relocate(MEM_Minutes, .after = `MEMs: Clock Time`)

# Create a Minutes SINCE WAKE variable
wake_data_q3 <- wake_data_q3 %>%
  group_by(SubjectID, DAYNUMB) %>%
  mutate(wake_min = Booklet_Minutes[`Collection Sample` == 1][1],
         minutes_since_wake = Booklet_Minutes - wake_min) %>%
  ungroup() %>% 
  relocate(minutes_since_wake, .after = `Sleep Diary Wake Minutes`)



```


### DHEA Exploration

A summary of DHEA: 

```{r, echo = F}
summary(wake_data_q3$`DHEA (nmol/L)`)
```
How many values equal 5.205?

```{r, echo = F}
# Display subjects that have 5.205 DHEA levels
wake_data_q3 %>%
  filter((dplyr::near(`DHEA (nmol/L)`, 5.205) | SubjectID == "3037")) %>% 
  reframe(SubjectID, `DHEA (nmol/L)`)

# There are 2 obs besides 3037 that have DHEA of 5.205
# There are 12 obs for 3037

```
There appears to be one subject that has 4 values of 5.205 for DHEA (nmol/L).

```{r, echo = F}
# Remove these observations
wake_data_q3 <- wake_data_q3 %>%
  filter(
    is.na(`DHEA (nmol/L)`) |
    !(dplyr::near(`DHEA (nmol/L)`, 5.205) | SubjectID == "3037"))

# 14 observations removed - 358 observations left
```


Visualizing DHEA over time: 

```{r, echo = F, message = F, warning=F}
ggplot(wake_data_q3, aes(x = `Collection Sample`,
                      y = `DHEA (nmol/L)`,
                      color = factor(SubjectID))) +
  geom_point() + 
  geom_line() +
  facet_wrap(~DAYNUMB) + 
  theme_lucid()
```

Visualize the outcome:

```{r, echo = F}
# A plot of the distribution of the outcome
ggplot(wake_data_q3, aes(x = `DHEA (nmol/L)`)) +
         geom_histogram(bins = 20, fill = "palevioletred2",
                        col = "black") + 
  theme_lucid() 
```

```{r, echo = F}
# A plot of the distribution of the outcome
ggplot(wake_data_q3, aes(x = log(`DHEA (nmol/L)`))) +
         geom_histogram(bins = 20, fill = "palevioletred2",
                        col = "black") + 
  theme_lucid() 
```

### DHEA Model

```{r, echo = F}
#### Fit the LMM

# Create a logged DHEA variable
wake_data_q3$log_DHEA <- log(wake_data_q3$`DHEA (nmol/L)`)

# Knot at 30 minutes - Collection Sample 2
wake_data_q3 <- wake_data_q3 %>%
  mutate(
    # Slope from waking to 30 miutes
    time_pre30  = pmin(minutes_since_wake, 30),
    # Slope from after 30 minutes
    time_post30 = pmax(minutes_since_wake - 30,
                       0)
  )

# Center Sleep Diary wake minutes
wake_data_q3 <- wake_data_q3 %>% 
  mutate(Sleep_Centered = `Sleep Diary Wake Minutes` - 
           mean(`Sleep Diary Wake Minutes`, na.rm = TRUE),
         Book_MinC = Booklet_Minutes - mean(Booklet_Minutes, 
                                            na.rm = TRUE)) %>% 
  relocate(Sleep_Centered, .after = `Sleep Diary Wake Minutes`)
# Mean wake up time is - 409.4545 minutes


##### Model
model_q3 <- lmer(log_DHEA ~ time_pre30 + time_post30
                 + (1 | SubjectID),
                 REML = TRUE,
                 data = wake_data_q3)

summary(model_q3)
confint(model_q3)

```

```{r, echo = F}
#### Back transform betas 
# Pre 30
a <- -0.01724 * 30 
a_orig <- (exp(a) - 1) * 100 # -38.52871

# Confidence interval
a1 <- (exp(-0.02313 * 30) - 1) * 100 # -71.6346
a2 <- (exp(-0.0113 * 30) - 1) * 100 # -25.92

# Post 30
b <- -0.00153 * 600
b_org <- (exp(b) - 1) * 100 # -37.07332
b1 <- (exp(-0.00183* 600) - 1) * 100 
b2 <- (exp(-0.00122* 600) - 1) * 100 

```


Interpretations: 

Time Pre30: For the time period of wake time to 30 minutes, there is approximately a 40% decrease in DHEA (95% CI: (-50.04%, -28.75%)).
This is statistically significant with p < 0.001.
DHEA shows a statistically significant decrease, not the expected increase, from waking to 30 minutes after waking. 

Time Post30: From 30 minutes to 600 minutes, there is approimately a 60% decrease in DHEA levels (95% CI: (-66.65%, -51.91%)). This is statistically significant with p-val < 0.001. 
As expected, DHEA showed a statistically significant decline over 10 hours. 

Using a piecewise linear mixed-effects model with a knot at 30 minutes post-waking, DHEA concentrations showed a significant early decline - approximately 39% decrease over the first 30 minutes. 
After 30 minutes, DHEA declined more gradually across the day, corresponding to an additional ~60% decrease over 10 hours.

### Model Plots


```{r, echo = F, warning = F, message = F}
t_max <- quantile(wake_data_q3$minutes_since_wake, 0.95, na.rm = TRUE)

pred_grid <- tibble(
  minutes_since_wake = seq(0, as.numeric(t_max), by = 5)) %>%
  mutate(
    time_pre30  = pmin(minutes_since_wake, 30),
    time_post30 = pmax(minutes_since_wake - 30, 0)
  )

# Get fixed effects predictions
pred_grid$fit <- predict(model_q3, newdata = pred_grid, re.form = NA)

# Add a 95% CI
X <- model.matrix(~ time_pre30 + time_post30, data = pred_grid)
V <- vcov(model_q3)
se_fit <- sqrt(diag(X %*% V %*% t(X)))

pred_grid <- pred_grid %>%
  mutate(
    lwr = fit - 1.96 * se_fit,
    upr = fit + 1.96 * se_fit
  )

# Plot
ggplot() +
  geom_point(data = wake_data_q3,
             aes(x = minutes_since_wake, y = log_DHEA),
             alpha = 0.25, size = 1) +
  geom_ribbon(data = pred_grid,
              aes(x = minutes_since_wake, 
                  ymin = lwr,
                  ymax = upr),
              alpha = 0.2,
              fill = "blue") +
  geom_line(data = pred_grid,
            aes(x = minutes_since_wake, y = fit),
            linewidth = 1,
            col = "blue") +
  geom_vline(xintercept = 30, linetype = 2,
             col = "red") +
  labs(x = "Minutes since waking", y = "log(DHEA)",
       title = "Piecewise linear mixed model fit (fixed effects only)",
       subtitle = "Knot at 30 minutes") + 
  theme_lucid()


```



#### Model Diagnostics

```{r, echo = F}
# Obtain data used to fit model
model_dataq3 <- model_q3@frame

# Obtain residuals of model
model_dataq3$resids <- resid(model_q3)

# Obtain fitted values
model_dataq3$fitted <- fitted(model_q3)
```

```{r, echo = F}
# Residual histogram
ggplot(model_dataq3, aes(x = resids)) +
  geom_histogram(bins = 25,
                 fill = "cadetblue2",
                 color = "black") +
  theme_lucid()

```
The residual hist looks symmetric although we do see it has a slight left skew.  

```{r, echo = F}
ggplot(model_dataq3, aes(x = fitted, y = resids)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = 2) +
  theme_classic()

```
These look evenly scattered but we do see an interesting linear pattern along the bottom left. 

These diagnostics are sufficient for our purposes and we will not continue with adjusting the model. 


### Cortisol Exploration

A summary of Cortisol: 

```{r, echo = F}
summary(wake_data_q3$`Cortisol (nmol/L)`)
```

How many measurements are above 80?

```{r, echo = F}
wake_data_q3 %>% 
  filter(`Cortisol (nmol/L)` >= 26) %>% 
  reframe(SubjectID, `Collection Sample`, 
          DAYNUMB,`Cortisol (nmol/L)`)

# Exclude the observation with 89 cortisol
wake_data_q3 <- wake_data_q3 %>%
  filter(!dplyr::near(`Cortisol (nmol/L)`, 89.55714))
  
```


There appears to be one subject that has a value greater than 80. We wll remove this observation for this subject as this is most likely a lab error. 

Visualizing Cortisol over time: 

```{r, echo = F, message = F, warning=F}
ggplot(wake_data_q3, aes(x = `Collection Sample`,
                      y = `Cortisol (nmol/L)`,
                      color = factor(SubjectID))) +
  geom_point() + 
  geom_line() +
  facet_wrap(~DAYNUMB) + 
  theme_lucid()
```


### Cortisol Model

```{r, echo = F}
######### Modeling
##### Model
model_cortisol <- lmer(`Cortisol (nmol/L)` ~ time_pre30 + time_post30
                 + (1 | SubjectID),
                 REML = TRUE,
                 data = wake_data_q3)

summary(model_cortisol)
confint(model_cortisol)

```

```{r, echo = F}
#### Back transform betas 
# Pre 30
c <- 0.014546 * 30 
c_orig <- (exp(c) - 1) * 100 # -

# Confidence interval
c1 <- (exp(-0.0307 * 30) - 1) * 100 # 
c2 <- (exp(0.0599 * 30) - 1) * 100 # 

# Post 30
d <- -0.009 * 600
d_org <- (exp(d) - 1) * 100 # -37.07332
d1 <- (exp(-0.0109* 600) - 1) * 100 
d2 <- (exp(-0.006* 600) - 1) * 100 


```

Interpretations: 

Time Pre30: For the time period of wake time to 30 minutes, there is approximately a 54% increase in Cortisol (95% CI: (-60.19%, 503.15%)).
This is not statistically significant with p = 0.53.
Although there is an increase from waking to 30 minutes after waking, it is not a statistically significant increase.  

Time Post30: From 30 minutes to 600 minutes, there is approimately a 99% decrease in DHEA levels (95% CI: (-99.86%, -97.27%)). This is statistically significant with p-val < 0.001. 
As expected, Cortisol showed a statistically significant decline over 10 hours. 

Cortisol follows the expected pattern of an increase from waking to 30 minutes followed by a decline from 30 minutes and on. 


```{r, echo = F, warning = F, message = F}
t_max <- quantile(wake_data_q3$minutes_since_wake, 0.95, na.rm = TRUE)

pred_grid <- tibble(
  minutes_since_wake = seq(0, as.numeric(t_max), by = 5)) %>%
  mutate(
    time_pre30  = pmin(minutes_since_wake, 30),
    time_post30 = pmax(minutes_since_wake - 30, 0))

# Get fixed effects predictions
pred_grid$fit_cort <- predict(model_cortisol, 
                              newdata = pred_grid,
                              re.form = NA)

# Add a 95% CI
X_cortisol <- model.matrix(~ time_pre30 + time_post30, 
                           data = pred_grid)
V_cortisol <- vcov(model_cortisol)

se_fit_cort <- sqrt(diag(X_cortisol %*% V_cortisol %*% t(X_cortisol)))

pred_grid <- pred_grid %>%
  mutate(
    lwr_cort = fit_cort - 1.96 * se_fit_cort,
    upr_cort = fit_cort + 1.96 * se_fit_cort
  )

# Plot
ggplot() +
  geom_point(data = wake_data_q3,
             aes(x = minutes_since_wake, 
                 y = `Cortisol (nmol/L)`),
             alpha = 0.25, size = 1) +
  geom_ribbon(data = pred_grid,
              aes(x = minutes_since_wake, 
                  ymin = lwr_cort,
                  ymax = upr_cort),
              alpha = 0.2,
              fill = "blue") +
  geom_line(data = pred_grid,
            aes(x = minutes_since_wake, 
                y = fit_cort),
            linewidth = 1,
            col = "blue") +
  geom_vline(xintercept = 30, linetype = 2,
             col = "red") +
  labs(x = "Minutes since waking", y = "Cortisol",
       title = "Piecewise linear mixed model fit (fixed effects only)",
       subtitle = "Knot at 30 minutes") + 
  theme_lucid()


```





