---
title: Power Analysis Plan
author: Diana Felipe
output: 
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    latex_engine: xelatex
    extra_dependencies: ["longtable", "float"]
mainfont: Arial
fontsize: 12pt
documentclass: article
geometry: margin=1in
header-includes:
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage{float}
  - \usepackage{caption}
  - \usepackage{titling}
  - \setlength{\droptitle}{-4cm}
  - \preauthor{\vspace{-0.6cm}}
  - \postauthor{\vspace{-3cm}}
---

```{r, echo = F, warning=F, message = F}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(knitr)
library(ggplot2)
library(MASS)
library(emmeans)
library(multcomp)
library(kableExtra)
library(stats)
library(dplyr)
library(tibble)
library(labelled)
library(tableone)
library(lme4)
library(lmerTest)
library(nlme)
library(gtsummary)
library(gt)
library(ggeffects)
library(UpSetR)
library(visdat)
library(lubridate)
library(see)
library(data.table)
library(gridExtra)
library(kableExtra)
library(here)
library(readr)
library(powertools)
opts_chunk$set(tidy = F)
```

```{r, echo = F}
# Read in data
data <- read_csv(here("Project2", "Data", "PrelimData.csv"),
                 show_col_types = FALSE)


```

<!----------------------------------------------------------------------------->
<!------------------------------------- Methods ------------------------------->
<!----------------------------------------------------------------------------->


\vspace{-10mm}
# Methods
\vspace{-8mm}

<!-- No need to restate hypotheses!! -->
<!-- For Aim 1, we will be testing the hypothesis that higher baseline cytokine
and chemokine levels as well as increases after 1 year can predict decreases in memory consolidation and AD-signature corticol thickness. For Aim 2, we will test the hypothesis that the association of Amyloid Deposition with memory consolidation and AD-signature corticol thickness depends on the levels of cytokines/chemokines. -->
Assessments of raw data obtained for analysis include missingness, outliers, and any other abnormalities to create the final data set for analyses.
To assess the hypotheses of interest for Aims 1 & 2, we will use linear regression models for each combination of the 6 primary outcomes and 4 Primary explanatory variables (PEVs) - 24 models for each Aim. 
The primary outcomes of interest for Aims 1 & 2 consists of the 5 memory consolidation variables as well as gray matter corticol thickness.
The PEVs of interest for both Aims are the changes from baseline to 1 year for the 4 chemokines/cytokines of interest. 
Secondary analyses will be conducted with the remaning 6 chemokines/cytokines.
For Aim 1, the models models will additionally include baseline measures of the outcomes to allow for differences in baseline levels. 
<!-- To allow for baseline differences. -->
For Aim 2, the model will be exactly as the models used for Aim 1 with the difference of the inclusion of an interaction between AD and the given cytokine/chemokine.
We will dichotomize the Amyloid Deposition (AD) variable into AD+ and AD- using the threshold of 1.01 for SUVR from the literature. 
The primary PEV of interest for Aim 2 will be the interaction between significant AD (Yes/No) and the cytokine/chemokine of interest.
This model will additionally include AD as a covariate to assess if AD is associated with the outcomes.
In all primary analyses, we will adjust for age and sex. 
We will adjust for education, cardiovascular markers, 
immune health history variables,
Clinical Dementia Rating, Geriartric Depression Scale, use of Non-Steroidal Anti-Inflammatory drugs, 
routinely collected cognitive measures, and APOE genotype as necessary in secondary analyses.
Alternative approaches to the analysis plan we described considers 6 models that contain all 4 chemokines/cytokines in each model. However, preliminary data results revealed a correlation of ~0.93 between IL-6 and MCP-1 and with the assumption that all 4 PEVs are moderately to highly correlated, including all in the same model would lead to unstable estimates and obscure their individual contributions.

In regards to the all of the available covariates, we will take the following steps to assess inclusion in the final model for each outcome. 
First, we will assess if any of the variables act as confounders or precision variables. Variables identified as confounders will be retained in the final model.
Following this, a full model with the PEVs and candidate covariates will be fit. An omnibus F-test will be conducted to assess whether the model explains variation in the outcome compared to a model with no predictors.
We will then proceed with parital F-tests to assess the contribution of individual variables and groups of variables to the model.
Model fit will additionally be assessed using AIC.
Taking these steps will allow for the most simple and parsimonious model. 
Although we are conducting 24 analyses for each aim, we will adjust significance levels across the 6 primary outcomes. Specifically, we applied a significance threshold of 0.05/6 = 0.0083 for all power analyses for Aims 1 and 2. However, we will consider more appropriate and alternative methods of multiple testing correction when the full data for analysis is obtained. 
In the final report, we will report model estimates, Standard Errors, p-values, and confidence intervals.
A Table 1, relevant tables, and figures will also be included in the final report.
All analyses will be conducted with the R statistical software version 4.4.1

\vspace{-10mm}
## Power Analysis
\vspace{-8mm}

```{r, echo = F, include=F, results='hide'}
# Calculate the correlations from the preliminary data
# Correlations are used in power calculations when measuring the strength
# and direction of a relationship between two continuous variables instead of
# comparing group means

colnames(data)
# CVLT_CNG3 - Measure of episodic memory
#   - Recall/Recognition cognitive outcome variable
# CORT_CNG3 - Measure of corticol thickness
# IL_6 - A novel Cytokine/chemokine of interest
# MCP_1 - A tested cytokine/chemokine of interest

#### Correlations
# Correlation between IL6 and episodic memory
cor(data$IL_6, data$CVLT_CNG3)
# -0.2585617

# Correlation between MCP1 and episodic memory
cor(data$MCP_1, data$CVLT_CNG3)
# -0.3183585

# Correlation between IL6 and corticol thickness
cor(data$IL_6, data$CORT_CNG3)
# -0.5993366

# Correlation between MCP1 and corticol thickness
cor(data$MCP_1, data$CORT_CNG3)
# -0.685273

#### Correlation between cytokines/chemokines
# Correlation between IL6 and MCP1
cor(data$IL_6, data$MCP_1)
# 0.9335558
# - Pretty highly correlated so perhaps we can reduce the number of chemokines/
#cytokines to just 1 instead of examining all 4 which will also address the
# multiple testing issue

### OR JUST DO THIS
cor(data)
```


To determine achievable power for Aim 1 & 2 hypotheses, we obtained a range of correlations from the Preliminary data. Given that we are interested in detecting the the strength and direction of relationships, correlation is an appropriate measure of effect size in these analyses. 
We calculated correlations between IL-6 and MCP-1 with the two outcomes provided in the preliminary data and obtained a range of correlations from approximately -0.2 to -0.7.
For Aim 1, we used the corr.1samp() function from the powertools R package to conduct the power analysis.
We specified N = 175, a correlation coefficient of 0 for the null hypothesis, the aforementioned range of correlations for the alternative hypothesis, and a 2 sided test as we are looking to predict changes in the outcome with an emphasis in declines.
We visualize the results in Figure 1 below. 
We have 80% power to detect a minimum effect size of approximately 0.26 with 175 subjects. We are undepowered to detect any effect sizes


<!----------------------------------------------------------------------------->
<!-------------------------------------- Aim 1 -------------------------------->
<!----------------------------------------------------------------------------->

```{r, echo = F, fig.cap="Power Analysis Results for Aim 1. The vertical dashed blue line is the threshold to achieve 80% power. The red solid line is the effect size at which 80% power is achieved.", out.width="70%", fig.align='center', eval = T}
################################################################################
###                           Aim 1 Power Calculation                        ###
################################################################################

#### Power

# Define the original Vector of correlation values
aim1_r <- seq(0, 0.80, by = 0.02)

# Adjust the vector above based on what we observe with power
# Power reaches 1 after 0.46
aim1_r2 <- seq(0.15, 0.4, by = 0.01)

# Apply the corr.1samp() function to this range of effect sizes
# What are the possible values of power for the range of correlations with a set
# sample size of 175?
aim1_power <- vapply(aim1_r2, function(k) {
  corr.1samp(
    N = 175,
    rho0 = 0,
    rhoA = k,
    alpha = 0.05/6,
    power = NULL,
    sides = 2
  )}, numeric(1)
)

# Create a data frame of the powers
aim1_power_df <- data.frame(`Effect Size` = aim1_r2,
                      `Power` = round(aim1_power, 3),
                      check.names = FALSE)

# Create a table of the power results
# kable(aim1_power_df,
#       format = "latex",
#       booktabs = TRUE) %>% 
#   kable_styling(bootstrap_options = c("striped"))

# Create a plot instead
ggplot(aim1_power_df, 
       aes(x = `Effect Size`, 
           y = Power)) +
  geom_line() +
  geom_point(size = 2) +
  geom_hline(yintercept = 0.80, 
             linetype = "dashed", 
             color = "blue",
             linewidth = 1.2) +
  geom_vline(xintercept = 0.26,
             linetype = "solid",
             color = "red",
             linewidth = 0.3) + 
  theme_lucid() + 
  labs(title = "Effect Size for 80% Power") + 
  scale_y_continuous(breaks = seq(0, 1, by = 0.1)) +
  scale_x_continuous(breaks = seq(0.15, 0.45, by = 0.05)) +
  annotate("text", 
           x = 0.26, 
           y = 0.77, 
           label = "Effect Size = 0.26",
           hjust = -0.1, 
           vjust = 1, 
           angle = 360,
           color = "red")

```

<!----------------------------------------------------------------------------->
<!-------------------------------------- Aim 2 -------------------------------->
<!----------------------------------------------------------------------------->
smaller than 0.26 with the planned sample size.
For Aim 2, we used the corr.2samp() function from the powertools R package to conduct the power analysis.
As we are using a dichotimized AD variable for analysis, we approximated the interaction by comparing the predictor-outcome correlation across the AD positive (AD+) and AD negative (AD-) groups. 
We specified N = 175, varying sample sizes for AD + subjects ranging from 50 to 150 by 25, 
correlations from 0.2 to 0.5 for the AD- group, with additional increments from 0.1 to 0.4 for the AD+ group compared to the AD- group. We expect stronger correlations for the AD+ group.
In Figure 2 we display the results.
Overall, results indicate that higher baseline correlations in the AD− group and larger differences between the AD− and AD+ groups are more likely to achieve 80% power. In contrast, the analysis for Aim 2 appears underpowered to detect smaller differences between the groups. For example, 80% power is not attainable with a correlation of 0.2 for the AD- group and correlations of 0.3 - 0.6 for the AD+ group. We additionally observe that a sample size of 100 AD+ appears to be the optimal number for power across all scenarios. 

On a final note, we emphasize that these power analyses are a simplified version of the analysis plan outlined above as we are not adjusting for any covariates. However, these results can be applied to the analysis because if we are powered for these simplified approaches, we are certainly powered for the more complex and final analyses.

```{r, echo = F, fig.cap="Power analysis results for Aim 2. Power to detect differences in correlation (0.10–0.40) between the AD− and AD+ groups. The vertical dashed line is the threshold for 80% power. The x-axis shows the difference in correlations between groups, and each panel corresponds to a fixed correlation in the AD− group. Larger differences, especially at higher baseline AD- correlations, yield greater power and more easily reach 80%.", out.width="80%", fig.align='center', fig.pos="H", eval = T}
# Define the new vector of correlations
# We Assume that AD + group is hypothesized to have strong correlations

# Correlations for AD - group
adneg_r <- seq(0.2, 0.5, by = 0.1) 

# Differences between AD -  and AD + groups
rho_diff <- seq(0.1, 0.4, by = 0.01)

# Sample sizes for AD + groups
aim2_sampsizes <- seq(50, 150, by = 25)

# Create all combinations of these in a data frame
aim2_powerdf <- expand.grid(
  n1 = aim2_sampsizes,
  rho1 = adneg_r,
  diff = rho_diff
)

# Add a column for n2
aim2_powerdf$n2 <- 175 - aim2_powerdf$n1

# Define rho2 - Correlations for AD + group
aim2_powerdf$rho2 <- aim2_powerdf$rho1 + aim2_powerdf$diff

# Create ratio for corr.2samp
aim2_powerdf$ratio <- aim2_powerdf$n2 / aim2_powerdf$n1

# Calculate power
aim2_powerdf$power <- mapply(function(n1, ratio, r1, r2) {
  corr.2samp(
    n1 = n1,
    n.ratio = ratio,
    rho1 = r1,
    rho2 = r2,
    alpha = 0.05/6,
    power = NULL,
    sides = 2
  )
},
n1 = aim2_powerdf$n1,
ratio = aim2_powerdf$ratio,
r1 = aim2_powerdf$rho1,
r2 = aim2_powerdf$rho2
)


## Visual for report
ggplot(aim2_powerdf, aes(x = diff, 
                         y = power,
                         color = factor(n1))) + 
  geom_line() + 
  theme_lucid() + 
  scale_x_continuous(breaks = seq(0, 0.5, by = 0.1)) +
  scale_y_continuous(breaks = seq(0, 100, by = 0.2)) + 
  geom_hline(yintercept = 0.8, linetype = "dashed", color = "blue") + 
  labs(title = "Power to Detect Differences Between AD+ and AD- Groups", 
       x = "Difference in Correlations",
       y = "Power",
       color = "N for AD +") + 
  facet_wrap( ~ rho1) 

# In this plot, power is on the y-axis and the difference in effect sizes
# Between the AD + and the AD - are on the x-axis.
# Each panel corresponds to The effect size for the AD + group
# The x-axis is the additional change in effect size compared the AD - group.
# Each colored line represents the number of AD + in the total sample.
# We see that we achieve 80% power with 100/175 subjects being AD +, with an 
# Effect size of 0.4 and the AD - group having an effect size of 0.7.

```
\newpage

# Data and Code Availability

The data and statistical code used to generate the results and figures in this report are available in the following GitHub repository:

- GitHub repository: [https://github.com/DianaF65/BIOS6624](https://github.com/DianaF65/BIOS6624)

Directory structure:

- Background/ - A description of the background of the project.

- Code/ 
  - The "Project2.R" file contains the code used for all power analyses.

- Data/ – Raw data files. Because this data is not publicly available, it is not available on GitHub. The raw preliminary data was originally obtained from the BIOS 6624 P2 module. The data is on my local computer at this file path: 
"/Users/dianafelipe/Desktop/BIOS 6624/BIOS6624/Project2/Data/PrelimData.csv"

- Reports/ The "Project2_FinalReport.Rmd" R Markdown file was used to generate this final report of analyses.

All analyses and figures were generated using the R statistical programming language v. 4.4.1, and the repository contains the R scripts
and R Markdown files used to produce this report

