---
title: "\\Large Risk Factors for 10-year Stroke Risk in the Framingham Heart Study"
author: Diana Felipe
output: 
  bookdown::pdf_document2:
    urlcolor: magenta
    toc: false
    number_sections: false
    latex_engine: xelatex
    extra_dependencies: ["longtable", "float"]
    keep_tex: true
mainfont: Arial
fontsize: 12pt
documentclass: article
geometry: margin=1in
header-includes: 
  - \usepackage{setspace} 
  - \doublespacing 
  - \usepackage{float} 
  - \usepackage{caption} 
  - \usepackage{titling} 
  - \setlength{\droptitle}{-3cm} 
  - \setlength{\textfloatsep}{5pt} 
  - \preauthor{\vspace{-0.6cm}} 
  - \postauthor{\vspace{-2.5cm}}
---

```{r, echo = F, warning=F, message = F}
knitr::opts_chunk$set(echo = FALSE)
library(readr)
library(knitr)
library(tidyr)
library(ggplot2)
library(kableExtra)
library(dplyr)
library(tibble)
library(labelled)
library(tableone)
library(nlme)
library(gtsummary)
library(gt)
library(ggeffects)
library(see)
library(gridExtra)
library(regressinator)
library(patchwork)
library(purrr)
library(stringr)
library(survival)
library(survminer)
library(gtsummary)
library(here)
library(expss)
library(labelled)

opts_chunk$set(tidy = F)
```

```{r, echo = F}
# Read in the original data
frame_data <- read_csv(here("Project3", "Data", "frmgham2.csv"),
                       show_col_types = FALSE)

# Read in created survival df
surv_df <- read_csv(here("Project3", "Data", "frmg_survival_df.csv"),
                    show_col_types = FALSE)

# Convert factor vars to factors
surv_df <- surv_df %>% 
  mutate(across(c(RANDID, SEX, DIABETES, 
                  deathwithin10, PREVCHD, BPMEDS, CURSMOKE), 
                as.factor))
# Make sure STROKE10 is numeric
surv_df$STROKE10 <- as.numeric(as.character(surv_df$STROKE10))

```

# Introduction
\vspace{-6mm}

The Framington Heart Study is a longitudinal and prospective study with the overarching goal of investigating the etiology of cardiovascular disease in the community of Framingham, Massachusetts. The study was conducted on an initial population of 5,209 individuals on which data was collected for health behaviors, medication use, clinical measurements such as blood pressure, and more. The primary goal of this analysis is to identify risk factors of stroke within a 10-year period for five different risk profiles of interest: 1.) Subjects with no risk factors for stroke, 2.) Subjects with high blood pressure (BP) (> 60 (mmHg) ), 3.) Subjects with diabetes, 4.) Subjects with high blood pressure and diabetes, 5.) Subjects who were smokers. All risk factors are measured at baseline. Additionally, we summarize the change in the risk factors over the whole duration of the study period to assess whether a more complex time-varying covariate analysis is warranted. 
This report covers the preliminary findings of our analysis answering the primary question of interest. It fills in the results from the data analysis plan agreed upon in the Data Analysis Plan report.

\vspace{-10mm}
# Methods

\vspace{-6mm}
## Data Cleaning
\vspace{-6mm}

The raw data contained 11,627 observations for 4,434 subjects and 38 variables. Missingness was assessed in the whole data set thruogh summaries and visualizations. Regarding the primary variables of interest, BP Medication had ~5% missginness, total cholesterol had 3.5% missingness, and BMI had ~0.5% missingness. We verified that there were no relationships between missingness and any variables and therefore concluded that data was Missing Completely at Random.
We created new variables specifically for the 10-year window of interest for the primary analysis. We created an indicator variable for whether subjects had a stroke with 10 years (Yes/No) as well as the time to stroke for these events. 
The data for sex-specific analyses consisted of 2,378 observations for females and 1,897 observationsfor males. We used a complete case data frame, removing any rows with missing data. 

\vspace{-8mm}
## Data Analysis
\vspace{-6mm}

We performed a preliminary descriptive analysis of survival curves by sex with visualizations of the Kaplan Meier curves before fitting the primary Cox model.
The preliminary Kaplan-Meier analysis reveled frequent crossing between the separate male and female survival curves. This indicated a violation of the proportional hazards assumption and therefore necessitates stratified Cox models to obtain the sex-specific risk differences. 
To identify risk factors that are associated with stroke in a 10-year period, we conducted a Survival analysis with stroke as the event and days to stroke as the time to event. Specifically, we fit a Cox Proportional Hazards (PH) Survival Model to identify important risk factors of stroke and obtain the 10-year probability of stroke based on the five risk profiles. 
Age, Diabetes status, and Systolic Blood Pressure (BP) were included in all models for all analyses. 
The secondary variables of interest were presence of Coronary Heart Disease (CHD), use of anit-hypertensive medication, smoking status, total cholesterol, and BMI. 
We performed LASSO variable selection using cross validation for the secondary variables separately for each sex. 
For males, use of antihypertensive medication as well as smoker status was retained, suggesting these were the strongest predictors of stroke risk among the variables considered.
For females, no additional variables besides the primary variables that were included in the analysis were retained using LASSO. 
For the primary analysis, we kept the LASSO selected variables specific to each sex. 
In regard to model assumptions, Schoenfeld residiual test for males and females indicated no evidence of violation of the proportional hazards assumption for individual covariates for globally (global p = 0.714).
The LML Plot for Diabetes status and use of antihypertensive medication visually satisfied the PH assumption for categorical variables as the curves for the different strata were parallel. 
Lastly, we created a table describing how variables such as Systolic BP and BMI change over over time. 
We used a significance level of $\alpha = 0.05$ for the significance threshold. All analyses were conducted with R version 4.1.1 using the glmnet, survival, and survminer packages. 

\vspace{-8mm}
# Results
\vspace{-6mm}

<!------------------------------------ Table 1 -------------------------------->

```{r, echo = F, fig.pos='H'}
##################################### Table 1 ##################################

# Fix levels for categorical variables for Table1
# Data frame for table1
tab1_data <- surv_df

# Sex
tab1_data$factor_SEX <- factor(tab1_data$SEX,
                        levels = c(1, 2),
                        labels = c("Male", "Female"))

# Diabetes
tab1_data$factor_DIABETES <- factor(tab1_data$DIABETES,
                                    levels = c(0, 1),
                                    labels = c("Not a Diabetic", "Diabetic"))

# Prevchd
tab1_data$factor_PREVCHD <- factor(tab1_data$PREVCHD,
                                   levels = c(0, 1),
                                   labels = c("Free of CHD",
                                              "Prevalent CHD"))

# Cursmoke
tab1_data$factor_CURSMOKE <- factor(tab1_data$CURSMOKE,
                             levels = c(0, 1),
                             labels = c("Not a Current Smoker",
                                        "Current Smoker"))

# BPMeds
tab1_data$factor_BPMEDS <- factor(tab1_data$BPMEDS,
                                  levels = c(0, 1),
                                  labels = c("No Use of Medication",
                                             "Use of Medication"))

# Stroke 10
tab1_data$factor_STROKE <- factor(tab1_data$STROKE10,
                                  levels = c(0, 1),
                                  labels = c("No Stroke",
                                             "Stroke"))



# Relabel the variables for the Table 1
tab1_data <- apply_labels(tab1_data,
                     factor_STROKE = "Stroke Within 10 Yrs",
                     factor_SEX = "Sex",
                     factor_BPMEDS = "Use of Anti-hypertensive Medication",
                     factor_PREVCHD = "Prevalent Coronary Heart Disease",
                     factor_CURSMOKE = "Cigarrete Smoking Status",
                     AGE = "Age",
                     SYSBP = "Systolic Blood Pressure (mmHg)",
                     STROKE10TIME = "Time to Stroke (Days)",
                     TOTCHOL = "Serum Total Cholesterol (mg/dL)",
                     BMI = "Body Mass Index"
                     )

# Display the modified Table1
tab1_data %>% 
  tbl_summary(include = c(factor_STROKE,
                          STROKE10TIME,
                          AGE,
                          factor_BPMEDS,
                          factor_PREVCHD,
                          factor_CURSMOKE,
                          SYSBP,
                          TOTCHOL,
                          BMI),
              by = factor_SEX,
              statistic = list(c(all_continuous()) 
                               ~ "{mean}({sd})")) %>% 
  modify_header(label = "**Sex**") %>% 
  bold_labels() %>% 
  modify_caption(caption = "Patient Characteristics At Baseline")
  
  
  # as_kable_extra(booktabs = TRUE, linesep = "") %>%
  # kableExtra::kable_styling(latex_options = c("hold_position"))
```

Table 1 contains a description of the data used for all analyses. 
The data is stratified by sex, n and \% of patients is displayed for the categorical variables, and the mean and standard deviation is shown for each continuous variable.
We note the small number of subjects across both sexes that had a stroke within the 10 year time span. In addition, the use of Anti-hypertensive medication and prevalent coronary heart disease groups are noticeably unbalanced. 
Table 2 displays the Hazard Ratios (HRs), 95\% Confidence Intervals (CIs), and p-values for the sex-specific models.
Among females, Age and Systolic BP were significantly associated with a higher risk of stroke. 
Each additional year in age

<!----------------------------------------------------------------------------->
<!--------------------------------- KM Curves --------------------------------->
<!----------------------------------------------------------------------------->


```{r, echo = F, eval = F, results='hide'}
################################ Prelim Cox Analysis ###########################
# Create some Kaplan Meier curves by covariates

### Sex
# Fit model
km_sex <- survfit(Surv(STROKE10TIME, STROKE10) ~ SEX,
                   data = surv_df)

# Plot KM curves
ggsurvplot(km_sex, 
           conf.int = FALSE,
           censor = FALSE,
           xlim = c(0, max(surv_df$STROKE10TIME)),
           ylim = c(0.95, 1))
# For the most part, seems that survival curves are quite similar between sexes
# We do see some crossing of the curves though

```


```{r, echo = F, results='hide', warning=F, message=F, eval = F}
# LML Plots
ggsurvplot(km_sex,
           conf.int = F, 
           censor = F,
           fun="cloglog")
# We see that the lines are not parallel and this suggests that the PH assumption is not met
```

<!----------------------------------------------------------------------------->
<!------------------------------ Variable Selection --------------------------->
<!----------------------------------------------------------------------------->

```{r, echo = F, include=FALSE}
############################## Variable Selection ##############################
library(glmnet)
# Vignette: https://glmnet.stanford.edu/articles/Coxnet.html

#----------------------------------- Males ------------------------------------#

# Perform some variable selection prevchd, bpmeds, smokecur, total chol, and BMI

# Males only df
males_df <- surv_df %>% 
  filter(SEX == 1)

# Fit model
y_m <- Surv(time = males_df$STROKE10TIME, 
            event = males_df$STROKE10)

# Create the model matrix for glmnet
x_m <- model.matrix(~ AGE + DIABETES + SYSBP
  + PREVCHD + BPMEDS + CURSMOKE + TOTCHOL + BMI,
  data = males_df)[, -1]  # Remove the intercept

# Lasso - cross validation
cv_m <- cv.glmnet(x_m, y_m, 
                  family = "cox", 
                  alpha = 1,
                  # Force the first 3 variables to be kept
                  penalty.factor = c(0, 0, 0, 
                                     rep(1, ncol(x_m) - 3)))

# Plot to see the optimal lambda value
plot(cv_m)
# Looks pretty flat across all top and bottom values

# Which variables were kept with 1se
## The largest lambda within 1 SE of the minimum error 
## More penalty -> simplier model, more stable and more conservaitive,
## Provides the best simple model!

coef_1se_males <- coef(cv_m, s = "lambda.1se")
# AGE       0.05808631
# DIABETES1 1.39157280
# SYSBP     0.03271646
# PREVCHD1  .         
# BPMEDS1   .         
# CURSMOKE1 .         
# TOTCHOL   .         
# BMI       .  

# Get the nonzero variables
m_1se <- which(as.vector(coef_1se_males) != 0)
m_selected_1se <- rownames(coef_1se_males)[m_1se]
# "AGE"       "DIABETES1" "SYSBP" 
## None of the predictors add enough predictve value to justify including them

# Which variables were kept with lambda min
## The value of lambda that gives the lowest corss-validated error
## More flexble model -> keeps more variables
## Higher risk of overfitting though
## Provides the best fit
coef_min_males <- coef(cv_m, s = "lambda.min")
#                    1
# AGE       0.06165026
# DIABETES1 1.50155634
# SYSBP     0.03236087
# PREVCHD1  .         
# BPMEDS1   0.30363247
# CURSMOKE1 0.35060978
# TOTCHOL   .         
# BMI       .

# Get the nonzero variables
m_min <- which(as.vector(coef_min_males) != 0)
m_selected_min <- rownames(coef_min_males)[m_min]
# [1] "AGE"       "DIABETES1" "SYSBP"     "BPMEDS1"   "CURSMOKE1"

### Based on minimum, keep these vars for males - the only variable with meaningful
### predictive signal for males
```


```{r, echo = F, include=FALSE}
############################## Variable Selection ##############################
library(glmnet)
# Vignette: https://glmnet.stanford.edu/articles/Coxnet.html

#----------------------------------- Females ----------------------------------#

# Perform some variable selection prevchd, bpmeds, smokecur, total chol, and BMI

# Females only df
females_df <- surv_df %>% 
  filter(SEX == 2)

# Fit model
y_f <- Surv(time = females_df$STROKE10TIME, 
            event = females_df$STROKE10)

# Create the model matrix for glmnet
x_f <- model.matrix( ~ AGE + DIABETES + SYSBP 
  +  PREVCHD + BPMEDS + CURSMOKE + TOTCHOL + BMI,
  data = females_df)[, -1]  # Remove the intercept

# Lasso - cross validation
cv_f <- cv.glmnet(x_f, y_f, 
                  family = "cox", 
                  alpha = 1,
                  # Force the first 3 variables to be kept
                  penalty.factor = c(0, 0, 0, 
                                     rep(1, ncol(x_f) - 3)))

# Plot to see the optimal lambda value
plot(cv_f)
# Looks pretty flat across as well all top and bottom values

# Which variables were kept with 1se
## The largest lambda within 1 SE of the minimum error 
## More penalty -> simplier model, more stable and more conservaitive,
## Provides the best simple model!

coef_1se_females <- coef(cv_f, s = "lambda.1se")
#                    1
# AGE       0.07113834
# DIABETES1 0.56976402
# SYSBP     0.02570976
# PREVCHD1  .         
# BPMEDS1   .         
# CURSMOKE1 .         
# TOTCHOL   .         
# BMI       .

# Get the nonzero variables
f_1se <- which(as.vector(coef_1se_females) != 0)
f_selected_1se <- rownames(coef_1se_females)[f_1se]
# "AGE"       "DIABETES1" "SYSBP" 
## None of the predictors add enough predictve value to justify including them

# Which variables were kept with lambda min
## The value of lambda that gives the lowest corss-validated error
## More flexble model -> keeps more variables
## Higher risk of overfitting though
## Provides the best fit
coef_min_females <- coef(cv_f, s = "lambda.min")
#                    1
# AGE       0.07113834
# DIABETES1 0.56976402
# SYSBP     0.02570976
# PREVCHD1  .         
# BPMEDS1   .         
# CURSMOKE1 .         
# TOTCHOL   .         
# BMI       .

# Get the nonzero variables
f_min <- which(as.vector(coef_min_females) != 0)
f_selected_min <- rownames(coef_min_females)[f_min]
# "AGE"       "DIABETES1" "SYSBP" 

### Based on minimum, keep PRECHD, BPMEDS, TOTCHOL, and BMI
### Remove CURSMOKE
### The only variable with meaningful predictive signal for females
```


<!----------------------------------------------------------------------------->
<!----------------------------------- FEMALES --------------------------------->
<!----------------------------------------------------------------------------->

```{r, echo = F, warning=F, message=F}
################################## Model Fitting ###############################

##################################### Females ##################################

# Model for females
females_fit <- coxph(Surv(STROKE10TIME, STROKE10) ~ AGE + DIABETES 
                     + SYSBP,
                     data = females_df
)

# Summary of general model
females_sum1 <- summary(females_fit)
```


```{r, echo = F, results='hide', eval = F}
################################ Test PH assumptions ##########################
## Schoenfeld residuals :Do the effects of each covariate change over time?
## H0: PH holds - no time varying effect
## H1: Violation of PH

# Schoenfeld residuals
females_res <- cox.zph(females_fit)

# Plot
ggcoxzph(females_res,
         se = F)

```

```{r, echo = F, results='hide', eval = F}
# LML Plots for each categorical variable

# Fit model for DIABETES
females_diabetes <- survfit(coxph(Surv(STROKE10TIME, STROKE10) ~
                                    strata(DIABETES),
                                  data = females_df))
# LML Plot
plot(females_diabetes, 
     fun = "cloglog", 
     col = c("red", "blue"), 
     main = "LML Plot for Diabetes",
     xlab = "Log(Time)", 
     ylab = "Log(-Log(Survival))")
# Assumption met for Sex

# Fit model for PREVCHD
females_prevchd <- survfit(coxph(Surv(STROKE10TIME, STROKE10) ~ 
                                   strata(PREVCHD),
                           data = females_df))

# LML Plot - no need actually 
plot(females_prevchd, 
     fun = "cloglog", 
     col = c("red", "blue"), 
     main = "LML Plot for PREVCHD",
     xlab = "Log(Time)", 
     ylab = "Log(-Log(Survival))")
# Assumption met for prevchd

# Fit Model for BPMEDS
females_bpmeds <- survfit(coxph(Surv(STROKE10TIME, STROKE10) ~ 
                                  strata(BPMEDS),
                          data = females_df))

# LML Plot - no neede actually 
plot(females_bpmeds, 
     fun = "cloglog", 
     col = c("red", "blue"), 
     main = "LML Plot for BPMEDS",
     xlab = "Log(Time)", 
     ylab = "Log(-Log(Survival))")
# Assumption met for bpmeds
```


```{r, echo = F}
##### Create data frame of results
females_table <- females_fit %>% 
  tbl_regression(intercept = TRUE, 
                 conf.level = .95,
                 exponentiate = TRUE,
                 pvalue_fun = label_style_pvalue(digits = 2),
                 conf.int = TRUE,
                 show_single_row = c(DIABETES, SYSBP),
                 label = list(
                   AGE ~ "Age",
                   SYSBP ~ "Systolic BP",
                   DIABETES ~ "Diabetes Status"))
```


```{r, echo = F, eval = F, results='hide', warning=F, message=F}
# Plot survival curve - hmmm
ggsurvplot(survfit(females_fit),
           data = females_df,
           censor = F,
           conf.int = T,
           ylim = c(0.95, 1))
```



```{r, echo = F, results='hide'}
################################# Risk Profile 1 ###############################

# Create data frame for the different risk profiles
# Risk profile 1: 
## Average age, no diabetes, average sysbp, no prechd, no bpmeds, no cursmoke,
## average totchol, average bmi
females_rp1 <- data.frame(
    AGE = c(40, 50, 60),
    DIABETES = factor(rep(0, 3), 
                      levels = levels(females_df$DIABETES)),
    SYSBP = rep(mean(females_df$SYSBP), 3),
    CURSMOKE = factor(rep(0, 3), 
                    levels = levels(females_df$CURSMOKE))
)
#   AGE DIABETES    SYSBP CURSMOKE
# 1  40        0 133.5524        0
# 2  50        0 133.5524        0
# 3  60        0 133.5524        0

# Fit cox model for risk profile 1
females_fit_riskprofiles <- coxph(Surv(STROKE10TIME, STROKE10) ~ 
                                    AGE + DIABETES + SYSBP + CURSMOKE,
                                  data = females_df
)
        
# Use survfit                          
frp1_fit <- survfit(females_fit_riskprofiles, newdata = females_rp1)


# Estimates for 10 years
frp1_summary <- summary(frp1_fit, times = 365*10)
# time n.risk n.event survival1 survival2 survival3
#  3650   2177      57     0.994     0.988     0.975

# Data frame of survival 10 years and risk 10 years
frp1_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(frp1_summary$surv), 3),
                           Risk = paste0(round(1 - frp1_summary$surv, 3)*100,
                                         "%")
)

#   Age Survival Risk
# 1  40    0.994 0.6%
# 2  50    0.988 1.2%
# 3  60    0.975 2.5%
```


```{r, echo = F, results='hide'}
################################# Risk Profile 2 ###############################

# Risk profile 2: 
## High blood pressure and no other risk factors
females_rp2 <- data.frame(
  AGE = c(40, 50, 60),
  DIABETES = factor(rep(0, 3), 
                    levels = levels(females_df$DIABETES)),
  SYSBP = rep(200, 3),
  CURSMOKE = factor(rep(0, 3), 
                    levels = levels(males_df$CURSMOKE))
)

#   AGE DIABETES SYSBP CURSMOKE
# 1  40        0   200        0
# 2  50        0   200        0
# 3  60        0   200        0


# Fit cox model for risk profile 2
frp2_fit <- survfit(females_fit_riskprofiles, newdata = females_rp2)

# Estimates for 10 yearsf
frp2_summary <- summary(frp2_fit, times = 365*10)
 # time n.risk n.event survival1 survival2 survival3
 # 3650   2177      57     0.969     0.935     0.867

# Create data frame of survival 10 years and risk 10 years
frp2_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(frp2_summary$surv), 3),
                           Risk = paste0(round(1 - frp2_summary$surv, 3)*100,
                                         "%")
)
#   Age Survival  Risk
# 1  40    0.969  3.1%
# 2  50    0.935  6.5%
# 3  60    0.867 13.3%

```


```{r, echo = F, results='hide'}

################################# Risk Profile 3 ###############################

#### Risk profile 3:
## Diabetes and no other risk factors
females_rp3 <- data.frame(
  AGE = c(40, 50, 60),
  DIABETES = factor(rep(1, 3), 
                    levels = levels(females_df$DIABETES)),
  SYSBP = rep(mean(females_df$SYSBP), 3),
  CURSMOKE = factor(rep(0, 3), 
                    levels = levels(females_df$CURSMOKE))
)
#   AGE DIABETES    SYSBP CURSMOKE
# 1  40        1 133.5524        0
# 2  50        1 133.5524        0
# 3  60        1 133.5524        0



# Fit cox model for risk profile 3
frp3_fit <- survfit(females_fit_riskprofiles, newdata = females_rp3)

# Estimates for 10 years
frp3_probs <- summary(frp3_fit, times = 365*10)
 # time n.risk n.event survival1 survival2 survival3
 # 3650   2177      57      0.99     0.979     0.955


# Estimates for 10 years
frp3_summary <- summary(frp3_fit, times = 365*10)
 # time n.risk n.event survival1 survival2 survival3
 # 3650   2177      57      0.99     0.979     0.955

# Create data frame of survival 10 years and risk 10 years
frp3_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(frp3_summary$surv), 3),
                           Risk = paste0(round(1 - frp3_summary$surv, 3)*100,
                                         "%")
)

#   Age Survival Risk
# 1  40    0.988 1.2%
# 2  50    0.976 2.4%
# 3  60    0.951 4.9%

```


```{r, echo = F, results='hide'}
################################# Risk Profile 4 ###############################

#### Risk profile 4:
## Diabetes AND high BP
females_rp4 <- data.frame(
  AGE = c(40, 50, 60),
  DIABETES = factor(rep(1, 3), 
                    levels = levels(females_df$DIABETES)),
  SYSBP = rep(200, 3),
  CURSMOKE = factor(rep(0, 3), 
                    levels = levels(females_df$CURSMOKE))
)
# 
#   AGE DIABETES    SYSBP CURSMOKE
# 1  40        1 133.5524        0
# 2  50        1 133.5524        0
# 3  60        1 133.5524        0

# Fit cox model for risk profile 3
frp4_fit <- survfit(females_fit_riskprofiles, newdata = females_rp4)

# Estimates for 10 years
frp4_summary <- summary(frp4_fit, times = 365*10)
# time n.risk n.event survival1 survival2 survival3
# 3650   2177      57      0.99     0.979     0.955

# Create data frame of survival 10 years and risk 10 years
frp4_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(frp4_summary$surv), 3),
                           Risk = paste0(round(1 - frp4_summary$surv, 3)*100,
                                         "%")
)

# Age Survival  Risk
#   Age Survival  Risk
# 1  40    0.945  5.5%
# 2  50    0.886 11.4%
# 3  60    0.772 22.8%

```


```{r, echo = F, results='hide'}
################################# Risk Profile 5 ###############################

#### Risk profile 4:
## Smoker and no other risk factors
females_rp5 <- data.frame(
  AGE = c(40, 50, 60),
  DIABETES = factor(rep(0, 3), 
                    levels = levels(females_df$DIABETES)),
  SYSBP = rep(mean(females_df$SYSBP), 3),
  CURSMOKE = factor(rep(1, 3), 
                    levels = levels(females_df$CURSMOKE))
)
# 
#  AGE DIABETES    SYSBP CURSMOKE
# 1  40        0 133.5524        1
# 2  50        0 133.5524        1
# 3  60        0 133.5524        1

# Fit cox model for risk profile 3
frp5_fit <- survfit(females_fit_riskprofiles, newdata = females_rp5)

# Estimates for 10 years
frp5_summary <- summary(frp5_fit, times = 365*10)
 # time n.risk n.event survival1 survival2 survival3
 # 3650   2177      57     0.992     0.983     0.964

# Create data frame of survival 10 years and risk 10 years
frp5_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(frp5_summary$surv), 3),
                           Risk = paste0(round(1 - frp5_summary$surv, 3)*100,
                                         "%")
)

# Age Survival Risk
#   Age Survival Risk
# 1  40    0.992 0.8%
# 2  50    0.983 1.7%
# 3  60    0.964 3.6%
```


```{r, echo = F}
################### Combine these all for the report
library(knitr)
library(kableExtra)

# Create individual kables for each df
# Results 1
fk1 <- kable(frp1_results, caption = "meow1",
             format = "html") %>% 
  kable_styling()

# Results 2
fk2 <- kable(frp2_results, caption = "meow2",
             format = "html") %>% 
  kable_styling()

# Results 3
fk3 <- kable(frp3_results, caption = "meow3",
             format = "html") %>% 
  kable_styling()

# Results 4
fk4 <- kable(frp4_results, caption = "meow4",
             format = "html") %>% 
  kable_styling()

# Combimne results into one
female_results <- cbind(frp1_results, 
                       frp2_results[, 2:3],
                       frp3_results[,2:3],
                       frp4_results[,2:3],
                       frp5_results[,2:3]
                       )

```


<!----------------------------------------------------------------------------->
<!------------------------------------ MALES ---------------------------------->
<!----------------------------------------------------------------------------->


```{r, echo = F, warning=F, message=F, fig.cap="Model Results for Males"}
################################## Model Fitting ###############################

##################################### Males ##################################
# Model for males
males_fit <- coxph(
  Surv(STROKE10TIME, STROKE10) ~ AGE + DIABETES + SYSBP
  + BPMEDS + CURSMOKE,
  data = males_df
)

```

```{r, echo = F, eval = F, results='hide', warning=F, message=F}
# Plot survival curve - hmmm
ggsurvplot(survfit(males_fit),
           data = males_df,
           censor = F,
           conf.int = T)
```


```{r, echo = F, include=FALSE, results='hide'}
################################ Test PH assumptions ##########################
## Schoenfeld residuals :Do the effects of each covariate change over time?
## H0: PH holds - no time varying effect
## H1: Violation of PH

# Schoenfeld residuals
males_res <- cox.zph(males_fit)
#          chisq df    p
# AGE      0.467  1 0.49
# DIABETES 2.636  1 0.10
# SYSBP    1.653  1 0.20
# BPMEDS   0.101  1 0.75
# GLOBAL   4.204  4 0.38

# Plot
ggcoxzph(males_res,
         se = F)
# Looks good

```


```{r, echo = F, results='hide', eval = F}
# LML Plots for each categorical variable

# Fit model for DIABETES
males_diabetes <- survfit(coxph(Surv(STROKE10TIME, STROKE10) ~
                                    strata(DIABETES),
                                  data = males_df))
# LML Plot
plot(males_diabetes, 
     fun = "cloglog", 
     col = c("red", "blue"), 
     main = "LML Plot for Diabetes",
     xlab = "Log(Time)", 
     ylab = "Log(-Log(Survival))")
# Assumption met for Sex

# Fit model for PREVCHD - no
males_prevchd <- survfit(coxph(Surv(STROKE10TIME, STROKE10) ~ 
                                   strata(PREVCHD),
                           data = males_df))

# LML Plot
plot(males_prevchd, 
     fun = "cloglog", 
     col = c("red", "blue"), 
     main = "LML Plot for PREVCHD",
     xlab = "Log(Time)", 
     ylab = "Log(-Log(Survival))")
# Assumption met for prevchd

# Fit Model for BPMEDS - Yes
males_bpmeds <- survfit(coxph(Surv(STROKE10TIME, STROKE10) ~ 
                                  strata(BPMEDS),
                          data = males_df))

# LML Plot
plot(males_bpmeds, 
     fun = "cloglog", 
     col = c("red", "blue"), 
     main = "LML Plot for BPMEDS",
     xlab = "Log(Time)", 
     ylab = "Log(-Log(Survival))")
# Assumption met for bpmeds
```


```{r, echo = F}

# Create data frame of results
males_table <- males_fit %>% 
    tbl_regression(intercept = TRUE, 
                 conf.level = .95,
                 exponentiate = TRUE,
                 pvalue_fun = label_style_pvalue(digits = 2),
                 conf.int = TRUE,
                 show_single_row = c(SYSBP, BPMEDS, DIABETES, CURSMOKE),
                 label = list(
                   AGE ~ "Age",
                   SYSBP ~ "Systolic BP",
                   DIABETES ~ "Diabetes Status",
                   BPMEDS ~ "BP Medication",
                   CURSMOKE ~ "Smoker Status")) 
```

\newpage

```{r, echo = F}

tbl_side_by_side <- tbl_merge(
  tbls = list(females_table, males_table),
  tab_spanner = c("**Females**", "**Males**"),
  quiet = TRUE
)

tbl_side_by_side %>%
  modify_caption("Model Results by Sex") %>%
  as_gt() %>%
  gt::tab_source_note(
    gt::md("*BP = Blood Pressure; CHD = Coronary Heart Disease")
  ) %>%
  gt::tab_source_note(
    gt::md("**All factor variable HRs are compared to reference groups")
  ) %>%
  gt::tab_options(
    table.font.size = "small",
    data_row.padding = gt::px(2)
  )
```

```{r, echo = F, results='hide'}
################################# Risk Profile 1 ###############################

# Create data frame for the different risk profiles
# Risk profile 1: 
## Average age, no diabetes, average sysbp, no prechd, no bpmeds, no cursmoke,
## average totchol, average bmi
males_rp1 <- data.frame(
    AGE = c(40, 50, 60),
    DIABETES = factor(rep(0, 3), 
                      levels = levels(males_df$DIABETES)),
    SYSBP = rep(mean(males_df$SYSBP), 3),
    BPMEDS = factor(rep(0, 3), 
                    levels = levels(males_df$BPMEDS)),
    CURSMOKE = factor(rep(0, 3), 
                      levels = levels(males_df$CURSMOKE))
)
#   AGE DIABETES    SYSBP BPMEDS CURSMOKE
# 1  40        0 131.5422      0        0
# 2  50        0 131.5422      0        0
# 3  60        0 131.5422      0        0

# Fit cox model for risk profile 1
males_fit_riskprofiles <- coxph(Surv(STROKE10TIME, STROKE10) ~ 
                                    AGE + DIABETES + SYSBP + CURSMOKE,
                                  data = males_df
)
        
# Use survfit                          
mrp1_fit <- survfit(males_fit_riskprofiles, newdata = males_rp1)


# Estimates for 10 years
mrp1_summary <- summary(mrp1_fit, times = 365*10)
 # time n.risk n.event survival1 survival2 survival3
 # 3650   1647      48     0.995      0.99     0.981

# Data frame of survival 10 years and risk 10 years
mrp1_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(mrp1_summary$surv), 3),
                           Risk = paste0(round(1 - mrp1_summary$surv, 3)*100,
                                         "%")
)

```


```{r, echo = F, results='hide'}
################################# Risk Profile 2 ###############################

# Risk profile 2: 
## High blood pressure and no other risk factors
males_rp2 <- data.frame(
  AGE = c(40, 50, 60),
  DIABETES = factor(rep(0, 3), 
                    levels = levels(males_df$DIABETES)),
  SYSBP = rep(200, 3),
  BPMEDS = factor(rep(0, 3), 
                  levels = levels(males_df$BPMEDS)),
  CURSMOKE = factor(rep(0, 3), 
                    levels = levels(males_df$CURSMOKE))
)

# AGE DIABETES SYSBP PREVCHD BPMEDS CURSMOKE  TOTCHOL      BMI
# 1  40        0   180       0      0        0 233.7359 26.16627
# 2  50        0   230       0      0        0 233.7359 26.16627
# 3  60        0   280       0      0        0 233.7359 26.16627


# Fit cox model for risk profile 2
mrp2_fit <- survfit(males_fit_riskprofiles, newdata = males_rp2)

# Estimates for 10 years
mrp2_summary <- summary(mrp2_fit, times = 365*10)
 # time n.risk n.event survival1 survival2 survival3
 # 3650   1647      48     0.976     0.795     0.114

# Create data frame of survival 10 years and risk 10 years
mrp2_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(mrp2_summary$surv), 3),
                           Risk = paste0(round(1 - mrp2_summary$surv, 3)*100,
                                         "%")
)
# Age Survival  Risk
# 1  40    0.976  2.4%
# 2  50    0.795 20.5%
# 3  60    0.114 88.6%


```


```{r, echo = F, results='hide'}

################################# Risk Profile 3 ###############################

#### Risk profile 3:
## Diabetes and no other risk factors
males_rp3 <- data.frame(
  AGE = c(40, 50, 60),
  DIABETES = factor(rep(1, 3), 
                    levels = levels(males_df$DIABETES)),
  SYSBP = rep(mean(males_df$SYSBP), 3),
  BPMEDS = factor(rep(0, 3), 
                  levels = levels(males_df$BPMEDS)),
  CURSMOKE = factor(rep(0, 3), 
                    levels = levels(males_df$CURSMOKE))
)
# AGE DIABETES    SYSBP PREVCHD BPMEDS CURSMOKE  TOTCHOL      BMI
# 1  40        1 131.5422       0      0        0 233.7359 26.16627
# 2  50        1 131.5422       0      0        0 233.7359 26.16627
# 3  60        1 131.5422       0      0        0 233.7359 26.16627


# Fit cox model for risk profile 3
mrp3_fit <- survfit(males_fit_riskprofiles, newdata = males_rp3)

# Estimates for 10 years
mrp3_probs <- summary(mrp3_fit, times = 365*10)
 # time n.risk n.event survival1 survival2 survival3
 # 3650   1647      48     0.974      0.95     0.905


# Estimates for 10 years
mrp3_summary <- summary(mrp3_fit, times = 365*10)
 # time n.risk n.event survival1 survival2 survival3
 # 3650   1647      48     0.974      0.95     0.905

# Create data frame of survival 10 years and risk 10 years
mrp3_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(mrp3_summary$surv), 3),
                           Risk = paste0(round(1 - mrp3_summary$surv, 3)*100,
                                         "%")
)

# Age Survival Risk
# 1  40    0.974 2.6%
# 2  50    0.950   5%
# 3  60    0.905 9.5%

```


```{r, echo = F, results='hide'}
################################# Risk Profile 4 ###############################

#### Risk profile 4:
## High BP and Diabetes
males_rp4 <- data.frame(
  AGE = c(40, 50, 60),
  DIABETES = factor(rep(1, 3), 
                    levels = levels(males_df$DIABETES)),
  SYSBP = rep(200, 3),
  BPMEDS = factor(rep(0, 3), 
                  levels = levels(males_df$BPMEDS)),
  CURSMOKE = factor(rep(0, 3), 
                    levels = levels(males_df$CURSMOKE))
)
# 
# AGE DIABETES SYSBP PREVCHD BPMEDS CURSMOKE  TOTCHOL      BMI
# 1  40        1   180       0      0        0 233.7359 26.16627
# 2  50        1   230       0      0        0 233.7359 26.16627
# 3  60        1   280       0      0        0 233.7359 26.16627

# Fit cox model for risk profile 3
mrp4_fit <- survfit(males_fit_riskprofiles, newdata = males_rp4)

# Estimates for 10 years
mrp4_summary <- summary(mrp4_fit, times = 365*10)
 # time n.risk n.event survival1 survival2 survival3
 # 3650   1647      48     0.883     0.307  1.44e-05

# Create data frame of survival 10 years and risk 10 years
mrp4_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(mrp4_summary$surv), 3),
                           Risk = paste0(round(1 - mrp4_summary$surv, 3)*100,
                                         "%")
)

# Age Survival  Risk
#   Age Survival  Risk
# 1  40    0.883 11.7%
# 2  50    0.307 69.3%
# 3  60    0.000  100%

```


```{r, echo = F, results='hide'}
################################# Risk Profile 5 ###############################

#### Risk profile 4:
## Smoker and no other risk factors
males_rp5 <- data.frame(
  AGE = c(40, 50, 60),
  DIABETES = factor(rep(0, 3), 
                    levels = levels(males_df$DIABETES)),
  SYSBP = rep(mean(males_df$SYSBP), 3),
  BPMEDS = factor(rep(0, 3), 
                  levels = levels(males_df$BPMEDS)),
  CURSMOKE = factor(rep(1, 3)
))
# 
# AGE DIABETES    SYSBP PREVCHD BPMEDS CURSMOKE  TOTCHOL      BMI
# 1  40        0 131.5422       0      0        1 233.7359 26.16627
# 2  50        0 131.5422       0      0        1 233.7359 26.16627
# 3  60        0 131.5422       0      0        1 233.7359 26.16627

# Fit cox model for risk profile 3
mrp5_fit <- survfit(males_fit_riskprofiles, newdata = males_rp5)

# Estimates for 10 years
mrp5_summary <- summary(mrp5_fit, times = 365*10)
 # time n.risk n.event survival1 survival2 survival3
 # 3650   1647      48      0.99      0.98     0.962

# Create data frame of survival 10 years and risk 10 years
mrp5_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(mrp5_summary$surv), 3),
                           Risk = paste0(round(1 - mrp5_summary$surv, 3)*100,
                                         "%")
)

# Age Survival Risk
# 1  40    0.990   1%
# 2  50    0.980   2%
# 3  60    0.962 3.8%
```


```{r, echo = F}
################### Combine these all for the report
library(knitr)
library(kableExtra)

# Create individual kables for each df
# Results 1
rk1 <- kable(mrp1_results, caption = "meow1",
             format = "html") %>% 
  kable_styling()

# Results 2
rk2 <- kable(mrp2_results, caption = "meow2",
             format = "html") %>% 
  kable_styling()

# Results 3
rk3 <- kable(mrp3_results, caption = "meow3",
             format = "html") %>% 
  kable_styling()

# Results 4
rk4 <- kable(mrp4_results, caption = "meow4",
             format = "html") %>% 
  kable_styling()

# Combimne results into one
male_results <- cbind(mrp1_results, 
                       mrp2_results[, 2:3],
                       mrp3_results[,2:3],
                       mrp4_results[,2:3],
                       mrp5_results[,2:3]
                       )

```

```{r, echo = F}
# Display both male and female tables together

kable(female_results,
  format = "latex",
  booktabs = TRUE,
  digits = 3, 
  caption = "Survival Rates and Risk \\% Over 10 Years for Different 
  Risk Profiles") %>%
  kable_styling(latex_options = c("striped", "hold_position",
                                  "scale_down")) %>%
  add_header_above(c(" ", "Risk Profile 1\nNo Risk factors" = 2,
                     "Risk Profile 2\nHigh BP" = 2,
                     "Risk Profile 3\nDiabetes" = 2,
                     "Risk Profile 4\n High BP & Diabetes" = 2,
                     "Risk Profile 5\n Smoker" = 2)) %>%
  add_header_above(c("Females"
                     = 11),
                   escape = FALSE)


kable(male_results,
  format = "latex",
  booktabs = TRUE,
  digits = 3) %>%
  # caption = "Survival Rates and Risk \\% Over 10 Years for Males") %>%
  kable_styling(latex_options = c("striped", "hold_position",
                                  "scale_down")) %>%
  add_header_above(c(" ", "Risk Profile 1\nNo Risk factors" = 2,
                     "Risk Profile 2\nHigh BP" = 2,
                     "Risk Profile 3\nDiabetes" = 2,
                     "Risk Profile 4\n High BP & Diabetes" = 2,
                     "Risk Profile 5\n Smoker" = 2)) %>%
  add_header_above(c("Males"
                     = 11),
                   escape = FALSE) %>%
  footnote(number = c("BP = Blood Pressure & High BP = 200 (mmHg)",
                      "All Risk Factors are at Baseline Timepoint"))

```


<!------------------------------- Results continued -------------------------->

\vspace{-1mm}
was associated with a 7% higher hazard of stroke (95% CI: 3%, 12%) and each additional one-unit increase in systolic BP was associated with a 3%  higher hazard of stroke (95% CI: 2%, 3%).
For males, age, Diabetes status, and Systolic BP were significantly associated with a higher risk of stroke. 
Each one year increase in age was associated with a 7% higher hazard of stroke (95% CI: 3%, 11%), diabetes was associated with a 409% or 5.09 times increase in hazard (95% CI: 2.21, 11.7) compared to those without diabetes, and each one unit increase in systolic BP was associated with a 3.2% increase in hazard of stroke (95% CI: 2%, 4%). The use of antihypertensive medication was not significantly associated with risk of stroke for males. 

Table 3 displays the predicted survival probabilities and 10-year risk estimates for males and females aged 40, 50, and 60 across the five risk profiles of interest for this investigation. Figure 1 displays survival probability curves for males. Individuals with no risk factors had 10-year survival probabilities above 90% and corresponding risks below 3%. These estimates were similar to those observed among current smokers at baseline, although smokers did show slightly lower survival probabilities and higher risk.

```{r, echo = F, message = F, warning = F, fig.cap="Survival Curves for the 5 Risk Profiles for Males"}
library(colorspace)
# Plot the survival curves
m1 <- ggsurvplot(survfit(males_fit_riskprofiles, 
                   newdata = males_rp1),
           data = males_df,
           censor = FALSE,
           conf.int = FALSE,
           ylim = c(0.98, 1),
           xlim = c(0, 3650),
           # risk.table = TRUE,
           ggtheme = theme_minimal(),
           # legend = "none",
           legend.labs = c("40", "50", "60"),
           legend.title = "Age",
           title = "Risk Profile 1\nAverage Individual",
           font.main = 10,
           palette = "Set1"
           ) 
# m1

# Plot the survival curves
m2 <- ggsurvplot(survfit(males_fit_riskprofiles, 
                   newdata = males_rp2),
           data = males_df,
           censor = F,
           conf.int = FALSE,
           # Min is 0.35 and max is 0.99
           ylim = c(0.8, 1),
           xlim = c(0, 3650),
           ggtheme = theme_minimal(),
           legend.labs = c("Age 40", "Age 50", "Age 60"),
           legend = "none",
           title = "Risk Profile 2\nHigh BP",
           font.main = 10,
           palette = "Set1"
) 
# f2


# Plot the survival curves
m3 <- ggsurvplot(survfit(males_fit_riskprofiles, 
                   newdata = males_rp3),
           data = males_df,
           censor = F,
           conf.int = FALSE,
           # Min is 0.35 and max is 0.99
           ylim = c(0.90, 1),
           xlim = c(0, 3650),
           ggtheme = theme_minimal(),
           legend = "none",
           title = "Risk Profile 3\nDiabetes",
           font.main = 10,
           palette = "Set1"
) 
# f3

# Plot the survival curves
m4 <- ggsurvplot(survfit(males_fit_riskprofiles, 
                   newdata = males_rp4),
           data = males_df,
           censor = F,
           conf.int = FALSE,
           # Min is 0.35 and max is 0.99
           ylim = c(0.35, 1),
           xlim = c(0, 3650),
           ggtheme = theme_minimal(),
           legend.labs = c("Age 40", "Age 50", "Age 60"),
           legend = "none",
           title = "Risk Profile 4\nHigh BP & Diabetes",
           font.main = 10,
           palette = "Set1"
) 
# f4

# Plot the survival curves
m5 <- ggsurvplot(survfit(males_fit_riskprofiles, 
                         newdata = males_rp5),
                 data = males_df,
                 censor = F,
                 conf.int = FALSE,
                 # Min is 0.35 and max is 0.99
                 ylim = c(0.95, 1),
                 xlim = c(0, 3650),
                 ggtheme = theme_minimal(),
                 legend = "none",
                 title = "Risk Profile 5\nSmoker",
                 font.main = 10,
                 palette = "Set1"
) 
# f5

##### Display all plots together

library(ggpubr)
# Using ggarrange
plot2 <- ggarrange(m1$plot, m2$plot, m3$plot, m4$plot, m5$plot,
          ncol = 3, nrow = 2,
          common.legend = TRUE,
          legend = "top",
          legend.labs = c("Age", "Age 50", "Age 60"),
          vjust = 0.5
          ) 
# Add a note at the bottom of this figure
annotate_figure(plot2,
                top = text_grob("Estimated Survival Curves - Males",
                                size = 12,
                                face = "bold"),
                bottom = text_grob("**Y-axis is not the same for every plot",
                                   size = 10,
                                   hjust = 1.3,
                                   face = "italic"))
  

# With survival package - could not get one legend for all plots
# arrange_ggsurvplots(list(f1, f2, f3, f4, f5),
#                     ncol = 3, nrow = 2,
#                     title = "Estimated Survival Curves - males")
```

In contrast, survival probabilities decreased and risk increased substantially among individuals with high blood pressure, diabetes, and especially the combination of both conditions. The greatest differences were observed between Risk Profile 1 (no risk factors) and Risk Profile 4 (high blood pressure and diabetes). For example, among males aged 60, the predicted 10-year stroke risk increased from approximately 2% for individuals with no risk factors to approximately 62% for those with both high blood pressure and diabetes. Notable differences between males and females were observed across all age groups within this risk profile.
Overall, stroke risk increased with age for both sexes, and high blood pressure and diabetes appeared to be the strongest predictors of stroke risk. 

Table 4 summarizes descriptive statistics across the three study periods. As expected, the number of participants decreased over time, with increasing missingness observed for anti-hypertensive medication use, total cholesterol, and BMI. The prevalence of anti-hypertensive medication use and CHD increased over time, while the proportion of current smokers decreased. Among continuous variables, systolic blood pressure showed a steady increase across periods, whereas total cholesterol and BMI remained relatively stable.


```{r, echo = F}
##################################### Table 4 ##################################

# Table describing the change in the risk factors over time

# Fix levels for categorical variables for Table1
# Data frame for table1
tab3_data <- frame_data

# Remove those that had a stroke at baseline
tab3_data <- tab3_data %>% 
  group_by(RANDID) %>% 
  filter(!(STROKE == 1 & TIMESTRK == 0))


# Make sure the factor variables are factors

# Sex
tab3_data$factor_SEX <- factor(tab3_data$SEX,
                        levels = c(1, 2),
                        labels = c("Male", "Female"))

# Diabetes
tab3_data$factor_DIABETES <- factor(tab3_data$DIABETES,
                                    levels = c(0, 1),
                                    labels = c("Not a Diabetic", "Diabetic"))

# Prevchd
tab3_data$factor_PREVCHD <- factor(tab3_data$PREVCHD,
                                   levels = c(0, 1),
                                   labels = c("Free of CHD",
                                              "Prevalent CHD"))

# Cursmoke
tab3_data$factor_CURSMOKE <- factor(tab3_data$CURSMOKE,
                             levels = c(0, 1),
                             labels = c("Not a Current Smoker",
                                        "Current Smoker"))

# BPMeds
tab3_data$factor_BPMEDS <- factor(tab3_data$BPMEDS,
                                  levels = c(0, 1),
                                  labels = c("No Use of Medication",
                                             "Use of Medication"))

# Period
tab3_data$factor_PERIOD <- factor(tab3_data$PERIOD,
                                  levels = c(1, 2, 3),
                                  labels = c("Period 1\n",
                                             "Period 2\n",
                                             "Period 3\n"))


# Relabel the variables for the Table 1
tab3_data <- apply_labels(tab3_data,
                     factor_SEX = "Sex",
                     factor_BPMEDS = "Use of Anti-hypertensive Medication",
                     factor_PREVCHD = "Prevalent Coronary Heart Disease",
                     factor_CURSMOKE = "Cigarrete Smoking Status",
                     AGE = "Age",
                     SYSBP = "Systolic Blood Pressure (mmHg)",
                     TOTCHOL = "Serum Total Cholesterol (mg/dL)",
                     BMI = "Body Mass Index"
                     )

# Display the modified Table1
tab3_data %>% 
  tbl_summary(include = c(factor_SEX,
                          AGE,
                          factor_BPMEDS,
                          factor_PREVCHD,
                          factor_CURSMOKE,
                          SYSBP,
                          TOTCHOL,
                          BMI),
              missing_text = "Missing",
              by = factor_PERIOD,
              statistic = list(c(AGE, SYSBP, TOTCHOL, BMI) 
                               ~ "{mean}({sd})")) %>% 
  modify_header(label = "Study Period") %>% 
  # bold_labels() %>% 
  modify_caption(caption = "Patient Characteristics Across the 
                 Three Periods") %>% 
  # Remove the stars from the 3 study period col headers
  modify_header(all_stat_cols() ~ "{level}\n (N = {n}) ") %>% 
  # Modify to fit in report
  as_kable(
           booktabs = TRUE) %>% 
  kableExtra::kable_styling(latex_options = c("scale_down",
                                              "hold_position")) %>% 
  column_spec(1, bold = TRUE) %>% 
  row_spec(0, bold = TRUE) %>% 
  add_indent(c(2:3, 6:8, 10:11, 13:14, 17, 19))
```

\vspace{-6mm}
## Discussion
\vspace{-6mm}

Our results highlight both differences and similarities in stroke risk between males and females during the first 10 years of follow-up in the Framingham Heart Study. Across both sexes, age and systolic blood pressure were consistently associated with increased stroke risk. Specifically, each additional year of age was associated with approximately a 7% increase in hazard, while each 1 mmHg increase in systolic blood pressure was associated with an approximate 3% increase in hazard for both sexes.
Diabetes was strongly associated with stroke risk among males but not among females. Among males, diabetes was associated with a substantial increase in hazard, suggesting that its impact on stroke risk may differ by sex. Overall, the survival analysis identified key predictors of stroke, with some differences in significance across sex-specific models.

Evaluation of survival probabilities across the five risk profiles showed some clear patterns. Older individuals and males generally showed higher predicted stroke risk compared to younger individuals and females. Within each sex, stroke risk increased markedly with age, particularly when comparing individuals aged 40, 50, and 60 years. Across both sexes, high blood pressure and diabetes appear to be the most influential risk factors, with the combination of these conditions producing the highest predicted risks.
Interestingly, we did not observe a significant association between smoking status and stroke risk in this analysis. This may be due to limitations in the data or model specification instead of a true absence of effect, given that there is a well-established relationship between smoking and cardiovascular outcomes.
These findings underscore the importance of age and blood pressure in stroke risk, as well as potential sex-specific differences in the impact of diabetes. From a public health perspective, these results reinforce the importance of managing blood pressure and metabolic conditions, particularly among older individuals.

Several limitations should be considered. First, the small number of stroke events may have limited statistical power, particularly for detecting associations within subgroups. 
Second, the distributions of anti-hypertensive medication use and prevalent coronary heart disease were unbalanced, which may have affected the stability and interpretation of model estimates. 
Third, potential residual confounding may be present, as not all relevant risk factors (e.g., lifestyle factors, socioeconomic status, or treatment adherence) were included in the models.
Additionally, risk factors were assessed at baseline, and changes over time were not incorporated into the survival models, which may lead to misclassification of exposure. 
Given the information presented in Table 4, a survival model incorporating the time-varying covariates is warranted and should be considered in the future. 
Finally, the Framingham Heart Study population may limit generalizability,
as it does not fully represent more diverse populations.


```{r, echo = F, message=F, warning=F, results='hide', eval = F}
library(colorspace)
# Plot the survival curves
f1 <- ggsurvplot(survfit(females_fit_riskprofiles, 
                   newdata = females_rp1),
           data = females_df,
           censor = FALSE,
           conf.int = FALSE,
           ylim = c(0.95, 1),
           xlim = c(0, 3650),
           # risk.table = TRUE,
           ggtheme = theme_minimal(),
           # legend = "none",
           legend.labs = c("40", "50", "60"),
           legend.title = "Age",
           title = "Risk Profile 1 Estimated Survival\nAverage Individual",
           font.main = 10,
           palette = c("deeppink2", "turquoise3", "springgreen3")
           ) 
# f1

# Plot the survival curves
f2 <- ggsurvplot(survfit(females_fit_riskprofiles, 
                   newdata = females_rp2),
           data = females_df,
           censor = F,
           conf.int = FALSE,
           # Min is 0.35 and max is 0.99
           ylim = c(0.35, 1),
           xlim = c(0, 3650),
           ggtheme = theme_minimal(),
           legend.labs = c("Age 40", "Age 50", "Age 60"),
           legend = "none",
           title = "Risk Profile 2 Estimated Survival\nHigh BP",
           font.main = 10,
           palette = c("deeppink2", "turquoise3", "springgreen3")
) 
# f2


# Plot the survival curves
f3 <- ggsurvplot(survfit(females_fit_riskprofiles, 
                   newdata = females_rp3),
           data = females_df,
           censor = F,
           conf.int = FALSE,
           # Min is 0.35 and max is 0.99
           ylim = c(0.95, 1),
           xlim = c(0, 3650),
           ggtheme = theme_minimal(),
           legend = "none",
           title = "Risk Profile 2 Estimated Survival\nDiabetes",
           font.main = 10,
           palette = c("deeppink2", "turquoise3", "springgreen3")
) 
# f3

# Plot the survival curves
f4 <- ggsurvplot(survfit(females_fit_riskprofiles, 
                   newdata = females_rp4),
           data = females_df,
           censor = F,
           conf.int = FALSE,
           # Min is 0.35 and max is 0.99
           # ylim = c(0.95, 1),
           xlim = c(0, 3650),
           ggtheme = theme_minimal(),
           legend.labs = c("Age 40", "Age 50", "Age 60"),
           legend = "none",
           title = "Risk Profile 2 Estimated Survival\nHigh BP & Diabetes",
           font.main = 10,
           palette = c("deeppink2", "turquoise3", "springgreen3")
) 
# f4

# Plot the survival curves
f5 <- ggsurvplot(survfit(females_fit_riskprofiles, 
                         newdata = females_rp5),
                 data = females_df,
                 censor = F,
                 conf.int = FALSE,
                 # Min is 0.35 and max is 0.99
                 ylim = c(0.95, 1),
                 xlim = c(0, 3650),
                 ggtheme = theme_minimal(),
                 legend = "none",
                 title = "Risk Profile 2 Estimated Survival\nSmoker",
                 font.main = 10,
                 palette = c("deeppink2", "turquoise3", "springgreen3")
) 
# f5

##### Display all plots together

library(ggpubr)
# Using ggarrange
plot1 <- ggarrange(f1$plot, f2$plot, f3$plot, f4$plot, f5$plot,
          ncol = 3, nrow = 2,
          common.legend = TRUE,
          legend = "top",
          legend.labs = c("Age", "Age 50", "Age 60"),
          vjust = 0.5
          ) 
# Add a note at the bottom of this figure
annotate_figure(plot1,
                top = text_grob("Estimated Survival Curves - Females",
                                size = 12,
                                face = "bold"),
                bottom = text_grob("**Y-axis is not the same for every plot",
                                   size = 10,
                                   hjust = 1.3,
                                   face = "italic"))
  

# With survival package - could not get one legend for all plots
# arrange_ggsurvplots(list(f1, f2, f3, f4, f5),
#                     ncol = 3, nrow = 2,
#                     title = "Estimated Survival Curves - Females")
```


\newpage

# Data and Code Availability

The data and statistical code used to generate the results and figures in this report are available in the following GitHub repository:

- GitHub repository: [https://github.com/DianaF65/BIOS6624](https://github.com/DianaF65/BIOS6624)

Directory structure:

- Background/ - A description of the background of the project.

- Code/ 
  - The "Project3.R" file contains initial data exploration as well as code for construction of the final data set used for all analyses.
  - The R Markdown file 'Project3_FinalReport.Rmd' consists of all Code used for the Frequentist and Bayesian analyses.

- RawData/ – Raw data files. Because this data is not publicly available, it is not available on GitHub. The raw data was originally obtained from the BIOS 6624 P3 module. The raw and processed datasets to generate these results are available on my local computer at this file path: 
"/Users/dianafelipe/Desktop/BIOS 6624/BIOS6624/Project3/Data/frmghaml2.csv"

- Reports/ The "Project13_FinalReport.Rmd" R Markdown file was used to generate this final report of analyses.

All analyses and figures were generated using the R statistical programming language v. 4.4.1, and the repository contains the R scripts
and R Markdown files used to produce this report



