---
title: P3 meow
author: Diana Felipe
output: 
  bookdown::pdf_document2:
    urlcolor: magenta
    toc: false
    number_sections: false
    latex_engine: xelatex
    extra_dependencies: ["longtable", "float"]
mainfont: Arial
fontsize: 12pt
documentclass: article
geometry: margin=1in
header-includes:
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage{float}
  - \usepackage{caption}
  - \usepackage{titling}
  - \setlength{\droptitle}{-3cm}
  - \preauthor{\vspace{-0.6cm}}
  - \postauthor{\vspace{-3cm}}
---

```{r, echo = F, warning=F, message = F}
knitr::opts_chunk$set(echo = FALSE)
library(readr)
library(knitr)
library(tidyr)
library(ggplot2)
library(kableExtra)
library(dplyr)
library(tibble)
library(labelled)
library(tableone)
library(nlme)
library(gtsummary)
library(gt)
library(ggeffects)
library(see)
library(gridExtra)
library(regressinator)
library(patchwork)
library(purrr)
library(stringr)
library(survival)
library(survminer)
library(gtsummary)
library(here)

opts_chunk$set(tidy = F)
```

```{r, echo = F}
# Read in the Data
# Read in created survival df
surv_df <- read_csv(here("Project3", "Data", "frmg_survival_df.csv"))

# Convert factor vars to factors
surv_df <- surv_df %>% 
  mutate(across(c(RANDID, SEX, DIABETES, 
                  deathwithin10, PREVCHD, BPMEDS, CURSMOKE), 
                as.factor))
# Make sure STROKE10 is numeric
surv_df$STROKE10 <- as.numeric(as.character(surv_df$STROKE10))

```


\vspace{-8mm}
# Introduction
\vspace{-8mm}

And then i said i sai dmoe meow meow moew meow moewo jew


\vspace{-10mm}
# Methods
\vspace{-6mm}

and then they said meow moew moew moewmoewm oewmoew meow

\vspace{-6mm}
## Data Cleaning
\vspace{-6mm}

and then we said we said meow meow meomeow moew moew

```{r, echo = F, results='hide', eval = F, warning=F, message=F}
############################## Variable Selection ##############################
# Vignette: https://glmnet.stanford.edu/articles/Coxnet.html

# Perform some variable selection prevchd, bpmeds, smokecur, total chol, and BMI

##### First, going to assess whether sex should have different baseline hazards
# Unstratified model
fit_unstrat <- coxph(
  Surv(STROKE10TIME, STROKE10) ~ SEX + PREVCHD + 
    BPMEDS + CURSMOKE + TOTCHOL + BMI,
  data = surv_df
)

# Stratified by sex model
fit_strat <- coxph(
  Surv(STROKE10TIME, STROKE10) ~ PREVCHD + BPMEDS + 
    CURSMOKE + TOTCHOL + BMI + strata(SEX),
  data = surv_df
)

# Compare model fit with AIC
AIC(fit_unstrat, fit_strat)
# df      AIC
# fit_unstrat  6 1723.074
# fit_strat    5 1576.158

# We saw though that the KM curves by sex were basically identical
# Sex was not significant in the model
# The un and stratified models give almost the same results 
# Since we were getting convergence issues down the line with glmnet,
# we are going to NOT stratify by sex 

# Proceed with the unstratified pooled glmnet cox model

###### Glmnet

# Fit stratified cox models with elastic net penalty
# Prepare for glmnet
# Surv object of response variables
y <- Surv(time = surv_df$STROKE10TIME,
          event = surv_df$STROKE10)

# Create matrix of predictors
x <- model.matrix(~ PREVCHD + BPMEDS + CURSMOKE + TOTCHOL + BMI,
                  data = surv_df)[, -1] # This removes the intercept


# Fit stratified cox lasso
lasso_fit <- glmnet(x = x, 
                    y = y,
                    family = "cox",
                    alpha = 1) # LASSO penalty

# Plot fit 
plot(lasso_fit)

# K-fold Cross validation
set.seed(345)
cv_fit <- cv.glmnet(x = x, 
                    y = y, 
                    family = "cox",
                    alpha = 1)

# Look at plot - optimal lambda value
plot(cv_fit)
# The left vertical line is where the CV error curve hits its minimum
# The right vertical line is the most regularized model with CV error within 1 SD
# Basically we see that model performance is basically flat and increasing
# model complexity does not improve performance much


# Get the optimal lambdas
cv_fit$lambda.min
#  0.001777793
## This gives the best fit or lowest CV error where a small penalty means
# more variables are kept - risk of overfitting though

# See which variables are kept
coef_min <- coef(cv_fit, s = "lambda.min")
# PREVCHD1  0.827245643
# BPMEDS1   1.382781943
# CURSMOKE1 0.123849200
# TOTCHOL   0.002622194
# BMI       0.043796358

# Get the nonzero variables
nz_min <- which(as.vector(coef_min) != 0)
selected_min <- rownames(coef_min)[nz_min]
# "PREVCHD1"  "BPMEDS1"   "CURSMOKE1" "TOTCHOL"   "BMI" 

cv_fit$lambda.1se
# 0.01376481
## Largest lambda within 1 SE of the minimum
# More penalty more simpler model

# See which variables are kept
coef_1se <- coef(cv_fit, s = "lambda.1se")
# PREVCHD1  .
# BPMEDS1   .
# CURSMOKE1 .
# TOTCHOL   .
# BMI       .

# Get the nonzero variables
nz_1se <- which(as.vector(coef_1se) != 0)
selected_1se <- rownames(coef_1se)[nz_1se]
# character(0)

# Basically we found that lambda.min keeps all variables while
# lambda.1se drops everything
# We did already have pre-selected variables that were clinically driven so 
# perhaps further reduction is not necessary

# Using lambda.min, all five candidate predictors were retained.
# Using the more conservative lambda.1se rule, all coefficients were shrunk to zero, 
# indicating that the cross-validated error was fairly flat and that there was 
# limited evidence for a sparser predictive model with clear improvement over the null.

# We used LASSO with cross-validation to assess variable importance. 
# Because the cross-validated error curve was relatively flat, the more 
# conservative λ₁ₛₑ resulted in a null model. Therefore, we selected λ_min to 
# retain relevant predictors while minimizing prediction error.

# After lasso, fit a standard coxph model with ALL variables
```



\vspace{-8mm}
## Data Analysis
\vspace{-6mm}

Then we all got toegher and said mewo mewo moew moewmoew

\vspace{-10mm}
# Results
\vspace{-6mm}

churu churuc churu yummy churu chury

<!------------------------------------ Table 1 -------------------------------->

```{r, echo = F, eval = F}
# Fix levels for categorical variables for Table1
# Race
data$RACE_tab1 <- factor(data$collapse_RACE,
                         levels = c("Other", "White_NH"),
                         labels = c("Other", "White Non-Hispanic"))
# Education
data$EDU_tab1 <- factor(data$collapsed_EDUCBAS,
                        levels = c("Comp_Coll_Higher", "Less_Than_College"),
                        labels = c("Complete College Or Higher",
                                   "Less Than College"))
# Adherence


# Relabel the variables for the Table 1

tab1_data <- apply_labels(data,
                     VLOAD_yr0 = "Viral Load - (Yr 0)",
                     VLOAD_yr2 = "Viral Load - (Yr 2)",
                     LEU3N_yr0 = "CD4+ T-Cell Count (Yr 0)",
                     LEU3N_yr2 = "CD4+ T-Cell Count (Yr 2)",
                     AGG_MENT_yr0 = "Agg. Mental Quality of Life Score (Yr 0)",
                     AGG_MENT_yr2 = "Agg. Mental Quality of Life Score (Yr 2)",
                     AGG_PHYS_yr0 = "Agg. Physical Quality of Life Score (Yr 0)",
                     AGG_PHYS_yr2 = "Agg. Physical Quality of Life Score (Yr 2)",
                     collapse_SMOKE = "Smoking Status (Baseline)",
                     EDU_tab1 = "Education (Baseline)",
                     RACE_tab1 = "Race",
                     age = "Age (Baseline)",
                     hard_drugs_cat = "Hard Drug Use (Baseline)",
                     collapsed_ADH = "HAART Adherence")

# Display the modified Table1
tab1_data %>% 
  tbl_summary(include = c(VLOAD_yr0, VLOAD_yr2,
                          LEU3N_yr0, LEU3N_yr2,
                          AGG_MENT_yr0, AGG_MENT_yr2,
                          AGG_PHYS_yr0, AGG_PHYS_yr2,
                          BMI, collapse_SMOKE, 
                          age, RACE_tab1,
                          EDU_tab1,
                          collapsed_ADH),
              by = hard_drugs_cat,
              statistic = list(c("VLOAD_yr2", "VLOAD_yr0",
                          "LEU3N_yr2", "LEU3N_yr0",
                          "BMI") 
                               ~ "{mean}{sd}")) %>% 
  modify_header(label = "**Hard Drug Use at Baseline**") %>% 
  bold_labels() %>% 
  modify_caption(caption = "**Patient Characteristics by Hard Drug Use at Baseline**")
```



```{r, echo = F, results = 'hide', include=F, warning=F, message=F}
################################## Model Fitting ###############################

##################################### Females ##################################
# Data frame for females
females_df <- surv_df %>% 
  filter(SEX == 2)
# 2378 obs for 2378 subjects

# Model for females
females_fit <- coxph(
  Surv(STROKE10TIME, STROKE10) ~ AGE + DIABETES + SYSBP
  + PREVCHD + BPMEDS 
  + CURSMOKE + TOTCHOL + BMI,
  data = females_df
)

# Summary of general model
females_sum1 <- summary(females_fit)

# Create data frame of results
females_fit %>% 
  tbl_regression(intercept = TRUE, 
                 conf.level = .95,
                 exponentiate = TRUE,
                 pvalue_fun = label_style_pvalue(digits = 2),
                 conf.int = TRUE,
                 show_single_row = c(DIABETES, PREVCHD, BPMEDS, CURSMOKE)
                 )

# Plot survival curve - hmmm
ggsurvplot(survfit(females_fit),
           data = females_df,
           censor = F,
           conf.int = T)

```

```{r, echo = F, results='hide'}
################################# Risk Profile 1 ###############################

# Create data frame for the different risk profiles
# Risk profile 1: 
## Average age, no diabetes, average sysbp, no prechd, no bpmeds, no cursmoke,
## average totchol, average bmi
females_rp1 <- data.frame(
    AGE = c(40, 50, 60),
    DIABETES = factor(rep(0, 3), 
                      levels = levels(females_df$DIABETES)),
    SYSBP = rep(mean(females_df$SYSBP), 3),
    PREVCHD = factor(rep(0, 3), 
                     levels = levels(females_df$PREVCHD)),
    BPMEDS = factor(rep(0, 3), 
                    levels = levels(females_df$BPMEDS)),
    CURSMOKE = factor(rep(0, 3), 
                      levels = levels(females_df$CURSMOKE)),
    TOTCHOL = rep(mean(females_df$TOTCHOL, 
                       na.rm = TRUE), 3),
    BMI = rep(mean(females_df$BMI), 3)
)

library(see)
# Plot the survival curves
f1 <- ggsurvplot(survfit(females_fit, 
                   newdata = females_rp1),
           data = females_df,
           censor = FALSE,
           conf.int = TRUE,
           ylim = c(0.95, 1),
           xlim = c(0, 3650),
           # risk.table = TRUE,
           ggtheme = theme_minimal(),
           legend.labs = c("Age 40", "Age 50", "Age 60"),
           title = "Risk Profile 1 Estimated Survival\nAverage Individual"
           ) 
# f1

# Fit cox model for risk profile 1
frp1_fit <- survfit(females_fit, newdata = females_rp1)


# Estimates for 10 years
frp1_summary <- summary(frp1_fit, times = 365*10)
# time n.risk n.event survival1 survival2 survival3
# 3650   2177      57     0.995     0.988     0.976

# Data frame of survival 10 years and risk 10 years
frp1_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(frp1_summary$surv), 3),
                           Risk = paste0(round(1 - frp1_summary$surv, 3)*100,
                                         "%")
)



```


```{r, echo = F, results='hide'}
################################# Risk Profile 2 ###############################

# Risk profile 2: 
## High blood pressure and no other risk factors
females_rp2 <- data.frame(
  AGE = c(40, 50, 60),
  DIABETES = factor(rep(0, 3), 
                    levels = levels(females_df$DIABETES)),
  SYSBP = c(180, 230, 280),
  PREVCHD = factor(rep(0, 3), 
                   levels = levels(females_df$PREVCHD)),
  BPMEDS = factor(rep(0, 3), 
                  levels = levels(females_df$BPMEDS)),
  CURSMOKE = factor(rep(0, 3), 
                    levels = levels(females_df$CURSMOKE)),
  TOTCHOL = rep(mean(females_df$TOTCHOL, 
                     na.rm = TRUE), 3),
  BMI = rep(mean(females_df$BMI), 3)
)

# AGE DIABETES SYSBP PREVCHD BPMEDS CURSMOKE  TOTCHOL      BMI
# 1  40        0   180       0      0        0 239.5698 25.56664
# 2  50        0   230       0      0        0 239.5698 25.56664
# 3  60        0   280       0      0        0 239.5698 25.56664


# Plot the survival curves
f2 <- ggsurvplot(survfit(females_fit, 
                   newdata = females_rp2),
           data = females_df,
           censor = F,
           conf.int = FALSE,
           # Min is 0.35 and max is 0.99
           ylim = c(0.35, 1),
           xlim = c(0, 3650)
) 
# f2

# Fit cox model for risk profile 2
frp2_fit <- survfit(females_fit, newdata = females_rp2)

# Estimates for 10 years
frp2_summary <- summary(frp2_fit, times = 365*10)
# time n.risk n.event survival1 survival2 survival3
# 3650   2177      57     0.983     0.878     0.382

# Create data frame of survival 10 years and risk 10 years
frp2_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(frp2_summary$surv), 3),
                           Risk = paste0(round(1 - frp2_summary$surv, 3)*100,
                                         "%")
)
# Age Survival  Risk
# 1  40    0.983  1.7%
# 2  50    0.878 12.2%
# 3  60    0.382 61.8%


```


```{r, echo = F, results='hide'}

################################# Risk Profile 3 ###############################

#### Risk profile 3:
## Diabetes and no other risk factors
females_rp3 <- data.frame(
  AGE = c(40, 50, 60),
  DIABETES = factor(rep(1, 3), 
                    levels = levels(females_df$DIABETES)),
  SYSBP = rep(mean(females_df$SYSBP), 3),
  PREVCHD = factor(rep(0, 3), 
                   levels = levels(females_df$PREVCHD)),
  BPMEDS = factor(rep(0, 3), 
                  levels = levels(females_df$BPMEDS)),
  CURSMOKE = factor(rep(0, 3), 
                    levels = levels(females_df$CURSMOKE)),
  TOTCHOL = rep(mean(females_df$TOTCHOL, 
                     na.rm = TRUE), 3),
  BMI = rep(mean(females_df$BMI), 3)
)
# AGE DIABETES    SYSBP PREVCHD BPMEDS CURSMOKE  TOTCHOL      BMI
# 1  40        1 133.5524       0      0        0 239.5698 25.56664
# 2  50        1 133.5524       0      0        0 239.5698 25.56664
# 3  60        1 133.5524       0      0        0 239.5698 25.56664

# Plot the survival curves
f3 <- ggsurvplot(survfit(females_fit, 
                   newdata = females_rp3),
           data = females_df,
           censor = F,
           conf.int = FALSE,
           # Min is 0.35 and max is 0.99
           ylim = c(0.95, 1),
           xlim = c(0, 3650)
) 
# f3


# Fit cox model for risk profile 3
frp3_fit <- survfit(females_fit, newdata = females_rp3)

# Estimates for 10 years
frp3_probs <- summary(frp3_fit, times = 365*10)
# time n.risk n.event survival1 survival2 survival3
# 3650   2177      57      0.99     0.979     0.955


# Estimates for 10 years
frp3_summary <- summary(frp3_fit, times = 365*10)
# time n.risk n.event survival1 survival2 survival3
# 3650   2177      57      0.99     0.979     0.955

# Create data frame of survival 10 years and risk 10 years
frp3_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(frp3_summary$surv), 3),
                           Risk = paste0(round(1 - frp3_summary$surv, 3)*100,
                                         "%")
)

# Age Survival Risk
# 1  40    0.990   1%
# 2  50    0.979 2.1%
# 3  60    0.955 4.5%

```


```{r, echo = F, results='hide'}
################################# Risk Profile 4 ###############################

#### Risk profile 4:
## Diabetes and no other risk factors
females_rp4 <- data.frame(
  AGE = c(40, 50, 60),
  DIABETES = factor(rep(1, 3), 
                    levels = levels(females_df$DIABETES)),
  SYSBP = c(180, 230, 280),
  PREVCHD = factor(rep(0, 3), 
                   levels = levels(females_df$PREVCHD)),
  BPMEDS = factor(rep(0, 3), 
                  levels = levels(females_df$BPMEDS)),
  CURSMOKE = factor(rep(0, 3), 
                    levels = levels(females_df$CURSMOKE)),
  TOTCHOL = rep(mean(females_df$TOTCHOL, 
                     na.rm = TRUE), 3),
  BMI = rep(mean(females_df$BMI), 3)
)
# 
# AGE DIABETES SYSBP PREVCHD BPMEDS CURSMOKE  TOTCHOL      BMI
# 1  40        1   180       0      0        0 239.5698 25.56664
# 2  50        1   230       0      0        0 239.5698 25.56664
# 3  60        1   280       0      0        0 239.5698 25.56664


# Plot the survival curves
f4 <- ggsurvplot(survfit(females_fit, 
                   newdata = females_rp4),
           data = females_df,
           censor = F,
           conf.int = FALSE,
           # Min is 0.35 and max is 0.99
           # ylim = c(0.95, 1),
           xlim = c(0, 3650)
) 
# f4

# Fit cox model for risk profile 3
frp4_fit <- survfit(females_fit, newdata = females_rp4)

# Estimates for 10 years
frp4_summary <- summary(frp4_fit, times = 365*10)
# time n.risk n.event survival1 survival2 survival3
# 3650   2177      57      0.99     0.979     0.955

# Create data frame of survival 10 years and risk 10 years
frp4_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(frp4_summary$surv), 3),
                           Risk = paste0(round(1 - frp4_summary$surv, 3)*100,
                                         "%")
)

# Age Survival  Risk
# 1  40    0.968  3.2%
# 2  50    0.785 21.5%
# 3  60    0.167 83.3%

```


```{r, echo = F, results='hide'}
################################# Risk Profile 5 ###############################

#### Risk profile 4:
## Smoker and no other risk factors
females_rp5 <- data.frame(
  AGE = c(40, 50, 60),
  DIABETES = factor(rep(0, 3), 
                    levels = levels(females_df$DIABETES)),
  SYSBP = rep(mean(females_df$SYSBP), 3),
  PREVCHD = factor(rep(0, 3), 
                   levels = levels(females_df$PREVCHD)),
  BPMEDS = factor(rep(0, 3), 
                  levels = levels(females_df$BPMEDS)),
  CURSMOKE = factor(rep(1, 3), 
                    levels = levels(females_df$CURSMOKE)),
  TOTCHOL = rep(mean(females_df$TOTCHOL, 
                     na.rm = TRUE), 3),
  BMI = rep(mean(females_df$BMI), 3)
)
# 
# AGE DIABETES    SYSBP PREVCHD BPMEDS CURSMOKE  TOTCHOL      BMI
# 1  40        0 133.5524       0      0        1 239.5698 25.56664
# 2  50        0 133.5524       0      0        1 239.5698 25.56664
# 3  60        0 133.5524       0      0        1 239.5698 25.56664


# Plot the survival curves
f5 <- ggsurvplot(survfit(females_fit, 
                         newdata = females_rp5),
                 data = females_df,
                 censor = F,
                 conf.int = FALSE,
                 # Min is 0.35 and max is 0.99
                 ylim = c(0.95, 1),
                 xlim = c(0, 3650)
) 
# f5

# Fit cox model for risk profile 3
frp5_fit <- survfit(females_fit, newdata = females_rp5)

# Estimates for 10 years
frp5_summary <- summary(frp5_fit, times = 365*10)
# time n.risk n.event survival1 survival2 survival3
# 3650   2177      57     0.992     0.984     0.966

# Create data frame of survival 10 years and risk 10 years
frp5_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(frp5_summary$surv), 3),
                           Risk = paste0(round(1 - frp5_summary$surv, 3)*100,
                                         "%")
)

# Age Survival Risk
# 1  40    0.992 0.8%
# 2  50    0.984 1.6%
# 3  60    0.966 3.4%
```


```{r, echo = F}
################### Combine these all for the report
library(knitr)
library(kableExtra)

# Create individual kables for each df
# Results 1
fk1 <- kable(frp1_results, caption = "meow1",
             format = "html") %>% 
  kable_styling()

# Results 2
fk2 <- kable(frp2_results, caption = "meow2",
             format = "html") %>% 
  kable_styling()

# Results 3
fk3 <- kable(frp3_results, caption = "meow3",
             format = "html") %>% 
  kable_styling()

# Results 4
fk4 <- kable(frp4_results, caption = "meow4",
             format = "html") %>% 
  kable_styling()

# Combine these horizontally


### Try 2
# knitr::kable(list(frp1_results, 
#                   frp2_results[2:3],
#                   frp3_results[2:3]),
#              booktabs = TRUE,
#              caption = "Survival and Risk Rates for Females") |>
#   kable_styling(latex_options = "hold_position") |>
#   add_header_above(c("Meow1" = 2, "Meow2" = 2)) %>% 
#   footnote(general = "Meowsss")
  
# Combimne results into one
female_results <- cbind(frp1_results, 
                       frp2_results[, 2:3],
                       frp3_results[,2:3],
                       frp4_results[,2:3],
                       frp5_results[,2:3]
                       )

# Display with kable
kable(female_results,
  # format = "latex",
  booktabs = TRUE,
  digits = 3,
  caption = "Survival Rates and Risk % Over 10 Years for Females") |>
  kable_styling(latex_options = c("striped", "hold_position")) |>
  add_header_above(c(" ", "Risk Profile 1\nNo Risk factors" = 2, 
                     "Risk Profile 2\nHigh BP" = 2,
                     "Risk Profile 3\nDiabetes" = 2,
                     "Risk Profile 4\n High BP & Diabetes" = 2,
                     "Risk Profile 5\n Smoker" = 2)) %>% 
  footnote(number = c("BP = Blood Pressure ",
                      "All Risk Factors are at Baseline Timepoint"))

```

```{r, echo = F, eval = F}
##### Display all plots together

arrange_ggsurvplots(results_list)
```



\newpage

# Data and Code Availability

The data and statistical code used to generate the results and figures in this report are available in the following GitHub repository:

- GitHub repository: [https://github.com/DianaF65/BIOS6624](https://github.com/DianaF65/BIOS6624)

Directory structure:

- Background/ - A description of the background of the project.

- Code/ 
  - The "Project1_ExploratoryAnalysis.R" file contains initial data exploration as well as code for construction of the final data set used for all analyses.
  - The R Markdown file 'All_Analyses.Rmd' consists of all Code used for the Frequentist and Bayesian analyses.

- RawData/ – Raw data files. Because this data is not publicly available, it is not available on GitHub. The raw data was originally obtained from the BIOS 6624 P1 module. The raw and processed datasets to generate these results are available on my local computer at this file path: 
"/Users/dianafelipe/Desktop/BIOS 6624/BIOS6624/Project1/Data/hiv_6624_final.csv"

- Reports/ The "Project1_Report.Rmd" R Markdown file was used to generate this final report of analyses.

All analyses and figures were generated using the R statistical programming language v. 4.4.1, and the repository contains the R scripts
and R Markdown files used to produce this report


