---
title: Risk Factors of 10-year Stroke
author: Diana Felipe
output: 
  bookdown::pdf_document2:
    urlcolor: magenta
    toc: false
    number_sections: false
    latex_engine: xelatex
    extra_dependencies: ["longtable", "float"]
    keep_tex: true
mainfont: Arial
fontsize: 12pt
documentclass: article
geometry: margin=1in
header-includes: 
  - \usepackage{setspace} 
  - \doublespacing 
  - \usepackage{float} 
  - \usepackage{caption} 
  - \usepackage{titling} 
  - \setlength{\droptitle}{-3cm} 
  - \setlength{\textfloatsep}{5pt} 
  - \preauthor{\vspace{-0.6cm}} 
  - \postauthor{\vspace{-2.5cm}}
---

```{r, echo = F, warning=F, message = F}
knitr::opts_chunk$set(echo = FALSE)
library(readr)
library(knitr)
library(tidyr)
library(ggplot2)
library(kableExtra)
library(dplyr)
library(tibble)
library(labelled)
library(tableone)
library(nlme)
library(gtsummary)
library(gt)
library(ggeffects)
library(see)
library(gridExtra)
library(regressinator)
library(patchwork)
library(purrr)
library(stringr)
library(survival)
library(survminer)
library(gtsummary)
library(here)
library(expss)
library(labelled)

opts_chunk$set(tidy = F)
```

```{r, echo = F}
# Read in the original data
frame_data <- read_csv(here("Project3", "Data", "frmgham2.csv"),
                       show_col_types = FALSE)

# Read in created survival df
surv_df <- read_csv(here("Project3", "Data", "frmg_survival_df.csv"),
                    show_col_types = FALSE)

# Convert factor vars to factors
surv_df <- surv_df %>% 
  mutate(across(c(RANDID, SEX, DIABETES, 
                  deathwithin10, PREVCHD, BPMEDS, CURSMOKE), 
                as.factor))
# Make sure STROKE10 is numeric
surv_df$STROKE10 <- as.numeric(as.character(surv_df$STROKE10))

```

# Introduction
\vspace{-6mm}

The Framington Heart Study is a longitudinal and prospective study with the overarching goal of investigating the etiology of cardiovascular disease in the community of Framingham, Massachusetts. The study was conducted on an initial population of 5,209 individuals on which data was collected for health behaviors, medication use, clinical measurements such as blood pressure, and more. The primary goal of this analysis is to identify risk factors of stroke within a 10-year period for five different risk profiles of interest: 1.) Subjects with no risk factors for stroke, 2.) Subjects with high blood pressure (BP) (> 60 (mmHg) ), 3.) Subjects with diabetes, 4.) Subjects with high blood pressure and diabetes, 5.) Subjects who were smokers. All risk factors are measured at baseline. Additionally, we will summarize the change in the risk factors over the whole duration of the study period to determine whether a more complex time-varying covariate analysis is warranted. 
This report covers the preliminary findings of our analysis answering the primary question of interest. It fills in the results from the data analysis plan agreed upon in the Data Analysis Plan report.

\vspace{-10mm}
# Methods

\vspace{-6mm}
## Data Cleaning
\vspace{-6mm}

The raw data contained 11,627 observations for 4,434 subjects and 38 variables. Missingness was assessed in the whole data set thruogh summaries and visualizations. Regarding the primary variables of interest, BP Medication had ~5% missginness, total cholesterol had 3.5% missingness, and BMI had ~0.5% missingness. We verified that there were no relationships between missingness and any variables and therefore concluded that data was Missing Completely at Random.
We then proceeded with creating new variables specifically for the 10-year window of interest for the primary analysis. We created an indicator variable for whether subjects had a stroke with 10 years (Yes/No) as well as the time to stroke for these events. 
The data for sex-specific analyses consisted of 2,378 observations for females and 2,056 observationsfor males. We used a complete case data frame, removing any rows with missing data. 

\vspace{-8mm}
## Data Analysis
\vspace{-6mm}

We performed a preliminary descriptive analysis of survival curves by sex with visualizations of the Kaplan Meier curves before fitting the primary Cox model.
The preliminary Kaplan-Meier analysis reveled frequent crossing between the separate male and female survival curves. This indicated a violation of the proportional hazards assumption and therefore necessitates stratified Cox models to obtain the sex-specific risk differences. 
We performed LASSO variable selection using cross validation separately for each sex. For males, only antihypertensive medication use was retained, suggesting it is the strongest predictor of stroke risk among the variables considered.
For females, in addition to antihypertensive medication, additional predictors including prevalent CHD, total cholesterol, and BMI were retained in females. 
For the primary analysis, we kept the LASSO selected variables specific to each sex. 
In regard to model assumptions, Schoenfeld residiual test for males and females indicated no evidence of violation of the proportional hazards assumption for individual covariates for globally (global p = 0.714).
The LML Plot for Diabetes status, use of antihypertensive medication, and prevalent CHD visually satisfied the PH assumption for categorical variables as the curves for the different strata were parallel. 
To identify risk factors that are associated with stroke in a 10-year period, we conducted a Survival analysis with stroke as the event and days to stroke as the time to event. Specifically, we fit a Cox Proportional Hazards (PH) Survival Model to identify important risk factors of stroke and obtain the 10-year probability of stroke based on the five risk profiles. 
We adjusted for Age, Diabetes status (Yes/No), and Systolic BP in the models for both sexes and incorporated the LASSO selected variables specific to each sex. 
Lastly, we created a table describing how variables such as Systolic BP and BMI change over over time. 
We used a significance level of $\alpha = 0.05$ for the significance threshold. All analyses were conducted with R version 4.1.1 using the glmnet, survival, and survminer packages. 

# Results
\vspace{-6mm}

<!------------------------------------ Table 1 -------------------------------->

```{r, echo = F, fig.pos='H'}
##################################### Table 1 ##################################

# Fix levels for categorical variables for Table1
# Data frame for table1
tab1_data <- surv_df

# Sex
tab1_data$factor_SEX <- factor(tab1_data$SEX,
                        levels = c(1, 2),
                        labels = c("Male", "Female"))

# Diabetes
tab1_data$factor_DIABETES <- factor(tab1_data$DIABETES,
                                    levels = c(0, 1),
                                    labels = c("Not a Diabetic", "Diabetic"))

# Prevchd
tab1_data$factor_PREVCHD <- factor(tab1_data$PREVCHD,
                                   levels = c(0, 1),
                                   labels = c("Free of CHD",
                                              "Prevalent CHD"))

# Cursmoke
tab1_data$factor_CURSMOKE <- factor(tab1_data$CURSMOKE,
                             levels = c(0, 1),
                             labels = c("Not a Current Smoker",
                                        "Current Smoker"))

# BPMeds
tab1_data$factor_BPMEDS <- factor(tab1_data$BPMEDS,
                                  levels = c(0, 1),
                                  labels = c("No Use of Medication",
                                             "Use of Medication"))

# Stroke 10
tab1_data$factor_STROKE <- factor(tab1_data$STROKE10,
                                  levels = c(0, 1),
                                  labels = c("No Stroke",
                                             "Stroke"))



# Relabel the variables for the Table 1
tab1_data <- apply_labels(tab1_data,
                     factor_STROKE = "Stroke Within 10 Yrs",
                     factor_SEX = "Sex",
                     factor_BPMEDS = "Use of Anti-hypertensive Medication",
                     factor_PREVCHD = "Prevalent Coronary Heart Disease",
                     factor_CURSMOKE = "Cigarrete Smoking Status",
                     AGE = "Age",
                     SYSBP = "Systolic Blood Pressure (mmHg)",
                     STROKE10TIME = "Time to Stroke (Days)",
                     TOTCHOL = "Serum Total Cholesterol (mg/dL)",
                     BMI = "Body Mass Index"
                     )

# Display the modified Table1
tab1_data %>% 
  tbl_summary(include = c(factor_STROKE,
                          STROKE10TIME,
                          AGE,
                          factor_BPMEDS,
                          factor_PREVCHD,
                          factor_CURSMOKE,
                          SYSBP,
                          TOTCHOL,
                          BMI),
              by = factor_SEX,
              statistic = list(c(all_continuous()) 
                               ~ "{mean}({sd})")) %>% 
  modify_header(label = "**Sex**") %>% 
  bold_labels() %>% 
  modify_caption(caption = "Patient Characteristics At Baseline")
  
  
  # as_kable_extra(booktabs = TRUE, linesep = "") %>%
  # kableExtra::kable_styling(latex_options = c("hold_position"))
```

Table 1 contains a description of the data used for all analyses. 
The data is stratified by sex, n and \% of patients is displayed for the categorical variables, and the mean and standard deviation is shown for each continuous variable.
We note the small number of subjects across both sexes that had a stroke within the 10 year time span. In addition, the use of Anti-hypertensive medication and prevalent coronary heart disease groups are noticeably unbalanced. 
Table 2 displays the Hazard Ratios (HRs), 95\% Confidence Intervals (CIs), and p-values for the sex-specific models.
For females, Age and Systolic BP were significantly associated with a higher risk of stroke. Additional increases in age were associated with a 7% higher risk while higher values of systolic BP were associated with a 3% higher risk of stroke. The remaining variables were not significantly associated with risk of stroke for the 10 year period. 
For males, Diabetes status and Systolic BP were significantly associated with a higher risk of stroke. Higher values for age were associated with an increased stroke risk of 6% and higher Systolic BP increased stroke risk by 3%. 

<!----------------------------------------------------------------------------->
<!--------------------------------- KM Curves --------------------------------->
<!----------------------------------------------------------------------------->


```{r, echo = F, eval = F, results='hide'}
################################ Prelim Cox Analysis ###########################
# Create some Kaplan Meier curves by covariates

### Sex
# Fit model
km_sex <- survfit(Surv(STROKE10TIME, STROKE10) ~ SEX,
                   data = surv_df)

# Plot KM curves
ggsurvplot(km_sex, 
           conf.int = FALSE,
           censor = FALSE,
           xlim = c(0, max(surv_df$STROKE10TIME)),
           ylim = c(0.95, 1))
# For the most part, seems that survival curves are quite similar between sexes
# We do see some crossing of the curves though

```


```{r, echo = F, results='hide', warning=F, message=F, eval = F}
# LML Plots
ggsurvplot(km_sex,
           conf.int = F, 
           censor = F,
           fun="cloglog")
# We see that the lines are not parallel and this suggests that the PH assumption is not met
```

<!----------------------------------------------------------------------------->
<!------------------------------ Variable Selection --------------------------->
<!----------------------------------------------------------------------------->

```{r, echo = F, include=FALSE}
############################## Variable Selection ##############################
library(glmnet)
# Vignette: https://glmnet.stanford.edu/articles/Coxnet.html

#----------------------------------- Males ------------------------------------#

# Perform some variable selection prevchd, bpmeds, smokecur, total chol, and BMI

# Males only df
males_df <- surv_df %>% 
  filter(SEX == 1)

# Fit model
y_m <- Surv(time = males_df$STROKE10TIME, 
            event = males_df$STROKE10)

# Create the model matrix for glmnet
x_m <- model.matrix(
  ~ PREVCHD + BPMEDS + CURSMOKE + TOTCHOL + BMI,
  data = males_df)[, -1]  # Remove the intercept

# Lasso - cross validation
cv_m <- cv.glmnet(x_m, y_m, 
                  family = "cox", 
                  alpha = 1)

# Plot to see the optimal lambda value
plot(cv_m)
# Looks pretty flat across all top and bottom values

# Which variables were kept with 1se
## The largest lambda within 1 SE of the minimum error 
## More penalty -> simplier model, more stable and more conservaitive,
## Provides the best simple model!

coef_1se_males <- coef(cv_m, s = "lambda.1se")
#           1
# PREVCHD1  .
# BPMEDS1   .
# CURSMOKE1 .
# TOTCHOL   .
# BMI       .

# Get the nonzero variables
m_1se <- which(as.vector(coef_1se_males) != 0)
m_selected_1se <- rownames(coef_1se_males)[m_1se]
# character(0)  -----  NO variables kept
## None of the predictors add enough predictve value to justify including them

# Which variables were kept with lambda min
## The value of lambda that gives the lowest corss-validated error
## More flexble model -> keeps more variables
## Higher risk of overfitting though
## Provides the best fit
coef_min_males <- coef(cv_m, s = "lambda.min")
# > #           1
# > # PREVCHD1  .
# > # BPMEDS1   .
# > # CURSMOKE1 .
# > # TOTCHOL   .
# > # BMI       .
# > coef(cv_m, s = "lambda.min")
# 5 x 1 sparse Matrix of class "dgCMatrix"
#                  1
# PREVCHD1  .       
# BPMEDS1   1.217418
# CURSMOKE1 .       
# TOTCHOL   .       
# BMI       .

# Get the nonzero variables
m_min <- which(as.vector(coef_min_males) != 0)
m_selected_min <- rownames(coef_min_males)[m_min]
# character(0)  -----  NO variables kept
# [1] "BPMEDS1"

### Based on minimum, keep BPMEDS for males - the only variable with meaningful
### predictive signal for males
```


```{r, echo = F, include=FALSE}
############################## Variable Selection ##############################
library(glmnet)
# Vignette: https://glmnet.stanford.edu/articles/Coxnet.html

#----------------------------------- Females ----------------------------------#

# Perform some variable selection prevchd, bpmeds, smokecur, total chol, and BMI

# Females only df
females_df <- surv_df %>% 
  filter(SEX == 2)

# Fit model
y_f <- Surv(time = females_df$STROKE10TIME, 
            event = females_df$STROKE10)

# Create the model matrix for glmnet
x_f <- model.matrix(
  ~ PREVCHD + BPMEDS + CURSMOKE + TOTCHOL + BMI,
  data = females_df)[, -1]  # Remove the intercept

# Lasso - cross validation
cv_f <- cv.glmnet(x_f, y_f, 
                  family = "cox", 
                  alpha = 1)

# Plot to see the optimal lambda value
plot(cv_f)
# Looks pretty flat across as well all top and bottom values

# Which variables were kept with 1se
## The largest lambda within 1 SE of the minimum error 
## More penalty -> simplier model, more stable and more conservaitive,
## Provides the best simple model!

coef_1se_females <- coef(cv_f, s = "lambda.1se")
#           1
# PREVCHD1  .
# BPMEDS1   .
# CURSMOKE1 .
# TOTCHOL   .
# BMI       .

# Get the nonzero variables
f_1se <- which(as.vector(coef_1se_females) != 0)
f_selected_1se <- rownames(coef_1se_females)[f_1se]
# character(0)  -----  NO variables kept
## None of the predictors add enough predictve value to justify including them

# Which variables were kept with lambda min
## The value of lambda that gives the lowest corss-validated error
## More flexble model -> keeps more variables
## Higher risk of overfitting though
## Provides the best fit
coef_min_females <- coef(cv_f, s = "lambda.min")
#                     1
# PREVCHD1  1.051070381
# BPMEDS1   1.235644137
# CURSMOKE1 .          
# TOTCHOL   0.002948321
# BMI       0.041851181

# Get the nonzero variables
f_min <- which(as.vector(coef_min_females) != 0)
f_selected_min <- rownames(coef_min_females)[f_min]
# "PREVCHD1" "BPMEDS1"  "TOTCHOL"  "BMI" 

### Based on minimum, keep PRECHD, BPMEDS, TOTCHOL, and BMI
### Remove CURSMOKE
### The only variable with meaningful predictive signal for females
```


<!----------------------------------------------------------------------------->
<!----------------------------------- FEMALES --------------------------------->
<!----------------------------------------------------------------------------->

```{r, echo = F, warning=F, message=F}
################################## Model Fitting ###############################

##################################### Females ##################################

# Model for females
females_fit <- coxph(Surv(STROKE10TIME, STROKE10) ~ AGE + DIABETES 
                     + SYSBP + PREVCHD + BPMEDS + TOTCHOL + BMI,
  data = females_df
)

# Summary of general model
females_sum1 <- summary(females_fit)
```


```{r, echo = F, results='hide', eval = F}
################################ Test PH assumptions ##########################
## Schoenfeld residuals :Do the effects of each covariate change over time?
## H0: PH holds - no time varying effect
## H1: Violation of PH

# Schoenfeld residuals
females_res <- cox.zph(females_fit)

# Plot
ggcoxzph(females_res,
         se = F)

```

```{r, echo = F, results='hide', eval = F}
# LML Plots for each categorical variable

# Fit model for DIABETES
females_diabetes <- survfit(coxph(Surv(STROKE10TIME, STROKE10) ~
                                    strata(DIABETES),
                                  data = females_df))
# LML Plot
plot(females_diabetes, 
     fun = "cloglog", 
     col = c("red", "blue"), 
     main = "LML Plot for Diabetes",
     xlab = "Log(Time)", 
     ylab = "Log(-Log(Survival))")
# Assumption met for Sex

# Fit model for PREVCHD
females_prevchd <- survfit(coxph(Surv(STROKE10TIME, STROKE10) ~ 
                                   strata(PREVCHD),
                           data = females_df))

# LML Plot
plot(females_prevchd, 
     fun = "cloglog", 
     col = c("red", "blue"), 
     main = "LML Plot for PREVCHD",
     xlab = "Log(Time)", 
     ylab = "Log(-Log(Survival))")
# Assumption met for prevchd

# Fit Model for BPMEDS
females_bpmeds <- survfit(coxph(Surv(STROKE10TIME, STROKE10) ~ 
                                  strata(BPMEDS),
                          data = females_df))

# LML Plot
plot(females_bpmeds, 
     fun = "cloglog", 
     col = c("red", "blue"), 
     main = "LML Plot for BPMEDS",
     xlab = "Log(Time)", 
     ylab = "Log(-Log(Survival))")
# Assumption met for bpmeds
```


```{r, echo = F}
##### Create data frame of results
females_table <- females_fit %>% 
  tbl_regression(intercept = TRUE, 
                 conf.level = .95,
                 exponentiate = TRUE,
                 pvalue_fun = label_style_pvalue(digits = 2),
                 conf.int = TRUE,
                 show_single_row = c(DIABETES, PREVCHD, BPMEDS),
                 label = list(
                   AGE ~ "Age",
                   DIABETES ~ "Diabetes Status",
                   PREVCHD ~ "Prevalent CHD",
                   BPMEDS ~ "BP Medication",
                   SYSBP ~ "Systolic BP",
                   TOTCHOL ~ "Total Cholesterol"))
```


```{r, echo = F, eval = F, results='hide', warning=F, message=F}
# Plot survival curve - hmmm
ggsurvplot(survfit(females_fit),
           data = females_df,
           censor = F,
           conf.int = T,
           ylim = c(0.95, 1))
```



```{r, echo = F, results='hide'}
################################# Risk Profile 1 ###############################

# Create data frame for the different risk profiles
# Risk profile 1: 
## Average age, no diabetes, average sysbp, no prechd, no bpmeds, no cursmoke,
## average totchol, average bmi
females_rp1 <- data.frame(
    AGE = c(40, 50, 60),
    DIABETES = factor(rep(0, 3), 
                      levels = levels(females_df$DIABETES)),
    SYSBP = rep(mean(females_df$SYSBP), 3),
    PREVCHD = factor(rep(0, 3), 
                     levels = levels(females_df$PREVCHD)),
    BPMEDS = factor(rep(0, 3), 
                    levels = levels(females_df$BPMEDS)),
    CURSMOKE = factor(rep(0, 3), 
                      levels = levels(females_df$CURSMOKE)),
    TOTCHOL = rep(mean(females_df$TOTCHOL, 
                       na.rm = TRUE), 3),
    BMI = rep(mean(females_df$BMI), 3)
)

# Fit cox model for risk profile 1
mrp1_fit <- survfit(females_fit, newdata = females_rp1)


# Estimates for 10 years
mrp1_summary <- summary(mrp1_fit, times = 365*10)
# time n.risk n.event survival1 survival2 survival3
# 3650   2177      57     0.995     0.988     0.976

# Data frame of survival 10 years and risk 10 years
mrp1_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(mrp1_summary$surv), 3),
                           Risk = paste0(round(1 - mrp1_summary$surv, 3)*100,
                                         "%")
)

```


```{r, echo = F, results='hide'}
################################# Risk Profile 2 ###############################

# Risk profile 2: 
## High blood pressure and no other risk factors
females_rp2 <- data.frame(
  AGE = c(40, 50, 60),
  DIABETES = factor(rep(0, 3), 
                    levels = levels(females_df$DIABETES)),
  SYSBP = rep(200, 3),
  PREVCHD = factor(rep(0, 3), 
                   levels = levels(females_df$PREVCHD)),
  BPMEDS = factor(rep(0, 3), 
                  levels = levels(females_df$BPMEDS)),
  CURSMOKE = factor(rep(0, 3), 
                    levels = levels(females_df$CURSMOKE)),
  TOTCHOL = rep(mean(females_df$TOTCHOL, 
                     na.rm = TRUE), 3),
  BMI = rep(mean(females_df$BMI), 3)
)

# AGE DIABETES SYSBP PREVCHD BPMEDS CURSMOKE  TOTCHOL      BMI
# 1  40        0   180       0      0        0 239.5698 25.56664
# 2  50        0   230       0      0        0 239.5698 25.56664
# 3  60        0   280       0      0        0 239.5698 25.56664


# Fit cox model for risk profile 2
mrp2_fit <- survfit(females_fit, newdata = females_rp2)

# Estimates for 10 years
mrp2_summary <- summary(mrp2_fit, times = 365*10)
# time n.risk n.event survival1 survival2 survival3
# 3650   2177      57     0.983     0.878     0.382

# Create data frame of survival 10 years and risk 10 years
mrp2_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(mrp2_summary$surv), 3),
                           Risk = paste0(round(1 - mrp2_summary$surv, 3)*100,
                                         "%")
)
# Age Survival  Risk
# 1  40    0.983  1.7%
# 2  50    0.878 12.2%
# 3  60    0.382 61.8%


```


```{r, echo = F, results='hide'}

################################# Risk Profile 3 ###############################

#### Risk profile 3:
## Diabetes and no other risk factors
females_rp3 <- data.frame(
  AGE = c(40, 50, 60),
  DIABETES = factor(rep(1, 3), 
                    levels = levels(females_df$DIABETES)),
  SYSBP = rep(mean(females_df$SYSBP), 3),
  PREVCHD = factor(rep(0, 3), 
                   levels = levels(females_df$PREVCHD)),
  BPMEDS = factor(rep(0, 3), 
                  levels = levels(females_df$BPMEDS)),
  CURSMOKE = factor(rep(0, 3), 
                    levels = levels(females_df$CURSMOKE)),
  TOTCHOL = rep(mean(females_df$TOTCHOL, 
                     na.rm = TRUE), 3),
  BMI = rep(mean(females_df$BMI), 3)
)
# AGE DIABETES    SYSBP PREVCHD BPMEDS CURSMOKE  TOTCHOL      BMI
# 1  40        1 133.5524       0      0        0 239.5698 25.56664
# 2  50        1 133.5524       0      0        0 239.5698 25.56664
# 3  60        1 133.5524       0      0        0 239.5698 25.56664


# Fit cox model for risk profile 3
mrp3_fit <- survfit(females_fit, newdata = females_rp3)

# Estimates for 10 years
mrp3_probs <- summary(mrp3_fit, times = 365*10)
# time n.risk n.event survival1 survival2 survival3
# 3650   2177      57      0.99     0.979     0.955


# Estimates for 10 years
mrp3_summary <- summary(mrp3_fit, times = 365*10)
# time n.risk n.event survival1 survival2 survival3
# 3650   2177      57      0.99     0.979     0.955

# Create data frame of survival 10 years and risk 10 years
mrp3_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(mrp3_summary$surv), 3),
                           Risk = paste0(round(1 - mrp3_summary$surv, 3)*100,
                                         "%")
)

# Age Survival Risk
# 1  40    0.990   1%
# 2  50    0.979 2.1%
# 3  60    0.955 4.5%

```


```{r, echo = F, results='hide'}
################################# Risk Profile 4 ###############################

#### Risk profile 4:
## Diabetes and no other risk factors
females_rp4 <- data.frame(
  AGE = c(40, 50, 60),
  DIABETES = factor(rep(1, 3), 
                    levels = levels(females_df$DIABETES)),
  SYSBP = rep(200, 3),
  PREVCHD = factor(rep(0, 3), 
                   levels = levels(females_df$PREVCHD)),
  BPMEDS = factor(rep(0, 3), 
                  levels = levels(females_df$BPMEDS)),
  CURSMOKE = factor(rep(0, 3), 
                    levels = levels(females_df$CURSMOKE)),
  TOTCHOL = rep(mean(females_df$TOTCHOL, 
                     na.rm = TRUE), 3),
  BMI = rep(mean(females_df$BMI), 3)
)
# 
# AGE DIABETES SYSBP PREVCHD BPMEDS CURSMOKE  TOTCHOL      BMI
# 1  40        1   180       0      0        0 239.5698 25.56664
# 2  50        1   230       0      0        0 239.5698 25.56664
# 3  60        1   280       0      0        0 239.5698 25.56664

# Fit cox model for risk profile 3
mrp4_fit <- survfit(females_fit, newdata = females_rp4)

# Estimates for 10 years
mrp4_summary <- summary(mrp4_fit, times = 365*10)
# time n.risk n.event survival1 survival2 survival3
# 3650   2177      57      0.99     0.979     0.955

# Create data frame of survival 10 years and risk 10 years
mrp4_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(mrp4_summary$surv), 3),
                           Risk = paste0(round(1 - mrp4_summary$surv, 3)*100,
                                         "%")
)

# Age Survival  Risk
# 1  40    0.968  3.2%
# 2  50    0.785 21.5%
# 3  60    0.167 83.3%

```


```{r, echo = F, results='hide'}
################################# Risk Profile 5 ###############################

#### Risk profile 4:
## Smoker and no other risk factors
females_rp5 <- data.frame(
  AGE = c(40, 50, 60),
  DIABETES = factor(rep(0, 3), 
                    levels = levels(females_df$DIABETES)),
  SYSBP = rep(mean(females_df$SYSBP), 3),
  PREVCHD = factor(rep(0, 3), 
                   levels = levels(females_df$PREVCHD)),
  BPMEDS = factor(rep(0, 3), 
                  levels = levels(females_df$BPMEDS)),
  CURSMOKE = factor(rep(1, 3), 
                    levels = levels(females_df$CURSMOKE)),
  TOTCHOL = rep(mean(females_df$TOTCHOL, 
                     na.rm = TRUE), 3),
  BMI = rep(mean(females_df$BMI), 3)
)
# 
# AGE DIABETES    SYSBP PREVCHD BPMEDS CURSMOKE  TOTCHOL      BMI
# 1  40        0 133.5524       0      0        1 239.5698 25.56664
# 2  50        0 133.5524       0      0        1 239.5698 25.56664
# 3  60        0 133.5524       0      0        1 239.5698 25.56664

# Fit cox model for risk profile 3
mrp5_fit <- survfit(females_fit, newdata = females_rp5)

# Estimates for 10 years
mrp5_summary <- summary(mrp5_fit, times = 365*10)
# time n.risk n.event survival1 survival2 survival3
# 3650   2177      57     0.992     0.984     0.966

# Create data frame of survival 10 years and risk 10 years
mrp5_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(mrp5_summary$surv), 3),
                           Risk = paste0(round(1 - mrp5_summary$surv, 3)*100,
                                         "%")
)

# Age Survival Risk
# 1  40    0.992 0.8%
# 2  50    0.984 1.6%
# 3  60    0.966 3.4%
```


```{r, echo = F}
################### Combine these all for the report
library(knitr)
library(kableExtra)

# Create individual kables for each df
# Results 1
fk1 <- kable(mrp1_results, caption = "meow1",
             format = "html") %>% 
  kable_styling()

# Results 2
fk2 <- kable(mrp2_results, caption = "meow2",
             format = "html") %>% 
  kable_styling()

# Results 3
fk3 <- kable(mrp3_results, caption = "meow3",
             format = "html") %>% 
  kable_styling()

# Results 4
fk4 <- kable(mrp4_results, caption = "meow4",
             format = "html") %>% 
  kable_styling()

# Combimne results into one
female_results <- cbind(mrp1_results, 
                       mrp2_results[, 2:3],
                       mrp3_results[,2:3],
                       mrp4_results[,2:3],
                       mrp5_results[,2:3]
                       )

```


<!----------------------------------------------------------------------------->
<!------------------------------------ MALES ---------------------------------->
<!----------------------------------------------------------------------------->


```{r, echo = F, warning=F, message=F, fig.cap="Model Results for Males"}
################################## Model Fitting ###############################

##################################### Males ##################################
# Model for males
males_fit <- coxph(
  Surv(STROKE10TIME, STROKE10) ~ AGE + DIABETES + SYSBP
  + BPMEDS,
  data = males_df
)

```

```{r, echo = F, eval = F, results='hide', warning=F, message=F}
# Plot survival curve - hmmm
ggsurvplot(survfit(males_fit),
           data = males_df,
           censor = F,
           conf.int = T)
```


```{r, echo = F, include=FALSE, results='hide'}
################################ Test PH assumptions ##########################
## Schoenfeld residuals :Do the effects of each covariate change over time?
## H0: PH holds - no time varying effect
## H1: Violation of PH

# Schoenfeld residuals
males_res <- cox.zph(males_fit)
#          chisq df    p
# AGE      0.467  1 0.49
# DIABETES 2.636  1 0.10
# SYSBP    1.653  1 0.20
# BPMEDS   0.101  1 0.75
# GLOBAL   4.204  4 0.38

# Plot
ggcoxzph(males_res,
         se = F)
# Looks good

```


```{r, echo = F, results='hide', eval = F}
# LML Plots for each categorical variable

# Fit model for DIABETES
males_diabetes <- survfit(coxph(Surv(STROKE10TIME, STROKE10) ~
                                    strata(DIABETES),
                                  data = males_df))
# LML Plot
plot(males_diabetes, 
     fun = "cloglog", 
     col = c("red", "blue"), 
     main = "LML Plot for Diabetes",
     xlab = "Log(Time)", 
     ylab = "Log(-Log(Survival))")
# Assumption met for Sex

# Fit model for PREVCHD
males_prevchd <- survfit(coxph(Surv(STROKE10TIME, STROKE10) ~ 
                                   strata(PREVCHD),
                           data = males_df))

# LML Plot
plot(males_prevchd, 
     fun = "cloglog", 
     col = c("red", "blue"), 
     main = "LML Plot for PREVCHD",
     xlab = "Log(Time)", 
     ylab = "Log(-Log(Survival))")
# Assumption met for prevchd

# Fit Model for BPMEDS
males_bpmeds <- survfit(coxph(Surv(STROKE10TIME, STROKE10) ~ 
                                  strata(BPMEDS),
                          data = males_df))

# LML Plot
plot(males_bpmeds, 
     fun = "cloglog", 
     col = c("red", "blue"), 
     main = "LML Plot for BPMEDS",
     xlab = "Log(Time)", 
     ylab = "Log(-Log(Survival))")
# Assumption met for bpmeds
```


```{r, echo = F}
##### Create data frame of results
females_table <- females_fit %>% 
  tbl_regression(intercept = TRUE, 
                 conf.level = .95,
                 exponentiate = TRUE,
                 pvalue_fun = label_style_pvalue(digits = 2),
                 conf.int = TRUE,
                 show_single_row = c(DIABETES, PREVCHD, BPMEDS),
                 label = list(
                   AGE ~ "Age",
                   DIABETES ~ "Diabetes Status",
                   PREVCHD ~ "Prevalent CHD",
                   BPMEDS ~ "BP Medication",
                   SYSBP ~ "Systolic BP",
                   TOTCHOL ~ "Total Cholesterol"))
```


```{r, echo = F}
# Summary of general model
males_sum1 <- summary(males_fit)

# Create data frame of results
males_table <- males_fit %>% 
    tbl_regression(intercept = TRUE, 
                 conf.level = .95,
                 exponentiate = TRUE,
                 pvalue_fun = label_style_pvalue(digits = 2),
                 conf.int = TRUE,
                 show_single_row = c(DIABETES, BPMEDS),
                 label = list(
                   AGE ~ "Age",
                   DIABETES ~ "Diabetes Status",
                   SYSBP ~ "Systolic BP",
                   BPMEDS ~ "BP Medication")) 
```


```{r, echo = F}

tbl_side_by_side <- tbl_merge(
  tbls = list(females_table, males_table),
  tab_spanner = c("**Females**", "**Males**"),
  quiet = TRUE
)

tbl_side_by_side %>%
  modify_caption("Model Results by Sex") %>%
  as_gt() %>%
  gt::tab_source_note(
    gt::md("*BP = Blood Pressure; CHD = Coronary Heart Disease")
  ) %>%
  gt::tab_source_note(
    gt::md("**All factor variable HRs are compared to reference groups")
  ) %>%
  gt::tab_options(
    table.font.size = "small",
    data_row.padding = gt::px(2)
  )
```

```{r, echo = F, results='hide'}
################################# Risk Profile 1 ###############################

# Create data frame for the different risk profiles
# Risk profile 1: 
## Average age, no diabetes, average sysbp, no prechd, no bpmeds, no cursmoke,
## average totchol, average bmi
males_rp1 <- data.frame(
    AGE = c(40, 50, 60),
    DIABETES = factor(rep(0, 3), 
                      levels = levels(males_df$DIABETES)),
    SYSBP = rep(mean(males_df$SYSBP), 3),
    PREVCHD = factor(rep(0, 3), 
                     levels = levels(males_df$PREVCHD)),
    BPMEDS = factor(rep(0, 3), 
                    levels = levels(males_df$BPMEDS)),
    CURSMOKE = factor(rep(0, 3), 
                      levels = levels(males_df$CURSMOKE)),
    TOTCHOL = rep(mean(males_df$TOTCHOL, 
                       na.rm = TRUE), 3),
    BMI = rep(mean(males_df$BMI), 3)
)

# Fit cox model for risk profile 1
mrp1_fit <- survfit(males_fit, newdata = males_rp1)


# Estimates for 10 years
mrp1_summary <- summary(mrp1_fit, times = 365*10)
 # time n.risk n.event survival1 survival2 survival3
 # 3650   1647      48     0.995      0.99     0.981

# Data frame of survival 10 years and risk 10 years
mrp1_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(mrp1_summary$surv), 3),
                           Risk = paste0(round(1 - mrp1_summary$surv, 3)*100,
                                         "%")
)

```


```{r, echo = F, results='hide'}
################################# Risk Profile 2 ###############################

# Risk profile 2: 
## High blood pressure and no other risk factors
males_rp2 <- data.frame(
  AGE = c(40, 50, 60),
  DIABETES = factor(rep(0, 3), 
                    levels = levels(males_df$DIABETES)),
  SYSBP = rep(200, 3),
  PREVCHD = factor(rep(0, 3), 
                   levels = levels(males_df$PREVCHD)),
  BPMEDS = factor(rep(0, 3), 
                  levels = levels(males_df$BPMEDS)),
  CURSMOKE = factor(rep(0, 3), 
                    levels = levels(males_df$CURSMOKE)),
  TOTCHOL = rep(mean(males_df$TOTCHOL, 
                     na.rm = TRUE), 3),
  BMI = rep(mean(males_df$BMI), 3)
)

# AGE DIABETES SYSBP PREVCHD BPMEDS CURSMOKE  TOTCHOL      BMI
# 1  40        0   180       0      0        0 233.7359 26.16627
# 2  50        0   230       0      0        0 233.7359 26.16627
# 3  60        0   280       0      0        0 233.7359 26.16627


# Fit cox model for risk profile 2
mrp2_fit <- survfit(males_fit, newdata = males_rp2)

# Estimates for 10 years
mrp2_summary <- summary(mrp2_fit, times = 365*10)
 # time n.risk n.event survival1 survival2 survival3
 # 3650   1647      48     0.976     0.795     0.114

# Create data frame of survival 10 years and risk 10 years
mrp2_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(mrp2_summary$surv), 3),
                           Risk = paste0(round(1 - mrp2_summary$surv, 3)*100,
                                         "%")
)
# Age Survival  Risk
# 1  40    0.976  2.4%
# 2  50    0.795 20.5%
# 3  60    0.114 88.6%


```


```{r, echo = F, results='hide'}

################################# Risk Profile 3 ###############################

#### Risk profile 3:
## Diabetes and no other risk factors
males_rp3 <- data.frame(
  AGE = c(40, 50, 60),
  DIABETES = factor(rep(1, 3), 
                    levels = levels(males_df$DIABETES)),
  SYSBP = rep(mean(males_df$SYSBP), 3),
  PREVCHD = factor(rep(0, 3), 
                   levels = levels(males_df$PREVCHD)),
  BPMEDS = factor(rep(0, 3), 
                  levels = levels(males_df$BPMEDS)),
  CURSMOKE = factor(rep(0, 3), 
                    levels = levels(males_df$CURSMOKE)),
  TOTCHOL = rep(mean(males_df$TOTCHOL, 
                     na.rm = TRUE), 3),
  BMI = rep(mean(males_df$BMI), 3)
)
# AGE DIABETES    SYSBP PREVCHD BPMEDS CURSMOKE  TOTCHOL      BMI
# 1  40        1 131.5422       0      0        0 233.7359 26.16627
# 2  50        1 131.5422       0      0        0 233.7359 26.16627
# 3  60        1 131.5422       0      0        0 233.7359 26.16627


# Fit cox model for risk profile 3
mrp3_fit <- survfit(males_fit, newdata = males_rp3)

# Estimates for 10 years
mrp3_probs <- summary(mrp3_fit, times = 365*10)
 # time n.risk n.event survival1 survival2 survival3
 # 3650   1647      48     0.974      0.95     0.905


# Estimates for 10 years
mrp3_summary <- summary(mrp3_fit, times = 365*10)
 # time n.risk n.event survival1 survival2 survival3
 # 3650   1647      48     0.974      0.95     0.905

# Create data frame of survival 10 years and risk 10 years
mrp3_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(mrp3_summary$surv), 3),
                           Risk = paste0(round(1 - mrp3_summary$surv, 3)*100,
                                         "%")
)

# Age Survival Risk
# 1  40    0.974 2.6%
# 2  50    0.950   5%
# 3  60    0.905 9.5%

```


```{r, echo = F, results='hide'}
################################# Risk Profile 4 ###############################

#### Risk profile 4:
## High BP and Diabetes
males_rp4 <- data.frame(
  AGE = c(40, 50, 60),
  DIABETES = factor(rep(1, 3), 
                    levels = levels(males_df$DIABETES)),
  SYSBP = rep(200, 3),
  PREVCHD = factor(rep(0, 3), 
                   levels = levels(males_df$PREVCHD)),
  BPMEDS = factor(rep(0, 3), 
                  levels = levels(males_df$BPMEDS)),
  CURSMOKE = factor(rep(0, 3), 
                    levels = levels(males_df$CURSMOKE)),
  TOTCHOL = rep(mean(males_df$TOTCHOL, 
                     na.rm = TRUE), 3),
  BMI = rep(mean(males_df$BMI), 3)
)
# 
# AGE DIABETES SYSBP PREVCHD BPMEDS CURSMOKE  TOTCHOL      BMI
# 1  40        1   180       0      0        0 233.7359 26.16627
# 2  50        1   230       0      0        0 233.7359 26.16627
# 3  60        1   280       0      0        0 233.7359 26.16627

# Fit cox model for risk profile 3
mrp4_fit <- survfit(males_fit, newdata = males_rp4)

# Estimates for 10 years
mrp4_summary <- summary(mrp4_fit, times = 365*10)
 # time n.risk n.event survival1 survival2 survival3
 # 3650   1647      48     0.883     0.307  1.44e-05

# Create data frame of survival 10 years and risk 10 years
mrp4_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(mrp4_summary$surv), 3),
                           Risk = paste0(round(1 - mrp4_summary$surv, 3)*100,
                                         "%")
)

# Age Survival  Risk
#   Age Survival  Risk
# 1  40    0.883 11.7%
# 2  50    0.307 69.3%
# 3  60    0.000  100%

```


```{r, echo = F, results='hide'}
################################# Risk Profile 5 ###############################

#### Risk profile 4:
## Smoker and no other risk factors
males_rp5 <- data.frame(
  AGE = c(40, 50, 60),
  DIABETES = factor(rep(0, 3), 
                    levels = levels(males_df$DIABETES)),
  SYSBP = rep(mean(males_df$SYSBP), 3),
  PREVCHD = factor(rep(0, 3), 
                   levels = levels(males_df$PREVCHD)),
  BPMEDS = factor(rep(0, 3), 
                  levels = levels(males_df$BPMEDS)),
  CURSMOKE = factor(rep(1, 3), 
                    levels = levels(males_df$CURSMOKE)),
  TOTCHOL = rep(mean(males_df$TOTCHOL, 
                     na.rm = TRUE), 3),
  BMI = rep(mean(males_df$BMI), 3)
)
# 
# AGE DIABETES    SYSBP PREVCHD BPMEDS CURSMOKE  TOTCHOL      BMI
# 1  40        0 131.5422       0      0        1 233.7359 26.16627
# 2  50        0 131.5422       0      0        1 233.7359 26.16627
# 3  60        0 131.5422       0      0        1 233.7359 26.16627

# Fit cox model for risk profile 3
mrp5_fit <- survfit(males_fit, newdata = males_rp5)

# Estimates for 10 years
mrp5_summary <- summary(mrp5_fit, times = 365*10)
 # time n.risk n.event survival1 survival2 survival3
 # 3650   1647      48      0.99      0.98     0.962

# Create data frame of survival 10 years and risk 10 years
mrp5_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(mrp5_summary$surv), 3),
                           Risk = paste0(round(1 - mrp5_summary$surv, 3)*100,
                                         "%")
)

# Age Survival Risk
# 1  40    0.990   1%
# 2  50    0.980   2%
# 3  60    0.962 3.8%
```


```{r, echo = F}
################### Combine these all for the report
library(knitr)
library(kableExtra)

# Create individual kables for each df
# Results 1
rk1 <- kable(mrp1_results, caption = "meow1",
             format = "html") %>% 
  kable_styling()

# Results 2
rk2 <- kable(mrp2_results, caption = "meow2",
             format = "html") %>% 
  kable_styling()

# Results 3
rk3 <- kable(mrp3_results, caption = "meow3",
             format = "html") %>% 
  kable_styling()

# Results 4
rk4 <- kable(mrp4_results, caption = "meow4",
             format = "html") %>% 
  kable_styling()

# Combimne results into one
male_results <- cbind(mrp1_results, 
                       mrp2_results[, 2:3],
                       mrp3_results[,2:3],
                       mrp4_results[,2:3],
                       mrp5_results[,2:3]
                       )

```

```{r, echo = F}
# Display both male and female tables together

kable(female_results,
  format = "latex",
  booktabs = TRUE,
  digits = 3, 
  caption = "Survival Rates and Risk \\% Over 10 Years for Different 
  Risk Profiles") %>%
  kable_styling(latex_options = c("striped", "hold_position",
                                  "scale_down")) %>%
  add_header_above(c(" ", "Risk Profile 1\nNo Risk factors" = 2,
                     "Risk Profile 2\nHigh BP" = 2,
                     "Risk Profile 3\nDiabetes" = 2,
                     "Risk Profile 4\n High BP & Diabetes" = 2,
                     "Risk Profile 5\n Smoker" = 2)) %>%
  add_header_above(c("Females"
                     = 11),
                   escape = FALSE)


kable(male_results,
  format = "latex",
  booktabs = TRUE,
  digits = 3) %>%
  # caption = "Survival Rates and Risk \\% Over 10 Years for Males") %>%
  kable_styling(latex_options = c("striped", "hold_position",
                                  "scale_down")) %>%
  add_header_above(c(" ", "Risk Profile 1\nNo Risk factors" = 2,
                     "Risk Profile 2\nHigh BP" = 2,
                     "Risk Profile 3\nDiabetes" = 2,
                     "Risk Profile 4\n High BP & Diabetes" = 2,
                     "Risk Profile 5\n Smoker" = 2)) %>%
  add_header_above(c("Males"
                     = 11),
                   escape = FALSE) %>%
  footnote(number = c("BP = Blood Pressure & High BP = 200 (mmHg)",
                      "All Risk Factors are at Baseline Timepoint"))

```

\newpage

<!------------------------------- Results continued -------------------------->

\vspace*{0.6mm}

Age and Diabetes status were not significantly associated with risk of stroke for males. 
Table 3 displays the predicted survival and $\%$ risk over 10 years for males and females aged 40, 50, and 60 for the 5 risk profiles of interest. Figure 1 visualizes the survival probability curves for males.
Comparing the survival probabilities and risks for males and females, 
we see that individuals with no risk factors have survival probabilities greater than 97$\%$ and risks lower than $3\%$. 
The probabilities and risk $\%$ is identical to that of subjects who identifed as being current smokers at baseline. 
The survival probabilities decrease and risk percentages increase for both sexes with high blood pressure, diabetes, or high BP and diabetes.  We observe the sharpest contrast from Risk Profile 1 with Risk Profile 4 with noticeable decreases in survival probabilites and increases in risk. 
Overall, sex differences in survival probabilites and risk are apparent, each additional increment of age from 40 to 60 evidently affects these numbers, and high blood pressure and diabetes appear to be the biggest risk factors for stroke. 

```{r, echo = F, message = F, warning = F, fig.pos="H", eval =F}
library(colorspace)
# Plot the survival curves
m1 <- ggsurvplot(survfit(males_fit, 
                   newdata = males_rp1),
           data = males_df,
           censor = FALSE,
           conf.int = FALSE,
           # ylim = c(0.98, 1),
           xlim = c(0, 3650),
           # risk.table = TRUE,
           ggtheme = theme_minimal(),
           # legend = "none",
           legend.labs = c("40", "50", "60"),
           legend.title = "Age",
           title = "Risk Profile 1 Estimated Survival\nAverage Individual",
           font.main = 10,
           palette = "Set1"
           ) 
# m1

# Plot the survival curves
m2 <- ggsurvplot(survfit(males_fit, 
                   newdata = males_rp2),
           data = males_df,
           censor = F,
           conf.int = FALSE,
           # Min is 0.35 and max is 0.99
           # ylim = c(0.35, 1),
           xlim = c(0, 3650),
           ggtheme = theme_minimal(),
           legend.labs = c("Age 40", "Age 50", "Age 60"),
           legend = "none",
           title = "Risk Profile 2 Estimated Survival\nHigh BP",
           font.main = 10,
           palette = "Set1"
) 
# f2


# Plot the survival curves
m3 <- ggsurvplot(survfit(males_fit, 
                   newdata = males_rp3),
           data = males_df,
           censor = F,
           conf.int = FALSE,
           # Min is 0.35 and max is 0.99
           # ylim = c(0.90, 1),
           xlim = c(0, 3650),
           ggtheme = theme_minimal(),
           legend = "none",
           title = "Risk Profile 2 Estimated Survival\nDiabetes",
           font.main = 10,
           palette = "Set1"
) 
# f3

# Plot the survival curves
m4 <- ggsurvplot(survfit(males_fit, 
                   newdata = males_rp4),
           data = males_df,
           censor = F,
           conf.int = FALSE,
           # Min is 0.35 and max is 0.99
           # ylim = c(0.95, 1),
           xlim = c(0, 3650),
           ggtheme = theme_minimal(),
           legend.labs = c("Age 40", "Age 50", "Age 60"),
           legend = "none",
           title = "Risk Profile 2 Estimated Survival\nHigh BP & Diabetes",
           font.main = 10,
           palette = "Set1"
) 
# f4

# Plot the survival curves
m5 <- ggsurvplot(survfit(males_fit, 
                         newdata = males_rp5),
                 data = males_df,
                 censor = F,
                 conf.int = FALSE,
                 # Min is 0.35 and max is 0.99
                 # ylim = c(0.95, 1),
                 xlim = c(0, 3650),
                 ggtheme = theme_minimal(),
                 legend = "none",
                 title = "Risk Profile 2 Estimated Survival\nSmoker",
                 font.main = 10,
                 palette = "Set1"
) 
# f5

##### Display all plots together

library(ggpubr)
# Using ggarrange
plot2 <- ggarrange(m1$plot, m2$plot, m3$plot, m4$plot, m5$plot,
          ncol = 3, nrow = 2,
          common.legend = TRUE,
          legend = "top",
          legend.labs = c("Age", "Age 50", "Age 60"),
          vjust = 0.5
          ) 
# Add a note at the bottom of this figure
annotate_figure(plot2,
                top = text_grob("Estimated Survival Curves - Males",
                                size = 12,
                                face = "bold"),
                bottom = text_grob("**Y-axis is not the same for every plot",
                                   size = 10,
                                   hjust = 1.3,
                                   face = "italic"))
  

# With survival package - could not get one legend for all plots
# arrange_ggsurvplots(list(f1, f2, f3, f4, f5),
#                     ncol = 3, nrow = 2,
#                     title = "Estimated Survival Curves - males")
```

In regard to how much change there is in the risk factors over the 3 periods of data collection, Table 4 displays the N, mean, and SDs for the variables for the three time periods. 
As to be expected, we do observe a steady decline in the number of study participants for both males and females over time. 
We also note the increase in missing values over time with anti-hypertensive medication, total cholesterol, and BMI. 
We observe increases in the number of subjects using anti-hypertensive medication and prevalent CHD as well as decreases in the number of subjects who smoke over time. Regarding the continuous covariates, Systolic BP steadily increases with time whereas total cholesterol and BMI have relatively flat trends across time. 

```{r, echo = F, fig.pos='H'}
##################################### Table 4 ##################################

# Table describing the change in the risk factors over time

# Fix levels for categorical variables for Table1
# Data frame for table1
tab3_data <- frame_data

# Make sure the factor variables are factors

# Sex
tab3_data$factor_SEX <- factor(tab3_data$SEX,
                        levels = c(1, 2),
                        labels = c("Male", "Female"))

# Diabetes
tab3_data$factor_DIABETES <- factor(tab3_data$DIABETES,
                                    levels = c(0, 1),
                                    labels = c("Not a Diabetic", "Diabetic"))

# Prevchd
tab3_data$factor_PREVCHD <- factor(tab3_data$PREVCHD,
                                   levels = c(0, 1),
                                   labels = c("Free of CHD",
                                              "Prevalent CHD"))

# Cursmoke
tab3_data$factor_CURSMOKE <- factor(tab3_data$CURSMOKE,
                             levels = c(0, 1),
                             labels = c("Not a Current Smoker",
                                        "Current Smoker"))

# BPMeds
tab3_data$factor_BPMEDS <- factor(tab3_data$BPMEDS,
                                  levels = c(0, 1),
                                  labels = c("No Use of Medication",
                                             "Use of Medication"))

# Period
tab3_data$factor_PERIOD <- factor(tab3_data$PERIOD,
                                  levels = c(1, 2, 3),
                                  labels = c("Period 1\n",
                                             "Period 2\n",
                                             "Period 3\n"))


# Relabel the variables for the Table 1
tab3_data <- apply_labels(tab3_data,
                     factor_SEX = "Sex",
                     factor_BPMEDS = "Use of Anti-hypertensive Medication",
                     factor_PREVCHD = "Prevalent Coronary Heart Disease",
                     factor_CURSMOKE = "Cigarrete Smoking Status",
                     AGE = "Age",
                     SYSBP = "Systolic Blood Pressure (mmHg)",
                     TOTCHOL = "Serum Total Cholesterol (mg/dL)",
                     BMI = "Body Mass Index"
                     )

# Display the modified Table1
tab3_data %>% 
  tbl_summary(include = c(factor_SEX,
                          AGE,
                          factor_BPMEDS,
                          factor_PREVCHD,
                          factor_CURSMOKE,
                          SYSBP,
                          TOTCHOL,
                          BMI),
              missing_text = "Missing",
              by = factor_PERIOD,
              statistic = list(c(AGE, SYSBP, TOTCHOL, BMI) 
                               ~ "{mean}({sd})")) %>% 
  modify_header(label = "Study Period") %>% 
  # bold_labels() %>% 
  modify_caption(caption = "Patient Characteristics Across the 
                 Three Periods") %>% 
  # Remove the stars from the 3 study period col headers
  modify_header(all_stat_cols() ~ "{level}\n (N = {n}) ") %>% 
  # Modify to fit in report
  as_kable(
           booktabs = TRUE) %>% 
  kableExtra::kable_styling(latex_options = c("scale_down",
                                              "hold_position")) %>% 
  column_spec(1, bold = TRUE) %>% 
  row_spec(0, bold = TRUE) %>% 
  add_indent(c(2:3, 6:8, 10:11, 13:14, 17, 19))
```

\vspace{-6mm}
## Discussion
\vspace{-6mm}

And then meow meow moew moew moew moew moew mowem ewomeow meow moew

```{r, echo = F, eval = F, message=F, warning=F}
library(colorspace)
# Plot the survival curves
f1 <- ggsurvplot(survfit(females_fit, 
                   newdata = females_rp1),
           data = females_df,
           censor = FALSE,
           conf.int = FALSE,
           ylim = c(0.95, 1),
           xlim = c(0, 3650),
           # risk.table = TRUE,
           ggtheme = theme_minimal(),
           # legend = "none",
           legend.labs = c("40", "50", "60"),
           legend.title = "Age",
           title = "Risk Profile 1 Estimated Survival\nAverage Individual",
           font.main = 10,
           palette = c("deeppink2", "turquoise3", "springgreen3")
           ) 
# f1

# Plot the survival curves
f2 <- ggsurvplot(survfit(females_fit, 
                   newdata = females_rp2),
           data = females_df,
           censor = F,
           conf.int = FALSE,
           # Min is 0.35 and max is 0.99
           ylim = c(0.35, 1),
           xlim = c(0, 3650),
           ggtheme = theme_minimal(),
           legend.labs = c("Age 40", "Age 50", "Age 60"),
           legend = "none",
           title = "Risk Profile 2 Estimated Survival\nHigh BP",
           font.main = 10,
           palette = c("deeppink2", "turquoise3", "springgreen3")
) 
# f2


# Plot the survival curves
f3 <- ggsurvplot(survfit(females_fit, 
                   newdata = females_rp3),
           data = females_df,
           censor = F,
           conf.int = FALSE,
           # Min is 0.35 and max is 0.99
           ylim = c(0.95, 1),
           xlim = c(0, 3650),
           ggtheme = theme_minimal(),
           legend = "none",
           title = "Risk Profile 2 Estimated Survival\nDiabetes",
           font.main = 10,
           palette = c("deeppink2", "turquoise3", "springgreen3")
) 
# f3

# Plot the survival curves
f4 <- ggsurvplot(survfit(females_fit, 
                   newdata = females_rp4),
           data = females_df,
           censor = F,
           conf.int = FALSE,
           # Min is 0.35 and max is 0.99
           # ylim = c(0.95, 1),
           xlim = c(0, 3650),
           ggtheme = theme_minimal(),
           legend.labs = c("Age 40", "Age 50", "Age 60"),
           legend = "none",
           title = "Risk Profile 2 Estimated Survival\nHigh BP & Diabetes",
           font.main = 10,
           palette = c("deeppink2", "turquoise3", "springgreen3")
) 
# f4

# Plot the survival curves
f5 <- ggsurvplot(survfit(females_fit, 
                         newdata = females_rp5),
                 data = females_df,
                 censor = F,
                 conf.int = FALSE,
                 # Min is 0.35 and max is 0.99
                 ylim = c(0.95, 1),
                 xlim = c(0, 3650),
                 ggtheme = theme_minimal(),
                 legend = "none",
                 title = "Risk Profile 2 Estimated Survival\nSmoker",
                 font.main = 10,
                 palette = c("deeppink2", "turquoise3", "springgreen3")
) 
# f5

##### Display all plots together

library(ggpubr)
# Using ggarrange
plot1 <- ggarrange(f1$plot, f2$plot, f3$plot, f4$plot, f5$plot,
          ncol = 3, nrow = 2,
          common.legend = TRUE,
          legend = "top",
          legend.labs = c("Age", "Age 50", "Age 60"),
          vjust = 0.5
          ) 
# Add a note at the bottom of this figure
annotate_figure(plot1,
                top = text_grob("Estimated Survival Curves - Females",
                                size = 12,
                                face = "bold"),
                bottom = text_grob("**Y-axis is not the same for every plot",
                                   size = 10,
                                   hjust = 1.3,
                                   face = "italic"))
  

# With survival package - could not get one legend for all plots
# arrange_ggsurvplots(list(f1, f2, f3, f4, f5),
#                     ncol = 3, nrow = 2,
#                     title = "Estimated Survival Curves - Females")
```





