---
title: Risk Factors of Stroke in a 10-year Period
author: Diana Felipe
output: 
  bookdown::pdf_document2:
    urlcolor: magenta
    toc: false
    number_sections: false
    latex_engine: xelatex
    extra_dependencies: ["longtable", "float"]
    keep_tex: true
mainfont: Arial
fontsize: 12pt
documentclass: article
geometry: margin=1in
header-includes: 
  - \usepackage{setspace} 
  - \doublespacing 
  - \usepackage{float} 
  - \usepackage{caption} 
  - \usepackage{titling} 
  - \setlength{\droptitle}{-3cm} 
  - \setlength{\textfloatsep}{5pt} 
  - \preauthor{\vspace{-0.6cm}} 
  - \postauthor{\vspace{-2.5cm}}
---

```{r, echo = F, warning=F, message = F}
knitr::opts_chunk$set(echo = FALSE)
library(readr)
library(knitr)
library(tidyr)
library(ggplot2)
library(kableExtra)
library(dplyr)
library(tibble)
library(labelled)
library(tableone)
library(nlme)
library(gtsummary)
library(gt)
library(ggeffects)
library(see)
library(gridExtra)
library(regressinator)
library(patchwork)
library(purrr)
library(stringr)
library(survival)
library(survminer)
library(gtsummary)
library(here)
library(expss)
library(labelled)

opts_chunk$set(tidy = F)
```

```{r, echo = F}
# Read in the Data
# Read in created survival df
surv_df <- read_csv(here("Project3", "Data", "frmg_survival_df.csv"),
                    show_col_types = FALSE)

# Convert factor vars to factors
surv_df <- surv_df %>% 
  mutate(across(c(RANDID, SEX, DIABETES, 
                  deathwithin10, PREVCHD, BPMEDS, CURSMOKE), 
                as.factor))
# Make sure STROKE10 is numeric
surv_df$STROKE10 <- as.numeric(as.character(surv_df$STROKE10))

```

# Introduction
\vspace{-6mm}

The Framington Heart Study is a longitudinal and prospective study with the overarching goal of investigating the etiology of cardiovascular disease in the community of Framingham, Massachusetts. The study was conducted on an initial population of 5,209 individuals on which data was collected for health behaviors, medication use, clinical measurements such as blood pressure, and more. The primary goal of this analysis is to identify risk factors of stroke within a 10-year period for five different risk profiles of interest: 1.) Subjects with no risk factors for stroke, 2.) Subjects with high blood pressure (BP) (> 60 (mmHg) ), 3.) Subjects with diabetes, 4.) Subjects with high blood pressure and diabetes, 5.) Subjects who were smokers. All risk factors are measured at baseline. Additionally, we will summarize the change in the risk factors over the whole duration of the study period to determine whether a more complex time-varying covariate analysis is warranted. 
This report covers the preliminary findings of our analysis answering the primary question of interest. It fills in the results from the data analysis plan agreed upon in the Data Analysis Plan report.

\vspace{-10mm}
# Methods

\vspace{-6mm}
## Data Cleaning
\vspace{-6mm}

The raw data contained 11,627 observations for 4,434 subjects and 38 variables. Missingness was assessed in the whole data set thruogh summaries and visualizations. Regarding the primary variables of interest, BP Medication had ~5% missginness, total cholesterol had 3.5% missingness, and BMI had ~0.5% missingness. We verified that there were no relationships between missingness and any variables and therefore concluded that data was Missing Completely at Random.
We then proceeded with creating new variables specifically for the 10-year window of interest for the primary analysis. We created an indicator variable for whether subjects had a stroke with 10 years (Yes/No) as well as the time to stroke for these events. 
The data for all analyses consisted of 4,434 observations with 2,378 for females and 2,056 for males. We used a complete case data frame, removing any missing variables mentioned previously mentioned for the analyses. 

\vspace{-8mm}
## Data Analysis
\vspace{-6mm}

We conducted analyses separately for male and females. 
We also performed a preliminary descriptive analysis of survival curves by sex with visualizations of the Kaplan Meier curves before fitting the primary Cox model. This allowed us to visually compare the curves for the two sexes.
To identify risk factors that are associated with stroke in a 10-year period, we conducted a Survival analysis with stroke as the event and days to stroke as the time to event. Specifically, we fit a Cox Proportional Hazards (PH) Survival Model to identify important risk factors of stroke and obtain the 10-year probability of stroke based on the five risk profiles. 
We adjusted for Age, Diabetes status (Yes/No), and Systolic BP in all analyses. 
We performed variable selection with the Least Absolute Shrinkage Selection Operator (LASSO) on Presence of Coronary Heart Disease (CHD), whether subjects were on BP medications, smoking status, total cholesterol, and Body Mass Index (BMI). In particular, we used LASSO with K-fold cross validation to assess variable importance which retained all 5 variables. 
Therefore, we included a total of 8 variables for all analyses for each sex-stratified analysis. 
We note that the cross-validated error was fairly flat indicating limited evidence for a sparser predictive model with clear improvement over a model with no covariates. 
We assessed the proportional hazards assumptions for all models using the Schoenfeld Residuals Test as well as visually through the Log Minus Log (LML) plots for each covariate. 
Lastly, we created a table describing how variables such as Systolic BP and BMI change over over time. 

We used a significance level of $\alpha = 0.05$ for the significance threshold. All analyses were conducted with R version 4.1.1 using the glmnet, survival, and survminer packages. 

\vspace{-10mm}
# Results
\vspace{-6mm}

Table 1 contains a description of the data used for all analyses. 
The data is stratified by sex, n and \% of patients is displayed for the categorical variables, and the mean and standard deviation is shown for each continuous variable.
We note the small number of subjects across both sexes that had a stroke within the 10 year time span. In addition, the use of Anti-hypertensive medication and prevalent coronary heart disease groups are noticeably unbalanced. 
The preliminary analysis of the survival curves for sex showed...

In Table 2 we compare the predicted survival and $\%$ risk over 10 years for males and females for ages 40, 50, and 60 for the 5 risk profiles of interest.
<!------------------------------------ Table 1 -------------------------------->

```{r, echo = F, fig.pos='H', eval = F}
# Fix levels for categorical variables for Table1
# Data frame for table1
tab1_data <- surv_df

# Sex
tab1_data$factor_SEX <- factor(tab1_data$SEX,
                        levels = c(1, 2),
                        labels = c("Male", "Female"))

# Diabetes
tab1_data$factor_DIABETES <- factor(tab1_data$DIABETES,
                                    levels = c(0, 1),
                                    labels = c("Not a Diabetic", "Diabetic"))

# Prevchd
tab1_data$factor_PREVCHD <- factor(tab1_data$PREVCHD,
                                   levels = c(0, 1),
                                   labels = c("Free of CHD",
                                              "Prevalent CHD"))

# Cursmoke
tab1_data$factor_CURSMOKE <- factor(tab1_data$CURSMOKE,
                             levels = c(0, 1),
                             labels = c("Not a Current Smoker",
                                        "Current Smoker"))

# BPMeds
tab1_data$factor_BPMEDS <- factor(tab1_data$BPMEDS,
                                  levels = c(0, 1),
                                  labels = c("No Use of Medication",
                                             "Use of Medication"))

# Stroke 10
tab1_data$factor_STROKE <- factor(tab1_data$STROKE10,
                                  levels = c(0, 1),
                                  labels = c("No Stroke",
                                             "Stroke"))



# Relabel the variables for the Table 1
tab1_data <- apply_labels(tab1_data,
                     factor_STROKE = "Stroke Within 10 Yrs",
                     factor_SEX = "Sex",
                     factor_BPMEDS = "Use of Anti-hypertensive Medication",
                     factor_PREVCHD = "Prevalent Coronary Heart Disease",
                     factor_CURSMOKE = "Cigarrete Smoking Status",
                     AGE = "Age",
                     SYSBP = "Systolic Blood Pressure (mmHg)",
                     STROKE10TIME = "Time to Stroke (Days)",
                     TOTCHOL = "Serum Total Cholesterol (mg/dL)",
                     BMI = "Body Mass Index"
                     )

# Display the modified Table1
tab1_data %>% 
  tbl_summary(include = c(factor_STROKE,
                          STROKE10TIME,
                          AGE,
                          factor_BPMEDS,
                          factor_PREVCHD,
                          factor_CURSMOKE,
                          SYSBP,
                          TOTCHOL,
                          BMI),
              by = factor_SEX,
              statistic = list(c(all_continuous()) 
                               ~ "{mean}({sd})")) %>% 
  modify_header(label = "**Sex**") %>% 
  bold_labels() %>% 
  modify_caption(caption = "Patient Characteristics At Baseline")
  
  
  # as_kable_extra(booktabs = TRUE, linesep = "") %>%
  # kableExtra::kable_styling(latex_options = c("hold_position"))
```

<!----------------------------------------------------------------------------->
<!--------------------------------- KM Curves --------------------------------->
<!----------------------------------------------------------------------------->

The preliminary Kaplan-Meier analysis reveled frequent crossing between the separate male and female survival curves. This indicates a violation of the proportional hazards assumption and therefore necessitates stratified Cox models to obtain the sex-specific risk differences. 

```{r, echo = F, eval = F, results='hide'}
################################ Prelim Cox Analysis ###########################
# Create some Kaplan Meier curves by covariates

### Sex
# Fit model
km_sex <- survfit(Surv(STROKE10TIME, STROKE10) ~ SEX,
                   data = surv_df)

# Plot KM curves
ggsurvplot(km_sex, 
           conf.int = FALSE,
           censor = FALSE,
           xlim = c(0, max(surv_df$STROKE10TIME)),
           ylim = c(0.95, 1))
# For the most part, seems that survival curves are quite similar between sexes
# We do see some crossing of the curves though

```


<!----------------------------------------------------------------------------->
<!------------------------------ Variable Selection --------------------------->
<!----------------------------------------------------------------------------->

```{r, echo = F, eval = F, results='hide'}
############################## Variable Selection ##############################
# Vignette: https://glmnet.stanford.edu/articles/Coxnet.html

# Perform some variable selection prevchd, bpmeds, smokecur, total chol, and BMI

##### First, going to assess whether sex should have different baseline hazards
# Unstratified model
fit_unstrat <- coxph(
  Surv(STROKE10TIME, STROKE10) ~ SEX
    + PREVCHD + 
    BPMEDS + CURSMOKE + TOTCHOL + BMI,
  data = surv_df
)

# Stratified by sex model
fit_strat <- coxph(
  Surv(STROKE10TIME, STROKE10) ~ PREVCHD + BPMEDS + 
    CURSMOKE + TOTCHOL + BMI + strata(SEX),
  data = surv_df
)

# Compare model fit with AIC
AIC(fit_unstrat, fit_strat)
# df      AIC
# fit_unstrat  6 1723.074
# fit_strat    5 1576.158

# We saw though that the KM curves by sex did cross a lot
# Sex was not significant in the model
# The un and stratified models give almost the same results 
# Since we were getting convergence issues down the line with glmnet,
# we are going to NOT stratify by sex 

# Proceed with the unstratified pooled glmnet cox model

###### Glmnet

# Fit stratified cox models with elastic net penalty
# Prepare for glmnet
# Surv object of response variables
y <- Surv(time = surv_df$STROKE10TIME,
          event = surv_df$STROKE10)

# Create matrix of predictors
x <- model.matrix(~ PREVCHD + BPMEDS + CURSMOKE + TOTCHOL + BMI,
                  data = surv_df)[, -1] # This removes the intercept


# Fit stratified cox lasso
lasso_fit <- glmnet(x = x, 
                    y = y,
                    family = "cox",
                    alpha = 1) # LASSO penalty

# Plot fit 
plot(lasso_fit)

# K-fold Cross validation
set.seed(345)
cv_fit <- cv.glmnet(x = x, 
                    y = y, 
                    family = "cox",
                    alpha = 1)

# Look at plot - optimal lambda value
plot(cv_fit)
# The left vertical line is where the CV error curve hits its minimum
# The right vertical line is the most regularized model with CV error within 1 SD
# Basically we see that model performance is basically flat and increasing
# model complexity does not improve performance much


# Get the optimal lambdas
cv_fit$lambda.min
#  0.001777793
## This gives the best fit or lowest CV error where a small penalty means
# more variables are kept - risk of overfitting though

# See which variables are kept
coef_min <- coef(cv_fit, s = "lambda.min")
# PREVCHD1  0.827245643
# BPMEDS1   1.382781943
# CURSMOKE1 0.123849200
# TOTCHOL   0.002622194
# BMI       0.043796358

# Get the nonzero variables
nz_min <- which(as.vector(coef_min) != 0)
selected_min <- rownames(coef_min)[nz_min]
# "PREVCHD1"  "BPMEDS1"   "CURSMOKE1" "TOTCHOL"   "BMI" 

cv_fit$lambda.1se
# 0.01376481
## Largest lambda within 1 SE of the minimum
# More penalty more simpler model

# See which variables are kept
coef_1se <- coef(cv_fit, s = "lambda.1se")
# PREVCHD1  .
# BPMEDS1   .
# CURSMOKE1 .
# TOTCHOL   .
# BMI       .

# Get the nonzero variables
nz_1se <- which(as.vector(coef_1se) != 0)
selected_1se <- rownames(coef_1se)[nz_1se]
# character(0)

# Basically we found that lambda.min keeps all variables while
# lambda.1se drops everything
# We did already have pre-selected variables that were clinically driven so 
# perhaps further reduction is not necessary

# Using lambda.min, all five candidate predictors were retained.
# Using the more conservative lambda.1se rule, all coefficients were shrunk to zero, 
# indicating that the cross-validated error was fairly flat and that there was 
# limited evidence for a sparser predictive model with clear improvement over the null.

# We used LASSO with cross-validation to assess variable importance. 
# Because the cross-validated error curve was relatively flat, the more 
# conservative λ₁ₛₑ resulted in a null model. Therefore, we selected λ_min to 
# retain relevant predictors while minimizing prediction error.

# After lasso, fit a standard coxph model with ALL variables
```



<!----------------------------------------------------------------------------->
<!----------------------------------- FEMALES --------------------------------->
<!----------------------------------------------------------------------------->

```{r, echo = F, warning=F, message=F}
################################## Model Fitting ###############################

##################################### Females ##################################
# Data frame for females
females_df <- surv_df %>% 
  filter(SEX == 2)
# 2378 obs for 2378 subjects

# Model for females
females_fit <- coxph(
  Surv(STROKE10TIME, STROKE10) ~ AGE + DIABETES + SYSBP
  + PREVCHD + BPMEDS 
  + CURSMOKE + TOTCHOL + BMI,
  data = females_df
)

# Summary of general model
females_sum1 <- summary(females_fit)

# Create data frame of results
females_table <- females_fit %>% 
  tbl_regression(intercept = TRUE, 
                 conf.level = .95,
                 exponentiate = TRUE,
                 pvalue_fun = label_style_pvalue(digits = 2),
                 conf.int = TRUE,
                 show_single_row = c(DIABETES, PREVCHD, BPMEDS, CURSMOKE),
                 label = list(
                   AGE ~ "Age",
                   DIABETES ~ "Diabetes Status",
                   PREVCHD ~ "Prevalent CHD",
                   BPMEDS ~ "BP Medication",
                   CURSMOKE ~ "Smoking Status",
                   SYSBP ~ "Systolic BP",
                   TOTCHOL ~ "Total Cholesterol"))

```


```{r, echo = F}
#### Test PH assumptions
```



```{r, echo = F, eval = F, results='hide', warning=F, message=F}
# Plot survival curve - hmmm
ggsurvplot(survfit(females_fit),
           data = females_df,
           censor = F,
           conf.int = T,
           ylim = c(0.95, 1))
```


```{r, echo = F, results='hide'}
################################# Risk Profile 1 ###############################

# Create data frame for the different risk profiles
# Risk profile 1: 
## Average age, no diabetes, average sysbp, no prechd, no bpmeds, no cursmoke,
## average totchol, average bmi
females_rp1 <- data.frame(
    AGE = c(40, 50, 60),
    DIABETES = factor(rep(0, 3), 
                      levels = levels(females_df$DIABETES)),
    SYSBP = rep(mean(females_df$SYSBP), 3),
    PREVCHD = factor(rep(0, 3), 
                     levels = levels(females_df$PREVCHD)),
    BPMEDS = factor(rep(0, 3), 
                    levels = levels(females_df$BPMEDS)),
    CURSMOKE = factor(rep(0, 3), 
                      levels = levels(females_df$CURSMOKE)),
    TOTCHOL = rep(mean(females_df$TOTCHOL, 
                       na.rm = TRUE), 3),
    BMI = rep(mean(females_df$BMI), 3)
)

# Fit cox model for risk profile 1
mrp1_fit <- survfit(females_fit, newdata = females_rp1)


# Estimates for 10 years
mrp1_summary <- summary(mrp1_fit, times = 365*10)
# time n.risk n.event survival1 survival2 survival3
# 3650   2177      57     0.995     0.988     0.976

# Data frame of survival 10 years and risk 10 years
mrp1_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(mrp1_summary$surv), 3),
                           Risk = paste0(round(1 - mrp1_summary$surv, 3)*100,
                                         "%")
)



```


```{r, echo = F, results='hide'}
################################# Risk Profile 2 ###############################

# Risk profile 2: 
## High blood pressure and no other risk factors
females_rp2 <- data.frame(
  AGE = c(40, 50, 60),
  DIABETES = factor(rep(0, 3), 
                    levels = levels(females_df$DIABETES)),
  SYSBP = rep(200, 3),
  PREVCHD = factor(rep(0, 3), 
                   levels = levels(females_df$PREVCHD)),
  BPMEDS = factor(rep(0, 3), 
                  levels = levels(females_df$BPMEDS)),
  CURSMOKE = factor(rep(0, 3), 
                    levels = levels(females_df$CURSMOKE)),
  TOTCHOL = rep(mean(females_df$TOTCHOL, 
                     na.rm = TRUE), 3),
  BMI = rep(mean(females_df$BMI), 3)
)

# AGE DIABETES SYSBP PREVCHD BPMEDS CURSMOKE  TOTCHOL      BMI
# 1  40        0   180       0      0        0 239.5698 25.56664
# 2  50        0   230       0      0        0 239.5698 25.56664
# 3  60        0   280       0      0        0 239.5698 25.56664


# Fit cox model for risk profile 2
mrp2_fit <- survfit(females_fit, newdata = females_rp2)

# Estimates for 10 years
mrp2_summary <- summary(mrp2_fit, times = 365*10)
# time n.risk n.event survival1 survival2 survival3
# 3650   2177      57     0.983     0.878     0.382

# Create data frame of survival 10 years and risk 10 years
mrp2_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(mrp2_summary$surv), 3),
                           Risk = paste0(round(1 - mrp2_summary$surv, 3)*100,
                                         "%")
)
# Age Survival  Risk
# 1  40    0.983  1.7%
# 2  50    0.878 12.2%
# 3  60    0.382 61.8%


```


```{r, echo = F, results='hide'}

################################# Risk Profile 3 ###############################

#### Risk profile 3:
## Diabetes and no other risk factors
females_rp3 <- data.frame(
  AGE = c(40, 50, 60),
  DIABETES = factor(rep(1, 3), 
                    levels = levels(females_df$DIABETES)),
  SYSBP = rep(mean(females_df$SYSBP), 3),
  PREVCHD = factor(rep(0, 3), 
                   levels = levels(females_df$PREVCHD)),
  BPMEDS = factor(rep(0, 3), 
                  levels = levels(females_df$BPMEDS)),
  CURSMOKE = factor(rep(0, 3), 
                    levels = levels(females_df$CURSMOKE)),
  TOTCHOL = rep(mean(females_df$TOTCHOL, 
                     na.rm = TRUE), 3),
  BMI = rep(mean(females_df$BMI), 3)
)
# AGE DIABETES    SYSBP PREVCHD BPMEDS CURSMOKE  TOTCHOL      BMI
# 1  40        1 133.5524       0      0        0 239.5698 25.56664
# 2  50        1 133.5524       0      0        0 239.5698 25.56664
# 3  60        1 133.5524       0      0        0 239.5698 25.56664


# Fit cox model for risk profile 3
mrp3_fit <- survfit(females_fit, newdata = females_rp3)

# Estimates for 10 years
mrp3_probs <- summary(mrp3_fit, times = 365*10)
# time n.risk n.event survival1 survival2 survival3
# 3650   2177      57      0.99     0.979     0.955


# Estimates for 10 years
mrp3_summary <- summary(mrp3_fit, times = 365*10)
# time n.risk n.event survival1 survival2 survival3
# 3650   2177      57      0.99     0.979     0.955

# Create data frame of survival 10 years and risk 10 years
mrp3_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(mrp3_summary$surv), 3),
                           Risk = paste0(round(1 - mrp3_summary$surv, 3)*100,
                                         "%")
)

# Age Survival Risk
# 1  40    0.990   1%
# 2  50    0.979 2.1%
# 3  60    0.955 4.5%

```


```{r, echo = F, results='hide'}
################################# Risk Profile 4 ###############################

#### Risk profile 4:
## Diabetes and no other risk factors
females_rp4 <- data.frame(
  AGE = c(40, 50, 60),
  DIABETES = factor(rep(1, 3), 
                    levels = levels(females_df$DIABETES)),
  SYSBP = rep(200, 3),
  PREVCHD = factor(rep(0, 3), 
                   levels = levels(females_df$PREVCHD)),
  BPMEDS = factor(rep(0, 3), 
                  levels = levels(females_df$BPMEDS)),
  CURSMOKE = factor(rep(0, 3), 
                    levels = levels(females_df$CURSMOKE)),
  TOTCHOL = rep(mean(females_df$TOTCHOL, 
                     na.rm = TRUE), 3),
  BMI = rep(mean(females_df$BMI), 3)
)
# 
# AGE DIABETES SYSBP PREVCHD BPMEDS CURSMOKE  TOTCHOL      BMI
# 1  40        1   180       0      0        0 239.5698 25.56664
# 2  50        1   230       0      0        0 239.5698 25.56664
# 3  60        1   280       0      0        0 239.5698 25.56664

# Fit cox model for risk profile 3
mrp4_fit <- survfit(females_fit, newdata = females_rp4)

# Estimates for 10 years
mrp4_summary <- summary(mrp4_fit, times = 365*10)
# time n.risk n.event survival1 survival2 survival3
# 3650   2177      57      0.99     0.979     0.955

# Create data frame of survival 10 years and risk 10 years
mrp4_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(mrp4_summary$surv), 3),
                           Risk = paste0(round(1 - mrp4_summary$surv, 3)*100,
                                         "%")
)

# Age Survival  Risk
# 1  40    0.968  3.2%
# 2  50    0.785 21.5%
# 3  60    0.167 83.3%

```


```{r, echo = F, results='hide'}
################################# Risk Profile 5 ###############################

#### Risk profile 4:
## Smoker and no other risk factors
females_rp5 <- data.frame(
  AGE = c(40, 50, 60),
  DIABETES = factor(rep(0, 3), 
                    levels = levels(females_df$DIABETES)),
  SYSBP = rep(mean(females_df$SYSBP), 3),
  PREVCHD = factor(rep(0, 3), 
                   levels = levels(females_df$PREVCHD)),
  BPMEDS = factor(rep(0, 3), 
                  levels = levels(females_df$BPMEDS)),
  CURSMOKE = factor(rep(1, 3), 
                    levels = levels(females_df$CURSMOKE)),
  TOTCHOL = rep(mean(females_df$TOTCHOL, 
                     na.rm = TRUE), 3),
  BMI = rep(mean(females_df$BMI), 3)
)
# 
# AGE DIABETES    SYSBP PREVCHD BPMEDS CURSMOKE  TOTCHOL      BMI
# 1  40        0 133.5524       0      0        1 239.5698 25.56664
# 2  50        0 133.5524       0      0        1 239.5698 25.56664
# 3  60        0 133.5524       0      0        1 239.5698 25.56664

# Fit cox model for risk profile 3
mrp5_fit <- survfit(females_fit, newdata = females_rp5)

# Estimates for 10 years
mrp5_summary <- summary(mrp5_fit, times = 365*10)
# time n.risk n.event survival1 survival2 survival3
# 3650   2177      57     0.992     0.984     0.966

# Create data frame of survival 10 years and risk 10 years
mrp5_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(mrp5_summary$surv), 3),
                           Risk = paste0(round(1 - mrp5_summary$surv, 3)*100,
                                         "%")
)

# Age Survival Risk
# 1  40    0.992 0.8%
# 2  50    0.984 1.6%
# 3  60    0.966 3.4%
```


```{r, echo = F}
################### Combine these all for the report
library(knitr)
library(kableExtra)

# Create individual kables for each df
# Results 1
fk1 <- kable(mrp1_results, caption = "meow1",
             format = "html") %>% 
  kable_styling()

# Results 2
fk2 <- kable(mrp2_results, caption = "meow2",
             format = "html") %>% 
  kable_styling()

# Results 3
fk3 <- kable(mrp3_results, caption = "meow3",
             format = "html") %>% 
  kable_styling()

# Results 4
fk4 <- kable(mrp4_results, caption = "meow4",
             format = "html") %>% 
  kable_styling()

# Combimne results into one
female_results <- cbind(mrp1_results, 
                       mrp2_results[, 2:3],
                       mrp3_results[,2:3],
                       mrp4_results[,2:3],
                       mrp5_results[,2:3]
                       )

```


<!----------------------------------------------------------------------------->
<!------------------------------------ MALES ---------------------------------->
<!----------------------------------------------------------------------------->


```{r, echo = F, warning=F, message=F, fig.cap="Model Results for Males"}
################################## Model Fitting ###############################

##################################### Males ##################################
# Data frame for males
males_df <- surv_df %>% 
  filter(SEX == 1)
# 1897 obs for 1897 subjects

# Model for males
males_fit <- coxph(
  Surv(STROKE10TIME, STROKE10) ~ AGE + DIABETES + SYSBP
  + PREVCHD + BPMEDS 
  + CURSMOKE + TOTCHOL + BMI,
  data = males_df
)

# Summary of general model
males_sum1 <- summary(males_fit)

# Create data frame of results
males_table <- males_fit %>% 
    tbl_regression(intercept = TRUE, 
                 conf.level = .95,
                 exponentiate = TRUE,
                 pvalue_fun = label_style_pvalue(digits = 2),
                 conf.int = TRUE,
                 show_single_row = c(DIABETES, PREVCHD, BPMEDS, CURSMOKE),
                 label = list(
                   AGE ~ "Age",
                   DIABETES ~ "Diabetes Status",
                   PREVCHD ~ "Prevalent CHD",
                   BPMEDS ~ "BP Medication",
                   CURSMOKE ~ "Smoking Status",
                   SYSBP ~ "Systolic BP",
                   TOTCHOL ~ "Total Cholesterol")) 

```

```{r, echo = F, eval = F, results='hide', warning=F, message=F}
# Plot survival curve - hmmm
ggsurvplot(survfit(males_fit),
           data = males_df,
           censor = F,
           conf.int = T)
```



```{r, echo = F, eval = F}

tbl_side_by_side <- tbl_merge(
  tbls = list(females_table, males_table),
  tab_spanner = c("**Females**", "**Males**")
)

tbl_side_by_side %>%
  modify_caption("Model Results by Sex") %>%
  as_gt() %>%
  gt::tab_source_note(
    gt::md("*BP = Blood Pressure; CHD = Coronary Heart Disease")
  ) %>%
  gt::tab_source_note(
    gt::md("**All factor variable HRs are compared to reference groups")
  ) %>%
  gt::tab_options(
    table.font.size = "small",
    data_row.padding = gt::px(2)
  )
```

```{r, echo = F, results='hide'}
################################# Risk Profile 1 ###############################

# Create data frame for the different risk profiles
# Risk profile 1: 
## Average age, no diabetes, average sysbp, no prechd, no bpmeds, no cursmoke,
## average totchol, average bmi
males_rp1 <- data.frame(
    AGE = c(40, 50, 60),
    DIABETES = factor(rep(0, 3), 
                      levels = levels(males_df$DIABETES)),
    SYSBP = rep(mean(males_df$SYSBP), 3),
    PREVCHD = factor(rep(0, 3), 
                     levels = levels(males_df$PREVCHD)),
    BPMEDS = factor(rep(0, 3), 
                    levels = levels(males_df$BPMEDS)),
    CURSMOKE = factor(rep(0, 3), 
                      levels = levels(males_df$CURSMOKE)),
    TOTCHOL = rep(mean(males_df$TOTCHOL, 
                       na.rm = TRUE), 3),
    BMI = rep(mean(males_df$BMI), 3)
)

# Fit cox model for risk profile 1
mrp1_fit <- survfit(males_fit, newdata = males_rp1)


# Estimates for 10 years
mrp1_summary <- summary(mrp1_fit, times = 365*10)
 # time n.risk n.event survival1 survival2 survival3
 # 3650   1647      48     0.995      0.99     0.981

# Data frame of survival 10 years and risk 10 years
mrp1_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(mrp1_summary$surv), 3),
                           Risk = paste0(round(1 - mrp1_summary$surv, 3)*100,
                                         "%")
)

```


```{r, echo = F, results='hide'}
################################# Risk Profile 2 ###############################

# Risk profile 2: 
## High blood pressure and no other risk factors
males_rp2 <- data.frame(
  AGE = c(40, 50, 60),
  DIABETES = factor(rep(0, 3), 
                    levels = levels(males_df$DIABETES)),
  SYSBP = rep(200, 3),
  PREVCHD = factor(rep(0, 3), 
                   levels = levels(males_df$PREVCHD)),
  BPMEDS = factor(rep(0, 3), 
                  levels = levels(males_df$BPMEDS)),
  CURSMOKE = factor(rep(0, 3), 
                    levels = levels(males_df$CURSMOKE)),
  TOTCHOL = rep(mean(males_df$TOTCHOL, 
                     na.rm = TRUE), 3),
  BMI = rep(mean(males_df$BMI), 3)
)

# AGE DIABETES SYSBP PREVCHD BPMEDS CURSMOKE  TOTCHOL      BMI
# 1  40        0   180       0      0        0 233.7359 26.16627
# 2  50        0   230       0      0        0 233.7359 26.16627
# 3  60        0   280       0      0        0 233.7359 26.16627


# Fit cox model for risk profile 2
mrp2_fit <- survfit(males_fit, newdata = males_rp2)

# Estimates for 10 years
mrp2_summary <- summary(mrp2_fit, times = 365*10)
 # time n.risk n.event survival1 survival2 survival3
 # 3650   1647      48     0.976     0.795     0.114

# Create data frame of survival 10 years and risk 10 years
mrp2_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(mrp2_summary$surv), 3),
                           Risk = paste0(round(1 - mrp2_summary$surv, 3)*100,
                                         "%")
)
# Age Survival  Risk
# 1  40    0.976  2.4%
# 2  50    0.795 20.5%
# 3  60    0.114 88.6%


```


```{r, echo = F, results='hide'}

################################# Risk Profile 3 ###############################

#### Risk profile 3:
## Diabetes and no other risk factors
males_rp3 <- data.frame(
  AGE = c(40, 50, 60),
  DIABETES = factor(rep(1, 3), 
                    levels = levels(males_df$DIABETES)),
  SYSBP = rep(mean(males_df$SYSBP), 3),
  PREVCHD = factor(rep(0, 3), 
                   levels = levels(males_df$PREVCHD)),
  BPMEDS = factor(rep(0, 3), 
                  levels = levels(males_df$BPMEDS)),
  CURSMOKE = factor(rep(0, 3), 
                    levels = levels(males_df$CURSMOKE)),
  TOTCHOL = rep(mean(males_df$TOTCHOL, 
                     na.rm = TRUE), 3),
  BMI = rep(mean(males_df$BMI), 3)
)
# AGE DIABETES    SYSBP PREVCHD BPMEDS CURSMOKE  TOTCHOL      BMI
# 1  40        1 131.5422       0      0        0 233.7359 26.16627
# 2  50        1 131.5422       0      0        0 233.7359 26.16627
# 3  60        1 131.5422       0      0        0 233.7359 26.16627


# Fit cox model for risk profile 3
mrp3_fit <- survfit(males_fit, newdata = males_rp3)

# Estimates for 10 years
mrp3_probs <- summary(mrp3_fit, times = 365*10)
 # time n.risk n.event survival1 survival2 survival3
 # 3650   1647      48     0.974      0.95     0.905


# Estimates for 10 years
mrp3_summary <- summary(mrp3_fit, times = 365*10)
 # time n.risk n.event survival1 survival2 survival3
 # 3650   1647      48     0.974      0.95     0.905

# Create data frame of survival 10 years and risk 10 years
mrp3_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(mrp3_summary$surv), 3),
                           Risk = paste0(round(1 - mrp3_summary$surv, 3)*100,
                                         "%")
)

# Age Survival Risk
# 1  40    0.974 2.6%
# 2  50    0.950   5%
# 3  60    0.905 9.5%

```


```{r, echo = F, results='hide'}
################################# Risk Profile 4 ###############################

#### Risk profile 4:
## High BP and Diabetes
males_rp4 <- data.frame(
  AGE = c(40, 50, 60),
  DIABETES = factor(rep(1, 3), 
                    levels = levels(males_df$DIABETES)),
  SYSBP = rep(200, 3),
  PREVCHD = factor(rep(0, 3), 
                   levels = levels(males_df$PREVCHD)),
  BPMEDS = factor(rep(0, 3), 
                  levels = levels(males_df$BPMEDS)),
  CURSMOKE = factor(rep(0, 3), 
                    levels = levels(males_df$CURSMOKE)),
  TOTCHOL = rep(mean(males_df$TOTCHOL, 
                     na.rm = TRUE), 3),
  BMI = rep(mean(males_df$BMI), 3)
)
# 
# AGE DIABETES SYSBP PREVCHD BPMEDS CURSMOKE  TOTCHOL      BMI
# 1  40        1   180       0      0        0 233.7359 26.16627
# 2  50        1   230       0      0        0 233.7359 26.16627
# 3  60        1   280       0      0        0 233.7359 26.16627

# Fit cox model for risk profile 3
mrp4_fit <- survfit(males_fit, newdata = males_rp4)

# Estimates for 10 years
mrp4_summary <- summary(mrp4_fit, times = 365*10)
 # time n.risk n.event survival1 survival2 survival3
 # 3650   1647      48     0.883     0.307  1.44e-05

# Create data frame of survival 10 years and risk 10 years
mrp4_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(mrp4_summary$surv), 3),
                           Risk = paste0(round(1 - mrp4_summary$surv, 3)*100,
                                         "%")
)

# Age Survival  Risk
#   Age Survival  Risk
# 1  40    0.883 11.7%
# 2  50    0.307 69.3%
# 3  60    0.000  100%

```


```{r, echo = F, results='hide'}
################################# Risk Profile 5 ###############################

#### Risk profile 4:
## Smoker and no other risk factors
males_rp5 <- data.frame(
  AGE = c(40, 50, 60),
  DIABETES = factor(rep(0, 3), 
                    levels = levels(males_df$DIABETES)),
  SYSBP = rep(mean(males_df$SYSBP), 3),
  PREVCHD = factor(rep(0, 3), 
                   levels = levels(males_df$PREVCHD)),
  BPMEDS = factor(rep(0, 3), 
                  levels = levels(males_df$BPMEDS)),
  CURSMOKE = factor(rep(1, 3), 
                    levels = levels(males_df$CURSMOKE)),
  TOTCHOL = rep(mean(males_df$TOTCHOL, 
                     na.rm = TRUE), 3),
  BMI = rep(mean(males_df$BMI), 3)
)
# 
# AGE DIABETES    SYSBP PREVCHD BPMEDS CURSMOKE  TOTCHOL      BMI
# 1  40        0 131.5422       0      0        1 233.7359 26.16627
# 2  50        0 131.5422       0      0        1 233.7359 26.16627
# 3  60        0 131.5422       0      0        1 233.7359 26.16627

# Fit cox model for risk profile 3
mrp5_fit <- survfit(males_fit, newdata = males_rp5)

# Estimates for 10 years
mrp5_summary <- summary(mrp5_fit, times = 365*10)
 # time n.risk n.event survival1 survival2 survival3
 # 3650   1647      48      0.99      0.98     0.962

# Create data frame of survival 10 years and risk 10 years
mrp5_results <- data.frame(Age = c(40, 50, 60),
                           Survival = round(as.vector(mrp5_summary$surv), 3),
                           Risk = paste0(round(1 - mrp5_summary$surv, 3)*100,
                                         "%")
)

# Age Survival Risk
# 1  40    0.990   1%
# 2  50    0.980   2%
# 3  60    0.962 3.8%
```


```{r, echo = F, fig.pos='H'}
################### Combine these all for the report
library(knitr)
library(kableExtra)

# Create individual kables for each df
# Results 1
rk1 <- kable(mrp1_results, caption = "meow1",
             format = "html") %>% 
  kable_styling()

# Results 2
rk2 <- kable(mrp2_results, caption = "meow2",
             format = "html") %>% 
  kable_styling()

# Results 3
rk3 <- kable(mrp3_results, caption = "meow3",
             format = "html") %>% 
  kable_styling()

# Results 4
rk4 <- kable(mrp4_results, caption = "meow4",
             format = "html") %>% 
  kable_styling()

# Combimne results into one
male_results <- cbind(mrp1_results, 
                       mrp2_results[, 2:3],
                       mrp3_results[,2:3],
                       mrp4_results[,2:3],
                       mrp5_results[,2:3]
                       )

# Display with kable
# kable(male_results,
#   format = "latex",
#   booktabs = TRUE,
#   digits = 3,
#   caption = "Survival Rates and Risk \\% Over 10 Years for Males") %>% 
#   kable_styling(latex_options = c("striped", "hold_position", 
#                                   "scale_down")) %>% 
#   add_header_above(c(" ", "Risk Profile 1\nNo Risk factors" = 2, 
#                      "Risk Profile 2\nHigh BP" = 2,
#                      "Risk Profile 3\nDiabetes" = 2,
#                      "Risk Profile 4\n High BP & Diabetes" = 2,
#                      "Risk Profile 5\n Smoker" = 2)) %>% 
#   footnote(number = c("BP = Blood Pressure ",
#                       "All Risk Factors are at Baseline Timepoint"))

```

```{r, echo = F, eval = F}
# Display both male and female tables together

kable(female_results,
  format = "latex",
  booktabs = TRUE,
  digits = 3, 
  caption = "Survival Rates and Risk \\% Over 10 Years") %>%
  kable_styling(latex_options = c("striped", "hold_position",
                                  "scale_down")) %>%
  add_header_above(c(" ", "Risk Profile 1\nNo Risk factors" = 2,
                     "Risk Profile 2\nHigh BP" = 2,
                     "Risk Profile 3\nDiabetes" = 2,
                     "Risk Profile 4\n High BP & Diabetes" = 2,
                     "Risk Profile 5\n Smoker" = 2)) %>%
  add_header_above(c("Females"
                     = 11),
                   escape = FALSE)


kable(male_results,
  format = "latex",
  booktabs = TRUE,
  digits = 3) %>%
  # caption = "Survival Rates and Risk \\% Over 10 Years for Males") %>%
  kable_styling(latex_options = c("striped", "hold_position",
                                  "scale_down")) %>%
  add_header_above(c(" ", "Risk Profile 1\nNo Risk factors" = 2,
                     "Risk Profile 2\nHigh BP" = 2,
                     "Risk Profile 3\nDiabetes" = 2,
                     "Risk Profile 4\n High BP & Diabetes" = 2,
                     "Risk Profile 5\n Smoker" = 2)) %>%
  add_header_above(c("Males"
                     = 11),
                   escape = FALSE) %>%
  footnote(number = c("BP = Blood Pressure & High BP = 200 (mmHg)",
                      "All Risk Factors are at Baseline Timepoint"))

```

```{r, echo = F, eval = F, message=F, warning=F}
library(colorspace)
# Plot the survival curves
f1 <- ggsurvplot(survfit(females_fit, 
                   newdata = females_rp1),
           data = females_df,
           censor = FALSE,
           conf.int = FALSE,
           ylim = c(0.95, 1),
           xlim = c(0, 3650),
           # risk.table = TRUE,
           ggtheme = theme_minimal(),
           # legend = "none",
           legend.labs = c("40", "50", "60"),
           legend.title = "Age",
           title = "Risk Profile 1 Estimated Survival\nAverage Individual",
           font.main = 10,
           palette = c("deeppink2", "turquoise3", "springgreen3")
           ) 
# f1

# Plot the survival curves
f2 <- ggsurvplot(survfit(females_fit, 
                   newdata = females_rp2),
           data = females_df,
           censor = F,
           conf.int = FALSE,
           # Min is 0.35 and max is 0.99
           ylim = c(0.35, 1),
           xlim = c(0, 3650),
           ggtheme = theme_minimal(),
           legend.labs = c("Age 40", "Age 50", "Age 60"),
           legend = "none",
           title = "Risk Profile 2 Estimated Survival\nHigh BP",
           font.main = 10,
           palette = c("deeppink2", "turquoise3", "springgreen3")
) 
# f2


# Plot the survival curves
f3 <- ggsurvplot(survfit(females_fit, 
                   newdata = females_rp3),
           data = females_df,
           censor = F,
           conf.int = FALSE,
           # Min is 0.35 and max is 0.99
           ylim = c(0.95, 1),
           xlim = c(0, 3650),
           ggtheme = theme_minimal(),
           legend = "none",
           title = "Risk Profile 2 Estimated Survival\nDiabetes",
           font.main = 10,
           palette = c("deeppink2", "turquoise3", "springgreen3")
) 
# f3

# Plot the survival curves
f4 <- ggsurvplot(survfit(females_fit, 
                   newdata = females_rp4),
           data = females_df,
           censor = F,
           conf.int = FALSE,
           # Min is 0.35 and max is 0.99
           # ylim = c(0.95, 1),
           xlim = c(0, 3650),
           ggtheme = theme_minimal(),
           legend.labs = c("Age 40", "Age 50", "Age 60"),
           legend = "none",
           title = "Risk Profile 2 Estimated Survival\nHigh BP & Diabetes",
           font.main = 10,
           palette = c("deeppink2", "turquoise3", "springgreen3")
) 
# f4

# Plot the survival curves
f5 <- ggsurvplot(survfit(females_fit, 
                         newdata = females_rp5),
                 data = females_df,
                 censor = F,
                 conf.int = FALSE,
                 # Min is 0.35 and max is 0.99
                 ylim = c(0.95, 1),
                 xlim = c(0, 3650),
                 ggtheme = theme_minimal(),
                 legend = "none",
                 title = "Risk Profile 2 Estimated Survival\nSmoker",
                 font.main = 10,
                 palette = c("deeppink2", "turquoise3", "springgreen3")
) 
# f5

##### Display all plots together

library(ggpubr)
# Using ggarrange
plot1 <- ggarrange(f1$plot, f2$plot, f3$plot, f4$plot, f5$plot,
          ncol = 3, nrow = 2,
          common.legend = TRUE,
          legend = "top",
          legend.labs = c("Age", "Age 50", "Age 60"),
          vjust = 0.5
          ) 
# Add a note at the bottom of this figure
annotate_figure(plot1,
                top = text_grob("Estimated Survival Curves - Females",
                                size = 12,
                                face = "bold"),
                bottom = text_grob("**Y-axis is not the same for every plot",
                                   size = 10,
                                   hjust = 1.3,
                                   face = "italic"))
  

# With survival package - could not get one legend for all plots
# arrange_ggsurvplots(list(f1, f2, f3, f4, f5),
#                     ncol = 3, nrow = 2,
#                     title = "Estimated Survival Curves - Females")
```


```{r, echo = F, message = F, warning = F, fig.pos="H", eval =F}
library(colorspace)
# Plot the survival curves
m1 <- ggsurvplot(survfit(males_fit, 
                   newdata = males_rp1),
           data = males_df,
           censor = FALSE,
           conf.int = FALSE,
           # ylim = c(0.98, 1),
           xlim = c(0, 3650),
           # risk.table = TRUE,
           ggtheme = theme_minimal(),
           # legend = "none",
           legend.labs = c("40", "50", "60"),
           legend.title = "Age",
           title = "Risk Profile 1 Estimated Survival\nAverage Individual",
           font.main = 10,
           palette = "Set1"
           ) 
# m1

# Plot the survival curves
m2 <- ggsurvplot(survfit(males_fit, 
                   newdata = males_rp2),
           data = males_df,
           censor = F,
           conf.int = FALSE,
           # Min is 0.35 and max is 0.99
           # ylim = c(0.35, 1),
           xlim = c(0, 3650),
           ggtheme = theme_minimal(),
           legend.labs = c("Age 40", "Age 50", "Age 60"),
           legend = "none",
           title = "Risk Profile 2 Estimated Survival\nHigh BP",
           font.main = 10,
           palette = "Set1"
) 
# f2


# Plot the survival curves
m3 <- ggsurvplot(survfit(males_fit, 
                   newdata = males_rp3),
           data = males_df,
           censor = F,
           conf.int = FALSE,
           # Min is 0.35 and max is 0.99
           # ylim = c(0.90, 1),
           xlim = c(0, 3650),
           ggtheme = theme_minimal(),
           legend = "none",
           title = "Risk Profile 2 Estimated Survival\nDiabetes",
           font.main = 10,
           palette = "Set1"
) 
# f3

# Plot the survival curves
m4 <- ggsurvplot(survfit(males_fit, 
                   newdata = males_rp4),
           data = males_df,
           censor = F,
           conf.int = FALSE,
           # Min is 0.35 and max is 0.99
           # ylim = c(0.95, 1),
           xlim = c(0, 3650),
           ggtheme = theme_minimal(),
           legend.labs = c("Age 40", "Age 50", "Age 60"),
           legend = "none",
           title = "Risk Profile 2 Estimated Survival\nHigh BP & Diabetes",
           font.main = 10,
           palette = "Set1"
) 
# f4

# Plot the survival curves
m5 <- ggsurvplot(survfit(males_fit, 
                         newdata = males_rp5),
                 data = males_df,
                 censor = F,
                 conf.int = FALSE,
                 # Min is 0.35 and max is 0.99
                 # ylim = c(0.95, 1),
                 xlim = c(0, 3650),
                 ggtheme = theme_minimal(),
                 legend = "none",
                 title = "Risk Profile 2 Estimated Survival\nSmoker",
                 font.main = 10,
                 palette = "Set1"
) 
# f5

##### Display all plots together

library(ggpubr)
# Using ggarrange
plot2 <- ggarrange(m1$plot, m2$plot, m3$plot, m4$plot, m5$plot,
          ncol = 3, nrow = 2,
          common.legend = TRUE,
          legend = "top",
          legend.labs = c("Age", "Age 50", "Age 60"),
          vjust = 0.5
          ) 
# Add a note at the bottom of this figure
annotate_figure(plot2,
                top = text_grob("Estimated Survival Curves - Males",
                                size = 12,
                                face = "bold"),
                bottom = text_grob("**Y-axis is not the same for every plot",
                                   size = 10,
                                   hjust = 1.3,
                                   face = "italic"))
  

# With survival package - could not get one legend for all plots
# arrange_ggsurvplots(list(f1, f2, f3, f4, f5),
#                     ncol = 3, nrow = 2,
#                     title = "Estimated Survival Curves - males")
```




